### 2016-11-07 paper/main.tex
Added eurovis template

% ---------------------------------------------------------------------
% EG author guidelines plus sample file for EG publication using LaTeX2e input
% D.Fellner, v1.17, Sep 23, 2010

\title[EG \LaTeX\ Author Guidelines]%
      {\LaTeX\ Author Guidelines for EUROGRAPHICS Proceedings Manuscripts}

% for anonymous conference submission please enter your SUBMISSION ID
% instead of the author's name (and leave the affiliation blank) !!
\author[D. Fellner \& S. Behnke]
       {D.\,W. Fellner\thanks{Chairman Eurographics Publications Board}$^{1,2}$
        and S. Behnke$^{2}$
%        S. Spencer$^2$\thanks{Chairman Siggraph Publications Board}
        \\
% For Computer Graphics Forum: Please use the abbreviation of your first name.
         $^1$TU Darmstadt \& Fraunhofer IGD, Germany\\
         $^2$Institut f{\"u}r ComputerGraphik \& Wissensvisualisierung, TU Graz, Austria
%        $^2$ Another Department to illustrate the use in papers from authors
%             with different affiliations
       }

% ------------------------------------------------------------------------

% if the Editors-in-Chief have given you the data, you may uncomment
% the following five lines and insert it here
%
% \volume{27}   % the volume in which the issue will be published;
% \issue{1}     % the issue number of the publication
% \pStartPage{1}      % set starting page

%-------------------------------------------------------------------------
\begin{document}

% \teaser{
%  \includegraphics[width=\linewidth]{eg_new}
%  \centering
%   \caption{New EG Logo}
% \label{fig:teaser}
% }

\maketitle

\begin{abstract}
   The ABSTRACT is to be in fully-justified italicized text, 
   between two horizontal lines,
   in one-column format, 
   below the author and affiliation information. 
   Use the word ``Abstract'' as the title, in 9-point Times, boldface type, 
   left-aligned to the text, initially capitalized. 
   The abstract is to be in 9-point, single-spaced type.
   The abstract may be up to 3 inches (7.62 cm) long. \\
   Leave one blank line after the abstract, 
   then add the subject categories according to the ACM Classification Index 
   (see http://www.acm.org/class/1998/).

\begin{classification} % according to http://www.acm.org/class/1998/
\CCScat{Computer Graphics}{I.3.3}{Picture/Image Generation}{Line and curve generation}
\end{classification}

\end{abstract}

%-------------------------------------------------------------------------
\section{Introduction}

Please follow the steps outlined in this document very carefully when
submitting your manuscript to Eurographics.

You may as well use the \LaTeX\ source as a template to typeset your own
paper. In this case we encourage you to also read the \LaTeX\ comments
embedded in the document.

%-------------------------------------------------------------------------
\section{Instructions}

Please read the following carefully.

%-------------------------------------------------------------------------
\subsection{Language}

All manuscripts must be in English.

%-------------------------------------------------------------------------
\subsection{Margins and page numbering}

All printed material, including text, illustrations, and charts,
must be kept within a print area %6.31 inches (16.03 cm) 
7 inches (17.7 cm) wide by
%9.10 inches (23.13 cm) 
9.44 inches (24 cm) high. Do not write or print anything
outside the print area. Number your pages on odd sites right
above, on even sites left above, no page number on the first site.
Do not use page numbering within the final version of your paper.

%------------------------------------------------------------------------
\subsection{Formatting your paper}

% All text with the exception of the abstract must be in a two-column format.
% The total allowable width of the text area -- including header and footer
% lines -- is 161\,mm (6.34 inch) wide by 231\,mm (9.10 inch) high.
% 
% Columns are to be 76\,mm (3.0 inch) wide, with a 8\,mm (0.315 inch) space 
% between them.

All text with the exception of the abstract must be in a two-column format.
The total allowable width of the text area -- including header and footer
lines -- is 177\,mm (7 inch) wide by 245\,mm (9.64 inch) high.

Columns are to be 84\,mm (3.3 inch) wide, with a 8\,mm (0.315 inch) space 
between them.

The space between the header line and the first line of the text body and
between the last line of the text body and the footer line is 5\,mm
(0.196 inch) each.

%-------------------------------------------------------------------------
\subsection{Type-style and fonts}

Wherever Times is specified, Times Roman may also be used. If
neither is available on your word processor, please use the font
closest in appearance to Times that you have access to. Only
Type-1 fonts will be accepted.

MAIN TITLE. The title should be in Times 17-point, boldface type and
centered. Capitalize the first letter of nouns, pronouns, verbs, adjectives,
and adverbs; do not capitalize articles, coordinate conjunctions, or
prepositions (unless the title begins with such a word). Leave two blank
lines after the title.

AUTHOR NAME(s) and AFFILIATION(s) are to be centered beneath the title and
printed in Times 9-point, non-boldface type. This information is to be
followed by two blank lines.

The ABSTRACT is to be in a one-column format. The MAIN TEXT is to be in a
two-column format.

MAIN TEXT. Type main text in 9-point Times, single-spaced. Do \emph{not} use
double-spacing. All paragraphs should be indented 1 em (the length of the
dash in the actual font). Make sure your text is fully justified -- that is,
flush left and flush right. Please do not place any additional blank lines
between paragraphs. Figure and table captions should be 9-point Times
boldface type as in Figure~\ref{fig:firstExample}.

\noindent Long captions should be set as in Figure~\ref{fig:ex1} or
Figure~\ref{fig:ex3}.

\begin{figure}[htb]
   % an empty figure just consisting of the caption lines
   \caption{\label{fig:ex1}
     'Empty' figure only to serve as an example of long caption requiring 
     more than one line. It is not typed centered but aligned on both sides.}
\end{figure}

\noindent
Figures which need the full textwidth can be typeset as Figure~\ref{fig:ex3}.

\noindent Callouts should be 9-point Times, non-boldface type. Initially
capitalize only the first word of section titles and first-, second-, and
third-order headings.

FIRST-ORDER HEADINGS. (For example, \textbf{1. Introduction}) should be Times
9-point boldface, initially capitalized, flush left, with one blank line
before, and one blank line after.

SECOND-ORDER HEADINGS. (For example, \textbf{2.1. Language}) should be Times
9-point boldface, initially capitalized, flush left, with one blank line
before, and one after. If you require a third-order heading (we discourage
it), use 9-point Times, boldface, initially capitalized, flush left, preceded
by one blank line, followed by a period and your text on the same line.

The headline \emph{(authors / title)} must be shortened if it uses the full 
two column width of the main text.
There must be enough space for the page numbers. Please use ``et al.'' if 
there are more than three authors and specify a shortened version for your title.
%-------------------------------------------------------------------------
\subsection{Footnotes}

Please do \emph{not} use footnotes at all!

%-------------------------------------------------------------------------
\subsection{References}

List all bibliographical references in 9-point Times, single-spaced, at the
end of your paper in alphabetical order. When referenced in the text, enclose
the citation index in square brackets, for example~\cite{Lous90}. Where
appropriate, include the name(s) of editors of referenced books.

For your references please use the following algorithm:
\begin{itemize} 
\item \textbf{one} author: first 3 chars plus year -- 
      e.g.\ \cite{Lous90}
\item \textbf{two}, \textbf{three} or \textbf{four} authors: first char
      of each family name plus year --  e.g.\ \cite{Fellner-Helmberg93} 
      or \cite{Kobbelt97-USHDR} or \cite{Lafortune97-NARF}
\item \textbf{more than 4} authors: first char of family name from 
      first 3 authors followed by a '*' followed by the year -- 
      e.g.\ \cite{Buhmann:1998:DCQ} or \cite{FolDamFeiHug.etal93} 
\end{itemize}

For BibTeX users a style file \ \texttt{eg-alpha.bst} \ is available which
uses the above algorithm.

%-------------------------------------------------------------------------
\subsection{Illustrations, graphs, and photographs}

All graphics should be centered.

%%%
%%% Figure 1
%%%
\begin{figure}[htb]
  \centering
  % the following command controls the width of the embedded PS file
  % (relative to the width of the current column)
  \includegraphics[width=.8\linewidth]{sampleFig}
  % replacing the above command with the one below will explicitly set
  % the bounding box of the PS figure to the rectangle (xl,yl),(xh,yh).
  % It will also prevent LaTeX from reading the PS file to determine
  % the bounding box (i.e., it will speed up the compilation process)
  % \includegraphics[width=.95\linewidth, bb=39 696 126 756]{sampleFig}
  %
  \parbox[t]{.9\columnwidth}{\relax
           For all figures please keep in mind that you \textbf{must not}
           use images with transparent background! 
           }
  %
  \caption{\label{fig:firstExample}
           Here is a sample figure.}
\end{figure}

If your paper includes images, it is very important that they are of
sufficient resolution to be faithfully reproduced.

To determine the optimum size (width and height) of an image, measure
the image's size as it appears in your document (in millimeters), and
then multiply those two values by 12. The resulting values are the
optimum $x$ and $y$ resolution, in pixels, of the image. Image quality
will suffer if these guidelines are not followed.

Example 1: 
%
An image measures 50\,mm by 75\,mm when placed in a document. This
image should have a resolution of no less than 600 pixels by 900
pixels in order to be reproduced faithfully.

Example 2: 
%
Capturing a screenshot of your entire $1024 \times 768$ pixel display
monitor may be useful in illustrating a concept from your research. In
order to be reproduced faithfully, that $1024 \times 768$ image should
be no larger than 85 mm by 64 mm (approximately) when placed in your
document.

%-------------------------------------------------------------------------
\subsection{Color}

\textbf{Please observe:} as of 2003 publications in the proceedings of the
Eurographics Conference can use color images throughout the paper. No
separate color tables are necessary.

However, workshop proceedings might have different agreements! 
Figure~\ref{fig:ex3} is an example for creating color plates.

%------------------------------------------------------------------------
\subsection{Embedding of Hyperlinks / Typesetting of URLs}

Due to the use of the package \texttt{hyperref} the original behavior
of the command $\backslash$\texttt{url} from the package \texttt{url}
is not available. To circumvent this problem we either recommend to
use the command $\backslash$\texttt{httpAddr} from the 
included package \texttt{egweblnk} (see below) or to replace the
command $\backslash$\texttt{url} by the command $\backslash$\texttt{webLink} 
-- e.g. in cases where $\backslash$\texttt{url} has been used
widely in BibTeX-References. In the latter case we suggest to run
BibTeX as usual and then replace all occurences of $\backslash$\texttt{url}  by
$\backslash$\texttt{webLink}

\noindent
The provided commands for hyperlinks, in a nutshell, are:

\begin{description} \itemsep 1ex
\item [\webLinkFont $\backslash$httpAddr \{URL without leading 'http:'\}]
      \mbox{}\\
      e.g. \  \httpAddr{//diglib.eg.org/EG/DL/WS}

\item[\webLinkFont $\backslash$ftpAddr \{URL without leading 'ftp:'\}]
      \mbox{}\\
      e.g. \  \ftpAddr{//www.eg.org/EG/DL/ftpupload}

\item[\webLinkFont $\backslash$URL \{url\}]
      \mbox{}\\
      e.g. \  \URL{http://www.eg.org/EG/DL/WS}

\item[\webLinkFont $\backslash$MailTo \{Email addr\}]
      \mbox{}\\
      e.g. \  \MailTo{publishing@eg.org}

\item[\webLinkFont $\backslash$MailToNA \{emailName\}\{@emailSiteAddress\}]
      \mbox{}\\
      e.g. \  \MailToNA{publishing}{@eg.org}

\item[\webLinkFont $\backslash$webLink\{URL without hyperlink creation\}]
      \mbox{}\\
      e.g. \  \webLink{http://www.eg.org/some_arbitrary_long/but_useless/URL}

\end{description}

%------------------------------------------------------------------------
\subsection{PDF Generation}

Your final paper should be delivered as a PDF document with all typefaces
embedded. \LaTeX{} users should use \texttt{dvips} and \texttt{ps2pdf} to
create this PDF document. Adobe Acrobat Distiller may be used in place of
\texttt{ps2pdf}.

Adobe PDFWriter is \emph{not} acceptable for use. Documents created with
PDFWriter will be returned to the author for revision. \texttt{pdftex} and
\texttt{pdflatex} (and its variants) can be used only if the author can
make certain that all typefaces are embedded and images are not downsampled
or subsampled during the PDF creation process.

Users with no access to these PDF creation tools should make available a
PostScript file and we will make a PDF document from it.

The PDF file \emph{must not} be change protected.

%------------------------------------------------------------------------
\subsubsection*{Configuration Notes: dvips / ps2pdf / etc.}

\noindent
\texttt{dvips} should be invoked with the \texttt{-Ppdf} and \texttt{-G0}
flags in order to use Type 1 PostScript typefaces:

\begin{verbatim}
    dvips -t a4 -Ppdf -G0 -o my.ps my.dvi
\end{verbatim}

\noindent
If you are using version 7.x of GhostScript, please use the following method of invoking \texttt{ps2pdf}, in
order to embed all typefaces and ensure that images are not downsampled or subsampled in the PDF
creation process:

\begin{verbatim}
  ps2pdf -dMaxSubsetPct=100 \
         -dCompatibilityLevel=1.3 \
         -dSubsetFonts=true \
         -dEmbedAllFonts=true \
         -dAutoFilterColorImages=false \
         -dAutoFilterGrayImages=false \
         -dColorImageFilter=/FlateEncode \
         -dGrayImageFilter=/FlateEncode \
         -dMonoImageFilter=/FlateEncode \
         mypaper.ps mypaper.pdf
\end{verbatim}

If you are using version 8.x of GhostScript, please use this method in place of the example above:
\begin{verbatim}
  ps2pdf -dPDFSETTINGS=/prepress \
         -dCompatibilityLevel=1.3 \
         -dAutoFilterColorImages=false \
         -dAutoFilterGrayImages=false \
         -dColorImageFilter=/FlateEncode \
         -dGrayImageFilter=/FlateEncode \
         -dMonoImageFilter=/FlateEncode \
         -dDownsampleColorImages=false \
         -dDownsampleGrayImages=false \
         mypaper.ps mypaper.pdf
\end{verbatim}

%------------------------------------------------------------------------
\subsubsection*{Configuration Notes: pdftex / pdflatex / etc.}

\noindent
Configuration of these tools to embed all typefaces can be accomplished by editing the \texttt{updmap.cfg} file
to enable inclusion of the standard (or base) 14 typefaces.

Linux users can run the \texttt{updmap} script to do this:
\begin{verbatim}
updmap --setoption pdftexDownloadBase14 true
\end{verbatim}

Windows users should edit the \texttt{updmap.cfg} files found in their TeX installation directories (one or both
of the following may be present):
\begin{verbatim}
  INSTALLDIR\texmf\web2c\updmap.cfg
  INSTALLDIR\localtexmf\miktex\config\updmap.cfg
\end{verbatim}

Ensure the value for \texttt{pdftexDownloadBase14} is "true," and then follow the instructions found here:
\httpAddr{//docs.miktex.org/manual/} to update your MikTeX installation.

%------------------------------------------------------------------------
\subsubsection*{Configuration Notes: Acrobat Distiller}

We recommend to download and install the version of the ``CMW'' Adobe Acrobat Distiller job options file
appropriate for your operating system and version of Acrobat from the following URL:

\httpAddr{//www.cadmusmediaworks.com/index2.html}\\
in the ``(Operating System)/Applications/Distiller Settings'' folder. The ``CMW'' job options file embeds
all typefaces and does not downsample or subsample images when creating the PDF document.
%------------------------------------------------------------------------
\subsection{Copyright forms}

You must include your signed Eurographics copyright release form
when you submit your finished paper. We MUST have this form before
your paper can be published in the proceedings.

%-------------------------------------------------------------------------
\subsection{Conclusions}

Please direct any questions to the production editor in charge of
these proceedings.

%-------------------------------------------------------------------------

%\bibliographystyle{eg-alpha}
\bibliographystyle{eg-alpha-doi}

\bibliography{egbibsample}

%-------------------------------------------------------------------------
\newpage

\begin{figure*}[tbp]
  \centering
  \mbox{} \hfill
  % the following command controls the width of the embedded PS file
  % (relative to the width of the current column)
  \includegraphics[width=.3\linewidth]{sampleFig}
  % replacing the above command with the one below will explicitly set
  % the bounding box of the PS figure to the rectangle (xl,yl),(xh,yh).
  % It will also prevent LaTeX from reading the PS file to determine
  % the bounding box (i.e., it will speed up the compilation process)
  % \includegraphics[width=.3\linewidth, bb=39 696 126 756]{sampleFig}
  \hfill
  \includegraphics[width=.3\linewidth]{sampleFig}
  \hfill \mbox{}
  \caption{\label{fig:ex3}%
           For publications with color tables (i.e., publications not offering
           color throughout the paper) please \textbf{observe}: 
           for the printed version -- and ONLY for the printed
           version -- color figures have to be placed in the last page.
           \newline
           For the electronic version, which will be converted to PDF before
           making it available electronically, the color images should be
           embedded within the document. Optionally, other multimedia
           material may be attached to the electronic version. }
\end{figure*}

\end{document}

### 2017-03-03 heatmap/heatmap.js
Switching to d3

@@ -1,34 +1,60 @@
-function preload(){
+var map = colorbrewer.YlGnBu[9];
 
+var width = 300;
+var height = 300;
+
+var x = d3.scale.linear().range([0,width]);
+var y = d3.scale.linear().range([0,height]);
+var z = d3.scale.quantize().range(map);
+
+var data = [
+  [1,2,3],
+  [4,5,6],
+  [7,8,9]
+];
+
+var uncertainty = [
+  [1,1,1],
+  [1,1,1],
+  [1,1,1]
+];
+
+x.domain([0,data[0].length]);
+y.domain([0,data.length]);
+z.domain([d3.min(d3.min(data)),d3.max(d3.max(data))]);
+
+var w = width/data[0].length;
+var h = height/data.length;
+
+var svg = d3.select("body").append("svg")
+  .attr("width",width)
+  .attr("height",height);
+
+var cells = svg.selectAll("g")
+    .data(data)
+    .enter()
+    .append("g")
+    .selectAll("rect")
+      .data(function(d,i){ return d.map(function(val){ return {r:i, v:val};});})
+      .enter()
+      .append("rect")
+      .datum(function(d,i){ d.c = i; d.u = uncertainty[d.r][d.c]; return d; })
+      .attr("x",function(d){ return x(d.c);})
+      .attr("y",function(d){ return y(d.r);})
+      .attr("width",w)
+      .attr("height",h)
+      .attr("fill",function(d){ return z(d.v);});
+
+
+function lerp(v1,v2,amount){
+       amount = Math.min(Math.max(amount,0),1);
+       return v1 + (v2 - v1) * amount;
 }
 
-function setup(){
-  pixelDensity(displayDensity());
-  createCanvas(500,500);
-  background(255);
-  drawHeatmap([0,0.25,0.5,0.75,1.0,0.1,0.9,0.3]);
-  noLoop();
-}
+function labLerp(c1,c2,amount){
+  //Linearly interpolates two colors in CIELAB space
+  var lc1 = d3.lab(c1);
+  var lc2 = d3.lab(c2);
+  return d3.lab( lerp(lc1.l,lc2.l,amount), lerp(lc1.a,lc2.a,amount), lerp(lc1.b,lc2.b,amount));
 
-function drawHeatmap(samples){
-  var delta, index, perRow, curY;
-  perRow = ceil(sqrt(samples.length));
[…]

### 2017-03-03 heatmap/heatmap.js
heatmap generation now modular

@@ -1,50 +1,35 @@
-var map = colorbrewer.YlGnBu[9];
 
-var width = 300;
-var height = 300;
+var svg = d3.select("body").append("svg");
 
-var x = d3.scale.linear().range([0,width]);
-var y = d3.scale.linear().range([0,height]);
-var z = d3.scale.quantize().range(map);
+function makeHeatmap(x,y,size,data,z){
+  //creates an svg heatmap with a 2d matrix of data, and a mapping function z
+  var w = size/data[0].length;
+  var h = size/data.length;
 
-var data = [
-  [1,2,3],
-  [4,5,6],
-  [7,8,9]
-];
+  var heatmap = svg.append("g")
+             .attr("transform","translate("+x+","+y+")");
 
-var uncertainty = [
-  [1,1,1],
-  [1,1,1],
-  [1,1,1]
-];
-
-x.domain([0,data[0].length]);
-y.domain([0,data.length]);
-z.domain([d3.min(d3.min(data)),d3.max(d3.max(data))]);
-
-var w = width/data[0].length;
-var h = height/data.length;
-
-var svg = d3.select("body").append("svg")
-  .attr("width",width)
-  .attr("height",height);
-
-var cells = svg.selectAll("g")
-    .data(data)
+  heatmap.selectAll("g")
+  .data(data)
+  .enter()
+  .append("g")
+  .selectAll("rect")
+    .data(function(d,i){ return d.map(function(val){ return {r:i, v:val};});})
     .enter()
-    .append("g")
-    .selectAll("rect")
-      .data(function(d,i){ return d.map(function(val){ return {r:i, v:val};});})
-      .enter()
-      .append("rect")
-      .datum(function(d,i){ d.c = i; d.u = uncertainty[d.r][d.c]; return d; })
-      .attr("x",function(d){ return x(d.c);})
-      .attr("y",function(d){ return y(d.r);})
-      .attr("width",w)
-      .attr("height",h)
-      .attr("fill",function(d){ return z(d.v);});
+    .append("rect")
+    .datum(function(d,i){ d.c = i; return d; })
+    .attr("x", function(d){ return d.c*w;})
+    .attr("y", function(d){ return d.r*h;})
+    .attr("width",w)
+    .attr("height",h)
+    .attr("fill", function(d){ return z(d);});
+}
 
[…]

### 2017-03-03 heatmap/heatmap.js
Heatmaps can be id'd.

@@ -1,13 +1,17 @@
 
 var svg = d3.select("body").append("svg");
 
-function makeHeatmap(x,y,size,data,z){
+function makeHeatmap(x,y,size,data,z,name){
   //creates an svg heatmap with a 2d matrix of data, and a mapping function z
   var w = size/data[0].length;
   var h = size/data.length;
 
   var heatmap = svg.append("g")
              .attr("transform","translate("+x+","+y+")");
+
+  if(name){
+    heatmap.attr("id",name);
+  }
 
   heatmap.selectAll("g")
   .data(data)
@@ -74,8 +78,8 @@ function main(){
   z.domain([d3.min(d3.min(data)),d3.max(d3.max(data))]);
   u.domain([0,1]);
 
-  makeHeatmap(0,0,250,data,function(d){ return z(d.v);});
-  makeHeatmap(300,0,250,uncertainty,function(d){ return u(d.v);});
+  makeHeatmap(0,0,250,data,function(d){ return z(d.v);},"value");
+  makeHeatmap(300,0,250,uncertainty,function(d){ return u(d.v);},"uncertainty");
 }
 
 main();

### 2017-03-03 paper/template.tex
switch to VIS 2017 format

% $Id: template.tex 11 2007-04-03 22:25:53Z jpeltier $

\documentclass{vgtc}                          % final (conference style)
%\documentclass[review]{vgtc}                 % review
%\documentclass[widereview]{vgtc}             % wide-spaced review
%\documentclass[preprint]{vgtc}               % preprint
%\documentclass[electronic]{vgtc}             % electronic version

%% Uncomment one of the lines above depending on where your paper is
%% in the conference process. ``review'' and ``widereview'' are for review
%% submission, ``preprint'' is for pre-publication, and the final version
%% doesn't use a specific qualifier. Further, ``electronic'' includes
%% hyperreferences for more convenient online viewing.

%% Please use one of the ``review'' options in combination with the
%% assigned online id (see below) ONLY if your paper uses a double blind
%% review process. Some conferences, like IEEE Vis and InfoVis, have NOT
%% in the past.

%% Figures should be in CMYK or Grey scale format, otherwise, colour 
%% shifting may occur during the printing process.

%% These few lines make a distinction between latex and pdflatex calls and they
%% bring in essential packages for graphics and font handling.
%% Note that due to the \DeclareGraphicsExtensions{} call it is no longer necessary
%% to provide the the path and extension of a graphics file:
%% \includegraphics{diamondrule} is completely sufficient.
%%
\ifpdf%                                % if we use pdflatex
  \pdfoutput=1\relax                   % create PDFs from pdfLaTeX
  \pdfcompresslevel=9                  % PDF Compression
  \pdfoptionpdfminorversion=7          % create PDF 1.7
  \ExecuteOptions{pdftex}
  \usepackage{graphicx}                % allow us to embed graphics files
  \DeclareGraphicsExtensions{.pdf,.png,.jpg,.jpeg} % for pdflatex we expect .pdf, .png, or .jpg files
\else%                                 % else we use pure latex
  \ExecuteOptions{dvips}
  \usepackage{graphicx}                % allow us to embed graphics files
  \DeclareGraphicsExtensions{.eps}     % for pure latex we expect eps files
\fi%

%% it is recomended to use ``\autoref{sec:bla}'' instead of ``Fig.~\ref{sec:bla}''
\graphicspath{{figures/}{pictures/}{images/}{./}} % where to search for the images

\usepackage{microtype}                 % use micro-typography (slightly more compact, better to read)
\PassOptionsToPackage{warn}{textcomp}  % to address font issues with \textrightarrow
\usepackage{textcomp}                  % use better special symbols
\usepackage{mathptmx}                  % use matching math font
\usepackage{times}                     % we use Times as the main font
\renewcommand*\ttdefault{txtt}         % a nicer typewriter font
\usepackage{cite}                      % needed to automatically sort the references
\usepackage{tabu}                      % only used for the table example
\usepackage{booktabs}                  % only used for the table example
%% We encourage the use of mathptmx for consistent usage of times font
%% throughout the proceedings. However, if you encounter conflicts
%% with other math-related packages, you may want to disable it.

%% If you are submitting a paper to a conference for review with a double
%% blind reviewing process, please replace the value ``0'' below with your
%% OnlineID. Otherwise, you may safely leave it at ``0''.
\onlineid{0}

%% declare the category of your paper, only shown in review mode
\vgtccategory{Research}

%% allow for this line if you want the electronic option to work properly
\vgtcinsertpkg

%% In preprint mode you may define your own headline.
%\preprinttext{To appear in an IEEE VGTC sponsored conference.}

%% Paper title.

\title{Global Illumination for Fun and Profit}

%% This is how authors are specified in the conference style

%% Author and Affiliation (single author).
%%\author{Roy G. Biv\thanks{e-mail: roy.g.biv@aol.com}}
%%\affiliation{\scriptsize Allied Widgets Research}

%% Author and Affiliation (multiple authors with single affiliations).
%%\author{Roy G. Biv\thanks{e-mail: roy.g.biv@aol.com} %
%%\and Ed Grimley\thanks{e-mail:ed.grimley@aol.com} %
%%\and Martha Stewart\thanks{e-mail:martha.stewart@marthastewart.com}}
%%\affiliation{\scriptsize Martha Stewart Enterprises \\ Microsoft Research}

%% Author and Affiliation (multiple authors with multiple affiliations)
\author{Roy G. Biv\thanks{e-mail: roy.g.biv@aol.com}\\ %
        \scriptsize Starbucks Research %
\and Ed Grimley\thanks{e-mail:ed.grimley@aol.com}\\ %
     \scriptsize Grimley Widgets, Inc. %
\and Martha Stewart\thanks{e-mail:martha.stewart@marthastewart.com}\\ %
     \parbox{1.4in}{\scriptsize \centering Martha Stewart Enterprises \\ Microsoft Research}}

%% A teaser figure can be included as follows, but is not recommended since
%% the space is now taken up by a full width abstract.
%\teaser{
%  \includegraphics[width=1.5in]{sample.eps}
%  \caption{Lookit! Lookit!}
%}

%% Abstract section.
\abstract{Duis autem vel eum iriure dolor in hendrerit in vulputate
velit esse molestie consequat, vel illum dolore eu feugiat nulla
facilisis at vero eros et accumsan et iusto odio dignissim qui blandit
praesent luptatum zzril delenit augue duis dolore te feugait nulla
facilisi. Lorem ipsum dolor sit amet, consectetuer adipiscing elit,
sed diam nonummy nibh euismod tincidunt ut laoreet dolore magna
aliquam erat volutpat. Ut wisi enim ad minim veniam, quis nostrud exerci tation ullamcorper
suscipit lobortis nisl ut aliquip ex ea commodo consequat. Duis autem
vel eum iriure dolor in hendrerit in vulputate velit esse molestie
consequat, vel illum dolore eu feugiat nulla facilisis at vero eros et
accumsan et iusto odio dignissim qui blandit praesent luptatum zzril
delenit augue duis dolore te feugait nulla facilisi.%
} % end of abstract

%% ACM Computing Classification System (CCS). 
%% See <http://www.acm.org/class/1998/> for details.
%% The ``\CCScat'' command takes four arguments.

\CCScatlist{ 
  \CCScat{K.6.1}{Management of Computing and Information Systems}%
{Project and People Management}{Life Cycle};
  \CCScat{K.7.m}{The Computing Profession}{Miscellaneous}{Ethics}
}

%% Copyright space is enabled by default as required by guidelines.
%% It is disabled by the 'review' option or via the following command:
% \nocopyrightspace

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%% START OF THE PAPER %%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

%% The ``\maketitle'' command must be the first command after the
%% ``\begin{document}'' command. It prepares and prints the title block.

%% the only exception to this rule is the \firstsection command
\firstsection{Introduction}

\maketitle

%% \section{Introduction} %for journal use above \firstsection{..} instead
This template is for papers of VGTC-sponsored conferences which are \emph{\textbf{not}} published in a special issue of TVCG.

\section{Using the Style Template}

\begin{itemize}
\item The style uses the hyperref package, thus turns references into internal links. We thus recommend to make use of the ``\texttt{\textbackslash autoref\{reference\}}'' call (instead of ``\texttt{Figure\~{}\textbackslash ref\{reference\}}'' or similar) since ``\texttt{\textbackslash autoref\{reference\}}'' turns the entire reference into an internal link, not just the number. Examples: \autoref{fig:sample} and \autoref{tab:vis_papers}.
\item The style automatically looks for image files with the correct extension (eps for regular \LaTeX; pdf, png, and jpg for pdf\LaTeX), in a set of given subfolders (figures/, pictures/, images/). It is thus sufficient to use ``\texttt{\textbackslash includegraphics\{CypressView\}}'' (instead of ``\texttt{\textbackslash includegraphics\{pictures/CypressView.jpg\}}'').
\item For adding hyperlinks and DOIs to the list of references, you can use ``\texttt{\textbackslash bibliographystyle\{abbrv-doi-hyperref-narrow\}}'' (instead of ``\texttt{\textbackslash bibliographystyle\{abbrv\}}''). It uses the doi and url fields in a bib\TeX\ entry and turns the entire reference into a link, giving priority to the doi. The doi can be entered with or without the ``\texttt{http://dx.doi.org/}'' url part. See the examples in the bib\TeX\ file and the bibliography at the end of this template.\\[1em]
\textbf{Note 1:} occasionally (for some \LaTeX\ distributions) this hyper-linked bib\TeX\ style may lead to \textbf{compilation errors} (``\texttt{pdfendlink ended up in different nesting level ...}'') if a reference entry is broken across two pages (due to a bug in hyperref). In this case make sure you have the latest version of the hyperref package (i.\,e., update your \LaTeX\ installation/packages) or, alternatively, revert back to ``\texttt{\textbackslash bibliographystyle\{abbrv-doi-narrow\}}'' (at the expense of removing hyperlinks from the bibliography) and try ``\texttt{\textbackslash bibliographystyle\{abbrv-doi-hyperref-narrow\}}'' again after some more editing.\\[1em]
\textbf{Note 2:} the ``\texttt{-narrow}'' versions of the bibliography style use the font ``PTSansNarrow-TLF'' for typesetting the DOIs in a compact way. This font needs to be available on your \LaTeX\ system. It is part of the \href{https://www.ctan.org/pkg/paratype}{``paratype'' package}, and many distributions (such as MikTeX) have it automatically installed. If you do not have this package yet and want to use a ``\texttt{-narrow}'' bibliography style then use your \LaTeX\ system's package installer to add it. If this is not possible you can also revert to the respective bibliography styles without the ``\texttt{-narrow}'' in the file name.\\[1em]
DVI-based processes to compile the template apparently cannot handle the different font so, by default, the template file uses the \texttt{abbrv-doi} bibliography style but the compiled PDF shows you the effect of the \texttt{abbrv-doi-hyperref-narrow} style.
\end{itemize}

\section{Bibliography Instructions}

\begin{itemize}
\item Sort all bibliographic entries alphabetically but the last name of the first author. This \LaTeX/bib\TeX\ template takes care of this sorting automatically.
\item Merge multiple references into one; e.\,g., use \cite{Max:1995:OMF,Kitware:2003} (not \cite{Kitware:2003}\cite{Max:1995:OMF}). Within each set of multiple references, the references should be sorted in ascending order. This \LaTeX/bib\TeX\ template takes care of both the merging and the sorting automatically.
\item Verify all data obtained from digital libraries, even ACM's DL and IEEE Xplore  etc.\ are sometimes wrong or incomplete.
\item Do not trust bibliographic data from other services such as Mendeley.com, Google Scholar, or similar; these are even more likely to be incorrect or incomplete.
\item Articles in journal---items to include:
  \begin{itemize}
  \item author names
	\item title
	\item journal name
	\item year
	\item volume
	\item number
	\item month of publication as variable name (i.\,e., \{jan\} for January, etc.; month ranges using \{jan \#\{/\}\# feb\} or \{jan \#\{-{}-\}\# feb\})
  \end{itemize}
\item use journal names in proper style: correct: ``IEEE Transactions on Visualization and Computer Graphics'', incorrect: ``Visualization and Computer Graphics, IEEE Transactions on''
\item Papers in proceedings---items to include:
  \begin{itemize}
  \item author names
	\item title
	\item abbreviated proceedings name: e.\,g., ``Proc.\textbackslash{} CONF\_ACRONYNM'' without the year; example: ``Proc.\textbackslash{} CHI'', ``Proc.\textbackslash{} 3DUI'', ``Proc.\textbackslash{} Eurographics'', ``Proc.\textbackslash{} EuroVis''
	\item year
	\item publisher
	\item town with country of publisher (the town can be abbreviated for well-known towns such as New York or Berlin)
  \end{itemize}
\item article/paper title convention: refrain from using curly brackets, except for acronyms/proper names/words following dashes/question marks etc.; example:
\begin{itemize}
	\item paper ``Marching Cubes: A High Resolution 3D Surface Construction Algorithm''
	\item should be entered as ``\{M\}arching \{C\}ubes: A High Resolution \{3D\} Surface Construction Algorithm'' or  ``\{M\}arching \{C\}ubes: A high resolution \{3D\} surface construction algorithm''
	\item will be typeset as ``Marching Cubes: A high resolution 3D surface construction algorithm''
\end{itemize}
\item for all entries
\begin{itemize}
	\item DOI can be entered in the DOI field as plain DOI number or as DOI url; alternative: a url in the URL field
	\item provide full page ranges AA-{}-BB
\end{itemize}
\item when citing references, do not use the reference as a sentence object; e.\,g., wrong: ``In \cite{Lorensen:1987:MCA} the authors describe \dots'', correct: ``Lorensen and Cline \cite{Lorensen:1987:MCA} describe \dots''
\end{itemize}

\section{Example Section}

Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
dolore magna aliquyam erat, sed diam
voluptua~\cite{Kitware:2003,Max:1995:OMF}. At vero eos et accusam et
justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit
amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor
invidunt ut labore et dolore magna aliquyam erat, sed diam
voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est.

\section{Exposition}

Duis autem vel eum iriure dolor in hendrerit in vulputate velit esse
molestie consequat, vel illum dolore eu feugiat nulla facilisis at
vero eros et accumsan et iusto odio dignissim qui blandit praesent
luptatum zzril delenit augue duis dolore te feugait nulla
facilisi. Lorem ipsum dolor sit amet, consectetuer adipiscing elit,
sed diam nonummy nibh euismod tincidunt ut laoreet dolore magna
aliquam erat volutpat~\cite{Kindlmann:1999:SAG}.

\begin{equation}
\sum_{j=1}^{z} j = \frac{z(z+1)}{2}
\end{equation}

Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
takimata sanctus est Lorem ipsum dolor sit amet.

\subsection{Lorem ipsum}

Lorem ipsum dolor sit amet (see \autoref{tab:vis_papers}), consetetur sadipscing elitr, sed diam
nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit
amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor
invidunt ut labore et dolore magna aliquyam erat, sed diam
voluptua. At vero eos et accusam et justo duo dolores et ea
rebum.

\begin{table}[tb]
  \caption{VIS/VisWeek accepted/presented papers: 1990--2015.}
  \label{tab:vis_papers}
  \scriptsize%
	\centering%
  \begin{tabu}{%
	r%
	*{7}{c}%
	*{2}{r}%
	}
  \toprule
   year & \rotatebox{90}{Vis/SciVis} &   \rotatebox{90}{SciVis conf} &   \rotatebox{90}{InfoVis} &   \rotatebox{90}{VAST} &   \rotatebox{90}{VAST conf} &   \rotatebox{90}{TVCG @ VIS} &   \rotatebox{90}{CG\&A @ VIS} &   \rotatebox{90}{VIS/VisWeek} \rotatebox{90}{incl. TVCG/CG\&A}   &   \rotatebox{90}{VIS/VisWeek} \rotatebox{90}{w/o TVCG/CG\&A}   \\
  \midrule
  2015 & 33 & 9 & 38 & 33 & 14 & 17 & 15 & 159 & 127 \\
  2014 & 34 &   & 45 & 33 & 21 & 20 &   & 153 & 133 \\
  2013 & 31 &   & 38 & 32 &   & 20 &   & 121 & 101 \\
  2012 & 42 &   & 44 & 30 &   & 23 &   & 139 & 116 \\
  2011 & 49 &   & 44 & 26 &   & 20 &   & 139 & 119 \\
  2010 & 48 &   & 35 & 26 &   &   &   & 109 & 109 \\
  2009 & 54 &   & 37 & 26 &   &   &   & 117 & 117 \\
  2008 & 50 &   & 28 & 21 &   &   &   & 99 & 99 \\
  2007 & 56 &   & 27 & 24 &   &   &   & 107 & 107 \\
  2006 & 63 &   & 24 & 26 &   &   &   & 113 & 113 \\
  2005 & 88 &   & 31 &   &   &   &   & 119 & 119 \\
  2004 & 70 &   & 27 &   &   &   &   & 97 & 97 \\
  2003 & 74 &   & 29 &   &   &   &   & 103 & 103 \\
  2002 & 78 &   & 23 &   &   &   &   & 101 & 101 \\
  2001 & 74 &   & 22 &   &   &   &   & 96 & 96 \\
  2000 & 73 &   & 20 &   &   &   &   & 93 & 93 \\
  1999 & 69 &   & 19 &   &   &   &   & 88 & 88 \\
  1998 & 72 &   & 18 &   &   &   &   & 90 & 90 \\
  1997 & 72 &   & 16 &   &   &   &   & 88 & 88 \\
  1996 & 65 &   & 12 &   &   &   &   & 77 & 77 \\
  1995 & 56 &   & 18 &   &   &   &   & 74 & 74 \\
  1994 & 53 &   &   &   &   &   &   & 53 & 53 \\
  1993 & 55 &   &   &   &   &   &   & 55 & 55 \\
  1992 & 53 &   &   &   &   &   &   & 53 & 53 \\
  1991 & 50 &   &   &   &   &   &   & 50 & 50 \\
  1990 & 53 &   &   &   &   &   &   & 53 & 53 \\
  \midrule
  \textbf{sum} & \textbf{1515} & \textbf{9} & \textbf{595} & \textbf{277} & \textbf{35} & \textbf{100} & \textbf{15} & \textbf{2546} & \textbf{2431} \\
  \bottomrule
  \end{tabu}%
\end{table}

\subsection{Mezcal Head}

Lorem ipsum dolor sit amet (see \autoref{fig:sample}), consetetur sadipscing elitr, sed diam
nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
takimata sanctus est Lorem ipsum dolor sit amet.

\subsubsection{Duis Autem}

Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit
amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor
invidunt ut labore et dolore magna aliquyam erat, sed diam
voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est. Lorem
ipsum dolor sit amet.

\begin{figure}[tb]
 \centering % avoid the use of \begin{center}...\end{center} and use \centering instead (more compact)
 \includegraphics[width=\columnwidth]{paper-count-w-2015-new}
 \caption{A visualization of the data from \autoref{tab:vis_papers}. The image is from \cite{Isenberg:2017:VMC} and is in the public domain.}
 \label{fig:sample}
\end{figure}

\subsubsection{Ejector Seat Reservation}

Duis autem~\cite{Lorensen:1987:MCA}\footnote{The algorithm behind
Marching Cubes \cite{Lorensen:1987:MCA} had already been
described by Wyvill et al. \cite{Wyvill:1986:DSS} a year
earlier.} vel eum iriure dolor in hendrerit
in vulputate velit esse molestie consequat,\footnote{Footnotes
appear at the bottom of the column.} vel illum dolore eu
feugiat nulla facilisis at vero eros et accumsan et iusto odio
dignissim qui blandit praesent luptatum zzril delenit augue duis
dolore te feugait nulla facilisi. Lorem ipsum dolor sit amet,
consectetuer adipiscing elit, sed diam nonummy nibh euismod tincidunt
ut laoreet dolore magna aliquam erat volutpat.

\paragraph{Confirmed Ejector Seat Reservation}

Ut wisi enim ad minim veniam, quis nostrud exerci tation ullamcorper
suscipit lobortis nisl ut aliquip ex ea commodo
consequat~\cite{Nielson:1991:TAD}. Duis autem vel eum iriure dolor in
hendrerit in vulputate velit esse molestie consequat, vel illum dolore
eu feugiat nulla facilisis at vero eros et accumsan et iusto odio
dignissim qui blandit praesent luptatum zzril delenit augue duis
dolore te feugait nulla facilisi.

\paragraph{Rejected Ejector Seat Reservation}

Ut wisi enim ad minim veniam, quis nostrud exerci tation ullamcorper
suscipit lobortis nisl ut aliquip ex ea commodo consequat. Duis autem
vel eum iriure dolor in hendrerit in vulputate velit esse molestie

\section{Conclusion}

Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit
amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor
invidunt ut labore et dolore magna aliquyam erat, sed diam
voluptua. At vero eos et accusam et justo duo dolores et ea
rebum.

%% if specified like this the section will be committed in review mode
\acknowledgments{
The authors wish to thank A, B, C. This work was supported in part by
a grant from XYZ.}

%\bibliographystyle{abbrv}
\bibliographystyle{abbrv-doi}
%\bibliographystyle{abbrv-doi-narrow}
%\bibliographystyle{abbrv-doi-hyperref}
%\bibliographystyle{abbrv-doi-hyperref-narrow}

\bibliography{template}
\end{document}

### 2017-03-04 heatmap/heatmap.js
Added 1.5D heatmaps

@@ -29,10 +29,43 @@ function makeHeatmap(x,y,size,data,z,name){
     .attr("fill", function(d){ return z(d);});
 }
 
-function makeArcmap(x,y,size,data,z){
+function makeArcmap(x,y,size,data,z,name){
   //creates an svg "wedge" map
-  //TODO finish arcmap function
+  var w = size/data[0].length;
+  var h = size/data.length;
 
+  var arcmap = svg.append("g")
+             .attr("transform","translate("+(x+size/2)+","+(y+size)+")");
+
+  if(name){
+    arcmap.attr("id",name);
+  }
+
+  arcmap.selectAll("g")
+  .data(data)
+  .enter()
+  .append("g")
+  .selectAll("path")
+    .data(function(d,i){ return d.map(function(val){ return {r:i, v:val};});})
+    .enter()
+    .append("path")
+    .datum(function(d,i){ d.c = i; return d; })
+    .attr("d", function(d){ return makeArc(d,size,data.length,data[d.r].length)();})
+    .attr("fill", function(d){ return z(d);});
+
+}
+
+function makeArc(d,size,rows,cols){
+  var angle = d3.scale.linear().domain([0,cols]).range([-Math.PI/6,Math.PI/6]);
+  var radius = d3.scale.linear().domain([0,rows]).range([size,0]);
+
+  var arc = d3.svg.arc()
+  .innerRadius(radius(d.r+1))
+  .outerRadius(radius(d.r))
+  .startAngle(angle(d.c))
+  .endAngle(angle(d.c+1));
+
+  return arc;
 }
 
 function lerp(v1,v2,amount){
@@ -75,11 +108,21 @@ function main(){
     [0.6,0.7,0.8]
   ];
 
+  var arcData = [
+    [1,2,3],
+    [4,5],
+    [6]
+  ];
+
   z.domain([d3.min(d3.min(data)),d3.max(d3.max(data))]);
   u.domain([0,1]);
 
   makeHeatmap(0,0,250,data,function(d){ return z(d.v);},"value");
   makeHeatmap(300,0,250,uncertainty,function(d){ return u(d.v);},"uncertainty");
+
+  makeArcmap(0,300,250,arcData,function(d){ return z(d.v);},"arc");
+
+  makeArcmap(300,300,250,arcData,function(d){ return u((d.v)/6.0);},"arcUncertainty");
 }
 
 main();

### 2017-03-07 paper/template.tex
Adding functions for colormap verification, references from google doc.
Bare bones version for plane trip hacking.

@@
 %% Author and Affiliation (multiple authors with multiple affiliations)
-\author{Roy G. Biv\thanks{e-mail: roy.g.biv@aol.com}\\ %
-        \scriptsize Starbucks Research %
-\and Ed Grimley\thanks{e-mail:ed.grimley@aol.com}\\ %
-     \scriptsize Grimley Widgets, Inc. %
-\and Martha Stewart\thanks{e-mail:martha.stewart@marthastewart.com}\\ %
-     \parbox{1.4in}{\scriptsize \centering Martha Stewart Enterprises \\ Microsoft Research}}
+\author{Michael Correll\\ %
+        \scriptsize University of Washington %
+\and Dominik Moritz\\ %
+     \scriptsize University of Washington %
+\and Jeffrey Heer\\ %
+     \scriptsize University of Washington}
 
@@
 
-\section{Using the Style Template}
-
-\begin{itemize}
-\item The style uses the hyperref package, thus turns references into internal links.
-We thus recommend to make use of the ``\texttt{\textbackslash autoref\{reference\}}'' call (instead of ``\texttt{Figure\~{}\textbackslash ref\{reference\}}'' or similar) since ``\texttt{\textbackslash autoref\{reference\}}'' turns the entire reference into an internal link, not just the number.
-Examples: \autoref{fig:sample} and \autoref{tab:vis_papers}.
-\item The style automatically looks for image files with the correct extension (eps for regular \LaTeX; pdf, png, and jpg for pdf\LaTeX), in a set of given subfolders (figures/, pictures/, images/).
-It is thus sufficient to use ``\texttt{\textbackslash includegraphics\{CypressView\}}'' (instead of ``\texttt{\textbackslash includegraphics\{pictures/CypressView.jpg\}}'').
-\item For adding hyperlinks and DOIs to the list of references, you can use ``\texttt{\textbackslash bibliographystyle\{abbrv-doi-hyperref-narrow\}}'' (instead of ``\texttt{\textbackslash bibliographystyle\{abbrv\}}'').
-It uses the doi and url fields in a bib\TeX\ entry and turns the entire reference into a link, giving priority to the doi.
-The doi can be entered with or without the ``\texttt{http://dx.doi.org/}'' url part.
-See the examples in the bib\TeX\ file and the bibliography at the end of this template.\\[1em]
-\textbf{Note 1:} occasionally (for some \LaTeX\ distributions) this hyper-linked bib\TeX\ style may lead to \textbf{compilation errors} (``\texttt{pdfendlink ended up in different nesting level ...}'') if a reference entry is broken across two pages (due to a bug in hyperref).
-In this case make sure you have the latest version of the hyperref package (i.\,e., update your \LaTeX\ installation/packages) or, alternatively, revert back to ``\texttt{\textbackslash bibliographystyle\{abbrv-doi-narrow\}}'' (at the expense of removing hyperlinks from the bibliography) and try ``\texttt{\textbackslash bibliographystyle\{abbrv-doi-hyperref-narrow\}}'' again after some more editing.\\[1em]
-\textbf{Note 2:} the ``\texttt{-narrow}'' versions of the bibliography style use the font ``PTSansNarrow-TLF'' for typesetting the DOIs in a compact way.
-This font needs to be available on your \LaTeX\ system.
-It is part of the \href{https://www.ctan.org/pkg/paratype}{``paratype'' package}, and many distributions (such as MikTeX) have it automatically installed.
-If you do not have this package yet and want to use a ``\texttt{-narrow}'' bibliography style then use your \LaTeX\ system's package installer to add it.
-If this is not possible you can also revert to the respective bibliography styles without the ``\texttt{-narrow}'' in the file name.\\[1em]
-DVI-based processes to compile the template apparently cannot handle the different font so, by default, the template file uses the \texttt{abbrv-doi} bibliography style but the compiled PDF shows you the effect of the \texttt{abbrv-doi-hyperref-narrow} style.
-\end{itemize}
-
-\section{Bibliography Instructions}
-
-\begin{itemize}
-\item Sort all bibliographic entries alphabetically but the last name of the first author. This \LaTeX/bib\TeX\ template takes care of this sorting automatically.
-\item Merge multiple references into one; e.\,g., use \cite{Max:1995:OMF,Kitware:2003} (not \cite{Kitware:2003}\cite{Max:1995:OMF}).
-Within each set of multiple references, the references should be sorted in ascending order.
-This \LaTeX/bib\TeX\ template takes care of both the merging and the sorting automatically.
-\item Verify all data obtained from digital libraries, even ACM's DL and IEEE Xplore  etc.\ are sometimes wrong or incomplete.
-\item Do not trust bibliographic data from other services such as Mendeley.com, Google Scholar, or similar; these are even more likely to be incorrect or incomplete.
-\item Articles in journal---items to include:
-  \begin{itemize}
-  \item author names
-	\item title
-	\item journal name
-	\item year
-	\item volume
-	\item number
-	\item month of publication as variable name (i.\,e., \{jan\} for January, etc.; month ranges using \{jan \#\{/\}\# feb\} or \{jan \#\{-{}-\}\# feb\})
-  \end{itemize}
-\item use journal names in proper style: correct: ``IEEE Transactions on Visualization and Computer Graphics'', incorrect: ``Visualization and Computer Graphics, IEEE Transactions on''
-\item Papers in proceedings---items to include:
-  \begin{itemize}
-  \item author names
-	\item title
-	\item abbreviated proceedings name: e.\,g., ``Proc.\textbackslash{} CONF\_ACRONYNM'' without the year; example: ``Proc.\textbackslash{} CHI'', ``Proc.\textbackslash{} 3DUI'', ``Proc.\textbackslash{} Eurographics'', ``Proc.\textbackslash{} EuroVis''
-	\item year
-	\item publisher
-	\item town with country of publisher (the town can be abbreviated for well-known towns such as New York or Berlin)
-  \end{itemize}
-\item article/paper title convention: refrain from using curly brackets, except for acronyms/proper names/words following dashes/question marks etc.; example:
-\begin{itemize}
-	\item paper ``Marching Cubes: A High Resolution 3D Surface Construction Algorithm''
-	\item should be entered as ``\{M\}arching \{C\}ubes: A High Resolution \{3D\} Surface Construction Algorithm'' or  ``\{M\}arching \{C\}ubes: A high resolution \{3D\} surface construction algorithm''
-	\item will be typeset as ``Marching Cubes: A high resolution 3D surface construction algorithm''
-\end{itemize}
-\item for all entries
-\begin{itemize}
-	\item DOI can be entered in the DOI field as plain DOI number or as DOI url; alternative: a url in the URL field
-	\item provide full page ranges AA-{}-BB
-\end{itemize}
-\item when citing references, do not use the reference as a sentence object; e.\,g., wrong: ``In \cite{Lorensen:1987:MCA} the authors describe \dots'', correct: ``Lorensen and Cline \cite{Lorensen:1987:MCA} describe \dots''
-\end{itemize}
-
-\section{Example Section}
-
-Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
-nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
-sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
-rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
-ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
-sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
-dolore magna aliquyam erat, sed diam
-voluptua~\cite{Kitware:2003,Max:1995:OMF}. At vero eos et accusam et
-justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
-takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit
-amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor
-invidunt ut labore et dolore magna aliquyam erat, sed diam
-voluptua. At vero eos et accusam et justo duo dolores et ea
-rebum. Stet clita kasd gubergren, no sea takimata sanctus est.
-
-\section{Exposition}
-
-Duis autem vel eum iriure dolor in hendrerit in vulputate velit esse
-molestie consequat, vel illum dolore eu feugiat nulla facilisis at
-vero eros et accumsan et iusto odio dignissim qui blandit praesent
-luptatum zzril delenit augue duis dolore te feugait nulla
-facilisi. Lorem ipsum dolor sit amet, consectetuer adipiscing elit,
-sed diam nonummy nibh euismod tincidunt ut laoreet dolore magna
-aliquam erat volutpat~\cite{Kindlmann:1999:SAG}.
-
-\begin{equation}
-\sum_{j=1}^{z} j = \frac{z(z+1)}{2}
-\end{equation}
-
-Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
-nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
-sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
-rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
-ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
-sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
-dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
-et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
-takimata sanctus est Lorem ipsum dolor sit amet.
-
-\subsection{Lorem ipsum}
-
-Lorem ipsum dolor sit amet (see \autoref{tab:vis_papers}), consetetur sadipscing elitr, sed diam
-nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
-sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
-rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
-ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
-sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
-dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
-et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
-takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit
-amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor
-invidunt ut labore et dolore magna aliquyam erat, sed diam
-voluptua. At vero eos et accusam et justo duo dolores et ea
-rebum.
-
-\begin{table}[tb]
-  \caption{VIS/VisWeek accepted/presented papers: 1990--2015.}
-  \label{tab:vis_papers}
-  \scriptsize%
-	\centering%
-  \begin{tabu}{%
-	r%
-	*{7}{c}%
-	*{2}{r}%
-	}
-  \toprule
-   year & \rotatebox{90}{Vis/SciVis} &   \rotatebox{90}{SciVis conf} &   \rotatebox{90}{InfoVis} &   \rotatebox{90}{VAST} &   \rotatebox{90}{VAST conf} &   \rotatebox{90}{TVCG @ VIS} &   \rotatebox{90}{CG\&A @ VIS} &   \rotatebox{90}{VIS/VisWeek} \rotatebox{90}{incl.
-TVCG/CG\&A}   &   \rotatebox{90}{VIS/VisWeek} \rotatebox{90}{w/o TVCG/CG\&A}   \\
-  \midrule
-  2015 & 33 & 9 & 38 & 33 & 14 & 17 & 15 & 159 & 127 \\
-  2014 & 34 &   & 45 & 33 & 21 & 20 &   & 153 & 133 \\
-  2013 & 31 &   & 38 & 32 &   & 20 &   & 121 & 101 \\
-  2012 & 42 &   & 44 & 30 &   & 23 &   & 139 & 116 \\
-  2011 & 49 &   & 44 & 26 &   & 20 &   & 139 & 119 \\
-  2010 & 48 &   & 35 & 26 &   &   &   & 109 & 109 \\
-  2009 & 54 &   & 37 & 26 &   &   &   & 117 & 117 \\
-  2008 & 50 &   & 28 & 21 &   &   &   & 99 & 99 \\
-  2007 & 56 &   & 27 & 24 &   &   &   & 107 & 107 \\
-  2006 & 63 &   & 24 & 26 &   &   &   & 113 & 113 \\
-  2005 & 88 &   & 31 &   &   &   &   & 119 & 119 \\
-  2004 & 70 &   & 27 &   &   &   &   & 97 & 97 \\
-  2003 & 74 &   & 29 &   &   &   &   & 103 & 103 \\
-  2002 & 78 &   & 23 &   &   &   &   & 101 & 101 \\
-  2001 & 74 &   & 22 &   &   &   &   & 96 & 96 \\
-  2000 & 73 &   & 20 &   &   &   &   & 93 & 93 \\
-  1999 & 69 &   & 19 &   &   &   &   & 88 & 88 \\
-  1998 & 72 &   & 18 &   &   &   &   & 90 & 90 \\
-  1997 & 72 &   & 16 &   &   &   &   & 88 & 88 \\
-  1996 & 65 &   & 12 &   &   &   &   & 77 & 77 \\
-  1995 & 56 &   & 18 &   &   &   &   & 74 & 74 \\
-  1994 & 53 &   &   &   &   &   &   & 53 & 53 \\
-  1993 & 55 &   &   &   &   &   &   & 55 & 55 \\
-  1992 & 53 &   &   &   &   &   &   & 53 & 53 \\
-  1991 & 50 &   &   &   &   &   &   & 50 & 50 \\
-  1990 & 53 &   &   &   &   &   &   & 53 & 53 \\
-  \midrule
-  \textbf{sum} & \textbf{1515} & \textbf{9} & \textbf{595} & \textbf{277} & \textbf{35} & \textbf{100} & \textbf{15} & \textbf{2546} & \textbf{2431} \\
-  \bottomrule
-  \end{tabu}%
-\end{table}
-
-\subsection{Mezcal Head}
-
-Lorem ipsum dolor sit amet (see \autoref{fig:sample}), consetetur sadipscing elitr, sed diam
-nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
-sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
-rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
-ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
-sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
-dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
-et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
-takimata sanctus est Lorem ipsum dolor sit amet.
-
-\subsubsection{Duis Autem}
-
-Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
-nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
-sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
-rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
-ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
-sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
-dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
-et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
-takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit
-amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor
-invidunt ut labore et dolore magna aliquyam erat, sed diam
-voluptua. At vero eos et accusam et justo duo dolores et ea
-rebum. Stet clita kasd gubergren, no sea takimata sanctus est. Lorem
-ipsum dolor sit amet.
-
-\begin{figure}[tb]
- \centering % avoid the use of \begin{center}...\end{center} and use \centering instead (more compact)
- \includegraphics[width=\columnwidth]{paper-count-w-2015-new}
- \caption{A visualization of the data from \autoref{tab:vis_papers}. The image is from \cite{Isenberg:2017:VMC} and is in the public domain.}
- \label{fig:sample}
-\end{figure}
-
-\subsubsection{Ejector Seat Reservation}
-
-Duis autem~\cite{Lorensen:1987:MCA}\footnote{The algorithm behind
-Marching Cubes \cite{Lorensen:1987:MCA} had already been
-described by Wyvill et al. \cite{Wyvill:1986:DSS} a year
-earlier.} vel eum iriure dolor in hendrerit
-in vulputate velit esse molestie consequat,\footnote{Footnotes
-appear at the bottom of the column.} vel illum dolore eu
-feugiat nulla facilisis at vero eros et accumsan et iusto odio
-dignissim qui blandit praesent luptatum zzril delenit augue duis
-dolore te feugait nulla facilisi. Lorem ipsum dolor sit amet,
-consectetuer adipiscing elit, sed diam nonummy nibh euismod tincidunt
-ut laoreet dolore magna aliquam erat volutpat.
-
-\paragraph{Confirmed Ejector Seat Reservation}
-
-Ut wisi enim ad minim veniam, quis nostrud exerci tation ullamcorper
-suscipit lobortis nisl ut aliquip ex ea commodo
-consequat~\cite{Nielson:1991:TAD}. Duis autem vel eum iriure dolor in
-hendrerit in vulputate velit esse molestie consequat, vel illum dolore
-eu feugiat nulla facilisis at vero eros et accumsan et iusto odio
-dignissim qui blandit praesent luptatum zzril delenit augue duis
-dolore te feugait nulla facilisi.
-
-\paragraph{Rejected Ejector Seat Reservation}
-
-Ut wisi enim ad minim veniam, quis nostrud exerci tation ullamcorper
-suscipit lobortis nisl ut aliquip ex ea commodo consequat. Duis autem
-vel eum iriure dolor in hendrerit in vulputate velit esse molestie
+\section{Related Works}
 

### 2017-03-07 heatmap/heatmap.js

@@ -70,6 +70,30 @@ function makeArc(d,size,rows,cols){
 
 //should use d3.interpolateLab for colors.
 
+function cDist(color1,color2){
+  //Euclidean distance between two colors in Lab space.
+  var c1 = d3.lab(c1);
+  var c2 = d3.lab(c2);
+
+  return Math.sqrt( Math.pow( (c1.l-c2.l),2) + Math.pow(c1.a-c2.a,2) + Math.pow(c1.b - c2.b,2));
+}
+
+function minDist(colorRamp){
+  //What's the closest distance between colors, in our array of colors?
+  var minD, D;
+
+  for(var i = 0;i<colorRamp.length;i++){
+    for(var j = i+1;j<colorRamp.length;j++){
+        D = cDist(colorRamp[i],colorRamp[j]);
+        if((i==0 && j==1) || D<minD){
+          minD = D;
+        }
+    }
+  }
+
+  return minD;
+}
+
 function main(){
   //Create all relevant maps
 

### 2017-03-10 paper/figures.tex
Added simple examples to code, outline of paper to paper.

\newcommand{\teaserFig}{
  \teaser{
		\centering
		\includegraphics[width=1.5in]{sample.eps}
		\caption{Lookit! Lookit!}
		\label{fig:teaser}
	}
}

### 2017-03-10 paper/template.tex
Added simple examples to code, outline of paper to paper.

@@
 
-\title{Global Illumination for Fun and Profit}
+\title{Value-Suppressing Maps for Uncertain Data}
 
@@
 
+\input{figures.tex}
+
+%\teaserFig
 %% A teaser figure can be included as follows, but is not recommended since
@@
 %% Abstract section.
-\abstract{Duis autem vel eum iriure dolor in hendrerit in vulputate
-velit esse molestie consequat, vel illum dolore eu feugiat nulla
-facilisis at vero eros et accumsan et iusto odio dignissim qui blandit
-praesent luptatum zzril delenit augue duis dolore te feugait nulla
-facilisi. Lorem ipsum dolor sit amet, consectetuer adipiscing elit,
-sed diam nonummy nibh euismod tincidunt ut laoreet dolore magna
-aliquam erat volutpat. Ut wisi enim ad minim veniam, quis nostrud exerci tation ullamcorper
-suscipit lobortis nisl ut aliquip ex ea commodo consequat. Duis autem
-vel eum iriure dolor in hendrerit in vulputate velit esse molestie
-consequat, vel illum dolore eu feugiat nulla facilisis at vero eros et
-accumsan et iusto odio dignissim qui blandit praesent luptatum zzril
-delenit augue duis dolore te feugait nulla facilisi.%
+\abstract{
+
+	% uncertainty-aware decisions.
+	% refrain from making a decision
+	% integrate uncertainty with the display
+	% value-suppressing encodings
+	% provide greater visual discriminability for values with high certainty, but alias together uncertain values
+	% ran a study
+	% results indicate X
+
+	Uncertainty information is a vital part of many datasets, and should be accurately conveyed to the analyst.
+	Directly integrating uncertainty information with data requires that designers construct bivariate maps that encode value and uncertainty simultaneously.
+The design of such bivariate maps can be difficult, and the impact of different choices of encodings on decision-making under uncertainty understudied.
+    In this paper, we present a category of bivariate maps for uncertainty, ``value-suppressing maps.'' As opposed to traditional bivariate maps, value-suppressing maps make the informed decision to alias together values when uncertainty is high, in order to afford greater discriminability when uncertainty is low.
+That is, certain data is intrinsically easier to visually disambiguate than uncertain data.
+    We present several examples of these value-suppressing maps, as well as a crowd-sourced evaluation showing X. We conclude that Y.
+
+	%However, in thematic maps, strong visual variables such as position and hue are already reserved to encode location and value.
+
+	%Remaining visual variables with semiotic connection to uncertainty, such as transparency and saturation, are comparatively more difficult to discriminate, and can interfere with color discrimination.
+
+	%Uncertainty, when it is directly encoded at all, is often visualized using visual variables with lower discriminability, or relegated to tooltips or separate charts.
+
 } % end of abstract
@@
 
-\CCScatlist{ 
-  \CCScat{K.6.1}{Management of Computing and Information Systems}%
-{Project and People Management}{Life Cycle};
-  \CCScat{K.7.m}{The Computing Profession}{Miscellaneous}{Ethics}
-}
+\keywords{Uncertainty Visualization, Color Perception, Thematic Maps, Semiotics.}
+
+%\CCScatlist{ 
+%  \CCScat{K.6.1}{Management of Computing and Information Systems}%
+%{Project and People Management}{Life Cycle};
+%  \CCScat{K.7.m}{The Computing Profession}{Miscellaneous}{Ethics}
+%}
 
@@
 %% \section{Introduction} %for journal use above \firstsection{..} instead
-This template is for papers of VGTC-sponsored conferences which are \emph{\textbf{not}} published in a special issue of TVCG.
+
+% What this section needs to do:
+% Uncertainty is important!
+% What do we mean by ``well-integrated'' uncertainty?
+% Scope down to thematic maps
+% Juxtapose/superimpost distinction?
+% Position bivariate maps (tacitly or explicitly) as state of the art
+% Teaser image showing what we mean
+
+Uncertainty is an inescapable component of collecting, presenting, and using data.
+A common goal in the communication of uncertainty is promoting \textbf{uncertainty-aware decisions}--- that is, the audience should be aware of the risks and rewards of certain decisions, modulate their confidence in their conclusions, and perhaps refrain from making a decision at all if there is too much uncertainty.
+A way that designers can contribute to this goal is by ensuring that uncertainty information is \textbf{well-integrated} with the rest of the data.
+That is, it should be difficult to discount or ignore the uncertainty in a dataset.
+A designer of a visualization could instantiate this principle through the simultaneous presentation of uncertainty and value.
+This simultaneous presentation necessitates the construction of a bivariate map--- a relation, in terms of visual variables, between a 2-tuple $(value$,$uncertainty)$ and a set of marks.
+
+In this paper, we present \textbf{Value-Suppressing Maps} (VSMs) for integrating data and uncertainty information in thematic maps.
+VSMs intentionally alias together data values with high uncertainty, affording greater discriminability as certainty increases.
+Rather than a square bivariate map, with unique outputs for each relevant combination of data and uncertainty, VSMs can be thought of as arcs: as uncertainty increases, values are mapped to smaller and smaller sets of outputs, culminating in a singularity where all inputs are mapped to an identical, highly uncertain glyph regardless of their data value.
+Figure X shows an example of a VSM, compared to a more traditional bivariate map.
+In this work, we describe the motivations behind the use of VSMs, examples of their utility for decision-making under uncertainty, and validate VDMs in a crowd-sourced experiment.
 
@@
 
+% What this section needs to do
+% State of the art in uncertainty vis, with heavy emphasis on MacEachren & co. 
+% Bivariate maps are hard!
+% Binning colormaps is de rigueur
+% The variables we'd want to use for uncertainty (like value or alpha or size) mess with color discriminability
+
+\subsection{Visual Variables for Representing Uncertainty}
+
+MacEachren et al. \cite{maceachren2012visual} evaluate a number of visual variables for representing uncertainty. 
+\subsection{Interactions Between Visual Variables}
+
+\section{Value-Suppressing Maps}
+
+\subsection{Design}
+
+\subsection{Examples}
+
+\section{Evaluation}
+
+%Show:
+% 1) Juxtaposition makes people ignore/underweight uncertainty (even when they don't screw up the search task)
+% 2) VSMs make people, surprise surprise, bad at distinguishing highly uncertain values
+% 3) But, they make uncertainty better-integrated
+%    i) lower weights on uncertain values in decision-making
+%    ii) a priori better discrimination
+
+\subsection{Methods}
+
+%conditions:
+% 1) juxtaposed d map, u map
+% 2) binned d/u map as per a priori jnds
+% 3) vsm
+% 4) continuous d/u map?
+
+\subsubsection{Participants}
+\subsection{Results}
+
+\section{Discussion}
+\subsection{Limitations and Future Work}
+% When should you use these things?
+
 \section{Conclusion}
-
-Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
-nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
-sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
-rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
-ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
-sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
-dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
-et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
-takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit
-amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor
-invidunt ut labore et dolore magna aliquyam erat, sed diam
-voluptua. At vero eos et accusam et justo duo dolores et ea
-rebum.
 
@@
 \acknowledgments{
-The authors wish to thank A, B, C. This work was supported in part by
-a grant from XYZ.}
+Omitted for review.}
 

### 2017-03-10 heatmap/heatmap.js

@@ -1,5 +1,8 @@
 
 var svg = d3.select("body").append("svg");
+var map = colorbrewer.YlGnBu[9];
+var z = d3.scale.quantize().range(map);
+var u = d3.scale.quantize().range(colorbrewer.Greys[9]);
 
 function makeHeatmap(x,y,size,data,z,name){
   //creates an svg heatmap with a 2d matrix of data, and a mapping function z
@@ -72,8 +75,8 @@ function makeArc(d,size,rows,cols){
 
 function cDist(color1,color2){
   //Euclidean distance between two colors in Lab space.
-  var c1 = d3.lab(c1);
-  var c2 = d3.lab(c2);
+  var c1 = d3.lab(color1);
+  var c2 = d3.lab(color2);
 
   return Math.sqrt( Math.pow( (c1.l-c2.l),2) + Math.pow(c1.a-c2.a,2) + Math.pow(c1.b - c2.b,2));
 }
@@ -94,6 +97,25 @@ function minDist(colorRamp){
   return minD;
 }
 
+function checkMap(mapName){
+  //Takes a d3 selection containing all the marks we care about
+  //Checks to make sure that colors it contains are sufficiently far apart.
+  var colorList = [];
+  var aColor;
+
+  mapName
+    .each(function(d){
+      aColor = d3.select(this).attr("fill");
+      if(!colorList.includes(aColor)){
+        colorList.push(aColor);
+      }
+    });
+
+    var closest = minDist(colorList);
+    console.log( "The two closest colors are "+closest+" apart in CIELAB.");
+    return closest>5;
+}
+
 function main(){
   //Create all relevant maps
 
@@ -104,10 +126,6 @@ function main(){
   3) Integrated map with "arc" mapping
   4) Legends for both square and arc maps.
   */
-
-  var map = colorbrewer.YlGnBu[9];
-  var z = d3.scale.quantize().range(map);
-  var u = d3.scale.quantize().range(colorbrewer.Greys[9]);
 
   var data = [
     [1,2,3],
@@ -127,15 +145,68 @@ function main(){
     [6]
   ];
 
+  var scaleData = [
+    [{v:0.25,u:0.0},{v:0.5,u:0.0},{v:0.75,u:0.0},{v:1.0,u:0.0}],
[…]

### 2017-03-10 heatmap/heatmap.js
Some more structural comments.

@@ -1,8 +1,12 @@
 
+//Global Variables
+
 var svg = d3.select("body").append("svg");
 var map = colorbrewer.YlGnBu[9];
 var z = d3.scale.quantize().range(map);
-var u = d3.scale.quantize().range(colorbrewer.Greys[9]);
+
+
+//Chart creation functions
 
 function makeHeatmap(x,y,size,data,z,name){
   //creates an svg heatmap with a 2d matrix of data, and a mapping function z
@@ -71,6 +75,21 @@ function makeArc(d,size,rows,cols){
   return arc;
 }
 
+//Data loading and generation
+
+function randomData(rows,cols){
+  //uniform random data
+  var data = [];
+  for(var i = 0;i<rows;i++){
+    data[i] = [];
+    for(var j=0;j<cols;j++){
+      data[i].push({ u: Math.random(), v: Math.random()});
+    }
+  }
+
+  return data;
+}
+
 //should use d3.interpolateLab for colors.
 
 function cDist(color1,color2){
@@ -83,18 +102,20 @@ function cDist(color1,color2){
 
 function minDist(colorRamp){
   //What's the closest distance between colors, in our array of colors?
-  var minD, D;
+  var minD, D,c1,c2;
 
   for(var i = 0;i<colorRamp.length;i++){
     for(var j = i+1;j<colorRamp.length;j++){
         D = cDist(colorRamp[i],colorRamp[j]);
         if((i==0 && j==1) || D<minD){
           minD = D;
+          c1 = colorRamp[i];
+          c2 = colorRamp[j];
         }
     }
   }
 
-  return minD;
+  return {"minD": minD, "c1": c1, "c2": c2};
 }
 
 function checkMap(mapName){
@@ -112,8 +133,8 @@ function checkMap(mapName){
     });
 
     var closest = minDist(colorList);
-    console.log( "The two closest colors are "+closest+" apart in CIELAB.");
-    return closest>5;
+    console.log( "The two closest colors:(" + closest.c1 +"," + closest.c2 +") are "+closest.minD+" apart in CIELAB.");
+    return closest.minD>5;
 }
 
 function main(){
@@ -160,13 +181,15 @@ function main(){
   ];
 
   z.domain([d3.min(d3.min(data)),d3.max(d3.max(data))]);
-  u.domain([0,1]);
 
[…]

### 2017-03-10 heatmap/heatmap.js
Part of viridis colormap
TODO: Need to switch to d3 v4 to get more scale goodies

@@ -2,7 +2,12 @@
 //Global Variables
 
 var svg = d3.select("body").append("svg");
-var map = colorbrewer.YlGnBu[9];
+//var map = colorbrewer.Spectral[9];
+
+//part of viridis
+//d3.range(10).map(d => d3.interpolateViridis(d/9));
+var map = ["#440154", "#482878", "#3e4989", "#31688e", "#26828e", "#1f9e89", "#35b779", "#6ece58", "#b5de2b"];
+
 var z = d3.scale.quantize().range(map);
 
 

### 2017-03-18 paper/template.tex
Added methods notes from JH meeting

@@
 
-\title{Value-Suppressing Maps for Uncertain Data}
+\title{Value-Suppressing Uncertainty Maps}
 
@@
 The design of such bivariate maps can be difficult, and the impact of different choices of encodings on decision-making under uncertainty understudied.
-    In this paper, we present a category of bivariate maps for uncertainty, ``value-suppressing maps.'' As opposed to traditional bivariate maps, value-suppressing maps make the informed decision to alias together values when uncertainty is high, in order to afford greater discriminability when uncertainty is low.
+    In this paper, we present a category of bivariate maps for uncertainty, ``value-suppressing uncertainty maps,'' (VSUMS).
+As opposed to traditional bivariate maps, VSUMS make the informed decision to alias together values when uncertainty is high, in order to afford greater discriminability when uncertainty is low.
 That is, certain data is intrinsically easier to visually disambiguate than uncertain data.
-    We present several examples of these value-suppressing maps, as well as a crowd-sourced evaluation showing X. We conclude that Y.
+    We present several examples of these maps, as well as a crowd-sourced evaluation showing X. We conclude that Y.
 
@@
 
-In this paper, we present \textbf{Value-Suppressing Maps} (VSMs) for integrating data and uncertainty information in thematic maps.
-VSMs intentionally alias together data values with high uncertainty, affording greater discriminability as certainty increases.
-Rather than a square bivariate map, with unique outputs for each relevant combination of data and uncertainty, VSMs can be thought of as arcs: as uncertainty increases, values are mapped to smaller and smaller sets of outputs, culminating in a singularity where all inputs are mapped to an identical, highly uncertain glyph regardless of their data value.
-Figure X shows an example of a VSM, compared to a more traditional bivariate map.
-In this work, we describe the motivations behind the use of VSMs, examples of their utility for decision-making under uncertainty, and validate VDMs in a crowd-sourced experiment.
+In this paper, we present \textbf{Value-Suppressing Uncertainty Maps} (VSUMs) for integrating data and uncertainty information in thematic maps.
+VSUMs intentionally alias together data values with high uncertainty, affording greater discriminability as certainty increases.
+Rather than a square bivariate map, with unique outputs for each relevant combination of data and uncertainty, VSUMs can be thought of as arcs: as uncertainty increases, values are mapped to smaller and smaller sets of outputs, culminating in a singularity where all inputs are mapped to an identical, highly uncertain mark regardless of their data value.
+Figure X shows an example of a VSUM, compared to a more traditional bivariate map.
+In this work, we describe the motivations behind the use of VSUMs, examples of their utility for decision-making under uncertainty, and validate VSUMs in a crowd-sourced experiment.
 
@@
 
-% What this section needs to do
+% What this section needs to do:
 % State of the art in uncertainty vis, with heavy emphasis on MacEachren & co. 
@@
 
-\section{Value-Suppressing Maps}
+\section{Value-Suppressing Uncertainty Maps}
 
@@
 % 1) Juxtaposition makes people ignore/underweight uncertainty (even when they don't screw up the search task)
-% 2) VSMs make people, surprise surprise, bad at distinguishing highly uncertain values
+% 2) VSUMs make people, surprise surprise, bad at distinguishing highly uncertain values
 % 3) But, they make uncertainty better-integrated

### 2017-03-18 study/src/script.js

@@ -1,7 +1,25 @@
 /*
-Code for Turk Experiments for analysis of Value Suppressing Maps for Uncertain Data
+Code for Turk Experiments for analysis of Value Suppressing Uncertainty Maps
 
+IVs:
 
+1. Map type
+A. Juxtaposed Map: Two heatmaps, one showing value, the other uncertainty.
+Color ramps have as small a bin size as possible, down to JND 5.0 in CIELAB.
+B. Superimposed Map: One heatmap, bivariate map down to as small a bin size as possible.
+Will be coarser than the juxtaposed map, since there's inteference between the two variables.
+C. VSUP: One heatmap, bivariate, but with uncertain values increasing aliased together.
+
+Procedure:
+1. Get Consent.
+2. Tutorial showing examples of tasks, as well as a general primer on uncertainty.
+   In this case, we'll be using variability as our uncertainty measure.
+3. Identification tasks. Given a heatmap, and a location (highlighted or indicated),
+   what is the value, and what is the uncertainty?
+   or: find min and max locations
+4. Roulette tasks: Given a heatmap, and n tokens, place them such that you maximize the E.V.
+   Include both gain and loss framings.
+5. Collect demographic information, including risk assay.
 */
 
 var startTime;
@@ -100,6 +118,10 @@ function finishTask(){
   //When we're finished with all tasks.
   main.selectAll("*").remove();
   postTest();
+}
+
+function riskAversion(fromElement){
+  //A risk aversion assay, from
 }
 
 function postTest(){

### 2017-03-20 heatmap/heatmap.js
Functionized map creation
Affords some metrics over time.

@@ -228,78 +228,66 @@ function makeScaleFunction(scaleData){
   }
 }
 
-function colorDiff(scaleData) {
+function numBins(scaleData){
+  var colors = scaleData.reduce(function(arr, curr) {
+    return arr.concat(curr.map(uSL));
+  }, []);
+  return colors.length;
+}
+
+function colorDiff(scaleData){
   var colors = scaleData.reduce(function(arr, curr) {
     return arr.concat(curr.map(uSL));
   }, []);
   return minDist(colors);
 }
 
+function makeMaps(threshold){
+    //Create all relevant maps
+    var DEFAULT_THRESHOLD = 5;
+    var THRESHOLD = threshold ? threshold : DEFAULT_THRESHOLD;
+    var scaleData, arcScaleData, closest, n;
+
+    n = 2;
+    while (true) {
+      var data = makeScaleData(n);
+      var c = colorDiff(data);
+      if (c.minD >= THRESHOLD) {
+        scaleData = data;
+        closest = c;
+      }
+      else {
+        break;
+      }
+      n++;
+    }
+    console.log("The two closest matrix colors:(" + closest.c1 +"," + closest.c2 +") are "+closest.minD+" apart in CIELAB.");
+    n = 2;
+    while (true) {
+      var data = makeArcScaleData(n);
+      var c = colorDiff(data);
+      if (c.minD >= THRESHOLD) {
+        arcScaleData = data;
+        closest = c;
+      } else {
+        break;
+      }
+      n++;
+    }
+    console.log("The two closest arc colors:(" + closest.c1 +"," + closest.c2 +") are "+closest.minD+" apart in CIELAB.");
+
+    return {square:scaleData, arc:arcScaleData};
+}
+
 function main(){
-  //Create all relevant maps
+  var maps = makeMaps(18);
 
-  /*TODO
-  We'll need:
-  1) Juxtaposed Data/Uncertainty maps
-  2) Integrated map with regular "square" mapping
-  3) Integrated map with "arc" mapping
-  4) Legends for both square and arc maps.
-  */
-
-  // var scaleData = [
[…]

### 2017-03-21 heatmap/heatmap.js
Debug heatmap

@@ -80,12 +80,24 @@ function makeArc(d,size,rows,cols){
 
 //Data loading and generation
 
+function gradientData(rows, cols) {
+  var data = [];
+  for(var i = 0; i<rows; i++){
+    data[i] = [];
+    for(var j = 0; j<cols; j++){
+      data[i].push({ u: i/(rows-1), v: j/(cols-1)});
+    }
+  }
+
+  return data;
+}
+
 function randomData(rows,cols){
   //uniform random data
   var data = [];
-  for(var i = 0;i<rows;i++){
+  for(var i = 0; i<rows; i++){
     data[i] = [];
-    for(var j=0;j<cols;j++){
+    for(var j = 0; j<cols; j++){
       data[i].push({ u: Math.random(), v: Math.random()});
     }
   }
@@ -297,7 +309,7 @@ function main(){
   makeHeatmap(0,0,250,scaleData, uSL, "SquareWhite");
   makeArcmap(300,0,250,arcScaleData, uSL, "ArcWhite");
 
-  var exampleData = randomData(10,10);
+  var exampleData = gradientData(10,10);
   makeHeatmap(0,300,250,exampleData, makeScaleFunction(scaleData), "exampleArc");
   makeHeatmap(300,300,250,exampleData,makeScaleFunction(arcScaleData),"exampleSquare");
 }

### 2017-03-21 paper/template.tex

@@
 
+%MC Packages and commands
+\usepackage{subcaption}
+\newcommand{\todo}[1]{\textbf{\textcolor{red}{TODO: #1}}\\}
+
 %% Paper title.
@@
 
-	Uncertainty information is a vital part of many datasets, and should be accurately conveyed to the analyst.
-	Directly integrating uncertainty information with data requires that designers construct bivariate maps that encode value and uncertainty simultaneously.
-The design of such bivariate maps can be difficult, and the impact of different choices of encodings on decision-making under uncertainty understudied.
-    In this paper, we present a category of bivariate maps for uncertainty, ``value-suppressing uncertainty maps,'' (VSUMS).
-As opposed to traditional bivariate maps, VSUMS make the informed decision to alias together values when uncertainty is high, in order to afford greater discriminability when uncertainty is low.
+	Uncertainty is a vital part of many datasets, and should be conveyed accurately to the analyst.
+	Directly incorporating uncertainty requires that designers construct bivariate mappings that encode value and uncertainty simultaneously.
+	In this paper, we present a technique for generating such bivariate maps for uncertainty, ``Value-suppressing uncertainty maps'' (VSUMs).
+	As opposed to traditional bivariate maps, VSUMs make the informed choice to alias together values when uncertainty is high, in order to afford greater discriminability between categories when uncertainty is low.
 That is, certain data is intrinsically easier to visually disambiguate than uncertain data.
-    We present several examples of these maps, as well as a crowd-sourced evaluation showing X. We conclude that Y.
-
-	%However, in thematic maps, strong visual variables such as position and hue are already reserved to encode location and value.
-
-	%Remaining visual variables with semiotic connection to uncertainty, such as transparency and saturation, are comparatively more difficult to discriminate, and can interfere with color discrimination.
-
-	%Uncertainty, when it is directly encoded at all, is often visualized using visual variables with lower discriminability, or relegated to tooltips or separate charts.
+	Values with low uncertainty can be more precisely compared, whereas similar values with higher uncertainty will be harder to differentiate in accordance with their variability.
+	By making informed decisions about where to create category bins, VSUMs make better use of the limited budget of discriminable marks in bivariate maps, and dissuade analysts from making decisions based on data with sufficiently high uncertainty.
+	We present several examples of VSUMs, as well as a crowd-sourced evaluation showing that, compared to more traditional bivariate maps, VSUMs promote decision-making that places greater weight on uncertainty information compared to traditional bivariate maps.
 
@@
 A common goal in the communication of uncertainty is promoting \textbf{uncertainty-aware decisions}--- that is, the audience should be aware of the risks and rewards of certain decisions, modulate their confidence in their conclusions, and perhaps refrain from making a decision at all if there is too much uncertainty.
+In general, a goal of designers of visualizations of uncertain data is the \emph{information fusion} of data and uncertainty~\cite{riveiro2007evaluation}.
 A way that designers can contribute to this goal is by ensuring that uncertainty information is \textbf{well-integrated} with the rest of the data.
 That is, it should be difficult to discount or ignore the uncertainty in a dataset.
-A designer of a visualization could instantiate this principle through the simultaneous presentation of uncertainty and value.
-This simultaneous presentation necessitates the construction of a bivariate map--- a relation, in terms of visual variables, between a 2-tuple $(value$,$uncertainty)$ and a set of marks.
+Simultaneous presentation of uncertainty and value necessitates the construction of a bivariate map--- a relation, in terms of visual variables, between a 2-tuple $(value$,$uncertainty)$ and a set of marks.
 
@@
 Figure X shows an example of a VSUM, compared to a more traditional bivariate map.
-In this work, we describe the motivations behind the use of VSUMs, examples of their utility for decision-making under uncertainty, and validate VSUMs in a crowd-sourced experiment.
+In this work, we describe the motivations behind the use of VSUMs, examples of their utility for decision-making under uncertainty, and validate that VSUMs in a crowd-sourced experiment.
 
@@
 % The variables we'd want to use for uncertainty (like value or alpha or size) mess with color discriminability
+Despite the acknowledged importance of uncertainty in understanding data, explicit representation of uncertainty often missing from visualizations~\cite{boukhelifa2009uncertainty}. The absence of
 
+The presentation Broadlie et al.~\cite{brodlie2012review}
 \subsection{Visual Variables for Representing Uncertainty}
 
-MacEachren et al. \cite{maceachren2012visual} evaluate a number of visual variables for representing uncertainty. 
+MacEachren et al.~\cite{maceachren2012visual} evaluate a number of visual variables for representing uncertainty. 
 \subsection{Interactions Between Visual Variables}

### 2017-03-21 heatmap/heatmap.js
Refactor code for scale generation.

@@ -70,10 +70,10 @@ function makeArc(d,size,rows,cols){
   var radius = d3.scaleLinear().domain([0,rows]).range([size,0]);
 
   var arc = d3.arc()
-  .innerRadius(radius(d.r+1))
-  .outerRadius(radius(d.r))
-  .startAngle(angle(d.c))
-  .endAngle(angle(d.c+1));
+    .innerRadius(radius(d.r+1))
+    .outerRadius(radius(d.r))
+    .startAngle(angle(d.c))
+    .endAngle(angle(d.c+1));
 
   return arc;
 }
@@ -213,6 +213,9 @@ function makeArcScaleData(n) {
 }
 
 function makeScaleFunction(scaleData){
+  var cScale = d3.scaleQuantize().domain([0,1]).range(map);
+  var iVal = d3.scaleLinear().domain([0,1]).range([0.0,1.0]);
+
   //a quantized scale function from a given scale
   return function f(d){
     var u,v;
@@ -220,23 +223,20 @@ function makeScaleFunction(scaleData){
     var j = 0;
 
     //find the right uncertainty row
-    while(scaleData[i][j].u<d.u && i<scaleData.length-1){
+    while(scaleData[i][j].u<d.u && i<scaleData.length - 1){
       i++;
     }
 
     //find the right value column
-    while(scaleData[i][j].v<d.v && j<scaleData[i].length-1){
+    while(scaleData[i][j].v<d.v && j<scaleData[i].length - 1){
       j++;
     }
 
     u = scaleData[i][j].u;
     v = scaleData[i][j].v;
 
-    var cScale = d3.scaleQuantize().domain([0,1]).range(map);
-    var steps = map.length;
     var c = (d3.hsl(cScale(v)));
-    var iVal = d3.scaleLinear().domain([0,1]).range([0.0,1.0]);
-    return d3.interpolateLab(c,"white")(iVal(u));
+    return d3.interpolateLab(c,"white")(u);  // iVal was identity so I removed it
   }
 }
 

### 2017-03-21 heatmap/heatmap.js
Indent

@@ -255,40 +255,40 @@ function colorDiff(scaleData){
 }
 
 function makeMaps(threshold){
-    //Create all relevant maps
-    var DEFAULT_THRESHOLD = 5;
-    var THRESHOLD = threshold ? threshold : DEFAULT_THRESHOLD;
-    var scaleData, arcScaleData, closest, n;
+  //Create all relevant maps
+  var DEFAULT_THRESHOLD = 5;
+  var THRESHOLD = threshold ? threshold : DEFAULT_THRESHOLD;
+  var scaleData, arcScaleData, closest, n;
 
-    n = 2;
-    while (true) {
-      var data = makeScaleData(n);
-      var c = colorDiff(data);
-      if (c.minD >= THRESHOLD) {
-        scaleData = data;
-        closest = c;
-      }
-      else {
-        break;
-      }
-      n++;
+  n = 2;
+  while (true) {
+    var data = makeScaleData(n);
+    var c = colorDiff(data);
+    if (c.minD >= THRESHOLD) {
+      scaleData = data;
+      closest = c;
     }
-    console.log("The two closest matrix colors:(" + closest.c1 +"," + closest.c2 +") are "+closest.minD+" apart in CIELAB.");
-    n = 2;
-    while (true) {
-      var data = makeArcScaleData(n);
-      var c = colorDiff(data);
-      if (c.minD >= THRESHOLD) {
-        arcScaleData = data;
-        closest = c;
-      } else {
-        break;
-      }
-      n++;
+    else {
+      break;
     }
-    console.log("The two closest arc colors:(" + closest.c1 +"," + closest.c2 +") are "+closest.minD+" apart in CIELAB.");
+    n++;
+  }
+  console.log("The two closest matrix colors:(" + closest.c1 +"," + closest.c2 +") are "+closest.minD+" apart in CIELAB.");
+  n = 2;
+  while (true) {
+    var data = makeArcScaleData(n);
+    var c = colorDiff(data);
+    if (c.minD >= THRESHOLD) {
+      arcScaleData = data;
+      closest = c;
+    } else {
+      break;
+    }
+    n++;
+  }
+  console.log("The two closest arc colors:(" + closest.c1 +"," + closest.c2 +") are "+closest.minD+" apart in CIELAB.");
 
-    return {square:scaleData, arc:arcScaleData};
+  return {square:scaleData, arc:arcScaleData};
 }
 
 function main(){

### 2017-03-21 heatmap/heatmap.js
Create scales only once.

@@ -293,12 +293,16 @@ function makeMaps(threshold){
 
 function main(){
   var maps = makeMaps(18);
-  makeHeatmap(0,0,250,maps.square, makeScaleFunction(maps.square), "legendSquare");
-  makeArcmap(300,0,250,maps.arc,makeScaleFunction(maps.arc),"legendArc");
 
-  var exampleData = gradientData(10,10);
-  makeHeatmap(0,300,250,exampleData, makeScaleFunction(maps.square), "exampleSquare");
-  makeHeatmap(300,300,250,exampleData,makeScaleFunction(maps.arc),"exampleArc");
+  var squareScale = makeScaleFunction(maps.square);
+  var arcScale = makeScaleFunction(maps.arc);
+
+  makeHeatmap(0,0,250,maps.square, squareScale, "legendSquare");
+  makeArcmap(300,0,250,maps.arc,arcScale,"legendArc");
+
+  var exampleData = gradientData(20,20);
+  makeHeatmap(0,300,250,exampleData, squareScale, "exampleSquare");
+  makeHeatmap(300,300,250,exampleData,arcScale,"exampleArc");
 }
 
 //Uncertainty maps

### 2017-03-21 heatmap/heatmap.js
Temporary fix for color mapping.

@@ -214,7 +214,6 @@ function makeArcScaleData(n) {
 
 function makeScaleFunction(scaleData){
   var cScale = d3.scaleQuantize().domain([0,1]).range(map);
-  var iVal = d3.scaleLinear().domain([0,1]).range([0.0,1.0]);
 
   //a quantized scale function from a given scale
   return function f(d){
@@ -222,21 +221,17 @@ function makeScaleFunction(scaleData){
     var i = 0;
     var j = 0;
 
-    //find the right uncertainty row
-    while(scaleData[i][j].u<d.u && i<scaleData.length - 1){
-      i++;
-    }
+    var rows = d3.scaleQuantize().domain([0,1]).range(scaleData);
+    var row = rows(d.u);
 
-    //find the right value column
-    while(scaleData[i][j].v<d.v && j<scaleData[i].length - 1){
-      j++;
-    }
+    var cols = d3.scaleQuantize().domain([0,1]).range(row);
+    var value = cols(d.v);
 
-    u = scaleData[i][j].u;
-    v = scaleData[i][j].v;
+    u = value.u;
+    v = value.v;
 
     var c = (d3.hsl(cScale(v)));
-    return d3.interpolateLab(c,"white")(u);  // iVal was identity so I removed it
+    return d3.interpolateLab(c,"white")(u);
   }
 }
 

### 2017-03-21 heatmap/heatmap.js
Use random instead of gradient

@@ -295,7 +295,7 @@ function main(){
   makeHeatmap(0,0,250,maps.square, squareScale, "legendSquare");
   makeArcmap(300,0,250,maps.arc,arcScale,"legendArc");
 
-  var exampleData = gradientData(20,20);
+  var exampleData = randomData(20,20);
   makeHeatmap(0,300,250,exampleData, squareScale, "exampleSquare");
   makeHeatmap(300,300,250,exampleData,arcScale,"exampleArc");
 }

### 2017-03-21 heatmap/heatmap.js
10x10

@@ -295,7 +295,7 @@ function main(){
   makeHeatmap(0,0,250,maps.square, squareScale, "legendSquare");
   makeArcmap(300,0,250,maps.arc,arcScale,"legendArc");
 
-  var exampleData = randomData(20,20);
+  var exampleData = randomData(10,10);
   makeHeatmap(0,300,250,exampleData, squareScale, "exampleSquare");
   makeHeatmap(300,300,250,exampleData,arcScale,"exampleArc");
 }

### 2017-03-21 heatmap/heatmap.js
Removed constraint that there be an even number of bins in the top of an arc.
Just go to the floor of the remaining number, which also removes issue
where number of bins is not a strict power of 2.

Also added in d3 colorbrewer schemes, just in case.

@@ -7,6 +7,12 @@ var svg = d3.select("body").append("svg");
 var NUM_STEPS = 1000;
 var map = d3.range(NUM_STEPS).map(d => d3.interpolateViridis(d/(NUM_STEPS-1)));
 
+//Other maps to test:
+//inferno
+//var map = d3.range(NUM_STEPS).map(d => d3.interpolateInferno(d/(NUM_STEPS-1)));
+//colorbrewer spectral
+//var map = d3.schemeSpectral[11];
+
 var z = d3.scaleQuantize().range(map);
 
 
@@ -193,7 +199,7 @@ function makeScaleData(n) {
  */
 function makeArcScaleData(n) {
   var arr = [];
-  var perRow = n%2==0 ? n : n-1;
+  var perRow = n;
   var numRows = Math.floor(Math.log2(n));
   var u = 0;
   while(perRow>=2){
@@ -203,7 +209,7 @@ function makeArcScaleData(n) {
     }
     arr.push(row);
     u++;
-    perRow/=2;
+    perRow= Math.floor(perRow/2);
   }
 
   // tip
@@ -295,7 +301,8 @@ function main(){
   makeHeatmap(0,0,250,maps.square, squareScale, "legendSquare");
   makeArcmap(300,0,250,maps.arc,arcScale,"legendArc");
 
-  var exampleData = gradientData(20,20);
+  //var exampleData = gradientData(100,100);
+  var exampleData = randomData(10,10);
   makeHeatmap(0,300,250,exampleData, squareScale, "exampleSquare");
   makeHeatmap(300,300,250,exampleData,arcScale,"exampleArc");
 }

### 2017-03-21 heatmap/heatmap.js
Some futureproofing
When we run full analyses we might hit non-valid color thresholds.

@@ -172,9 +172,15 @@ function checkMap(mapName){
       }
     });
 
-    var closest = minDist(colorList);
-    console.log( "The two closest colors:(" + closest.c1 +"," + closest.c2 +") are "+closest.minD+" apart in CIELAB.");
-    return closest.minD>5;
+    if(colorList.length==0){
+      console.log("No valid color map");
+      return false;
+    }
+    else{
+      var closest = minDist(colorList);
+      console.log( "The two closest colors:(" + closest.c1 +"," + closest.c2 +") are "+closest.minD+" apart in CIELAB.");
+      return closest.minD>5;
+    }
 }
 
 /**
@@ -274,7 +280,12 @@ function makeMaps(threshold){
     }
     n++;
   }
-  console.log("The two closest matrix colors:(" + closest.c1 +"," + closest.c2 +") are "+closest.minD+" apart in CIELAB.");
+  if(!closest){
+    console.log("No valid matrix color map at threshold "+THRESHOLD);
+  }
+  else{
+    console.log("The two closest matrix colors:(" + closest.c1 +"," + closest.c2 +") are "+closest.minD+" apart in CIELAB.");
+  }
   n = 2;
   while (true) {
     var data = makeArcScaleData(n);
@@ -287,7 +298,12 @@ function makeMaps(threshold){
     }
     n++;
   }
-  console.log("The two closest arc colors:(" + closest.c1 +"," + closest.c2 +") are "+closest.minD+" apart in CIELAB.");
+  if(!closest){
+    console.log("No valid arc color map at threshold "+THRESHOLD);
+  }
+  else{
+    console.log("The two closest arc colors:(" + closest.c1 +"," + closest.c2 +") are "+closest.minD+" apart in CIELAB.");
+  }
 
   return {square:scaleData, arc:arcScaleData};
 }

### 2017-03-21 paper/template.tex
JH abstract edits

@@
 
-	Uncertainty is a vital part of many datasets, and should be conveyed accurately to the analyst.
-	Directly incorporating uncertainty requires that designers construct bivariate mappings that encode value and uncertainty simultaneously.
-	In this paper, we present a technique for generating such bivariate maps for uncertainty, ``Value-suppressing uncertainty maps'' (VSUMs).
-	As opposed to traditional bivariate maps, VSUMs make the informed choice to alias together values when uncertainty is high, in order to afford greater discriminability between categories when uncertainty is low.
-That is, certain data is intrinsically easier to visually disambiguate than uncertain data.
-	Values with low uncertainty can be more precisely compared, whereas similar values with higher uncertainty will be harder to differentiate in accordance with their variability.
-	By making informed decisions about where to create category bins, VSUMs make better use of the limited budget of discriminable marks in bivariate maps, and dissuade analysts from making decisions based on data with sufficiently high uncertainty.
-	We present several examples of VSUMs, as well as a crowd-sourced evaluation showing that, compared to more traditional bivariate maps, VSUMs promote decision-making that places greater weight on uncertainty information compared to traditional bivariate maps.
+Uncertainty is a vital concern in many datasets.
+Visualizations often use bivariate mappings to encode value and uncertainty simultaneously.
+Due to interference between visual channels, these bivariate maps are often limited in resolution.
+We contribute Value-Suppressing Uncertainty Maps (VSUMs), an encoding technique that intentionally impairs value discrimination as uncertainty increases.
+In contrast to traditional bivariate maps, VSUMs alias values when uncertainty is high: expected values with lower uncertainty can be more precisely compared, whereas similar values with higher uncertainty are made harder to differentiate.
+By making informed binning decisions, VSUMs seek to better use the limited budget of discriminable marks in bivariate maps, and dissuade analysts from making decisions based on data with high uncertainty.
+We demonstrate several examples of VSUMs and present a crowdsourced evaluation showing that, compared to traditional bivariate maps, VSUMs bias analysts to more heavily weight uncertainty information in decision-making tasks.
 
@@
 VSUMs intentionally alias together data values with high uncertainty, affording greater discriminability as certainty increases.
-Rather than a square bivariate map, with unique outputs for each relevant combination of data and uncertainty, VSUMs can be thought of as arcs: as uncertainty increases, values are mapped to smaller and smaller sets of outputs, culminating in a singularity where all inputs are mapped to an identical, highly uncertain mark regardless of their data value.
+Rather than a traditional two dimensional bivariate map, with unique outputs for each relevant combination of data and uncertainty, VSUMs can be thought of as arcs: as uncertainty increases, values are mapped to smaller and smaller sets of outputs, culminating in a singularity where all inputs are mapped to an identical, highly uncertain mark regardless of their data value.
 Figure X shows an example of a VSUM, compared to a more traditional bivariate map.
@@
 \section{Value-Suppressing Uncertainty Maps}
+
+Correll et al.
+use a version of VSUMs in their LayerCake genomics visualization tool \cite{correll2015layercake,correll2011visualizing}, where the reads with insufficient certainty are mapped to a smaller and smaller set of increasing grey colors, creating the effect of uncertain values retreating into ``fog'' while highly certain values remain prominent.
 

### 2017-03-22 paper/figures.tex
Added uncertainty references
Submitted abstract version.

@@
+%Figure list:
+% Vsum vs. traditional 2D heatmap vs. juxtaposed maps
+% Chart of color bins w/r/t CIELAB threshold, with iconic maps at intervals
+% Process figure
+% Real examples with different color maps
+
 \newcommand{\teaserFig}{
@@
 		\centering
-		\includegraphics[width=1.5in]{sample.eps}
+		\includegraphics[width=0.9\textwidth]{example.pdf}
 		\caption{Lookit! Lookit!}
@@
 }
+
+\newcommand{\exampleFig}{
+\begin{figure}
+	\centering
+	\includegraphics[width=0.9\columnwidth]{example.pdf}
+	\caption{A comparison between a standard bivariate map and a VSUM.
+Both use the same visual channels to encode value (position along the Viridis color map) and uncertainty (lightness and saturation), and have an equal standard of perceptual discriminability (at least 18 units of distance in CIELAB color space between colors).
+However, highly uncertain values result in colors that are very close together in the bivariate case, meaning the full bivariate map is only 9 bins under these constraints--- a 4x4 map would result in colors that are perceptually too close together.
+By contrast, the VSUM intentionally reduces bins when uncertainty is high, eventually aliasing all highly uncertain values to the same color.
+This decision affords more distinct colors in other regions of the map, and so an increase in overall bins to 15.
+The resulting map suppresses the value of uncertain data, but increases the discriminability of highly certain data.}
+	\label{fig:example}
+\end{figure}
+
+}

### 2017-03-22 paper/template.tex
Submitted abstract version.

@@
 Visualizations often use bivariate mappings to encode value and uncertainty simultaneously.
-Due to interference between visual channels, these bivariate maps are often limited in resolution.
+Due to interference between visual channels, these bivariate maps are often limited in discriminability.
 We contribute Value-Suppressing Uncertainty Maps (VSUMs), an encoding technique that intentionally impairs value discrimination as uncertainty increases.
@@
 A common goal in the communication of uncertainty is promoting \textbf{uncertainty-aware decisions}--- that is, the audience should be aware of the risks and rewards of certain decisions, modulate their confidence in their conclusions, and perhaps refrain from making a decision at all if there is too much uncertainty.
-In general, a goal of designers of visualizations of uncertain data is the \emph{information fusion} of data and uncertainty~\cite{riveiro2007evaluation}.
 A way that designers can contribute to this goal is by ensuring that uncertainty information is \textbf{well-integrated} with the rest of the data.
@@
 Simultaneous presentation of uncertainty and value necessitates the construction of a bivariate map--- a relation, in terms of visual variables, between a 2-tuple $(value$,$uncertainty)$ and a set of marks.
+The design of bivariate maps is difficult and, due to the interference and interplay between different visual variables, often leads to maps with very few discriminable categories.
 
@@
 Rather than a traditional two dimensional bivariate map, with unique outputs for each relevant combination of data and uncertainty, VSUMs can be thought of as arcs: as uncertainty increases, values are mapped to smaller and smaller sets of outputs, culminating in a singularity where all inputs are mapped to an identical, highly uncertain mark regardless of their data value.
-Figure X shows an example of a VSUM, compared to a more traditional bivariate map.
+Figure \ref{fig:example} shows an example of a VSUM, compared to a more traditional bivariate map.
 In this work, we describe the motivations behind the use of VSUMs, examples of their utility for decision-making under uncertainty, and validate that VSUMs in a crowd-sourced experiment.
 
+\exampleFig 
 \section{Related Works}
@@
 % The variables we'd want to use for uncertainty (like value or alpha or size) mess with color discriminability
+
 Despite the acknowledged importance of uncertainty in understanding data, explicit representation of uncertainty often missing from visualizations~\cite{boukhelifa2009uncertainty}. The absence of

### 2017-03-22 heatmap/heatmap.js

@@ -318,7 +318,7 @@ function main(){
   makeArcmap(300,0,250,maps.arc,arcScale,"legendArc");
 
   //var exampleData = gradientData(100,100);
-  var exampleData = randomData(10,10);
+  var exampleData = randomData(5,5);
   makeHeatmap(0,300,250,exampleData, squareScale, "exampleSquare");
   makeHeatmap(300,300,250,exampleData,arcScale,"exampleArc");
 }

### 2017-03-22 paper/template.tex
Dealt with often.
Update template.tex

@@
 That is, it should be difficult to discount or ignore the uncertainty in a dataset.
-Simultaneous presentation of uncertainty and value necessitates the construction of a bivariate map--- a relation, in terms of visual variables, between a 2-tuple $(value$,$uncertainty)$ and a set of marks.
-The design of bivariate maps is difficult and, due to the interference and interplay between different visual variables, often leads to maps with very few discriminable categories.
+Simultaneous presentation of uncertainty and value necessitates the construction of a bivariate map--- a relation, in terms of visual variables, between 2-tuples $(\text{value}, \text{uncertainty})$ and mark properties.
+The design of bivariate maps is difficult and, due to the interference and interplay between different visual variables, often leads to maps with few discriminable categories.
 
@@
 VSUMs intentionally alias together data values with high uncertainty, affording greater discriminability as certainty increases.
-Rather than a traditional two dimensional bivariate map, with unique outputs for each relevant combination of data and uncertainty, VSUMs can be thought of as arcs: as uncertainty increases, values are mapped to smaller and smaller sets of outputs, culminating in a singularity where all inputs are mapped to an identical, highly uncertain mark regardless of their data value.
+Rather than a traditional two dimensional bivariate map, with unique outputs for each relevant combination of value and uncertainty, VSUMs can be thought of as arcs: as uncertainty increases, values are mapped to smaller and smaller sets of outputs, culminating in a singularity where all inputs are mapped to an identical, highly uncertain mark regardless of their data value.
 Figure \ref{fig:example} shows an example of a VSUM, compared to a more traditional bivariate map.
-In this work, we describe the motivations behind the use of VSUMs, examples of their utility for decision-making under uncertainty, and validate that VSUMs in a crowd-sourced experiment.
+In this work, we describe the motivations behind the use of VSUMs, examples of their utility for decision-making under uncertainty, and validate that VSUMs TODO in a crowd-sourced experiment.
 

### 2017-03-22 paper/template.tex
DM edits + clarifying phrases in intro.

--- 2017-03-22 paper/template.tex
@@
 Visualizations often use bivariate mappings to encode value and uncertainty simultaneously.
-Due to interference between visual channels, these bivariate maps are often limited in discriminability.
+Due to interference between visual channels, these bivariate maps are limited in discriminability.
 We contribute Value-Suppressing Uncertainty Maps (VSUMs), an encoding technique that intentionally impairs value discrimination as uncertainty increases.
@@
 Simultaneous presentation of uncertainty and value necessitates the construction of a bivariate map--- a relation, in terms of visual variables, between a 2-tuple $(value$,$uncertainty)$ and a set of marks.
-The design of bivariate maps is difficult and, due to the interference and interplay between different visual variables, often leads to maps with very few discriminable categories.
+The design of bivariate maps is difficult and, due to the interference and interplay between different visual variables, often leads to maps with few discriminable categories.
 
@@
 VSUMs intentionally alias together data values with high uncertainty, affording greater discriminability as certainty increases.
-Rather than a traditional two dimensional bivariate map, with unique outputs for each relevant combination of data and uncertainty, VSUMs can be thought of as arcs: as uncertainty increases, values are mapped to smaller and smaller sets of outputs, culminating in a singularity where all inputs are mapped to an identical, highly uncertain mark regardless of their data value.
+Rather than a traditional two-dimensional bivariate map, with differing outputs for each combination of data and uncertainty, VSUMs can be thought of as arcs: as uncertainty increases, values are mapped to smaller and smaller sets of outputs, culminating in a singularity where all inputs are mapped to an identical, highly uncertain mark regardless of their data value.
 Figure \ref{fig:example} shows an example of a VSUM, compared to a more traditional bivariate map.
-In this work, we describe the motivations behind the use of VSUMs, examples of their utility for decision-making under uncertainty, and validate that VSUMs in a crowd-sourced experiment.
+In this work, we describe the motivations behind the use of VSUMs, examples of their utility for decision-making under uncertainty, and assess VSUMs in a crowd-sourced experiment.
+Our experimental results indicate that VSUMs promote better integration between uncertainty and data: X.
 

### 2017-03-22 paper/template.tex
More DM fixes.

@@
 That is, it should be difficult to discount or ignore the uncertainty in a dataset.
-Simultaneous presentation of uncertainty and value necessitates the construction of a bivariate map--- a relation, in terms of visual variables, between a 2-tuple $(value$,$uncertainty)$ and a set of marks.
+Simultaneous presentation of uncertainty and value necessitates the construction of a bivariate map--- a relation, in terms of visual variables, between 2-tuples $(\text{value}, \text{uncertainty})$ and mark properties.
 The design of bivariate maps is difficult and, due to the interference and interplay between different visual variables, often leads to maps with few discriminable categories.

### 2017-03-22 paper/template.tex
Added amsmath

@@
 \usepackage{subcaption}
+\usepackage{amsmath}
 \newcommand{\todo}[1]{\textbf{\textcolor{red}{TODO: #1}}\\}
@@
 VSUMs intentionally alias together data values with high uncertainty, affording greater discriminability as certainty increases.
-Rather than a traditional two-dimensional bivariate map, with differing outputs for each combination of data and uncertainty, VSUMs can be thought of as arcs: as uncertainty increases, values are mapped to smaller and smaller sets of outputs, culminating in a singularity where all inputs are mapped to an identical, highly uncertain mark regardless of their data value.
+Rather than a traditional two-dimensional bivariate map, with differing outputs for each combination of value and uncertainty, VSUMs can be thought of as arcs: as uncertainty increases, values are mapped to smaller and smaller sets of outputs, culminating in a singularity where all inputs are mapped to an identical, highly uncertain mark regardless of their data value.
 Figure \ref{fig:example} shows an example of a VSUM, compared to a more traditional bivariate map.

### 2017-03-23 paper/template.tex
First draft of related works section

@@
 \usepackage{amsmath}
-\newcommand{\todo}[1]{\textbf{\textcolor{red}{TODO: #1}}\\}
-
+\usepackage{hyperref}
 %% Paper title.
@@
 
-Despite the acknowledged importance of uncertainty in understanding data, explicit representation of uncertainty often missing from visualizations~\cite{boukhelifa2009uncertainty}. The absence of
+Despite the acknowledged importance of uncertainty in understanding data, explicit representation of uncertainty often missing from visualizations~\cite{boukhelifa2009uncertainty}.
+This is partially due to the complexity of uncertainty as a concept.
+In typologies from Thomson et al~\cite{thomson2005typology} and Buttenfield \& Beard~\cite{buttenfield1994graphical}, the authors note that many, occasionally contradictory, concepts can fall under the category of ``uncertainty,'' including data quality, sampling error, credibility, and provenance.
+This complicates
 
-The presentation Broadlie et al.~\cite{brodlie2012review}
+In their reviews of the state of the art in uncertainty visualization in the field, Greithe et al.~\cite{griethe2006visualization} and Brodlie et al.~\cite{brodlie2012review} present lists of potential techniques for conveying uncertainty in concert with data.
+Some of these options (interaction, animation, and sonification) are not applicable for static charts; even for dynamic charts, many users do not interact with charts in sufficient detail to recover uncertainty information~\cite{nyt2016}.
+While we acknowledge the potential utility of these techniques in uncertainty visualization (for instance, the animation used to convey sampling error in Hullman et al.'s \emph{HOPs}~\cite{hullman2015hypothetical}), we limit the scope of our discussion to techniques which center on static charts.
+
+To support our goal of making uncertainty information well-integrated with data, we focus on the techniques of \textbf{juxtaposition} (where uncertainty is presented alongside the data), \textbf{superposition} (where additional glyphs are overlaid on top of the data), or \textbf{explicit encoding} (where a spare visual channel, such as color, is used to encode the uncertainty simultaneously with the data).
+Gleicher et al.~\cite{gleicher2011visual} explore these three strategies from the perspective of affording \textbf{comparisons}; here, we examine them from the perspective of \textbf{information fusion}.
+
+Juxtaposition creates two explicit visualizations: an ``uncertainty map'' and a ``data map.'' MacEachren~\cite{maceachren1992visualizing} portrays uncertainty visualization as the design problem of how to \emph{unify} these two maps, and so this explicit separation can be counterproductive to the goal of information fusion.
+Juxtaposition also creates an additional search task for viewers, who must register hot spots in one map with corresponding features of the other map.
+Nevertheless, especially in cases where the quantification of uncertainty is a separate analysis process from the measuring of data values, juxtaposition is a common design tactic for conveying uncertainty.
+
+Superposition, or the overlay of glyphs representing uncertainty on top of a data map, is other common strategy.
+Ware's~\cite{ware2009quantitative} \emph{textons} are an example of this strategy: monochrome glyphs that can encode an ordinal variable, or a binned quantitative variable.
+Greithe et al.~\cite{griethe2006visualization} and MacEachren~\cite{maceachren1992visualizing,maceachren1998visualizing} present other examples of this technique, including overlaid grids with ``fuzzy'' lines to indicate uncertainty, or overlaid treemaps with coarser and coarser resolution in uncertain regions.
+Costs to this approach include obscuring the underlying datamap (which can cause uncertainty information to dominate the display~\cite{brodlie2012review}), potential visual interactions (for instance, simultaneous contrast effects) that can impair the legibility of glyphs, and the necessity of performing spatial binning in order to ensure that there are a discrete number of uncertainty glyphs.
+
+VSUMs are intended for situations where designers explicitly encode uncertainty simultaneously with data.
+In addition to overcoming the drawbacks of the other techniques mentioned in this section, we believe that explicit encoding represents the most general form of the solution of how to fuse uncertainty information with data.
+The most pressing design decisions for explicit encoding are 1) which visual variable to use to encode uncertainty and 2) how to design bivariate maps which afford the encoding of both value and uncertainty.
+We deal with these two questions in greater detail below.
+
 \subsection{Visual Variables for Representing Uncertainty}
 
-MacEachren et al.~\cite{maceachren2012visual} evaluate a number of visual variables for representing uncertainty. 
-\subsection{Interactions Between Visual Variables}
+The decision to explicitly encode uncertainty increases the dimensionality of the data, and so requires the use of (at least one) additional visual channel ~\cite{brodlie2012review}.
+Uncertainty therefore inherently increases the visual complexity of a visualization.
+When the data are already complex to convey, and many of the more common or accurate visual variables are in use, allocating an additional dimension is non-trivial.
+As the number of dimensions increases, finding visual variables that are both perceptually accurate (in either estimation of quantity or discrimination of category) as well as perceptually separable from all the other encoding channels, becomes more and more difficult.
+
+A further hurdle is that not all visual variables are well-suited for conveying uncertainty.
+MacEachren et al.~\cite{maceachren2012visual} evaluate a number of visual variables with respect to their \emph{semiotic fit} for representing uncertainty.
+They observe that certain visual variables such as blurriness and transparency seem to have a more intuitive connection to uncertainty than other variables such as shape or hue.
+Unfortunately, Boukhelifa et al.~\cite{boukhelifa2012evaluating} discover that many visual variables habitually used for conveying uncertainty, such as blur and value, are also difficult to estimate.
+This results in a preference/performance gap where designers must choose between encoding uncertainty in a way that is intuitive but error prone, or use higher fidelity channels that are difficult to interpret.
+
+\subsection{Bivariate Maps}
+
+In for visualizations like choropleth maps, heatmaps, or treemaps, visual channels such as position and length are reserved for data variables other than value (such as geographic location or relative size).
+In these situations, data value is habitually encoded using color.
+Colors in such univariate quantitative color maps should be sufficiently far apart as to be perceptually distinguishable~\cite{ware1988color}, and vary in lightness as well as hue in order to afford an implicit ordering of value~\cite{borland2007rainbow,rogowitz2001blair}.
+Given these constraints, and the perceptual non-orthogonality of color channels, the construction of \emph{bivariate} color maps is difficult.
+Bivariate color maps out to be separable (the two variables can be distinguished independently) but also expressive and easy to interpret.
+Bivariate maps with these constraints are often limited to relatively few categories (say, a 3x3 matrix as in Figure \ref{fig:example}) \cite{robertson1986generation,trumbo1981theory}.
+Even so, the resulting bivariate maps are often difficult to interpret \cite{wainer1980empirical}.
+
+Dunn~\cite{dunn1989dynamic} attempts to circumvent these issues by dynamically allocating bins to the bivariate map.
+Since distinct, separable colors are limited, the dynamic approach allocates bins to regions of the highest utility.
+For instance, placing the dividing line between bins in areas aligned with a trend line allows the color to encode information about the sign of the residual with respect to each variable.
+This insight, that the bins in a bivariate map are a limited resource and so should be ``spent'' wisely, drives the creation of VSUMs, which allocate more bins in a bivariate map to regions with higher certainty.
 
@@
 
+We present Value-Suppressing Uncertainty Maps, a technique for creating bivariate maps of data \emph{value} and \emph{uncertainty}.
+These maps circumvent the issues in the creation of traditional bivariate maps by making the following assumptions:
+
+\begin{enumerate}
+	\item \textbf{Designers have a limited budget of discriminable marks.} That is, viewers cannot easily or reliably discriminate between an infinite number of categories.
+Rather than accept this perceptual ambiguity, designers ought to \emph{alias} together marks into a finite, but sufficiently discriminable, set.
+	\item \textbf{Uncertain values should have less weight than certain values.} Weight, in this case, might mean literal visual weight, or impact on the decision-making process. 
+\end{enumerate}
+
+%MC this paragraph is a little iffy, and sort of "the lady doth protest too much" territory.
+we might be able to get away with "people bin colors all the time.
+and in fact, you really should be binning your colors anyway."
+The first assumption seems to contradict the \emph{expressiveness principle}~\cite{mackinlay1986automating}, where the visual encoding of the data should reflect the composition of the data.
+For instance, if my data are a collection of measurements of some continuous variable, then it would make sense to encode these data using a continuous \emph{visual} variable, which would then mean an arbitrary number of potential marks.
+For instance, a bar can have an arbitrary, non-integer heights.
+However, for many visualizations, the perceptual ambiguity caused by having marks that cannot be reliably disambiguated is higher than the gain from having a visual encoding that maps readily to the backing data.
+Padilla et al.~\cite{padilla2017evaluating} present one such example, where binned heatmaps results in higher accuracy in tasks than in continuous heatmaps.
+
+In some cases, this second assumption is violated.
+For instance, an analyst might be interested in ``long tail risks'' or other ``black swan events,'' where the impact of an value, no matter how uncertain, must be considered and planned for~\cite{taleb2011black}.
+Other analysis tasks (such as filtering out outliers), require increased, rather than decreased, discrimbinability when uncertainty low.
+However, for many information fusion tasks, the assumption is that uncertainty is related to data quality, or the certainty of particular data values \cite{riveiro2007evaluation}.
+As with null hypothesis significance testing, the aim would then be to avoid Type I errors, where patterns arising from statistical noise or measurement error are interpreted as significant signals.
+
+If both of these assumptions hold, then it follows that the designer of a bivariate map should allocate more mark types to certain values, and fewer mark types to uncertain values.
+VSUMs codify this decision by \emph{reducing the number of mark categories for representing value as uncertainty increases.} The general VSUM algorithm is as follows:
+
+\begin{enumerate}
+	\item \textbf{Select a visual variable to encode value}. In most cases in this paper, this variable is position along some established color scale.
+	\item \textbf{Select a visual variable to encode uncertainty}.
+We choose variables that have a somewhat high semiotic connection with uncertainty (such as lightness and size), but still have a reasonable number of perceptually discriminable layers (as opposed to other, more intuitive channels such as blur or sketchiness).
+	\item \textbf{Select a discriminability threshold}.
+This is some minimum distance, in perceptual space, between marks.
+For many visual variables (such as length or hue), the perceptual discriminability between marks has been assessed in prior literature, and so the only step is to choose some value in terms of the just noticeable distance (JND).
+For less common visual variables, this perceptual space must be built up empirically (as in Demiralp et al.~\cite{demiralp2014learning}).
+	\item \textbf{Choose how quickly categories \emph{degrade}}.
+That is, if there are $n$ bins at some uncertainty level $m$, how many bins should there be when uncertainty increases to level $m+1$?
+In Figure \ref{fig:example}, this function is $\left \lfloor {\frac{n}{2}}\right \rfloor$, meaning that the number of value bins decreases by half for each level of uncertainty bins.
+If there is too much degradation, then there are fewer levels of uncertainty.
+If there is no degradation, the resulting map is just a traditional bivariate map.
+	\item \textbf{Choose the largest number of initial categories $n$ such that the discriminability threshold is met.} For each pair of marks in the map, perform a pairwise comparison of discriminability.
+If no marks are perceptually ambiguous by the standards of your threshold, then the map is complete.
+Otherwise, decrement $n$.
+\end{enumerate}
+
+Figure X shows this process in detail, and how it can be used to generate a bivariate map for presenting uncertainty.
+Javascript code for generating VSUMs given arbitrary d3 quantitative scale functions is available at \url{URL REMOVED FOR REVIEW}.
+
 Correll et al.
 use a version of VSUMs in their LayerCake genomics visualization tool \cite{correll2015layercake,correll2011visualizing}, where the reads with insufficient certainty are mapped to a smaller and smaller set of increasing grey colors, creating the effect of uncertain values retreating into ``fog'' while highly certain values remain prominent.
-
-\subsection{Design}
 

### 2017-03-23 paper/figures.tex
First draft of related works section
Added performance figure.

@@
 This decision affords more distinct colors in other regions of the map, and so an increase in overall bins to 15.
-The resulting map suppresses the value of uncertain data, but increases the discriminability of highly certain data.}
+The resulting map suppresses the value of uncertain data, but increases the discriminability of certain data.}
 	\label{fig:example}
 \end{figure}
+}
 
+\newcommand{\performanceFig}{
+	\begin{figure}
+		\centering
+		\includegraphics[width=0.9\columnwidth]{performance.pdf}
+		\caption{
+			The number of discrete color categories in traditional bivariate color maps versus VSUMs, assuming value is encoded by the Viridis color map~\protect\cite{garnier2015viridis}, and uncertainty is encoded by saturation/value (whiter values are more uncertain).
+Since uncertain colors are closer together than certain colors, and VSUMs intentionally require fewer color bins in those regions, VSUMs always have more color categories than the corresponding 2D bivariate map.
+The red lines denote:\\
+		    \textbf{1)} Theoretical 50\% JND from CIELAB specification. \\
+			\textbf{2)} 50\% JND from controlled lab study of Mahy et al.~\protect\cite{mahy1994evaluation}.\\
+			\textbf{3)} 50\% JND from Szafir et al.\protect\cite{szafir2014adapting}.\\
+			\textbf{4)} 50\% JND for small visual targets from Stone et al.~\protect\cite{stone2014engineering}.\\
+			\textbf{5)} 50\% The threshold resulting in the simplest possible (2x2) bivariate map. \\
 	    }
+		\label{fig:example}
+	\end{figure}
+}

### 2017-03-23 paper/template.tex
Typo in related works
Added performance figure.

@@
 
-Superposition, or the overlay of glyphs representing uncertainty on top of a data map, is other common strategy.
+Superposition, or the overlay of glyphs representing uncertainty on top of a data map, is another common strategy.
 Ware's~\cite{ware2009quantitative} \emph{textons} are an example of this strategy: monochrome glyphs that can encode an ordinal variable, or a binned quantitative variable.
@@
 Costs to this approach include obscuring the underlying datamap (which can cause uncertainty information to dominate the display~\cite{brodlie2012review}), potential visual interactions (for instance, simultaneous contrast effects) that can impair the legibility of glyphs, and the necessity of performing spatial binning in order to ensure that there are a discrete number of uncertainty glyphs.
+How this binning is performed can introduce undesirable artifacts on the resulting visualization, and suppress important signals, or highlight spurious ones \cite{battersby2016shapes}.
 
@@
 Correll et al.
-use a version of VSUMs in their LayerCake genomics visualization tool \cite{correll2015layercake,correll2011visualizing}, where the reads with insufficient certainty are mapped to a smaller and smaller set of increasing grey colors, creating the effect of uncertain values retreating into ``fog'' while highly certain values remain prominent.
+use a version of VSUMs in their LayerCake genomics visualization tool \cite{correll2015layercake,correll2011visualizing}, where marks representing genomic data with insufficient certainty are mapped to a smaller and smaller set of increasingly grey colors, creating the effect of uncertain values retreating into ``fog'' while highly certain values remain prominent.
+Other bivariate visualizations implicitly alias together uncertain values.
+For instance, if uncertainty is encoded by transparency, a maximally uncertain glyph may be entirely transparent, and so impossible to distinguish from any other maximally uncertain glyph.
+Other channels with a semiotic connection ot uncertainty, such as saturation, value, blur, or size, also have deleterious effects on the disambiguation of colors and shapes.
+In both of these cases, this property of aliasing is ad hoc, and places no guarantees on the discriminability of colors.
+The binning and degradation approach of VSUMs makes this choice to alias values explicit to both the designer and the viewer, and results in a bivariate mapping with known perceptual properties.
+
+\subsection{Design Considerations}
+
+\performanceFig
+
+There are multiple choices that designers must make before creating a VSUM.
+In particular, they must choose 1) which visual channels map to value and uncertainty, 2) a threshold of perceptual distinguishability (and, implicitly, a metric for evaluating discriminability), and 3) a degradation function to determine how quickly uncertain values are aliased.
+Often, the graphical conventions of a chart
 

### 2017-03-24 paper/template.tex
First draft minus experiments+examples+discussion

@@
 In particular, they must choose 1) which visual channels map to value and uncertainty, 2) a threshold of perceptual distinguishability (and, implicitly, a metric for evaluating discriminability), and 3) a degradation function to determine how quickly uncertain values are aliased.
-Often, the graphical conventions of a chart
+
+Often, the graphical conventions of a chart reserve variables such as position and area for other uses, and determine the visual channel to be used for value.
+In thematic maps, for instance, value is habitually encoded using color, while vertical and horizontal position are reserved for location.
+In this paper, we primarily deal with the case where value is encoded using color.
+While there are some similar conventions for representing uncertainty, generally designers have a greater degree of freedom in choosing a visual channel for uncertainty~\cite{maceachren1992visualizing}.
+We recommend channels with both a strong semiotic connection to uncertainty as well as a relatively large number of perceptually distinguishable levels.
+In this paper, we primarily discuss value/saturation (or, equivalently, transparency against a white background) and size.
+Correll \& Gleicher~\cite{correll2013error} show that even audiences without statistical backgrounds can interpret uncertainty information encoded in these channels.
+
+Another reason for limiting our discussion of bivariate mappings to color and size channels is that the perceptual distinguishability of these channels has been well-studied by prior work.
+CIELAB and other color spaces are designed with the intention of perceptual uniformity--- that is, distance in the color space should universally correspond with perceptual distance.
+Recent work~\cite{stone2014engineering,szafir2014adapting} has investigated the discriminability of colors in such spaces not only under laboratory conditions, but also among the disparate display conditions and color vision proficiency of crowdworkers.
+Similarly, formative work by Cleveland \& McGill~\cite{cleveland1984graphical} assesses the discriminability of marks of different height, and this too has been extended both to the more variable population of crowdworkers, and to the area channel specifically~\cite{heer2010crowdsourcing,talbot2014four}.
+We use values from these works as benchmarks for selecting thresholds of discriminability.
+The ``just noticeably difference'' (JND) of these channels provides a useful minimum; in practice, we would wish to create maps with marks that are far more reliably distinguishable than the 50\% or 75\% discriminability rate of a JND.
+Figure \ref{fig:performance} shows how these benchmarks can be used to select reasonable thresholds.
+
+How quickly the number of bins decreases is another parameter of VSUMs.
+If this degradation is too large, then we may only have a few discriminable levels of uncertainty.
+If this degradation is too small, then we can be overwhelmed with colors.
+For instance, if we only decrement by one bin per level of uncertainty, there would be \emph{1,485} different colors in a VSUM with a tolerance thresold of $1$ unit in CIELAB space.
+In the examples in this paper,  we halve the number of bins at each increasing level of uncertainty.
+When the initial bins at the highly certain region is a power of $2$ (as in Fig.
+\ref{fig:example}), this gives a tree structure to the resulting VSUM.
+As uncertainty increases, bins in the previous level map to one and only one ``parent'' in the next uncertainty level, resulting in fewer crossings of color boundaries among rows.
 

### 2017-03-24 paper/template.tex
Check in draft
First round of JH edits.

@@
 Uncertainty is an inescapable component of collecting, presenting, and using data.
-A common goal in the communication of uncertainty is promoting \textbf{uncertainty-aware decisions}--- that is, the audience should be aware of the risks and rewards of certain decisions, modulate their confidence in their conclusions, and perhaps refrain from making a decision at all if there is too much uncertainty.
-A way that designers can contribute to this goal is by ensuring that uncertainty information is \textbf{well-integrated} with the rest of the data.
+A common goal in the communication of uncertainty is promoting \emph{uncertainty-aware decisions}--- that is, the audience should be aware of the risks and rewards of certain decisions, modulate their confidence in their conclusions, and perhaps refrain from making a decision at all if there is too much uncertainty.
+A way that designers can contribute to this goal is by ensuring that uncertainty information is \emph{well-integrated} with the rest of the data.
 That is, it should be difficult to discount or ignore the uncertainty in a dataset.
@@
 In this paper, we present \textbf{Value-Suppressing Uncertainty Maps} (VSUMs) for integrating data and uncertainty information in thematic maps.
-VSUMs intentionally alias together data values with high uncertainty, affording greater discriminability as certainty increases.
+VSUMs intentionally alias together data values with high uncertainty, affording greater discriminability as uncertainty decreases.
 Rather than a traditional two-dimensional bivariate map, with differing outputs for each combination of value and uncertainty, VSUMs can be thought of as arcs: as uncertainty increases, values are mapped to smaller and smaller sets of outputs, culminating in a singularity where all inputs are mapped to an identical, highly uncertain mark regardless of their data value.
@@
 \exampleFig 
-\section{Related Works}
+\section{Related Work}
 
@@
 
-To support our goal of making uncertainty information well-integrated with data, we focus on the techniques of \textbf{juxtaposition} (where uncertainty is presented alongside the data), \textbf{superposition} (where additional glyphs are overlaid on top of the data), or \textbf{explicit encoding} (where a spare visual channel, such as color, is used to encode the uncertainty simultaneously with the data).
-Gleicher et al.~\cite{gleicher2011visual} explore these three strategies from the perspective of affording \textbf{comparisons}; here, we examine them from the perspective of \textbf{information fusion}.
+To support our goal of making uncertainty information well-integrated with data, we focus on the techniques of \emph{juxtaposition} (where uncertainty is presented alongside the data), \textbf{superposition} (where additional glyphs are overlaid on top of the data), or \emph{explicit encoding} (where a spare visual channel, such as color, is used to encode the uncertainty simultaneously with the data).
+Gleicher et al.~\cite{gleicher2011visual} explore these three strategies from the perspective of affording \emph{comparisons}; here, we examine them from the perspective of information fusion.
 
-Juxtaposition creates two explicit visualizations: an ``uncertainty map'' and a ``data map.'' MacEachren~\cite{maceachren1992visualizing} portrays uncertainty visualization as the design problem of how to \emph{unify} these two maps, and so this explicit separation can be counterproductive to the goal of information fusion.
+\subsection{Strategies for Encoding Uncertainty}
+
+\emph{Juxtaposition} creates two explicit visualizations: an ``uncertainty map'' and a ``data map.'' MacEachren~\cite{maceachren1992visualizing} portrays uncertainty visualization as the design problem of how to \emph{unify} these two maps, and so this explicit separation can be counterproductive to the goal of information fusion.
 Juxtaposition also creates an additional search task for viewers, who must register hot spots in one map with corresponding features of the other map.
@@
 
-Superposition, or the overlay of glyphs representing uncertainty on top of a data map, is another common strategy.
+\emph{Superposition}, or the overlay of glyphs representing uncertainty on top of a data map, is another common strategy.
 Ware's~\cite{ware2009quantitative} \emph{textons} are an example of this strategy: monochrome glyphs that can encode an ordinal variable, or a binned quantitative variable.
@@
 
-VSUMs are intended for situations where designers explicitly encode uncertainty simultaneously with data.
+\emph{Explicit encoding} directly represents uncertainty with a visual channel, simultaneously with data.
+VSUMs employ this strategy.
 In addition to overcoming the drawbacks of the other techniques mentioned in this section, we believe that explicit encoding represents the most general form of the solution of how to fuse uncertainty information with data.
@@
 Given these constraints, and the perceptual non-orthogonality of color channels, the construction of \emph{bivariate} color maps is difficult.
-Bivariate color maps out to be separable (the two variables can be distinguished independently) but also expressive and easy to interpret.
+Viewers need to be able to independently assess the variables of interest, but the resulting bivariate map should also be expressive and easy to interpret.
 Bivariate maps with these constraints are often limited to relatively few categories (say, a 3x3 matrix as in Figure \ref{fig:example}) \cite{robertson1986generation,trumbo1981theory}.
-Even so, the resulting bivariate maps are often difficult to interpret \cite{wainer1980empirical}.
+However, even simple bivariate maps can be difficult for the general audience to interpret \cite{wainer1980empirical}.
 
@@
 For instance, placing the dividing line between bins in areas aligned with a trend line allows the color to encode information about the sign of the residual with respect to each variable.
-This insight, that the bins in a bivariate map are a limited resource and so should be ``spent'' wisely, drives the creation of VSUMs, which allocate more bins in a bivariate map to regions with higher certainty.
+This insight, that the bins in a bivariate map are a limited resource and so should be ``spent'' wisely, drives the creation of VSUMs, which allocate more bins in a bivariate map to regions with less uncertainty.
 
@@
 Other analysis tasks (such as filtering out outliers), require increased, rather than decreased, discrimbinability when uncertainty low.
-However, for many information fusion tasks, the assumption is that uncertainty is related to data quality, or the certainty of particular data values \cite{riveiro2007evaluation}.
+However, for many information fusion tasks, the assumption is that uncertainty is related to data quality, or the uncertainty of particular data values \cite{riveiro2007evaluation}.
 As with null hypothesis significance testing, the aim would then be to avoid Type I errors, where patterns arising from statistical noise or measurement error are interpreted as significant signals.
@@
 Correll et al.
-use a version of VSUMs in their LayerCake genomics visualization tool \cite{correll2015layercake,correll2011visualizing}, where marks representing genomic data with insufficient certainty are mapped to a smaller and smaller set of increasingly grey colors, creating the effect of uncertain values retreating into ``fog'' while highly certain values remain prominent.
+use a version of VSUMs in their LayerCake genomics visualization tool \cite{correll2015layercake,correll2011visualizing}, where marks representing genomic data with increasing uncertainty are mapped to a smaller and smaller set of increasingly grey colors, creating the effect of uncertain values retreating into ``fog'' while highly certain values remain prominent.
 Other bivariate visualizations implicitly alias together uncertain values.
@@
 Figure \ref{fig:performance} shows how these benchmarks can be used to select reasonable thresholds.
+Other factors, such as nameability and aesthetic preference, can be added in as additional weights impacting the desirability of certain color palettes, as in Gramazio et al.~\cite{gramazio2017colorgorical}.
 

### 2017-03-24 paper/figures.tex
First draft minus experiments+examples+discussion
Added hex glyphs for size/color bivariate map

@@
 This decision affords more distinct colors in other regions of the map, and so an increase in overall bins to 15.
-The resulting map suppresses the value of uncertain data, but increases the discriminability of certain data.}
+The resulting map suppresses the value of uncertain data, but increases the discriminability of data with low uncertainty.}
 	\label{fig:example}
@@
 			The number of discrete color categories in traditional bivariate color maps versus VSUMs, assuming value is encoded by the Viridis color map~\protect\cite{garnier2015viridis}, and uncertainty is encoded by saturation/value (whiter values are more uncertain).
-Since uncertain colors are closer together than certain colors, and VSUMs intentionally require fewer color bins in those regions, VSUMs always have more color categories than the corresponding 2D bivariate map.
+Since these whiter uncertain colors are closer together in CIELAB space, and VSUMs intentionally require fewer color bins in those regions, VSUMs always have more color categories than the corresponding 2D bivariate map.
 The red lines denote:\\
@@
 			\textbf{4)} 50\% JND for small visual targets from Stone et al.~\protect\cite{stone2014engineering}.\\
-			\textbf{5)} 50\% The threshold resulting in the simplest possible (2x2) bivariate map. \\
+			\textbf{5)} The threshold resulting in the simplest possible (2x2) bivariate map. \\
 	    }
-		\label{fig:example}
+		\label{fig:performance}
 	\end{figure}

### 2017-03-24 heatmap/heatmap.js

@@ -84,6 +84,62 @@ function makeArc(d,size,rows,cols){
   return arc;
 }
 
+function makeArcHexmap(x,y,size,data,z,name){
+  //creates an svg "wedge" map
+  var w = size/data[0].length;
+  var h = size/data.length;
+
+  var arcmap = svg.append("g")
+             .attr("transform","translate("+(x+size/2)+","+(y+size)+")");
+
+  if(name){
+    arcmap.attr("id",name);
+  }
+
+  arcmap.selectAll("g")
+  .data(data)
+  .enter()
+  .append("g")
+  .selectAll("path")
+    .data(function(d,i){ return d.map(function(val){ return {r:i, v:val};});})
+    .enter()
+    .append("path")
+    .datum(function(d,i){ d.c = i; return d; })
+    .attr("d", function(d){ return makeHexagon(d,size,data.length,data[d.r].length);})
+    .attr("fill", function(d){ return z(d.v);});
+}
+
+function makeHexagon(d,size,rows,cols){
+  var angle = d3.scaleLinear().domain([0,cols]).range([-Math.PI/6,Math.PI/6]);
+  var radius = d3.scaleLinear().domain([0,rows]).range([size,0]);
+
+  var x,y,a,r;
+  a = angle(d.c+0.5);
+  //punt on determining the radius for now, it should really be a function of
+  // a) the number of bins in a rows
+  // b) some scale function passed by the user.
+  r = (radius(d.r+1)-radius(d.r))/2;
+
+  //convert from polar to cartesian
+  x = Math.cos(a-(Math.PI/2))*radius(d.r+0.5);
+  y = Math.sin(a-(Math.PI/2))*radius(d.r+0.5);
+  var h = Math.sqrt(3)/2;
+  var points = [
+    {"x": r+x, "y": y },
+    {"x": (r/2)+x, "y": y+(h*r) },
+    {"x": (-r/2)+x, "y": y+(h*r)},
+    {"x": -r+x, "y": y },
+    {"x": (-r/2)+x, "y": (-r*h)+y },
+    {"x": (r/2)+x, "y": (-r*h)+y }
+  ];
+
+  var line = d3.line()
+    .x(function(d){ return d.x;})
+    .y(function(d){ return d.y;});
+
+  return line(points);
+}
+
 //Data loading and generation
 
 function gradientData(rows, cols) {
@@ -315,7 +371,7 @@ function main(){
[…]

### 2017-03-25 paper/template.tex
Pre-size manipulating hex maps.

@@
 
-Despite the acknowledged importance of uncertainty in understanding data, explicit representation of uncertainty often missing from visualizations~\cite{boukhelifa2009uncertainty}.
+Despite the acknowledged importance of uncertainty in understanding data, explicit representations of uncertainty are often missing from visualizations~\cite{boukhelifa2009uncertainty}.
 This is partially due to the complexity of uncertainty as a concept.
 In typologies from Thomson et al~\cite{thomson2005typology} and Buttenfield \& Beard~\cite{buttenfield1994graphical}, the authors note that many, occasionally contradictory, concepts can fall under the category of ``uncertainty,'' including data quality, sampling error, credibility, and provenance.
-This complicates
 
@@
 \emph{Superposition}, or the overlay of glyphs representing uncertainty on top of a data map, is another common strategy.
-Ware's~\cite{ware2009quantitative} \emph{textons} are an example of this strategy: monochrome glyphs that can encode an ordinal variable, or a binned quantitative variable.
-Greithe et al.~\cite{griethe2006visualization} and MacEachren~\cite{maceachren1992visualizing,maceachren1998visualizing} present other examples of this technique, including overlaid grids with ``fuzzy'' lines to indicate uncertainty, or overlaid treemaps with coarser and coarser resolution in uncertain regions.
+Error bars are a common example of this strategy: a mark indicating measurement error is superimposed atop a mark representing a point value.
+Ware's~\cite{ware2009quantitative} \emph{textons} are an example of this strategy applied to thematic maps: monochrome glyphs that can encode an ordinal variable, or a binned quantitative variable.
+Greithe et al.~\cite{griethe2006visualization} and MacEachren~\cite{maceachren1992visualizing,maceachren1998visualizing} present other examples of superimposed uncertainty in thematic maps, including overlaid grids with ``fuzzy'' lines to indicate uncertainty, or overlaid treemaps with coarser and coarser resolution in uncertain regions.
 Costs to this approach include obscuring the underlying datamap (which can cause uncertainty information to dominate the display~\cite{brodlie2012review}), potential visual interactions (for instance, simultaneous contrast effects) that can impair the legibility of glyphs, and the necessity of performing spatial binning in order to ensure that there are a discrete number of uncertainty glyphs.
@@
 This insight, that the bins in a bivariate map are a limited resource and so should be ``spent'' wisely, drives the creation of VSUMs, which allocate more bins in a bivariate map to regions with less uncertainty.
+
+\subsection{}
 

### 2017-03-25 heatmap/heatmap.js

@@ -91,13 +91,16 @@ function makeArcHexmap(x,y,size,data,z,name){
   var h = size/data.length;
   r/=2;
 
-  var xPos = d3.scaleLinear().domain([0,data[0].length]).range([r,size-r]);
+  //other display options:
+  //var xPos = d3.scaleLinear().domain([0,data[0].length]).range([r,size-r]);
   //pack hexs as tightly as possible in y
   //var yPos = d3.scaleLinear().domain([0,data.length]).range([r,2*data.length*r]);
   //pack hexs as loosely as possible in y
-  var yPos = d3.scaleLinear().domain([0,data.length-1]).range([h/2,size-(h/2)]);
+  //var yPos = d3.scaleLinear().domain([0,data.length-1]).range([h/2,size-(h/2)]);
+
+  //currently, lay hexes out in radial fashion.
   var arcmap = svg.append("g")
-             .attr("transform","translate("+(x)+","+(y)+")");
+             .attr("transform","translate("+(x+size/2)+","+(y+size)+")");
 
   if(name){
     arcmap.attr("id",name);
@@ -112,11 +115,23 @@ function makeArcHexmap(x,y,size,data,z,name){
     .enter()
     .append("path")
     .datum(function(d,i){ d.c = i; return d; })
-    .attr("d", function(d){ return makeHexagon(d,r,xPos(d.c + ((data[0].length - data[d.r].length)/2)),yPos(d.r));})
+    .attr("d", function(d){ return makeArcHexagon(d,r,data.length,data[d.r].length,size);})
+  //  .attr("d", function(d){ return makeHexagon(r,xPos(d.c + ((data[0].length - data[d.r].length)/2)),yPos(d.r));})
     .attr("fill", function(d){ return z(d.v);});
 }
 
-function makeHexagon(d,r,x,y){
+function makeArcHexagon(d,r,rows,cols,size){
+  //radial layout
+  var x,y,a;
+  var angle = d3.scaleLinear().domain([0,cols]).range([-Math.PI/6,Math.PI/6]);
+  var radius = d3.scaleLinear().domain([0,rows]).range([size,0]);
+  a = angle(d.c+0.5);
+  x = Math.cos(a-(Math.PI/2))*radius(d.r+0.5);
+  y = Math.sin(a-(Math.PI/2))*radius(d.r+0.5);
+  return makeHexagon(r,x,y);
+}
+
+function makeHexagon(r,x,y){
[…]

### 2017-03-27 paper/template.tex
Typos

@@
 For instance, an analyst might be interested in ``long tail risks'' or other ``black swan events,'' where the impact of an value, no matter how uncertain, must be considered and planned for~\cite{taleb2011black}.
-Other analysis tasks (such as filtering out outliers), require increased, rather than decreased, discrimbinability when uncertainty low.
+Other analysis tasks (such as filtering out outliers), require increased, rather than decreased, discriminability when uncertainty low.
 However, for many information fusion tasks, the assumption is that uncertainty is related to data quality, or the uncertainty of particular data values \cite{riveiro2007evaluation}.
@@
 We use values from these works as benchmarks for selecting thresholds of discriminability.
-The ``just noticeably difference'' (JND) of these channels provides a useful minimum; in practice, we would wish to create maps with marks that are far more reliably distinguishable than the 50\% or 75\% discriminability rate of a JND.
+The ``just noticeable difference'' (JND) of these channels provides a useful minimum; in practice, we would wish to create maps with marks that are far more reliably distinguishable than the 50\% or 75\% discriminability rate of a JND.
 Figure \ref{fig:performance} shows how these benchmarks can be used to select reasonable thresholds.
@@
 If this degradation is too small, then we can be overwhelmed with colors.
-For instance, if we only decrement by one bin per level of uncertainty, there would be \emph{1,485} different colors in a VSUM with a tolerance thresold of $1$ unit in CIELAB space.
+For instance, if we only decrement by one bin per level of uncertainty, there would be \emph{1,485} different colors in a VSUM with a tolerance threshold of $1$ unit in CIELAB space.
 In the examples in this paper,  we halve the number of bins at each increasing level of uncertainty.

### 2017-03-27 heatmap/heatmap.js
No need to quantize viridis

@@ -3,18 +3,10 @@
 
 var svg = d3.select("body").append("svg");
 
-// viridis
-var NUM_STEPS = 1000;
-var map = d3.range(NUM_STEPS).map(d => d3.interpolateViridis(d/(NUM_STEPS-1)));
+var map = d3.interpolateViridis;
 
 //Other maps to test:
-//inferno
-//var map = d3.range(NUM_STEPS).map(d => d3.interpolateInferno(d/(NUM_STEPS-1)));
-//colorbrewer spectral
-//var map = d3.schemeSpectral[11];
-
-var z = d3.scaleQuantize().range(map);
-
+// interpolateInferno
 
 //Chart creation functions
 
@@ -183,7 +175,7 @@ function cDist(color1,color2){
   var c1 = d3.lab(color1);
   var c2 = d3.lab(color2);
 
-  return Math.sqrt( Math.pow( (c1.l-c2.l),2) + Math.pow(c1.a-c2.a,2) + Math.pow(c1.b - c2.b,2));
+  return Math.sqrt(Math.pow( (c1.l-c2.l),2) + Math.pow(c1.a-c2.a,2) + Math.pow(c1.b - c2.b,2));
 }
 
 function minDist(colorRamp){
@@ -237,11 +229,10 @@ function checkMap(mapName){
       }
     });
 
-    if(colorList.length==0){
+    if (colorList.length==0) {
       console.log("No valid color map");
       return false;
-    }
-    else{
+    } else {
       var closest = minDist(colorList);
       console.log( "The two closest colors:(" + closest.c1 +"," + closest.c2 +") are "+closest.minD+" apart in CIELAB.");
       return closest.minD>5;
@@ -290,8 +281,6 @@ function makeArcScaleData(n) {
 }
 
 function makeScaleFunction(scaleData){
-  var cScale = d3.scaleQuantize().domain([0,1]).range(map);
-
   //a quantized scale function from a given scale
   return function f(d){
     var u,v;
@@ -307,7 +296,7 @@ function makeScaleFunction(scaleData){
     u = value.u;
     v = value.v;
 
-    var c = (d3.hsl(cScale(v)));
+    var c = (d3.hsl(map(v)));
     return d3.interpolateLab(c,"white")(u);
   }
 }
@@ -345,7 +334,7 @@ function makeMaps(threshold){
     }
     n++;
   }
-  if(!closest){
+  if (!closest) {
[…]

### 2017-03-27 heatmap/heatmap.js
More interpolators

@@ -6,7 +6,11 @@ var svg = d3.select("body").append("svg");
 var map = d3.interpolateViridis;
 
 //Other maps to test:
+// interpolateViridis
 // interpolateInferno
+// interpolatePlasma
+// interpolateWarm
+// interpolateCool
 
 //Chart creation functions
 

### 2017-03-27 heatmap/heatmap.js
Delete unused function

@@ -165,7 +165,7 @@ function randomData(rows,cols){
   for(var i = 0; i<rows; i++){
     data[i] = [];
     for(var j = 0; j<cols; j++){
-      data[i].push({ u: Math.random(), v: Math.random()});
+      data[i].push({ u: Math.random(), v: Math.random() });
     }
   }
 
@@ -216,31 +216,6 @@ function distanceMatrix(scaleData){
     }
   }
   return matrix;
-}
-
-// TODO: delete
-function checkMap(mapName){
-  //Takes a d3 selection containing all the marks we care about
-  //Checks to make sure that colors it contains are sufficiently far apart.
-  var colorList = [];
-  var aColor;
-
-  mapName
-    .each(function(d){
-      aColor = d3.select(this).attr("fill");
-      if(!colorList.includes(aColor)){
-        colorList.push(aColor);
-      }
-    });
-
-    if (colorList.length==0) {
-      console.log("No valid color map");
-      return false;
-    } else {
-      var closest = minDist(colorList);
-      console.log( "The two closest colors:(" + closest.c1 +"," + closest.c2 +") are "+closest.minD+" apart in CIELAB.");
-      return closest.minD>5;
-    }
 }
 
 /**

### 2017-03-27 heatmap/heatmap.js
Move things around a bit

@@ -1,7 +1,7 @@
 
 //Global Variables
 
-var svg = d3.select("body").append("svg");
+var svg = d3.select("body").append("svg").append("g").attr("transform","translate(10, 10)");
 
 var map = d3.interpolateViridis;
 
@@ -348,11 +348,11 @@ function main(){
 
   makeHeatmap(0,0,250,maps.square, squareScale, "legendSquare");
   makeArcmap(300,0,250,maps.arc,arcScale,"legendArc");
+  makeArcHexmap(600,0,250,maps.arc,arcScale,"exampleArc");
 
   //var exampleData = gradientData(100,100);
   var exampleData = randomData(5,5);
   makeHeatmap(0,300,250,exampleData, squareScale, "exampleSquare");
-  makeArcHexmap(300,300,250,maps.arc,arcScale,"exampleArc");
 }
 
 //Uncertainty maps

### 2017-03-27 heatmap/heatmap.js
Demo different maps

@@ -350,9 +350,13 @@ function main(){
   makeArcmap(300,0,250,maps.arc,arcScale,"legendArc");
   makeArcHexmap(600,0,250,maps.arc,arcScale,"exampleArc");
 
-  //var exampleData = gradientData(100,100);
-  var exampleData = randomData(5,5);
-  makeHeatmap(0,300,250,exampleData, squareScale, "exampleSquare");
+  var gradient = gradientData(20,20);
+  makeHeatmap(0,300,250,gradient, squareScale);
+  makeHeatmap(300,300,250,gradient, arcScale);
+
+  var random = randomData(5,5);
+  makeHeatmap(0,600,250,random, squareScale);
+  makeHeatmap(300,600,250,random, arcScale); 
 }
 
 //Uncertainty maps

### 2017-03-27 heatmap/heatmap.js
Flight heatmap

@@ -1,7 +1,7 @@
 
 //Global Variables
 
-var svg = d3.select("body").append("svg").append("g").attr("transform","translate(10, 10)");
+var svg = d3.select("body").append("svg").append("g").attr("transform","translate(15, 15)");
 
 var map = d3.interpolateViridis;
 
@@ -356,7 +356,66 @@ function main(){
 
   var random = randomData(5,5);
   makeHeatmap(0,600,250,random, squareScale);
-  makeHeatmap(300,600,250,random, arcScale); 
+  makeHeatmap(300,600,250,random, arcScale);
+
+  d3.csv("data.csv", function(data) {
+    data = data.map(function(d) {
+      return {
+        DayOfWeek: +d.DayOfWeek,
+        DepDelay: +d.DepDelay,
+        DepTimeBlk: d.DepTimeBlk,
+        StdMeanErr: +d.StdMeanErr
+      }
+    });
+
+    var w = 600;
+    var h = 260;
+
+    var x = d3.scaleBand().range([0, w]).domain(data.map(function(d) { return d.DepTimeBlk; }));
+    var xAxis = d3.scalePoint().range([0, w]).domain([0, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20, 21, 22, 23, 24]);
+    var y = d3.scaleBand().range([0, h]).domain(data.map(function(d) { return d.DayOfWeek; }));
+    var yAxis = d3.scaleBand().range([0, h]).domain(["Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"]);
+   
+    var uScale = d3.scaleLinear().domain(d3.extent(data.map(function(d) { return d.StdMeanErr; }))).range([0,1]);
+    var vScale = d3.scaleLinear().domain(d3.extent(data.map(function(d) { return d.DepDelay; }))).range([0,1]);
+
+    var heatmap = svg.append("g")
+              .attr("transform","translate(0,900)");
+
+    heatmap.selectAll("rect")
+      .data(data)
+      .enter()
+      .append("rect")
+      .attr("x", function(d) { return x(d.DepTimeBlk); })
+      .attr("y", function(d) { return y(d.DayOfWeek); })
+      .attr("width", x.bandwidth())
+      .attr("height", y.bandwidth())
+      .attr("title", JSON.stringify)
+      .attr("fill", function(d) { return arcScale({
+        u: uScale(d.StdMeanErr),
[…]

### 2017-03-27 heatmap/heatmap.js
Comments

@@ -358,6 +358,7 @@ function main(){
   makeHeatmap(0,600,250,random, squareScale);
   makeHeatmap(300,600,250,random, arcScale);
 
+  // flight data example heatmap
   d3.csv("data.csv", function(data) {
     data = data.map(function(d) {
       return {
@@ -372,8 +373,10 @@ function main(){
     var h = 260;
 
     var x = d3.scaleBand().range([0, w]).domain(data.map(function(d) { return d.DepTimeBlk; }));
-    var xAxis = d3.scalePoint().range([0, w]).domain([0, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20, 21, 22, 23, 24]);
     var y = d3.scaleBand().range([0, h]).domain(data.map(function(d) { return d.DayOfWeek; }));
+
+    // special scales for axes
+    var xAxis = d3.scalePoint().range([0, w]).domain([0, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20, 21, 22, 23, 24]);
     var yAxis = d3.scaleBand().range([0, h]).domain(["Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"]);
    
     var uScale = d3.scaleLinear().domain(d3.extent(data.map(function(d) { return d.StdMeanErr; }))).range([0,1]);
@@ -391,11 +394,13 @@ function main(){
       .attr("width", x.bandwidth())
       .attr("height", y.bandwidth())
       .attr("title", JSON.stringify)
+      // set the color scale here
       .attr("fill", function(d) { return arcScale({
         u: uScale(d.StdMeanErr),
         v: vScale(d.DepDelay)
       });});
 
+    // axes
     heatmap.append("g")
       .attr("transform", "translate(0," + h + ")")
       .call(d3.axisBottom(xAxis));

### 2017-03-27 study/src/script.js
Study code to prep for merge

@@ -10,6 +10,9 @@ B. Superimposed Map: One heatmap, bivariate map down to as small a bin size as p
 Will be coarser than the juxtaposed map, since there's inteference between the two variables.
 C. VSUP: One heatmap, bivariate, but with uncertain values increasing aliased together.
 
+2. Map
+A. Value: (Viridis, Spectral?). Uncertainty: (Lightness, Size).
+
 Procedure:
 1. Get Consent.
 2. Tutorial showing examples of tasks, as well as a general primer on uncertainty.
@@ -68,6 +71,8 @@ function consent(){
 }
 
 function finishConsent(){
+  main.selectAll("*").remove();
+  postTest();
   //Thing to do when we've consented.
   //Ought to be setting up for a tutorial.
 }
@@ -120,8 +125,73 @@ function finishTask(){
   postTest();
 }
 
-function riskAversion(fromElement){
-  //A risk aversion assay, from
+function riskAversion(form){
+  //A risk aversion assay, from Mandrik and Bao 2005:
+
+  var items = [
+    "I do not feel comfortable taking chances.",
+    "I prefer situations that have foreseeable outcomes.",
+    "Before I make a decision, I like to be absolutely sure how things turn out.",
+    "I avoid situations that have uncertain outcomes",
+    "I feel comfortable improvising in new situations",
+    "I feel nervous when I have to make decisions in uncertain situations"
+  ];
+
+  var valences = [
+   1,
+   1,
+   1,
+   1,
+   -1,
+   1
+  ];
+
+  //Another option is the ROQ from Rohrmann 1997
+  //http://www.rohrmannresearch.net/pdfs/rohrmann-rac-roq.pdf
+
+  form.append("div")
+  .html("For the following items indicate how strongly you agree or disagree, where 1=\"Strongly Disagree\" and 7=\"Strongly Agree\":");
+
+  var dlist = form.append("ol");
+  var index = 0;
+
+  for(var item of items){
+    var question = dlist.append("li");
+
+    question.html(item);
+
+    var label = question.append("div");
+
+    label.classed("likert",true);
+
+    label.append("span")
+    .html("Strongly Disagree")
+    .attr("style","align-self: left");
+
+    label.append("span")
+    .html("Strongly Agree")
+    .attr("style","align-self: right");
+
+    var numbers = question.append("div");
+
+    numbers.classed("likert",true);
+
+    var questions = question.append("div");
+
+    questions.classed("likert",true);
+
+    for(var i = 1;i<=7;i++){
+      numbers.append("span")
[…]

### 2017-03-27 heatmap/heatmap.js
Separate svgs

@@ -1,7 +1,9 @@
 
 //Global Variables
 
-var svg = d3.select("body").append("svg").append("g").attr("transform","translate(15, 15)");
+var body = d3.select("body");
+
+var svg = body.append("svg").attr("height", 900).append("g").attr("transform","translate(10, 10)");
 
 var map = d3.interpolateViridis;
 
@@ -14,7 +16,7 @@ var map = d3.interpolateViridis;
 
 //Chart creation functions
 
-function makeHeatmap(x,y,size,data,z,name){
+function makeHeatmap(svg, x, y, size, data, z, name){
   //creates an svg heatmap with a 2d matrix of data, and a mapping function z
   var w = size/data[0].length;
   var h = size/data.length;
@@ -346,17 +348,17 @@ function main(){
   var squareScale = makeScaleFunction(maps.square);
   var arcScale = makeScaleFunction(maps.arc);
 
-  makeHeatmap(0,0,250,maps.square, squareScale, "legendSquare");
+  makeHeatmap(svg, 0,0,250,maps.square, squareScale, "legendSquare");
   makeArcmap(300,0,250,maps.arc,arcScale,"legendArc");
   makeArcHexmap(600,0,250,maps.arc,arcScale,"exampleArc");
 
   var gradient = gradientData(20,20);
-  makeHeatmap(0,300,250,gradient, squareScale);
-  makeHeatmap(300,300,250,gradient, arcScale);
+  makeHeatmap(svg, 0,300,250,gradient, squareScale);
+  makeHeatmap(svg, 300,300,250,gradient, arcScale);
 
   var random = randomData(5,5);
-  makeHeatmap(0,600,250,random, squareScale);
-  makeHeatmap(300,600,250,random, arcScale);
+  makeHeatmap(svg, 0,600,250,random, squareScale);
+  makeHeatmap(svg, 300,600,250,random, arcScale);
 
   // flight data example heatmap
   d3.csv("data.csv", function(data) {
@@ -369,8 +371,8 @@ function main(){
       }
     });
 
-    var w = 600;
-    var h = 260;
+    var w = 580;
+    var h = 240;
 
     var x = d3.scaleBand().range([0, w]).domain(data.map(function(d) { return d.DepTimeBlk; }));
     var y = d3.scaleBand().range([0, h]).domain(data.map(function(d) { return d.DayOfWeek; }));
@@ -382,8 +384,8 @@ function main(){
[…]

### 2017-03-27 heatmap/heatmap.js
Extract flight example

@@ -44,7 +44,7 @@ function makeHeatmap(svg, x, y, size, data, z, name){
     .attr("fill", function(d){ return z(d.v);});
 }
 
-function makeArcmap(x,y,size,data,z,name){
+function makeArcmap(svg, x, y, size, data, z, name){
   //creates an svg "wedge" map
   var w = size/data[0].length;
   var h = size/data.length;
@@ -82,7 +82,7 @@ function makeArc(d,size,rows,cols){
   return arc();
 }
 
-function makeArcHexmap(x,y,size,data,z,name){
+function makeArcHexmap(svg, x, y, size, data, z, name){
   //creates an svg "hexwedge" map of regular hexagons
 
   var r = size/data[0].length;
@@ -349,8 +349,8 @@ function main(){
   var arcScale = makeScaleFunction(maps.arc);
 
   makeHeatmap(svg, 0,0,250,maps.square, squareScale, "legendSquare");
-  makeArcmap(300,0,250,maps.arc,arcScale,"legendArc");
-  makeArcHexmap(600,0,250,maps.arc,arcScale,"exampleArc");
+  makeArcmap(svg, 300,0,250,maps.arc,arcScale,"legendArc");
+  makeArcHexmap(svg, 600,0,250,maps.arc,arcScale,"exampleArc");
 
   var gradient = gradientData(20,20);
   makeHeatmap(svg, 0,300,250,gradient, squareScale);
@@ -370,62 +370,64 @@ function main(){
         StdMeanErr: +d.StdMeanErr
       }
     });
-
-    var w = 580;
-    var h = 240;
-
-    var x = d3.scaleBand().range([0, w]).domain(data.map(function(d) { return d.DepTimeBlk; }));
-    var y = d3.scaleBand().range([0, h]).domain(data.map(function(d) { return d.DayOfWeek; }));
-
-    // special scales for axes
-    var xAxis = d3.scalePoint().range([0, w]).domain([0, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20, 21, 22, 23, 24]);
-    var yAxis = d3.scaleBand().range([0, h]).domain(["Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"]);
-   
-    var uScale = d3.scaleLinear().domain(d3.extent(data.map(function(d) { return d.StdMeanErr; }))).range([0,1]);
[…]

### 2017-03-27 paper/template.tex
Study code to prep for merge

@@
 Colors in such univariate quantitative color maps should be sufficiently far apart as to be perceptually distinguishable~\cite{ware1988color}, and vary in lightness as well as hue in order to afford an implicit ordering of value~\cite{borland2007rainbow,rogowitz2001blair}.
+Different choices of color maps can highlight different features of the data, and should be chosen with care based on the data types and important features to be visualized~\cite{rogowitz1996not}.
 Given these constraints, and the perceptual non-orthogonality of color channels, the construction of \emph{bivariate} color maps is difficult.
@@
 This insight, that the bins in a bivariate map are a limited resource and so should be ``spent'' wisely, drives the creation of VSUMs, which allocate more bins in a bivariate map to regions with less uncertainty.
-
-\subsection{}
 
@@
 If both of these assumptions hold, then it follows that the designer of a bivariate map should allocate more mark types to certain values, and fewer mark types to uncertain values.
-VSUMs codify this decision by \emph{reducing the number of mark categories for representing value as uncertainty increases.} The general VSUM algorithm is as follows:
+VSUMs codify this decision by \emph{reducing the number of mark categories for representing value as uncertainty increases.} This means that data encoded using a VSUM will make visible only the largest of differences in uncertain data, but highlight comparatively small differences in value when uncertainty is low.
+VSUMs therefore act as both a filtering mechanism (in that values with too much uncertainty are all mapped to the same glyph), as well as an implicit test of effect size (in that smaller and smaller changes in value are visible as uncertainty decreases).
+This strategy of dampening the low quality or high uncertainty values in maps in order to focus on more informative regions has measurable benefits, including the removal of statistically spurious visual patterns, and the highlighting of regions of interest~\cite{correll2017surprise}.
+
+The general VSUM algorithm is as follows:
 
@@
 For instance, if uncertainty is encoded by transparency, a maximally uncertain glyph may be entirely transparent, and so impossible to distinguish from any other maximally uncertain glyph.
-Other channels with a semiotic connection ot uncertainty, such as saturation, value, blur, or size, also have deleterious effects on the disambiguation of colors and shapes.
+Other channels with a semiotic connection to uncertainty, such as saturation, value, blur, or size, also have deleterious effects on the disambiguation of colors and shapes.
 In both of these cases, this property of aliasing is ad hoc, and places no guarantees on the discriminability of colors.

### 2017-03-27 heatmap/heatmap.js
Size/color glyphs now plugged into Stone et al. equations.

@@ -184,31 +184,115 @@ function cDist(color1,color2){
   return Math.sqrt(Math.pow( (c1.l-c2.l),2) + Math.pow(c1.a-c2.a,2) + Math.pow(c1.b - c2.b,2));
 }
 
-function cSizeDist(color1,color2,size){
-  //Based on Stone 2014, we expect the same color difference to
-  //"count" for less, as sizes decrease.
-  //since we are defining tolerances as initial distances, rather than
-  //JNDs, "size" is the [0,1] percentage of the original glyph size.
+function toVisualAngle(pixels){
+  //assume 96 dpi and 24" from screen,
+  // so each pixel is 0.0213 degrees of visual angle.
+  return Math.abs(pixels*0.0213);
+}
 
+function sizeThreshold(color1,color2,size,p){
+  //Are two colors sufficiently distinct such that
+  //p% of people will discriminate between them?
+  var c1 = d3.lab(color1);
+  var c2 = d3.lab(color2);
 
-  //if ND(p,s) = tolerance
-  //then ND(p,s*size) = tolerance + 1/size * K
-  //where K(L) = 0.75, K(a) = 1.54, K(b) = 2.87
+  //per axis values
+  dist = [
+    Math.abs(c1.l-c2.l),
+    Math.abs(c1.a-c2.a),
+    Math.abs(c1.b-c2.b)
+  ];
 
-  if(size==1){
-    return cDist(color1,color2);
+  var nd = ND(p,size);
+
+  return (dist[0]>=nd[0] || dist[1]>=nd[1] || dist[2]>=nd[2]);
+}
+
+function pGivenS(color1,color2,size){
+  //What's the probability that these colors will be distinguished,
+  //given their size?
+  var c1 = d3.lab(color1);
+  var c2 = d3.lab(color2);
+  var axes = ["l","a","b"];
+  A = [
+    10.16,
+    10.68,
+    10.70
+  ];
+
+  B = [
+    1.50,
+    3.08,
+    5.74
+  ];
+
+  var p,diff;
+  var maxP = 0;
+
+  for(var i = 0;i<axes.length;i++){
+    diff = Math.abs(c1[axes[i]]-c2[axes[i]]);
+    p = diff/(A[i] + (B[i]/size));
+    if(p>maxP){
+      maxP = p;
+    }
   }
-  else{
-    var c1 = d3.lab(color1);
-    var c2 = d3.lab(color2);
 
[…]

### 2017-03-27 heatmap/heatmap.js
Legend scaffold

@@ -444,7 +444,7 @@ function main(){
         StdMeanErr: +d.StdMeanErr
       }
     });
-    makeFlightExample(arcScale, maps.arcmap, data);
+    makeFlightExample(arcScale, maps.arc, data);
   }); 
 }
 
@@ -501,7 +501,15 @@ function makeFlightExample(colorScale, map, data) {
     .text("Day of the Week");
 
   // legend
-  // makeArcmap(heatmap, w + 60,0,200,map,colorScale);
+  var legendX = w + 140;
+  var legendY = 30;
+  var legendSize = 160;
+  makeArcmap(heatmap, legendX, legendY, legendSize,map,colorScale);
+  makeArcLegend(heatmap, legendX, legendY, legendSize, [0,1], [0,1]);
+}
+
+function makeArcLegend(svg, x, y, size, vRange, uRange) {
+  
 }
 
 //Uncertainty maps

### 2017-03-27 heatmap/heatmap.js
Made size and color maps equivalent by tweaking threshold parameters

@@ -480,7 +480,7 @@ function makeMaps(threshold){
 
   n = 2;
   while (true) {
-    var startsize = toVisualAngle(25);
+    var startsize = toVisualAngle(45);
     var data = makeArcScaleData(n);
     if(colorSizeDiff(data,THRESHOLD,startsize)){
       arcSizeScaleData = data;
@@ -529,7 +529,7 @@ function main(){
       }
     });
     makeFlightExample(arcScale, maps.arcmap, data);
-  }); 
+  });
 }
 
 function makeFlightExample(colorScale, map, data) {
@@ -542,7 +542,7 @@ function makeFlightExample(colorScale, map, data) {
   // special scales for axes
   var xAxis = d3.scalePoint().range([0, w]).domain([0, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20, 21, 22, 23, 24]);
   var yAxis = d3.scaleBand().range([0, h]).domain(["Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"]);
-  
+
   var uScale = d3.scaleLinear().domain(d3.extent(data.map(function(d) { return d.StdMeanErr; }))).range([0,1]);
   var vScale = d3.scaleLinear().domain(d3.extent(data.map(function(d) { return d.DepDelay; }))).range([0,1]);
 

### 2017-03-27 heatmap/heatmap.js
Fixed merge

@@ -480,7 +480,7 @@ function makeMaps(threshold){
 
   n = 2;
   while (true) {
-    var startsize = toVisualAngle(25);
+    var startsize = toVisualAngle(45);
     var data = makeArcScaleData(n);
     if(colorSizeDiff(data,THRESHOLD,startsize)){
       arcSizeScaleData = data;
@@ -529,7 +529,7 @@ function main(){
       }
     });
     makeFlightExample(arcScale, maps.arc, data);
-  }); 
+  });
 }
 
 function makeFlightExample(colorScale, map, data) {
@@ -593,7 +593,7 @@ function makeFlightExample(colorScale, map, data) {
 }
 
 function makeArcLegend(svg, x, y, size, vRange, uRange) {
-  
+
 }
 
 //Uncertainty maps

### 2017-03-28 heatmap/heatmap.js
Flight data with legend

@@ -533,7 +533,7 @@ function main(){
 }
 
 function makeFlightExample(colorScale, map, data) {
-  var w = 580;
+  var w = 560;
   var h = 240;
 
   var x = d3.scaleBand().range([0, w]).domain(data.map(function(d) { return d.DepTimeBlk; }));
@@ -543,8 +543,8 @@ function makeFlightExample(colorScale, map, data) {
   var xAxis = d3.scalePoint().range([0, w]).domain([0, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20, 21, 22, 23, 24]);
   var yAxis = d3.scaleBand().range([0, h]).domain(["Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"]);
   
-  var uScale = d3.scaleLinear().domain(d3.extent(data.map(function(d) { return d.StdMeanErr; }))).range([0,1]);
-  var vScale = d3.scaleLinear().domain(d3.extent(data.map(function(d) { return d.DepDelay; }))).range([0,1]);
+  var uScale = d3.scaleLinear().domain(d3.extent(data.map(function(d) { return d.StdMeanErr; }))).range([0,1]).nice();
+  var vScale = d3.scaleLinear().domain(d3.extent(data.map(function(d) { return d.DepDelay; }))).range([0,1]).nice(map[0].length);
 
   var heatmap = body.append("svg").attr("width", w + 100).attr("height", h + 60).append("g")
             .attr("transform","translate(10,10)");
@@ -586,14 +586,77 @@ function makeFlightExample(colorScale, map, data) {
 
   // legend
   var legendX = w + 140;
-  var legendY = 30;
-  var legendSize = 160;
+  var legendY = 40;
+  var legendSize = 180;
   makeArcmap(heatmap, legendX, legendY, legendSize,map,colorScale);
-  makeArcLegend(heatmap, legendX, legendY, legendSize, [0,1], [0,1]);
+
+  makeArcLegend(heatmap, legendX, legendY, legendSize, map, vScale.ticks(10), uScale.domain(), "Departure Delay", "Standard Mean Error");
 }
 
-function makeArcLegend(svg, x, y, size, vRange, uRange) {
+function makeArcLegend(svg, x, y, size, map, vTicks, uDom, vTitle, uTitle) {
+  var uStep = (uDom[0] - uDom[1]) / map.length;
[…]

### 2017-03-28 heatmap/heatmap.js
minutes

@@ -590,7 +590,7 @@ function makeFlightExample(colorScale, map, data) {
   var legendSize = 180;
   makeArcmap(heatmap, legendX, legendY, legendSize,map,colorScale);
 
-  makeArcLegend(heatmap, legendX, legendY, legendSize, map, vScale.ticks(10), uScale.domain(), "Departure Delay", "Standard Mean Error");
+  makeArcLegend(heatmap, legendX, legendY, legendSize, map, vScale.ticks(10), uScale.domain(), "Departure Delay (minutes)", "Standard Mean Error");
 }
 
 function makeArcLegend(svg, x, y, size, map, vTicks, uDom, vTitle, uTitle) {

### 2017-03-28 heatmap/heatmap.js
Reverse error

@@ -594,8 +594,8 @@ function makeFlightExample(colorScale, map, data) {
 }
 
 function makeArcLegend(svg, x, y, size, map, vTicks, uDom, vTitle, uTitle) {
-  var uStep = (uDom[0] - uDom[1]) / map.length;
-  var uDom = d3.range(uDom[1], uDom[0] + uStep, uStep);
+  var uStep = (uDom[1] - uDom[0]) / map.length;
+  var uDom = d3.range(uDom[0], uDom[1] + uStep, uStep);
 
   var legend = svg.append("g").attr("transform", "translate(" + x + "," + y + ")")
 

### 2017-03-29 paper/template.tex
Study check in

@@
 
+We performed a crowd-sourced experiment on Amazon's Mechanical Turk in order to evaluate the effectiveness of VSUMs for integrating uncertainty and value information in thematic maps.
+
 %Show:
@@
 
-\subsection{Methods}
-
 %conditions:
@@
 
+\subsection{Methods}
+\subsubsection{Hypotheses}
+
+\subsection{Results}
 \subsubsection{Participants}
-\subsection{Results}
 
+We limited our population to Turkers from the United States, with a prior task approval rate of at least 90\%.
+As the experimental tasks required multi-hue color perception, we presented participants with a set of Ishihara plates as a pretest, and excluded participants who either misidentified the values in the plates, or who self-reported as having a color vision deficiency (CVD) in the post-test.
+Based on piloting, we paid participants \$X dollars an hour, for a target rate of \$8/hour.
+
+Certain stimuli in the identification task had no uncertainty
 \section{Discussion}

### 2017-03-29 study/src/script.js

@@ -30,6 +30,15 @@ var startTime;
 var main = d3.select("#fcontainer");
 var done = false;
 
+
+var taskOneQuestions = [
+  "Click on the most uncertain location in the map",
+  "Click on the lowest uncertainty, highest value location in the map",
+  "Click on the lowest uncertainty, lowest value location in the map"
+];
+
+var attacking = false;
+
 function gup(name){
   var regexS = "[\\?&]"+name+"=([^&#]*)";
   var regex = new RegExp( regexS );
@@ -73,33 +82,128 @@ function consent(){
 function finishConsent(){
   //Thing to do when we've consented.
   //Ought to be setting up for a tutorial.
-  tutorial();
+  ishihara();
+}
+
+function ishihara(){
+  var plates = [5,6,8,12,42];
+  dl.permute(plates);
+
+  main.selectAll("iframe").remove();
+  main.select("#answer").remove();
+
+  main.append("p")
+  .html("In order to be eligible for this study, you must be capable of distinguishing colors. That is, you should be free from color vision deficiency (color blindness).");
+
+  main.append("p")
+  .html("To verify this, for each of the following images, input the number that you see in each circle:");
+
+  var numPlates = 3;
+
+  var platesDiv = main.append("div")
+    .attr("id","plates");
+
+  for(var i = 0;i<3;i++){
+    platesDiv.append("div")
+    .append("img")
+    .attr("width","300px")
+    .attr("src","img/ishihara/"+plates[i]+".png");
+
+    platesDiv.append("div")
+    .html("Visible number:")
+    .append("input")
+    .attr("type","number")
+    .attr("name","plate"+plates[i])
+    .attr("id","plate"+plates[i]);
+
+  }
+
+  main.append("p")
+  .append("input")
+  .attr("class","button")
+  .attr("id","answer")
+  .attr("name","answer")
+  .attr("value","Ready")
+  .on("click",finishIshihara);
+
+}
+
+function finishIshihara(){
+  var plates = [];
+  main.select("#plates").selectAll("input").each( function(){
+    plates.push({"right": d3.select(this).attr("name").replace("plate",""), "guess": d3.select(this).property("value")});
+  });
+  console.log(plates);
+  var correct = true;
+
+  for(plate of plates){
+    correct = correct && plate.right==(plate.guess+"");
+  }
+
+  if(correct){
+    tutorial();
+  }
+  else{
+    ineligible();
+  }
+
 }
 
 function tutorial(){
[…]

### 2017-03-30 paper/template.tex
Demo check in

@@
 Different choices of color maps can highlight different features of the data, and should be chosen with care based on the data types and important features to be visualized~\cite{rogowitz1996not}.
-Given these constraints, and the perceptual non-orthogonality of color channels, the construction of \emph{bivariate} color maps is difficult.
+Given these considerations, and the perceptual non-orthogonality of color channels, the construction of \emph{bivariate} color maps is difficult.
 Viewers need to be able to independently assess the variables of interest, but the resulting bivariate map should also be expressive and easy to interpret.
 Bivariate maps with these constraints are often limited to relatively few categories (say, a 3x3 matrix as in Figure \ref{fig:example}) \cite{robertson1986generation,trumbo1981theory}.
+In general, the quality of bivariate color maps is a multivariate measure involving consideration of not just the component color channels, but also the interpolation scheme and the color of the surround~\cite{bernard2015survey}.
 However, even simple bivariate maps can be difficult for the general audience to interpret \cite{wainer1980empirical}.
@@
 Dunn~\cite{dunn1989dynamic} attempts to circumvent these issues by dynamically allocating bins to the bivariate map.
-Since distinct, separable colors are limited, the dynamic approach allocates bins to regions of the highest utility.
+Since the amount of distinct, separable colors is limited, the dynamic approach allocates bins to regions of the highest utility.
 For instance, placing the dividing line between bins in areas aligned with a trend line allows the color to encode information about the sign of the residual with respect to each variable.
@@
 We performed a crowd-sourced experiment on Amazon's Mechanical Turk in order to evaluate the effectiveness of VSUMs for integrating uncertainty and value information in thematic maps.
+This focus on integration meant that we limited our experimental tasks to scenarios where the participants needed to consider both value and uncertainty before making a decision.
+After a pretest and short tutorial, we gave participants two main tasks:
+
+\begin{enumerate}
+	\item [T1] an \textbf{identification} task, where we gaves participants charts with value and uncertainty information, and asked them to locate specific regions.
+E.g., ``click on the region of the chart with highest uncertainty \emph{and} lowest value.''
+	\item [T2] a \textbf{prediction} task, similar to Battleship, where we gave participants charts with both \emph{forecast} and \emph{forecast uncertainty} information, and asked the participants to place tokens on the board in order to optimize expected value.
+E.g., ``place your $4$ ships on the safest locations on the board.''
+\end{enumerate}
+We discuss the two experimental tasks in more detail in the following sections.
+
+After completion of the main task, we solicited demographics information, and performed a domain-agnostic general risk aversion assay~\cite{mandrik2005exploring}.
+
+\subsubsection{Identification Task}
+
+\subsubsection{Prediction Task}
+
+For the prediction task, participants were given the rules of a game similar to Battleship.
+Greis et al.~\cite{greis2016decision} employ these game-like experimental tasks to frame uncertainty information, which can be abstract or complex, to the general audience.
+In our task, the participant and a (fictional) adversary have to place tokens representing ships on a spatial grid.
+Without direct access to the location of the opponent's ships, the player then has to place tokens representing missile strikes on their opponent's grid.
+If a player fires a missile at a square containing a ship, then that ship is sunk.
+The objective is the maximize the number of enemy ships hit, while minimizing the number of your own ships that are hit.
+In our task, participants were given a chart encoding predictions on the location of enemy ships (in the attacking case) or potential missile strikes (in the defending case).
+The \emph{value} component of the prediction was the likelihood of a item being present in a particular region.
+The \emph{uncertainty} component was the confidence in this prediction.
+
+We included both a \textbf{gain} framing (given predictions of where your opponent's ships are, where should you launch your missiles?), and a \textbf{loss} framing (given predictions of where your opponent's missiles are, where should you place your ships?) in order to account for systematic differences in human judgments under uncertainty.
+Tversky \& Kahneman~\cite{tversky1985framing} illustrate that framings in terms of gains or losses produce reliably different outcomes.
+In general, the function of the value of gains is concave, but is convex for losses~\cite{kahneman1979prospect}.
+That is, avoiding a large gain is often more valuable than striving for a similarly large gain.
 
@@
 
-\subsection{Methods}
 \subsubsection{Hypotheses}
@@
 
-Certain stimuli in the identification task had no uncertainty
 \section{Discussion}

### 2017-03-30 study/src/script.js

@@ -34,6 +34,9 @@ var questionNum = 1;
 var taskOneStimuli = [];
 var taskTwoStimuli = [];
 
+var taskTwoGrid = [];
+var taskTwoTokens = 0;
+
 //var map = d3.interpolateViridis;
 
 function scaleGenerator(map){
@@ -163,7 +166,6 @@ function finishIshihara(){
   main.select("#plates").selectAll("input").each( function(){
     plates.push({"right": d3.select(this).attr("name").replace("plate",""), "guess": d3.select(this).property("value")});
   });
-  console.log(plates);
   var correct = true;
 
   for(plate of plates){
@@ -180,7 +182,7 @@ function finishIshihara(){
 }
 
 function tutorial(){
-  //Set up tutorial. When we're done, start the main task
+  //Set up tutorial. When we're done, start the first task
   main.selectAll("*").remove();
 
   main.append("iframe")
@@ -204,7 +206,7 @@ function makeTaskOneStimuli(){
   var stimuli = [];
 
   var types = ["juxta","2D","vsum"];
-  var ramps = ["viridisLightness","viridisSize","plasmaLightness","plasmaSize"];
+  var ramps = ["viridisLightness","viridisSize"];
   var sizes = ["4","8"];
   var questions = [
     "Click on the map location with the <b>highest</b> uncertainty",
@@ -212,10 +214,42 @@ function makeTaskOneStimuli(){
     "Click on the map location with the <b>lowest</b> uncertainty AND <b>lowest</b> value"
   ];
 
+  var perLevel = 1;
+
   for(type of types){
     for(size of sizes){
       for(question of questions){
-        stimuli.push( {"type":type, "size": size, "question":question});
+        for(ramp of ramps){
+          for(var i = 0;i<perLevel;i++){
+            stimuli.push( {"type":type, "ramp": ramp, "size": size, "question":question});
+          }
+        }
+      }
+    }
+  }
+
+  dl.permute(stimuli);
+  return stimuli;
+}
+
+function makeTaskTwoStimuli(){
+  var stimuli = [];
+
+  var types = ["juxta","2D","vsum"];
+  var ramps = ["lightness","size"];
+  var sizes = ["4","8"];
+  var roles = ["att","def"];
+
+  var perLevel = 1;
+
+  for(type of types){
+    for(size of sizes){
+      for(ramp of ramps){
[…]

### 2017-03-30 study/src/script.js
Checkin

@@ -331,7 +331,7 @@ function taskOne(){
 
   main.append("svg")
   .attr("id","legend")
-  .attr("style","width: 150px; height: 200px;");
+  .attr("style","width: 400px; height: 200px;");
 
 }
 
@@ -362,8 +362,8 @@ function revealTaskOne(){
   switch(stim.type){
     case "vsum":
     makeHeatmap(mapSvg,100,0,200,makeTaskOneMap(stim.size),vMap.arcScale);
-    makeArcmap(legendSvg, 20, 60, 80,vMap.arc,vMap.arcScale);
-    makeArcLegend(legendSvg, 20, 60, 80, vMap.arc, [0,100], [0,100], "Value", "Uncertainty");
+    makeArcmap(legendSvg, 20, 60, 160,vMap.arc,vMap.arcScale);
+    makeArcLegend(legendSvg, 20, 60, 160, vMap.arc, [0,100], [0,100], "Value", "Uncertainty");
     break;
 
     case "juxta":
@@ -377,8 +377,8 @@ function revealTaskOne(){
     case "2D":
     default:
     makeHeatmap(mapSvg,100,0,200,makeTaskOneMap(stim.size),vMap.squareScale);
-    makeHeatmap(legendSvg, 20, 60, 80,vMap.square,vMap.squareScale);
-    makeHeatmapLegend(legendSvg, 20, 60, 80, vMap.square, [0,100], [0,100], "Value", "Uncertainty");
+    makeHeatmap(legendSvg, 20, 60, 160,vMap.square,vMap.squareScale);
+    makeHeatmapLegend(legendSvg, 20, 60, 160, vMap.square, [0,100], [0,100], "Value", "Uncertainty");
     break;
   }
 
@@ -452,7 +452,7 @@ function taskTwo(){
 
   main.append("svg")
   .attr("id","legend")
-  .attr("style","width: 150px; height: 200px;");
+  .attr("style","width: 400px; height: 200px;");
 
   main.append("input")
   .attr("class","button")
@@ -512,7 +512,7 @@ function initializeTaskTwo(){
     case "vsum":
     makeHeatmap(mapSvg,100,0,200,makeTaskOneMap(stim.size),vMap.arcScale);
     makeArcmap(legendSvg, 20, 60, 80,vMap.arc,vMap.arcScale);
[…]

### 2017-03-30 study/src/script.js
Button is button

@@ -103,6 +103,7 @@ function consent(){
   .attr("id","answer")
   .attr("name","answer")
   .attr("value","I Consent")
+  .attr("type", "button")
   .on("click",finishConsent);
 
   if(assignmentId=="ASSIGNMENT_ID_NOT_AVAILABLE"){

### 2017-03-30 study/src/script.js
Buttons are buttons, remove iframes border

@@ -98,12 +98,11 @@ function consent(){
   main.append("iframe")
   .attr("src","consent.html");
 
-  var readyBtn = main.append("input")
+  var readyBtn = main.append("button")
   .attr("class","button")
   .attr("id","answer")
   .attr("name","answer")
-  .attr("value","I Consent")
-  .attr("type", "button")
+  .text("I Consent")
   .on("click",finishConsent);
 
   if(assignmentId=="ASSIGNMENT_ID_NOT_AVAILABLE"){
@@ -153,11 +152,11 @@ function ishihara(){
   }
 
   main.append("p")
-  .append("input")
+  .append("button")
   .attr("class","button")
   .attr("id","answer")
   .attr("name","answer")
-  .attr("value","Ready")
+  .text("Ready")
   .on("click",finishIshihara);
 
 }
@@ -189,11 +188,11 @@ function tutorial(){
   main.append("iframe")
   .attr("src","tutorial.html");
 
-  main.append("input")
+  main.append("button")
   .attr("class","button")
   .attr("id","answer")
   .attr("name","answer")
-  .attr("value","Ready")
+  .text("Ready")
   .on("click",finishTutorial);
 }
 
@@ -318,11 +317,11 @@ function taskOne(){
 
   main.append("p")
   .attr("id","question")
-  .append("input")
+  .append("button")
   .attr("class","button")
   .attr("id","answer")
   .attr("name","answer")
-  .attr("value","Ready")
+  .text("Ready")
   .attr("style","position: relative; top: 225px;")
   .on("click",revealTaskOne);
 
@@ -343,11 +342,11 @@ function initializeTaskOne(){
   d3.select("#prompt")
   .html("Click the \"Ready\" button to begin.");
 
-  d3.select("#question").append("input")
+  d3.select("#question").append("button")
   .attr("class","button")
   .attr("id","answer")
   .attr("name","answer")
-  .attr("value","Ready")
+  .text("Ready")
   .attr("style","position: relative; top: 225px;")
   .on("click",revealTaskOne);
 

### 2017-03-30 heatmap/main.js
Viridis for heatmaps

@@ -1,15 +1,15 @@
-//Other maps to test:
-// interpolateViridis
-// interpolateInferno
-// interpolatePlasma
-// interpolateWarm
-// interpolateCool
-
 function main(){
   var body = d3.select("body");
   var svg = body.append("svg").attr("height", 900).append("g").attr("transform","translate(10, 10)");
 
-  var map = d3.interpolatePlasma;
+  var map = d3.interpolateViridis;
+  // Other maps to test:
+  //
+  // interpolateViridis
+  // interpolateInferno
+  // interpolatePlasma
+  // interpolateWarm
+  // interpolateCool
 
   var maps = makeMaps(map, 18);
 

### 2017-03-30 heatmap/heatmap.js
Legend scales.

@@ -549,12 +549,13 @@ function makeArcLegend(svg, x, y, size, map, vTicks, uDom, vTitle, uTitle) {
 
   var uScale = d3.scalePoint().domain(uDom).range([0, size]);
   var uAxis = d3.axisRight(uScale);
+  var px = size/180;
   legend.append("g")
-    .attr("transform", "translate(" + (size + 5) + "," + 24 + ")rotate(30)")
+    .attr("transform", "translate(" + (size + 6 * px) + "," + (28 * px) + ")rotate(30)")
     .call(uAxis);
 
   legend.append("text")
-    .attr("transform", "translate(" + (size + 10) + "," + (28 + size / 2) + ")rotate(-60)")
+    .attr("transform", "translate(" + (size + 10 * px) + "," + (28 * px + size / 2) + ")rotate(-60)")
     .style("text-anchor", "middle")
     .style("font-size", 13)
     .text(uTitle);
@@ -563,7 +564,7 @@ function makeArcLegend(svg, x, y, size, map, vTicks, uDom, vTitle, uTitle) {
     .domain([vTicks[0], vTicks[vTicks.length - 1]])
     .range([-30, 30]);
 
-  var offset = 3;
+  var offset = 3 * px;
 
   var arc = d3.arc()
     .innerRadius(size + offset)

### 2017-03-30 study/src/script.js
First round of DM study changes

@@ -40,7 +40,7 @@ var taskTwoTokens = 0;
 //var map = d3.interpolateViridis;
 
 function scaleGenerator(map){
-  var maps = makeMaps(map, 18);
+  var maps = makeMaps(map, 15);
   var uSL = makeuSL(map);
   var uSize = makeuSize(map);
 
@@ -100,6 +100,7 @@ function consent(){
 
   var readyBtn = main.append("button")
   .attr("class","button")
+  .attr("type","button")
   .attr("id","answer")
   .attr("name","answer")
   .text("I Consent")
@@ -154,6 +155,7 @@ function ishihara(){
   main.append("p")
   .append("button")
   .attr("class","button")
+  .attr("type","button")
   .attr("id","answer")
   .attr("name","answer")
   .text("Ready")
@@ -190,6 +192,7 @@ function tutorial(){
 
   main.append("button")
   .attr("class","button")
+  .attr("type","button")
   .attr("id","answer")
   .attr("name","answer")
   .text("Ready")
@@ -319,6 +322,7 @@ function taskOne(){
   .attr("id","question")
   .append("button")
   .attr("class","button")
+  .attr("type","button")
   .attr("id","answer")
   .attr("name","answer")
   .text("Ready")
@@ -327,11 +331,11 @@ function taskOne(){
 
   main.append("svg")
   .attr("id","map")
-  .attr("style","width: 400px; height: 200px;");
+  .attr("style","width: 450px; height: 200px;");
 
   main.append("svg")
   .attr("id","legend")
-  .attr("style","width: 400px; height: 200px;");
+  .attr("style","width: 250px; height: 250px;");
 
 }
 
@@ -344,6 +348,7 @@ function initializeTaskOne(){
 
   d3.select("#question").append("button")
   .attr("class","button")
+  .attr("type","button")
   .attr("id","answer")
   .attr("name","answer")
   .text("Ready")
@@ -359,17 +364,20 @@ function revealTaskOne(){
   var stim = taskOneStimuli[questionNum-1];
   //"viridisLightness","viridisSize","plasmaLightness","plasmaSize"
 
+  var vLabel = "Value";
+  var uLabel = "Uncertainty";
+
   switch(stim.type){
     case "vsum":
     makeHeatmap(mapSvg,100,0,200,makeTaskOneMap(stim.size),vMap.arcScale);
-    makeArcmap(legendSvg, 20, 60, 160,vMap.arc,vMap.arcScale);
[…]

### 2017-03-30 heatmap/heatmap.js
Hexmaps

@@ -28,6 +28,30 @@ function makeHeatmap(svg, x, y, size, data, z, name){
     .attr("fill", function(d){ return z(d.v);});
 }
 
+function makeHexmap(svg, x, y, size, data, z, maxSize, name){
+  var w = size/data[0].length;
+  var h = size/data.length;
+
+  var heatmap = svg.append("g")
+             .attr("transform","translate("+(x+maxSize)+","+(y+maxSize)+")");
+
+  if(name){
+    heatmap.attr("id",name);
+  }
+
+  heatmap.selectAll("g")
+  .data(data)
+  .enter()
+  .append("g")
+  .selectAll("path")
+    .data(function(d,i){ return d.map(function(val){ return {r:i, v:val};});})
+    .enter()
+    .append("path")
+    .datum(function(d,i){ d.c = i; return d; })
+    .attr("d", function(d){ return makeHexagon(z(d.v).s,d.c*w,d.r*h);})
+    .attr("fill", function(d){ return z(d.v).c;});
+}
+
 function makeArcmap(svg, x, y, size, data, z, name){
   //creates an svg "wedge" map
   var w = size/data[0].length;
@@ -71,7 +95,6 @@ function makeArcHexmap(svg, x, y, size, data, z, name){
 
   var r = size/data[0].length;
   var h = size/data.length;
-  r/=2;
 
   //other display options:
   //var xPos = d3.scaleLinear().domain([0,data[0].length]).range([r,size-r]);
@@ -97,7 +120,7 @@ function makeArcHexmap(svg, x, y, size, data, z, name){
     .enter()
     .append("path")
     .datum(function(d,i){ d.c = i; return d; })
-    .attr("d", function(d){ return makeArcHexagon(d,r*z(d.v).s,data.length,data[d.r].length,size);})
+    .attr("d", function(d){ return makeArcHexagon(d,z(d.v).s,data.length,data[d.r].length,size);})
   //  .attr("d", function(d){ return makeHexagon(r,xPos(d.c + ((data[0].length - data[d.r].length)/2)),yPos(d.r));})
     .attr("fill", function(d){ return z(d.v).c;});
 }
@@ -408,14 +431,14 @@ function colorSizeDiff(scaleData, threshold, startSize, uSize){
   return passesSizeThreshold(colors,p,startSize);
 }
 
-function makeMaps(map, threshold){
[…]

### 2017-03-30 heatmap/main.js

@@ -11,10 +11,11 @@ function main(){
   // interpolateWarm
   // interpolateCool
 
-  var maps = makeMaps(map, 18);
+  var MAX_SIZE = 14;
+  var maps = makeMaps(map, 18, MAX_SIZE);
 
   var uSL = makeuSL(map);
-  var uSize = makeuSize(map);
+  var uSize = makeuSize(map, MAX_SIZE);
 
   var squareScale = makeScaleFunction(maps.square, uSL);
   var arcScale = makeScaleFunction(maps.arc, uSL);
@@ -24,13 +25,15 @@ function main(){
   makeArcmap(svg, 300,0,250,maps.arc,arcScale,"legendArc");
   makeArcHexmap(svg, 600,0,250,maps.arcSize,arcSizeScale,"legendSizeArc");
 
-  var gradient = gradientData(20,20);
+  var gradient = gradientData(8,8);
   makeHeatmap(svg, 0,300,250,gradient, squareScale);
   makeHeatmap(svg, 300,300,250,gradient, arcScale);
+  makeHexmap(svg, 600,300,250,gradient, arcSizeScale, MAX_SIZE);
 
   var random = randomData(5,5);
   makeHeatmap(svg, 0,600,250,random, squareScale);
   makeHeatmap(svg, 300,600,250,random, arcScale);
+  makeHexmap(svg, 600,600,250,random, arcSizeScale, MAX_SIZE);
 
   // flight data example heatmap
   d3.csv("data.csv", function(data) {

### 2017-03-30 paper/figures.tex
Jeff's revisions.

@@
 	\includegraphics[width=0.9\columnwidth]{example.pdf}
-	\caption{A comparison between a standard bivariate map and a VSUM.
+	\caption{A standard bivariate map and a VSUM.
 Both use the same visual channels to encode value (position along the Viridis color map) and uncertainty (lightness and saturation), and have an equal standard of perceptual discriminability (at least 18 units of distance in CIELAB color space between colors).
-However, highly uncertain values result in colors that are very close together in the bivariate case, meaning the full bivariate map is only 9 bins under these constraints--- a 4x4 map would result in colors that are perceptually too close together.
+However, highly uncertain values result in colors that are very close together in the bivariate case, meaning the full bivariate map is only 9 bins under these constraints\,---\,a 4x4 map would result in colors that are perceptually too close together.
 By contrast, the VSUM intentionally reduces bins when uncertainty is high, eventually aliasing all highly uncertain values to the same color.
@@
 Since these whiter uncertain colors are closer together in CIELAB space, and VSUMs intentionally require fewer color bins in those regions, VSUMs always have more color categories than the corresponding 2D bivariate map.
-The red lines denote:\\
-		    \textbf{1)} Theoretical 50\% JND from CIELAB specification. \\
-			\textbf{2)} 50\% JND from controlled lab study of Mahy et al.~\protect\cite{mahy1994evaluation}.\\
-			\textbf{3)} 50\% JND from Szafir et al.\protect\cite{szafir2014adapting}.\\
-			\textbf{4)} 50\% JND for small visual targets from Stone et al.~\protect\cite{stone2014engineering}.\\
-			\textbf{5)} The threshold resulting in the simplest possible (2x2) bivariate map. \\
+The red lines denote:
+		  \textbf{1)} Theoretical 50\% JND from CIELAB specification.
+			\textbf{2)} 50\% JND from controlled lab study of Mahy et al.~\protect\cite{mahy1994evaluation}.
+			\textbf{3)} 50\% JND from Szafir et al.\protect\cite{szafir2014adapting}.
+			\textbf{4)} 50\% JND for small visual targets from Stone et al.~\protect\cite{stone2014engineering}.
+			\textbf{5)} The threshold resulting in the simplest possible (2x2) bivariate map.
 	    }

### 2017-03-30 paper/template.tex
Jeff's revisions.

@@
 Uncertainty is an inescapable component of collecting, presenting, and using data.
-A common goal in the communication of uncertainty is promoting \emph{uncertainty-aware decisions}--- that is, the audience should be aware of the risks and rewards of certain decisions, modulate their confidence in their conclusions, and perhaps refrain from making a decision at all if there is too much uncertainty.
+A common goal in the communication of uncertainty is promoting \emph{uncertainty-aware decisions}: the audience should be aware of the risks and rewards of certain decisions, modulate their confidence in their conclusions, and perhaps refrain from making a decision at all if there is too much uncertainty.
 A way that designers can contribute to this goal is by ensuring that uncertainty information is \emph{well-integrated} with the rest of the data.
 That is, it should be difficult to discount or ignore the uncertainty in a dataset.
-Simultaneous presentation of uncertainty and value necessitates the construction of a bivariate map--- a relation, in terms of visual variables, between 2-tuples $(\text{value}, \text{uncertainty})$ and mark properties.
+
+Simultaneous presentation of uncertainty and value necessitates the construction of a bivariate map\,---\,a relation, in terms of visual variables, between 2-tuples $(\text{value}, \text{uncertainty})$ and mark properties.
 The design of bivariate maps is difficult and, due to the interference and interplay between different visual variables, often leads to maps with few discriminable categories.
 
-In this paper, we present \textbf{Value-Suppressing Uncertainty Maps} (VSUMs) for integrating data and uncertainty information in thematic maps.
+In this paper, we contribute \textbf{Value-Suppressing Uncertainty Maps} (VSUMs) for integrating data and uncertainty information in thematic maps.
 VSUMs intentionally alias together data values with high uncertainty, affording greater discriminability as uncertainty decreases.
 Rather than a traditional two-dimensional bivariate map, with differing outputs for each combination of value and uncertainty, VSUMs can be thought of as arcs: as uncertainty increases, values are mapped to smaller and smaller sets of outputs, culminating in a singularity where all inputs are mapped to an identical, highly uncertain mark regardless of their data value.
-Figure \ref{fig:example} shows an example of a VSUM, compared to a more traditional bivariate map.
-In this work, we describe the motivations behind the use of VSUMs, examples of their utility for decision-making under uncertainty, and assess VSUMs in a crowd-sourced experiment.
+Figure~\ref{fig:example} shows examples of both a VSUM and a more traditional bivariate map.
+
+We describe the motivations behind the use of VSUMs, examples of their utility for decision-making under uncertainty, and assess VSUMs in a crowdsourced experiment.
 Our experimental results indicate that VSUMs promote better integration between uncertainty and data: X.
@@
 This is partially due to the complexity of uncertainty as a concept.
-In typologies from Thomson et al~\cite{thomson2005typology} and Buttenfield \& Beard~\cite{buttenfield1994graphical}, the authors note that many, occasionally contradictory, concepts can fall under the category of ``uncertainty,'' including data quality, sampling error, credibility, and provenance.
+In typologies from Thomson et al.~\cite{thomson2005typology} and Buttenfield \& Beard~\cite{buttenfield1994graphical}, the authors note that many, occasionally contradictory, concepts can fall under the category of ``uncertainty,'' including data quality, sampling error, credibility, and provenance.
 
-In their reviews of the state of the art in uncertainty visualization in the field, Greithe et al.~\cite{griethe2006visualization} and Brodlie et al.~\cite{brodlie2012review} present lists of potential techniques for conveying uncertainty in concert with data.
+Reviewing the state of the art in uncertainty visualization in the field, Greithe et al.~\cite{griethe2006visualization} and Brodlie et al.~\cite{brodlie2012review} present lists of potential techniques for conveying uncertainty in concert with data.
 Some of these options (interaction, animation, and sonification) are not applicable for static charts; even for dynamic charts, many users do not interact with charts in sufficient detail to recover uncertainty information~\cite{nyt2016}.
@@
 
-To support our goal of making uncertainty information well-integrated with data, we focus on the techniques of \emph{juxtaposition} (where uncertainty is presented alongside the data), \textbf{superposition} (where additional glyphs are overlaid on top of the data), or \emph{explicit encoding} (where a spare visual channel, such as color, is used to encode the uncertainty simultaneously with the data).
-Gleicher et al.~\cite{gleicher2011visual} explore these three strategies from the perspective of affording \emph{comparisons}; here, we examine them from the perspective of information fusion.
+To support our goal of making uncertainty information well-integrated with data, we focus on the techniques of \emph{juxtaposition} (where uncertainty is presented alongside the data), \textbf{superposition} (where additional glyphs are overlaid on top of the data), and \emph{explicit encoding} (where an additional visual channel is used to encode the uncertainty).
+Gleicher et al.~\cite{gleicher2011visual} explore these three strategies from the perspective of affording \emph{comparisons}; here, we examine them from the perspective of uncertainty visualization.
 
@@
 Greithe et al.~\cite{griethe2006visualization} and MacEachren~\cite{maceachren1992visualizing,maceachren1998visualizing} present other examples of superimposed uncertainty in thematic maps, including overlaid grids with ``fuzzy'' lines to indicate uncertainty, or overlaid treemaps with coarser and coarser resolution in uncertain regions.
-Costs to this approach include obscuring the underlying datamap (which can cause uncertainty information to dominate the display~\cite{brodlie2012review}), potential visual interactions (for instance, simultaneous contrast effects) that can impair the legibility of glyphs, and the necessity of performing spatial binning in order to ensure that there are a discrete number of uncertainty glyphs.
-How this binning is performed can introduce undesirable artifacts on the resulting visualization, and suppress important signals, or highlight spurious ones \cite{battersby2016shapes}.
+Costs to this approach include obscuring the underlying data map (which can cause uncertainty information to dominate the display~\cite{brodlie2012review}), potential visual interactions that can impair the legibility of glyphs (for instance, simultaneous contrast effects), and the necessity of performing spatial binning in order to ensure that there are a discrete number of uncertainty glyphs.
+How this binning is performed can introduce undesirable artifacts on the resulting visualization, suppressing important signals or highlighting spurious ones~\cite{battersby2016shapes}.
 
@@
 They observe that certain visual variables such as blurriness and transparency seem to have a more intuitive connection to uncertainty than other variables such as shape or hue.
-Unfortunately, Boukhelifa et al.~\cite{boukhelifa2012evaluating} discover that many visual variables habitually used for conveying uncertainty, such as blur and value, are also difficult to estimate.
-This results in a preference/performance gap where designers must choose between encoding uncertainty in a way that is intuitive but error prone, or use higher fidelity channels that are difficult to interpret.
+Unfortunately, Boukhelifa et al.~\cite{boukhelifa2012evaluating} find that many visual variables habitually used for conveying uncertainty, such as blur and value, are also difficult to estimate.
+This results in a preference/performance gap where designers must choose between encoding uncertainty in a way that is intuitive but error prone, or use higher fidelity channels that may be more difficult to interpret.
 
@@
 
-In for visualizations like choropleth maps, heatmaps, or treemaps, visual channels such as position and length are reserved for data variables other than value (such as geographic location or relative size).
-In these situations, data value is habitually encoded using color.
+For visualizations such as choropleth maps, heatmaps, or treemaps, visual channels such as position and length are reserved for data variables other than value (such as geographic location or relative size).
+In these situations, data value is often encoded using color.
 Colors in such univariate quantitative color maps should be sufficiently far apart as to be perceptually distinguishable~\cite{ware1988color}, and vary in lightness as well as hue in order to afford an implicit ordering of value~\cite{borland2007rainbow,rogowitz2001blair}.
 Different choices of color maps can highlight different features of the data, and should be chosen with care based on the data types and important features to be visualized~\cite{rogowitz1996not}.
-Given these considerations, and the perceptual non-orthogonality of color channels, the construction of \emph{bivariate} color maps is difficult.
+
+Given these considerations, and the perceptual integrality of color channels, the construction of \emph{bivariate} color maps is difficult.
 Viewers need to be able to independently assess the variables of interest, but the resulting bivariate map should also be expressive and easy to interpret.
@@
 In general, the quality of bivariate color maps is a multivariate measure involving consideration of not just the component color channels, but also the interpolation scheme and the color of the surround~\cite{bernard2015survey}.
-However, even simple bivariate maps can be difficult for the general audience to interpret \cite{wainer1980empirical}.
+However, even simple bivariate maps can be difficult for a general audience to interpret \cite{wainer1980empirical}.
 
 Dunn~\cite{dunn1989dynamic} attempts to circumvent these issues by dynamically allocating bins to the bivariate map.
-Since the amount of distinct, separable colors is limited, the dynamic approach allocates bins to regions of the highest utility.
+As the number of distinct, separable colors is limited, the dynamic approach allocates bins to regions of the highest utility.
 For instance, placing the dividing line between bins in areas aligned with a trend line allows the color to encode information about the sign of the residual with respect to each variable.
@@
 
-We present Value-Suppressing Uncertainty Maps, a technique for creating bivariate maps of data \emph{value} and \emph{uncertainty}.
+Value-Suppressing Uncertainty Maps (VSUMs) are a technique for creating bivariate maps of data \emph{value} and \emph{uncertainty}.
 These maps circumvent the issues in the creation of traditional bivariate maps by making the following assumptions:
@@
 and in fact, you really should be binning your colors anyway."
-The first assumption seems to contradict the \emph{expressiveness principle}~\cite{mackinlay1986automating}, where the visual encoding of the data should reflect the composition of the data.
+The first assumption seems to contradict Mackinlay's \emph{expressiveness principle}~\cite{mackinlay1986automating}, where the visual encoding of the data should reflect the composition of the data.
 For instance, if my data are a collection of measurements of some continuous variable, then it would make sense to encode these data using a continuous \emph{visual} variable, which would then mean an arbitrary number of potential marks.
-For instance, a bar can have an arbitrary, non-integer heights.
+For instance, a bar can have an arbitrary, non-integer height.
 However, for many visualizations, the perceptual ambiguity caused by having marks that cannot be reliably disambiguated is higher than the gain from having a visual encoding that maps readily to the backing data.
-Padilla et al.~\cite{padilla2017evaluating} present one such example, where binned heatmaps results in higher accuracy in tasks than in continuous heatmaps.
+Padilla et al.~\cite{padilla2017evaluating} present one such example, where binned heatmaps result in higher task accuracy than continuous heatmaps.
 
@@
 For instance, an analyst might be interested in ``long tail risks'' or other ``black swan events,'' where the impact of an value, no matter how uncertain, must be considered and planned for~\cite{taleb2011black}.
-Other analysis tasks (such as filtering out outliers), require increased, rather than decreased, discriminability when uncertainty low.
-However, for many information fusion tasks, the assumption is that uncertainty is related to data quality, or the uncertainty of particular data values \cite{riveiro2007evaluation}.
+Other analysis tasks (such as filtering out outliers), require increased, rather than decreased, discriminability when uncertainty is high.
+However, for many information fusion tasks, the assumption is that uncertainty is related to data quality, or the uncertainty of particular data values~\cite{riveiro2007evaluation}.
 As with null hypothesis significance testing, the aim would then be to avoid Type I errors, where patterns arising from statistical noise or measurement error are interpreted as significant signals.
@@
 
+\subsection{Method}
+
 The general VSUM algorithm is as follows:
@@
 	\item \textbf{Select a visual variable to encode uncertainty}.
-We choose variables that have a somewhat high semiotic connection with uncertainty (such as lightness and size), but still have a reasonable number of perceptually discriminable layers (as opposed to other, more intuitive channels such as blur or sketchiness).
+We prefer variables that have a somewhat high semiotic connection with uncertainty (such as lightness and size), but still have a reasonable number of perceptually discriminable layers (as opposed to other, more intuitive channels such as blur or sketchiness).
 	\item \textbf{Select a discriminability threshold}.
-This is some minimum distance, in perceptual space, between marks.
-For many visual variables (such as length or hue), the perceptual discriminability between marks has been assessed in prior literature, and so the only step is to choose some value in terms of the just noticeable distance (JND).
-For less common visual variables, this perceptual space must be built up empirically (as in Demiralp et al.~\cite{demiralp2014learning}).
+Choose some minimum distance, in a perceptual space, between marks.
+For many visual variables (such as length or hue), the perceptual discriminability between marks has been assessed in prior literature, and so the only step is to choose a value in terms of the just noticeable distance (JND).
+For less common visual variables, a perceptual model may be built empirically (as in Demiralp et al.~\cite{demiralp2014learning}).
 	\item \textbf{Choose how quickly categories \emph{degrade}}.
@@
 	\item \textbf{Choose the largest number of initial categories $n$ such that the discriminability threshold is met.} For each pair of marks in the map, perform a pairwise comparison of discriminability.
-If no marks are perceptually ambiguous by the standards of your threshold, then the map is complete.
+If no marks are perceptually ambiguous by the standards of your threshold, the map is complete.
 Otherwise, decrement $n$.
@@
 Figure X shows this process in detail, and how it can be used to generate a bivariate map for presenting uncertainty.
-Javascript code for generating VSUMs given arbitrary d3 quantitative scale functions is available at \url{URL REMOVED FOR REVIEW}.
+JavaScript code for generating VSUMs given arbitrary d3 quantitative scale functions is available at \url{URL REMOVED FOR REVIEW}.
 
-Correll et al.
-use a version of VSUMs in their LayerCake genomics visualization tool \cite{correll2015layercake,correll2011visualizing}, where marks representing genomic data with increasing uncertainty are mapped to a smaller and smaller set of increasingly grey colors, creating the effect of uncertain values retreating into ``fog'' while highly certain values remain prominent.
+Correll et al.~\cite{correll2015layercake,correll2011visualizing} use a pre-cursor of VSUMs in their LayerCake genomics visualization tool, where marks representing genomic data with increasing uncertainty are mapped to a smaller and smaller set of increasingly grey colors, creating the effect of uncertain values retreating into ``fog'' while highly certain values remain prominent.
 Other bivariate visualizations implicitly alias together uncertain values.
@@
 Other channels with a semiotic connection to uncertainty, such as saturation, value, blur, or size, also have deleterious effects on the disambiguation of colors and shapes.
-In both of these cases, this property of aliasing is ad hoc, and places no guarantees on the discriminability of colors.
-The binning and degradation approach of VSUMs makes this choice to alias values explicit to both the designer and the viewer, and results in a bivariate mapping with known perceptual properties.
+In both of these cases, the property of aliasing is \emph{ad hoc}, and places no guarantees on the discriminability of colors.
+The binning and degradation approach of VSUMs makes the choice to alias values explicit to both the designer and the viewer, and results in a bivariate mapping with known perceptual properties.
 
@@
 There are multiple choices that designers must make before creating a VSUM.
-In particular, they must choose 1) which visual channels map to value and uncertainty, 2) a threshold of perceptual distinguishability (and, implicitly, a metric for evaluating discriminability), and 3) a degradation function to determine how quickly uncertain values are aliased.
+In particular, they must choose 1) which visual channels map to value and uncertainty, 2) a threshold of perceptual distinguishability (and, implicitly, a corresponding metric), and 3) a degradation function to determine how quickly uncertain values are aliased.
 
 Often, the graphical conventions of a chart reserve variables such as position and area for other uses, and determine the visual channel to be used for value.
-In thematic maps, for instance, value is habitually encoded using color, while vertical and horizontal position are reserved for location.
+In thematic maps, for instance, value is regularly encoded using color, while vertical and horizontal position are reserved for location.
 In this paper, we primarily deal with the case where value is encoded using color.
@@
 We recommend channels with both a strong semiotic connection to uncertainty as well as a relatively large number of perceptually distinguishable levels.
-In this paper, we primarily discuss value/saturation (or, equivalently, transparency against a white background) and size.
+Here, we focus on value/saturation (or, equivalently, transparency against a white background) and size.
 Correll \& Gleicher~\cite{correll2013error} show that even audiences without statistical backgrounds can interpret uncertainty information encoded in these channels.
@@
 Another reason for limiting our discussion of bivariate mappings to color and size channels is that the perceptual distinguishability of these channels has been well-studied by prior work.
-CIELAB and other color spaces are designed with the intention of perceptual uniformity--- that is, distance in the color space should universally correspond with perceptual distance.
+CIELAB and other color spaces are designed to approximate perceptual uniformity\,---\,that is, distance in the color space should universally correspond with perceptual distance.
 Recent work~\cite{stone2014engineering,szafir2014adapting} has investigated the discriminability of colors in such spaces not only under laboratory conditions, but also among the disparate display conditions and color vision proficiency of crowdworkers.
-Similarly, formative work by Cleveland \& McGill~\cite{cleveland1984graphical} assesses the discriminability of marks of different height, and this too has been extended both to the more variable population of crowdworkers, and to the area channel specifically~\cite{heer2010crowdsourcing,talbot2014four}.
+Similarly, formative work by Cleveland \& McGill~\cite{cleveland1984graphical} assesses the discriminability of marks of different height, and this too has been extended both to a larger population of crowdworkers, and to the area channel specifically~\cite{heer2010crowdsourcing,talbot2014four}.
+
 We use values from these works as benchmarks for selecting thresholds of discriminability.
 The ``just noticeable difference'' (JND) of these channels provides a useful minimum; in practice, we would wish to create maps with marks that are far more reliably distinguishable than the 50\% or 75\% discriminability rate of a JND.
-Figure \ref{fig:performance} shows how these benchmarks can be used to select reasonable thresholds.
+Figure~\ref{fig:performance} shows how these benchmarks can be used to select reasonable thresholds.
 Other factors, such as nameability and aesthetic preference, can be added in as additional weights impacting the desirability of certain color palettes, as in Gramazio et al.~\cite{gramazio2017colorgorical}.
@@
 When the initial bins at the highly certain region is a power of $2$ (as in Fig.
-\ref{fig:example}), this gives a tree structure to the resulting VSUM.
+\ref{fig:example}), this imparts a tree structure to the resulting VSUM.
 As uncertainty increases, bins in the previous level map to one and only one ``parent'' in the next uncertainty level, resulting in fewer crossings of color boundaries among rows.
@@
 
+TODO.
+
 \section{Evaluation}
 
-We performed a crowd-sourced experiment on Amazon's Mechanical Turk in order to evaluate the effectiveness of VSUMs for integrating uncertainty and value information in thematic maps.
+We performed a crowdsourced experiment on Amazon's Mechanical Turk in order to evaluate the effectiveness of VSUMs for integrating uncertainty and value information in thematic maps.
 This focus on integration meant that we limited our experimental tasks to scenarios where the participants needed to consider both value and uncertainty before making a decision.
@@
 
-After completion of the main task, we solicited demographics information, and performed a domain-agnostic general risk aversion assay~\cite{mandrik2005exploring}.
+After completion of the main task, we solicited demographic information, and performed a domain-agnostic general risk aversion assay~\cite{mandrik2005exploring}.
 
@@
 In general, the function of the value of gains is concave, but is convex for losses~\cite{kahneman1979prospect}.
-That is, avoiding a large gain is often more valuable than striving for a similarly large gain.
+That is, avoiding a large loass is often more valuable than striving for a similarly large gain.
 

### 2017-03-30 heatmap/heatmap.js
Generate maps for juxtaposed

@@ -435,11 +435,57 @@ function makeMaps(map, threshold, maxSize){
   //Create all relevant maps
   var DEFAULT_THRESHOLD = 5;
   var THRESHOLD = threshold ? threshold : DEFAULT_THRESHOLD;
-  var scaleData, arcScaleData, arcSizeScaleData, closest, n;
+  var scaleData, arcScaleData, arcSizeScaleData, linearValueData, linearUncertaintyData, closest, n;
 
   var uSL = makeuSL(map);
   var uSize = makeuSize(map, maxSize);
 
+  // for juxtaposed
+  n = 2;
+  while (true) {
+    var data = d3.range(0, 1 + 1/n, 1/n);
+    var c = colorDiff([data], map);
+    if (c.minD >= THRESHOLD) {
+      linearValueData = data;
+      closest = c;
+    }
+    else {
+      break;
+    }
+    n++;
+  }
+  if (!closest) {
+    console.log("No valid linear color map at threshold "+THRESHOLD);
+  }
+  else{
+    console.log("The two closest linear value colors:(" + closest.c1 +"," + closest.c2 +") are "+closest.minD+" apart in CIELAB.");
+  }
+
+  var bw = d3.scaleLinear()
+    .domain([0,1])
+    .range(["white", "black"]);
+
+  n = 2;
+  while (true) {
+    var data = d3.range(0, 1 + 1/n, 1/n);
+    var c = colorDiff([data], bw);
+    if (c.minD >= THRESHOLD) {
+      linearUncertaintyData = data;
+      closest = c;
+    }
+    else {
+      break;
+    }
+    n++;
+  }
+  if (!closest) {
+    console.log("No valid linear color map at threshold "+THRESHOLD);
+  }
+  else{
+    console.log("The two closest linear uncertainty colors:(" + closest.c1 +"," + closest.c2 +") are "+closest.minD+" apart in CIELAB.");
+  }
+
+  // for matrix
   n = 2;
   while (true) {
     var data = makeScaleData(n);
@@ -457,8 +503,10 @@ function makeMaps(map, threshold, maxSize){
     console.log("No valid matrix color map at threshold "+THRESHOLD);
   }
   else{
-    console.log("The two closest arc size colors:(" + closest.c1 +"," + closest.c2 +") are "+closest.minD+" apart in CIELAB.");
+    console.log("The two closest matrix colors:(" + closest.c1 +"," + closest.c2 +") are "+closest.minD+" apart in CIELAB.");
   }
+
+  // for arc
[…]

### 2017-03-30 heatmap/heatmap.js
Juxtaposed scales

@@ -461,14 +461,10 @@ function makeMaps(map, threshold, maxSize){
     console.log("The two closest linear value colors:(" + closest.c1 +"," + closest.c2 +") are "+closest.minD+" apart in CIELAB.");
   }
 
-  var bw = d3.scaleLinear()
-    .domain([0,1])
-    .range(["white", "black"]);
-
   n = 2;
   while (true) {
     var data = d3.range(0, 1 + 1/n, 1/n);
-    var c = colorDiff([data], bw);
+    var c = colorDiff([data], d3.interpolateGreys);  // we always use this scale for juxtaposed
     if (c.minD >= THRESHOLD) {
       linearUncertaintyData = data;
       closest = c;

### 2017-03-30 heatmap/main.js

@@ -20,6 +20,12 @@ function main(){
   var squareScale = makeScaleFunction(maps.square, uSL);
   var arcScale = makeScaleFunction(maps.arc, uSL);
   var arcSizeScale = makeScaleFunction(maps.arcSize, uSize);
+  var juxtaposedValueScale = function(d) {
+    return map(d3.scaleQuantize().domain([0,1]).range(maps.linearValue)(d.v));
+  }
+  var juxtaposedUncertaintyScale = function(d) {
+    return d3.interpolateGreys(d3.scaleQuantize().domain([0,1]).range(maps.linearUncertainty)(d.u));
+  }
 
   makeHeatmap(svg, 0,0,250,maps.square, squareScale, "legendSquare");
   makeArcmap(svg, 300,0,250,maps.arc,arcScale,"legendArc");
@@ -29,11 +35,15 @@ function main(){
   makeHeatmap(svg, 0,300,250,gradient, squareScale);
   makeHeatmap(svg, 300,300,250,gradient, arcScale);
   makeHexmap(svg, 600,300,250,gradient, arcSizeScale, MAX_SIZE);
-
+  makeHeatmap(svg, 900,300,250,gradient, juxtaposedValueScale);
+  makeHeatmap(svg, 1200,300,250,gradient, juxtaposedUncertaintyScale);
+  
   var random = randomData(5,5);
   makeHeatmap(svg, 0,600,250,random, squareScale);
   makeHeatmap(svg, 300,600,250,random, arcScale);
   makeHexmap(svg, 600,600,250,random, arcSizeScale, MAX_SIZE);
+  makeHeatmap(svg, 900,600,250,random, juxtaposedValueScale);
+  makeHeatmap(svg, 1200,600,250,random, juxtaposedUncertaintyScale);
 
   // flight data example heatmap
   d3.csv("data.csv", function(data) {

### 2017-03-30 study/src/script.js
More DM study fixs.

@@ -37,12 +37,14 @@ var taskTwoStimuli = [];
 var taskTwoGrid = [];
 var taskTwoTokens = 0;
 
+var maxSize = 200/16;;
+
 //var map = d3.interpolateViridis;
 
 function scaleGenerator(map){
-  var maps = makeMaps(map, 15);
+  var maps = makeMaps(map, 15,maxSize);
   var uSL = makeuSL(map);
-  var uSize = makeuSize(map);
+  var uSize = makeuSize(map,maxSize);
 
   var squareScale = makeScaleFunction(maps.square, uSL);
   var arcScale = makeScaleFunction(maps.arc, uSL);
@@ -60,6 +62,10 @@ var vV = function(d){
   return d3.interpolateViridis(d.v);
 }
 
+var empty = function(d){
+  return "transparent";
+}
+
 var vP = function(d){
   return d3.interpolatePlasma(d.v);
 }
@@ -70,6 +76,10 @@ var vC = function(d){
 
 var u = function(d){
   return d3.interpolateGreys(1-d.u);
+}
+
+var uS = function(d){
+  return {"c":"black", "s":Math.max(0.1*maxSize,d.u*maxSize)};
 }
 
 function gup(name){
@@ -209,7 +219,7 @@ function makeTaskOneStimuli(){
   var stimuli = [];
 
   var types = ["juxta","2D","vsum"];
-  var ramps = ["viridisLightness","viridisSize"];
+  var ramps = ["Lightness","Size"];
   var sizes = ["4","8"];
   var questions = [
     "Click on the map location with the <b>highest</b> uncertainty",
@@ -239,7 +249,7 @@ function makeTaskTwoStimuli(){
   var stimuli = [];
 
   var types = ["juxta","2D","vsum"];
-  var ramps = ["lightness","size"];
+  var ramps = ["Lightness","Size"];
   var sizes = ["4","8"];
   var roles = ["att","def"];
 
@@ -333,6 +343,8 @@ function taskOne(){
   .attr("id","map")
   .attr("style","width: 450px; height: 200px;");
 
+  main.append("div").html("<br />");
+
   main.append("svg")
   .attr("id","legend")
   .attr("style","width: 250px; height: 250px;");
@@ -362,33 +374,71 @@ function revealTaskOne(){
   var mapSvg = d3.select("#map");
   var legendSvg = d3.select("#legend");
   var stim = taskOneStimuli[questionNum-1];
[…]

### 2017-03-30 study/src/script.js
Integrated new scale functions

@@ -50,8 +50,21 @@ function scaleGenerator(map){
   var arcScale = makeScaleFunction(maps.arc, uSL);
   var arcSizeScale = makeScaleFunction(maps.arcSize, uSize);
 
+  var juxtaposedValueScale = function(d) {
+    return map(d3.scaleQuantize().domain([0,1]).range(maps.linearValue)(d.v));
+  }
+
+  var juxtaposedUncertaintyScale = function(d) {
+    return d3.interpolateGreys(d3.scaleQuantize().domain([0,1]).range(maps.linearUncertainty)(d.u));
+  }
+
+  var juxtaposedSizeScale = function(d){
+    return {"c":"black", "s":Math.max(0.1*maxSize,d3.scaleQuantize().domain([0,1]).range(maps.linearUncertainty)(d.u)*maxSize)};
+  }
+
+
   return {"square":maps.square, "arc":maps.arc, "arcSize":maps.arcSize,
-  "cs":uSize, "cl":uSL, "squareScale": squareScale, "arcScale": arcScale, "arcSizeScale":arcSizeScale};
+  "cs":uSize, "cl":uSL, "squareScale": squareScale, "arcScale": arcScale, "arcSizeScale":arcSizeScale, "juxtaposedUncertaintyScale":juxtaposedUncertaintyScale, "juxtaposedValueScale":juxtaposedValueScale, "juxtaposedSizeScale":juxtaposedSizeScale};
 }
 
 var vMap = scaleGenerator(d3.interpolateViridis);
@@ -398,7 +411,7 @@ function revealTaskOne(){
     var tempMap = makeTaskOneMap(stim.size);
 
     if(stim.ramp=="Size"){
-      makeHeatmap(mapSvg,0,0,200,tempMap,vV);
+      makeHeatmap(mapSvg,0,0,200,tempMap,vMap.juxtaposedValueScale);
       mapSvg.append("rect")
         .attr("fill","transparent")
         .attr("stroke","black")
@@ -407,11 +420,11 @@ function revealTaskOne(){
         .attr("y","0px")
         .attr("width","200px")
         .attr("height","200px");
-      makeHexmap(mapSvg,225,0,200,tempMap,uS,maxSize);
+      makeHexmap(mapSvg,225,0,200,tempMap,vMap.juxtaposedSizeScale,maxSize);
       makeHeatmap(mapSvg,225,0,200,tempMap,empty);
     }
     else{
[…]

### 2017-03-30 study/src/script.js
Prepping for 2D size maps

@@ -442,7 +442,9 @@ function revealTaskOne(){
     case "2D":
     default:
     if(stim.ramp=="Size"){
-
+      //TODO Plug into 2D size maps
+      makeHeatmap(mapSvg,100,0,200,makeTaskOneMap(stim.size),vMap.squareScale);
+      makeHeatmap(legendSvg, 20, 50, 160,vMap.square,vMap.squareScale);
     }
     else{
       makeHeatmap(mapSvg,100,0,200,makeTaskOneMap(stim.size),vMap.squareScale);
@@ -644,8 +646,10 @@ function initializeTaskTwo(){
 
     case "2D":
     default:
-    if(stim.range=="Size"){
-
+    if(stim.ramp=="Size"){
+      //TODO plug this into 2d size scales
+      makeHeatmap(mapSvg,100,0,200,makeTaskOneMap(stim.size),taskMap.squareScale);
+      makeHeatmap(legendSvg, 20, 50, 160,taskMap.square,taskMap.squareScale);
     }
     else{
       makeHeatmap(mapSvg,100,0,200,makeTaskOneMap(stim.size),taskMap.squareScale);

### 2017-03-30 heatmap/heatmap.js
Legends

@@ -710,6 +710,32 @@ function makeHeatmapLegend(svg, x, y, size, map, vDom, uDom, vTitle, uTitle) {
     .text(uTitle);
 }
 
+function makeSimpleLegend(svg, x, y, height, size, data, map, title) {
+  var legend = svg.append("g").attr("transform", "translate(" + x + "," + y + ")");
+
+  var w = size / data.length;
+
+  legend.selectAll("rect").data(data).enter()
+    .append("rect")
+    .attr("x", function(d,i) {return i * w})
+    .attr("y", 0)
+    .attr("height", height)
+    .attr("width", w)
+    .attr("fill", map);
+
+  var xAxis = d3.scalePoint().range([0, size]).domain(data);
+
+  legend.append("g")
+    .attr("transform", "translate(0, " + height + ")")
+    .call(d3.axisBottom(xAxis).tickFormat(d3.format(".1f")));
+
+  legend.append("text")
+    .style("text-anchor", "middle")
+    .style("font-size", 13)
+    .attr("transform", "translate(" + (size / 2) + ", " + (height + 30) + ")")
+    .text(title);
+}
+
 //Uncertainty maps
 
 function makeuSL(map) {

### 2017-03-30 heatmap/main.js

@@ -30,6 +30,8 @@ function main(){
   makeHeatmap(svg, 0,0,250,maps.square, squareScale, "legendSquare");
   makeArcmap(svg, 300,0,250,maps.arc,arcScale,"legendArc");
   makeArcHexmap(svg, 600,0,250,maps.arcSize,arcSizeScale,"legendSizeArc");
+  makeSimpleLegend(svg, 900,0,40, 250,maps.linearValue, map, "Value");
+  makeSimpleLegend(svg, 900,100,40, 250,maps.linearUncertainty, d3.interpolateGreys, "Uncertainty");
 
   var gradient = gradientData(8,8);
   makeHeatmap(svg, 0,300,250,gradient, squareScale);

### 2017-03-30 heatmap/heatmap.js
Flip linear uncertainty scale

@@ -463,7 +463,7 @@ function makeMaps(map, threshold, maxSize){
 
   n = 2;
   while (true) {
-    var data = d3.range(0, 1 + 1/n, 1/n);
+    var data = d3.range(0, 1 + 1/n, 1/n).reverse();
     var c = colorDiff([data], d3.interpolateGreys);  // we always use this scale for juxtaposed
     if (c.minD >= THRESHOLD) {
       linearUncertaintyData = data;

### 2017-03-30 heatmap/heatmap.js
Color and size square

@@ -435,7 +435,7 @@ function makeMaps(map, threshold, maxSize){
   //Create all relevant maps
   var DEFAULT_THRESHOLD = 5;
   var THRESHOLD = threshold ? threshold : DEFAULT_THRESHOLD;
-  var scaleData, arcScaleData, arcSizeScaleData, linearValueData, linearUncertaintyData, closest, n;
+  var scaleData, arcScaleData, arcSizeScaleData, squareSizeScaleData, linearValueData, linearUncertaintyData, closest, n;
 
   var uSL = makeuSL(map);
   var uSize = makeuSize(map, maxSize);
@@ -525,6 +525,25 @@ function makeMaps(map, threshold, maxSize){
   n = 2;
   while (true) {
     var startsize = toVisualAngle(45);
+    var data = makeScaleData(n);
+    if(colorSizeDiff(data,THRESHOLD,startsize,uSize)){
+      squareSizeScaleData = data;
+      closest = colorDiff(data, uSL);
+    } else {
+      break;
+    }
+    n++;
+  }
+  if (!closest) {
+    console.log("No valid size color map at threshold "+THRESHOLD);
+  } else {
+    console.log("The two closest size colors:(" + closest.c1 +"," + closest.c2 +") are "+closest.minD+" units apart in CIELAB");
+  }
+
+  // for size and color arc
+  n = 2;
+  while (true) {
+    var startsize = toVisualAngle(45);
     var data = makeArcScaleData(n);
     if(colorSizeDiff(data,THRESHOLD,startsize,uSize)){
       arcSizeScaleData = data;
@@ -540,7 +559,7 @@ function makeMaps(map, threshold, maxSize){
     console.log("The two closest arc size colors:(" + closest.c1 +"," + closest.c2 +") are "+closest.minD+" units apart in CIELAB");
   }
 
-  return {square:scaleData, arc:arcScaleData, arcSize: arcSizeScaleData, linearValue: linearValueData, linearUncertainty: linearUncertaintyData};
+  return {square:scaleData, arc:arcScaleData, arcSize: arcSizeScaleData, squareSize: squareSizeScaleData, linearValue: linearValueData, linearUncertainty: linearUncertaintyData};
 }
 
 function makeFlightExample(svg, colorScale, map, data, type) {

### 2017-03-30 heatmap/main.js

@@ -19,6 +19,7 @@ function main(){
 
   var squareScale = makeScaleFunction(maps.square, uSL);
   var arcScale = makeScaleFunction(maps.arc, uSL);
+  var sizeScale = makeScaleFunction(maps.squareSize, uSize);
   var arcSizeScale = makeScaleFunction(maps.arcSize, uSize);
   var juxtaposedValueScale = function(d) {
     return map(d3.scaleQuantize().domain([0,1]).range(maps.linearValue)(d.v));
@@ -29,23 +30,26 @@ function main(){
 
   makeHeatmap(svg, 0,0,250,maps.square, squareScale, "legendSquare");
   makeArcmap(svg, 300,0,250,maps.arc,arcScale,"legendArc");
-  makeArcHexmap(svg, 600,0,250,maps.arcSize,arcSizeScale,"legendSizeArc");
-  makeSimpleLegend(svg, 900,0,40, 250,maps.linearValue, map, "Value");
-  makeSimpleLegend(svg, 900,100,40, 250,maps.linearUncertainty, d3.interpolateGreys, "Uncertainty");
+  makeHexmap(svg, 600,0,250,maps.squareSize, sizeScale, MAX_SIZE);
+  makeArcHexmap(svg, 900,0,250,maps.arcSize,arcSizeScale,"legendSizeArc");
+  makeSimpleLegend(svg, 1200,0,40, 250,maps.linearValue, map, "Value");
+  makeSimpleLegend(svg, 1200,100,40, 250,maps.linearUncertainty, d3.interpolateGreys, "Uncertainty");
 
   var gradient = gradientData(8,8);
   makeHeatmap(svg, 0,300,250,gradient, squareScale);
   makeHeatmap(svg, 300,300,250,gradient, arcScale);
-  makeHexmap(svg, 600,300,250,gradient, arcSizeScale, MAX_SIZE);
-  makeHeatmap(svg, 900,300,250,gradient, juxtaposedValueScale);
-  makeHeatmap(svg, 1200,300,250,gradient, juxtaposedUncertaintyScale);
+  makeHexmap(svg, 600,300,250,gradient, sizeScale, MAX_SIZE);
+  makeHexmap(svg, 900,300,250,gradient, arcSizeScale, MAX_SIZE);
+  makeHeatmap(svg, 1200,300,250,gradient, juxtaposedValueScale);
+  makeHeatmap(svg, 1500,300,250,gradient, juxtaposedUncertaintyScale);
   
[…]

### 2017-03-30 study/src/script.js
Integrating linear scales

@@ -64,7 +64,10 @@ function scaleGenerator(map){
 
 
   return {"square":maps.square, "arc":maps.arc, "arcSize":maps.arcSize,
-  "cs":uSize, "cl":uSL, "squareScale": squareScale, "arcScale": arcScale, "arcSizeScale":arcSizeScale, "juxtaposedUncertaintyScale":juxtaposedUncertaintyScale, "juxtaposedValueScale":juxtaposedValueScale, "juxtaposedSizeScale":juxtaposedSizeScale};
+  "cs":uSize, "cl":uSL, "squareScale": squareScale, "arcScale": arcScale,
+   "arcSizeScale":arcSizeScale, "juxtaposedUncertaintyScale":juxtaposedUncertaintyScale,
+    "juxtaposedValueScale":juxtaposedValueScale, "juxtaposedSizeScale":juxtaposedSizeScale,
+  "linearValue":maps.linearValue, "linearUncertainty":maps.linearUncertainty};
 }
 
 var vMap = scaleGenerator(d3.interpolateViridis);
@@ -422,6 +425,9 @@ function revealTaskOne(){
         .attr("height","200px");
       makeHexmap(mapSvg,225,0,200,tempMap,vMap.juxtaposedSizeScale,maxSize);
       makeHeatmap(mapSvg,225,0,200,tempMap,empty);
+
+      makeSimpleLegend(legendSvg, 20,20,20, 160,vMap.linearValue, d3.interpolateViridis,vLabel);
+      makeSimpleLegend(legendSvg, 20,100,20, 160,vMap.linearUncertainty,d3.interpolateGreys,uLabel);
     }
     else{
       makeHeatmap(mapSvg,0,0,200,tempMap,vMap.juxtaposedValueScale);
@@ -434,6 +440,9 @@ function revealTaskOne(){
         .attr("width","200px")
         .attr("height","200px");
       makeHeatmap(mapSvg,225,0,200,tempMap,vMap.juxtaposedUncertaintyScale);
+
+      makeSimpleLegend(legendSvg, 20,20,20, 160,vMap.linearValue, d3.interpolateViridis,vLabel);
+      makeSimpleLegend(legendSvg, 20,100,20, 160,vMap.linearUncertainty,d3.interpolateGreys,uLabel);
     }
     //makeHeatmap(legendSvg, 20, 60, 80,vMap.square,vMap.squareScale);
[…]

### 2017-03-30 study/src/script.js
JH study edits.

@@ -231,6 +231,28 @@ function finishTutorial(){
   taskOne();
 }
 
+function tutorialTwo(){
+  //Set up tutorial. When we're done, start the second task
+  main.selectAll("*").remove();
+
+  main.append("iframe")
+  .attr("src","tutorialTwo.html");
+
+  main.append("button")
+  .attr("class","button")
+  .attr("type","button")
+  .attr("id","answer")
+  .attr("name","answer")
+  .text("Ready")
+  .on("click",finishTutorialTwo);
+}
+
+function finishTutorialTwo(){
+  main.selectAll("#answer").remove();
+  main.selectAll("iframe").remove();
+  taskTwo();
+}
+
 function makeTaskOneStimuli(){
   var stimuli = [];
 
@@ -238,9 +260,15 @@ function makeTaskOneStimuli(){
   var ramps = ["Lightness"];
   var sizes = ["4","8"];
   var questions = [
-    "Click on the map location with the <b>highest</b> uncertainty",
-    "Click on the map location with the <b>lowest</b> uncertainty AND <b>highest</b> value",
-    "Click on the map location with the <b>lowest</b> uncertainty AND <b>lowest</b> value"
+    "Click on the map location with the <b>greatest</b> uncertainty",
+    "Click on the map location with the <b>least</b> uncertainty AND <b>greatest</b> value",
+    "Click on the map location with the <b>least</b> uncertainty AND <b>least</b> value"
+  ];
+
+  var qShort = [
+    "1u",
+    "0u1v",
+    "0u0v"
   ];
 
   var perLevel = 1;
@@ -250,7 +278,7 @@ function makeTaskOneStimuli(){
       for(question of questions){
         for(ramp of ramps){
           for(var i = 0;i<perLevel;i++){
-            stimuli.push( {"type":type, "ramp": ramp, "size": size, "question":question});
+            stimuli.push( {"type":type, "ramp": ramp, "size": size, "question":question, "qShort": qShort[questions.indexOf(question)]});
           }
         }
       }
@@ -480,6 +508,8 @@ function answerTaskOne(){
   for(stimProp in stim){
     answerData[stimProp] = stim[stimProp];
   }
+
+  delete answerData.question;
   writeAnswerTaskOne(answerData);
 }
 
@@ -506,7 +536,7 @@ function doneAnswerTaskOne(){
   d3.select("#questionNum").html(questionNum);
 
   if(done){
-    taskTwo();
+    tutorialTwo();
   }
   else{
[…]

### 2017-03-30 study/src/script.js
Minor turk fix.

@@ -134,7 +134,7 @@ function consent(){
 
   if(assignmentId=="ASSIGNMENT_ID_NOT_AVAILABLE"){
     readyBtn.attr("disabled","disabled");
-    readyBtn.txt("PREVIEW");
+    readyBtn.text("PREVIEW");
   }
 
 }

### 2017-03-30 study/src/script.js
DM study fixes.

@@ -55,7 +55,7 @@ function scaleGenerator(map){
   }
 
   var juxtaposedUncertaintyScale = function(d) {
-    return d3.interpolateGreys(d3.scaleQuantize().domain([0,1]).range(maps.linearUncertainty)(d.u));
+    return reverseGrey(d3.scaleQuantize().domain([0,1]).range(maps.linearUncertainty)(d.u));
   }
 
   var juxtaposedSizeScale = function(d){
@@ -91,7 +91,7 @@ var vC = function(d){
 }
 
 var u = function(d){
-  return d3.interpolateGreys(1-d.u);
+  return d3.reverseGrey(d.u);
 }
 
 var uS = function(d){
@@ -366,9 +366,6 @@ function taskOne(){
   .html("Question <span id=\"questionNum\">"+questionNum+"</span>/<span id=\"maxQuestions\">"+taskOneStimuli.length+"</span>");
 
   main.append("p")
-  .html("You will be asked to find a location on a map that fits a certain criteria. Multiple locations might fit the criteria, in which case choose the any valid location.");
-
-  main.append("p")
   .attr("id","prompt")
   .html("Click the \"Ready\" button to begin.");
 
@@ -385,13 +382,12 @@ function taskOne(){
 
   main.append("svg")
   .attr("id","map")
-  .attr("style","width: 450px; height: 200px;");
+  .attr("style","width: 700px; height: 250px;");
 
   main.append("div").html("<br />");
 
-  main.append("svg")
-  .attr("id","legend")
-  .attr("style","width: 250px; height: 250px;");
+  main.append("p")
+  .html("You will be asked to find a location on a map that fits a certain criteria. Multiple locations might fit the criteria, in which case choose the any valid location.");
 
 }
 
@@ -416,7 +412,6 @@ function initializeTaskOne(){
 function revealTaskOne(){
   main.select("#answer").remove();
   var mapSvg = d3.select("#map");
-  var legendSvg = d3.select("#legend");
   var stim = taskOneStimuli[questionNum-1];
 
   var vLabel = "Value";
@@ -428,14 +423,14 @@ function revealTaskOne(){
[…]

### 2017-03-30 study/src/script.js
Spelling fix

@@ -387,7 +387,7 @@ function taskOne(){
   main.append("div").html("<br />");
 
   main.append("p")
-  .html("You will be asked to find a location on a map that fits a certain criteria. Multiple locations might fit the criteria, in which case choose the any valid location.");
+  .html("You will be asked to find a location on a map that fits a certain criteria. Multiple locations might fit the criteria, in which case choose any valid location.");
 
 }
 
@@ -1118,4 +1118,4 @@ function ineligible(){
 
 }
 
-taskOne();
+consent();

### 2017-03-30 paper/template.tex
Typo
Add example figures

@@
 \graphicspath{{figures/}{pictures/}{images/}{./}} % where to search for the images
+
+\newcommand{\figref}[1]{\hyperref[#1]{Figure~\ref*{#1}}}
+
+\newcommand{\ie}{{i.e.}\xspace}
+\newcommand{\eg}{{e.g.,}\xspace}
+\newcommand{\cf}{{c.f.}\xspace}
+\newcommand{\ea}{{et~al.}\xspace}
+\newcommand{\aka}{{a.k.a.}\xspace}
+\newcommand{\etc}{{etc.}\xspace}
 
@@
 Rather than a traditional two-dimensional bivariate map, with differing outputs for each combination of value and uncertainty, VSUMs can be thought of as arcs: as uncertainty increases, values are mapped to smaller and smaller sets of outputs, culminating in a singularity where all inputs are mapped to an identical, highly uncertain mark regardless of their data value.
-Figure~\ref{fig:example} shows examples of both a VSUM and a more traditional bivariate map.
+\figref{fig:example} shows examples of both a VSUM and a more traditional bivariate map.
 
@@
 Viewers need to be able to independently assess the variables of interest, but the resulting bivariate map should also be expressive and easy to interpret.
-Bivariate maps with these constraints are often limited to relatively few categories (say, a 3x3 matrix as in Figure \ref{fig:example}) \cite{robertson1986generation,trumbo1981theory}.
+Bivariate maps with these constraints are often limited to relatively few categories (say, a 3x3 matrix as in \figref{fig:example}) \cite{robertson1986generation,trumbo1981theory}.
 In general, the quality of bivariate color maps is a multivariate measure involving consideration of not just the component color channels, but also the interpolation scheme and the color of the surround~\cite{bernard2015survey}.
@@
 That is, if there are $n$ bins at some uncertainty level $m$, how many bins should there be when uncertainty increases to level $m+1$?
-In Figure \ref{fig:example}, this function is $\left \lfloor {\frac{n}{2}}\right \rfloor$, meaning that the number of value bins decreases by half for each level of uncertainty bins.
+In \figref{fig:example}, this function is $\left \lfloor {\frac{n}{2}}\right \rfloor$, meaning that the number of value bins decreases by half for each level of uncertainty bins.
 If there is too much degradation, then there are fewer levels of uncertainty.
@@
 The ``just noticeable difference'' (JND) of these channels provides a useful minimum; in practice, we would wish to create maps with marks that are far more reliably distinguishable than the 50\% or 75\% discriminability rate of a JND.
-Figure~\ref{fig:performance} shows how these benchmarks can be used to select reasonable thresholds.
+\figref{fig:performance} shows how these benchmarks can be used to select reasonable thresholds.
 Other factors, such as nameability and aesthetic preference, can be added in as additional weights impacting the desirability of certain color palettes, as in Gramazio et al.~\cite{gramazio2017colorgorical}.
@@
 
-TODO.
+\airlineFig
+\viralFig
+
+\figref{fig:airlineVsum}
+\figref{fig:airline2d}
 
@@
 In general, the function of the value of gains is concave, but is convex for losses~\cite{kahneman1979prospect}.
-That is, avoiding a large loass is often more valuable than striving for a similarly large gain.
+That is, avoiding a large loss is often more valuable than striving for a similarly large gain.
 

### 2017-03-30 heatmap/heatmap.js

@@ -625,9 +625,9 @@ function makeFlightExample(svg, colorScale, map, data, type) {
     var legendY = 60;
     var legendSize = 160;
     makeArcmap(heatmap, legendX, legendY, legendSize,map,colorScale);
-    makeArcLegend(heatmap, legendX, legendY, legendSize, map, vScale.ticks(10), uScale.domain(), "Departure Delay (minutes)", "Standard Mean Error");
+    makeArcLegend(heatmap, legendX, legendY, legendSize, map, vScale.ticks(5), uScale.domain(), "Departure Delay (minutes)", "Standard Mean Error");
   } else {
-    var legendY = 80;
+    var legendY = 70;
     var legendSize = 140;
     makeHeatmap(heatmap, legendX, legendY, legendSize,map,colorScale);
     makeHeatmapLegend(heatmap, legendX, legendY, legendSize, map, vScale.domain(), uScale.domain(), "Departure Delay (minutes)", "Standard Mean Error");
@@ -681,21 +681,21 @@ function makeViralExample(svg, colorScale, map, data, type) {
   heatmap.append("text")
     .style("text-anchor", "middle")
     .style("font-size", 13)
-    .attr("transform", "translate(" + (w + 80) + ", " + (h / 2) + ")rotate(90)")
+    .attr("transform", "translate(" + (w + 65) + ", " + (h / 2) + ")rotate(90)")
     .text("Individual");
 
   // legend
-  var legendX = w + 140;
+  var legendX = w + 120;
   if (type === "arc") {
     var legendY = 60;
     var legendSize = 160;
     makeArcmap(heatmap, legendX, legendY, legendSize,map,colorScale);
-    makeArcLegend(heatmap, legendX, legendY, legendSize, map, vScale.ticks(10), uScale.domain(), "Mutations", "Bad Reads");
+    makeArcLegend(heatmap, legendX, legendY, legendSize, map, vScale.ticks(6), uScale.domain(), "% Mutations", "% Bad Reads");
   } else {
-    var legendY = 80;
+    var legendY = 70;
     var legendSize = 140;
     makeHeatmap(heatmap, legendX, legendY, legendSize,map,colorScale);
-    makeHeatmapLegend(heatmap, legendX, legendY, legendSize, map, vScale.domain(), uScale.domain(), "Mutations", "Bad Reads");
+    makeHeatmapLegend(heatmap, legendX, legendY, legendSize, map, vScale.domain(), uScale.domain(), "% Mutations", "% Bad Reads");
   }
 }
 

### 2017-03-30 heatmap/main.js

@@ -68,7 +68,7 @@ function main(){
   });
 
   // genome data example heatmap
-  d3.csv("viraldata.csv", function(data) {
+  d3.csv("viral_binned.csv", function(data) {
     data = data.map(function(d) {
       return {
         Individual: d.Individual,
@@ -76,8 +76,6 @@ function main(){
         Mutation: +d.Mutation,
         BadReads: +d.BadReads
       }
-    }).filter(function(d) {
-      return d.Position < 500 && d.Position < 700
     });
 
     makeViralExample(body.append("svg"), arcScale, maps.arc, data, "arc");

### 2017-03-30 heatmap/process_airline_data.py

@@ -1,0 +1,13 @@
+import sys
+import numpy as np
+import pandas as pd
+import scipy.stats as stats
+
+df = pd.read_csv(sys.stdin, dtype={ 'DepDelay': np.float64 }, usecols = ['DayOfWeek', 'DepTimeBlk', 'DepDelay'])
+
+gb = df.groupby(['DayOfWeek', 'DepTimeBlk'])['DepDelay'].agg({
+  'DepDelay': np.mean,
+  'StdMeanErr': lambda x: stats.sem(x, nan_policy='omit')
+})
+
+gb.to_csv(sys.stdout)

### 2017-03-30 heatmap/process_viral_data.py

@@ -1,0 +1,20 @@
+import sys
+import numpy as np
+import pandas as pd
+import scipy.stats as stats
+
+# Individual,Position,n_percent,dip_percent,Mutation,BadReads
+
+df = pd.read_csv(sys.stdin, usecols = ['Individual', 'Position', 'Mutation', 'BadReads'])
+
+# df = df[df.Position >= 6558][df.Position <= 6853]
+# df = df[df.Position >= 6784][df.Position <= 6853]
+# df = df[df.Position >= 6860][df.Position <= 9499]
+df = df[df.Position >= 1309][df.Position <= 10124]
+
+bins = np.linspace(df.Position.min(), df.Position.max(), 70)
+df.Position = np.digitize(df.Position, bins)
+df.Position = df.Position.apply(lambda x: bins[x-1])
+gb = df.groupby(['Individual', 'Position'], as_index=False).mean()
+
+gb.to_csv(sys.stdout)

### 2017-03-30 paper/figures.tex
Add example figures
checkin

@@
 
+\newcommand{\conditionFig}{
+	\begin{figure*}
+		\centering
+		\includegraphics[width=0.8\textwidth]{conditions.pdf}
+		\caption{The three graph types in our evaluation.
+Bivariate color ramps ensure that there are orthogonal mappings for each combination of value and uncertainty, but, since color channels are not separable, can afford only a few discrete colors before color categories become unacceptably close, perceptually.
+Juxtaposed maps, by keeping these channels separate, afford a larger range of colors, but require a visual search task to recover both value and uncertainty in a region.
+VSUMs have many of the advantages of bivariate maps, but assign more color categories to certain data, at the expense of ambiguity when uncertainty is high.}
+		\label{fig:conditions}
+	\end{figure*}
+}
+
 \newcommand{\performanceFig}{
@@
 }
+
+\newcommand{\airlineFig}{
+\begin{figure}
+	\centering
+	\includegraphics[width=\columnwidth]{airline-vsum.png}
+	\caption{Flight delay data visualized with VSUM.}
+	\label{fig:airlineVsum}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=\columnwidth]{airline-2d.png}
+	\caption{Flight delay data visualized with 2D uncertainty map.}
+	\label{fig:airline2d}
+\end{figure}
+}
+
+\newcommand{\viralFig}{
+\begin{figure}
+	\centering
+	\includegraphics[width=\columnwidth]{viral-vsum.png}
+	\caption{Virus genome data for reads between position $1309$ and $10124$ visualized with VSUM.}
+	\label{fig:viralVsum}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=\columnwidth]{viral-2d.png}
+	\caption{Virus genome data visualized with 2D uncertainty map.}
+	\label{fig:viral2d}
+\end{figure}
+}

### 2017-03-30 paper/template.tex
checkin

@@
 
+\conditionFig
+
 We performed a crowdsourced experiment on Amazon's Mechanical Turk in order to evaluate the effectiveness of VSUMs for integrating uncertainty and value information in thematic maps.
 This focus on integration meant that we limited our experimental tasks to scenarios where the participants needed to consider both value and uncertainty before making a decision.
-After a pretest and short tutorial, we gave participants two main tasks:
+We gave participants two main tasks:
 
@@
 \end{enumerate}
+
 We discuss the two experimental tasks in more detail in the following sections.
 
-After completion of the main task, we solicited demographic information, and performed a domain-agnostic general risk aversion assay~\cite{mandrik2005exploring}.
+For each task, participants saw one of three types of thematic maps.
+Either two \textbf{juxtaposed} maps side-by-side, with separate encodings for value and uncertainty, a traditional \textbf{bivariate} map, where value and uncertainty are orthogonally encoded, or a \textbf{VSUM}.
+Figure \ref{fig:stimuli} shows examples of these factor levels.
 
@@
 
+For the identification task, we gave participants bivariate representations of uncertainty, and asked them in a written prompt to identify a location meeting a given criteria, e.g., ``Click on the location with the greatest uncertainty.'' There were three potential criteria, mapping to the three vertices of a VSUM: 1) high value, low uncertainty, 2) low value, low uncertainty, and 3) high uncertainty.
+In each map,
+
 \subsubsection{Prediction Task}
 
-For the prediction task, participants were given the rules of a game similar to Battleship.
+For the prediction task, we gave participants the rules of a game similar to Battleship.
 Greis et al.~\cite{greis2016decision} employ these game-like experimental tasks to frame uncertainty information, which can be abstract or complex, to the general audience.
@@
 That is, avoiding a large loss is often more valuable than striving for a similarly large gain.
+
+To avoid potential (in this case, incorrect) transfer learning, the attacking and defending used different heatmaps than the identification tasks, and from each other.
 
@@
 As the experimental tasks required multi-hue color perception, we presented participants with a set of Ishihara plates as a pretest, and excluded participants who either misidentified the values in the plates, or who self-reported as having a color vision deficiency (CVD) in the post-test.
-Based on piloting, we paid participants \$X dollars an hour, for a target rate of \$8/hour.
+Based on piloting, we paid participants \$2 dollars an hour, for a target rate of \$8/hour.
+After completion of the main tasks, we solicited demographic information, and performed a domain-agnostic general risk aversion assay~\cite{mandrik2005exploring}.
 

### 2017-03-30 study/data/analysis.js

@@ -1,0 +1,123 @@
+var dataOne;
+var dataTwo;
+
+process();
+
+function process(){
+  //Steps:
+  // 1. Load in csvs of both raw data, and turk completion information.
+  // 2. Filter out rows from turkIDs who didn't complete the HIT.
+  // 5. Anonymize by replacing turkID with index in completion table.
+  // 6. save that sucker for the repo (in cleandata.csv)
+  // 7. do analysis.
+
+  var demoTable = dl.csv("./demo.csv");
+  var dataOneTable = dl.csv("./dataOne.csv");
+  var dataTwoTable = dl.csv("./dataTwo.csv");
+
+  var cleanedOne = [];
+  var cleanedTwo = [];
+  var valids = [];
+  var turkId;
+  var row;
+
+  for(var i = 0;i<demoTable.length;i++){
+    row = demoTable[i];
+    turkId = row.WorkerId;
+    valids = dataOneTable.filter(function(x){ return x.id == turkId;});
+    valids = cleanRows(valids,i);
+    cleanedOne = cleanedOne.concat(valids);
+
+    valids = dataTwoTable.filter(function(x){ return x.id == turkId;});
+    valids = cleanRows(valids,i);
+    cleanedOne = cleanedTwo.concat(valids);
+  }
+
+  for(var i = 0;i<cleanedOne.length;i++){
+    writeRow(cleanedOne[i]);
+  }
+
+  for(var i = 0;i<cleanedTwo.length;i++){
+    writeRow(cleanedTwo[i]);
+  }
+  
+  dataOne = cleanedOne;
+  dataTwo = cleanedTwo;
+}
+
+function writeRow(row){
+  var writeRequest = new XMLHttpRequest();
+  var writeString = "clean=true&answer="+JSON.stringify(row);
+  //console.log(writeString);
+  writeRequest.open("GET","writeJSON.php?"+writeString,true);
+  writeRequest.setRequestHeader("Content-Type", "application/json");
+  writeRequest.send();
+}
+
+function cleanRows(rows,index){
+  var cleanedRows = [];
+  var seen = dl.repeat(false,100);
+  for(let row of rows){
+    row.id = index;
+    if(!seen[row.index]){
+      seen[row.index]=true;
+      cleanedRows.push(row);
+    }
+  }
+  return cleanedRows;
+}
+
[…]

### 2017-03-30 paper/figures.tex
Filled out methods section

@@
 		\label{fig:conditions}
+	\end{figure*}
+}
+
+\newcommand{\taskTwoFig}{
+	\begin{figure*}
+		\centering
+		\begin{subfigure}{0.45\textwidth}
+			\includegraphics[width=\textwidth]{attack.png}
+			\caption{Attacking}
+			\label{fig:taskTwoAttack}
+		\end{subfigure}
+		~
+		\begin{subfigure}{0.45\textwidth}
+			\includegraphics[width=\textwidth]{defend.png}
+			\caption{Defending}
+			\label{fig:taskTwoDefend}
+		\end{subfigure}
+		\caption{The two framings of the Prediction task.
+In the ``attack'' framing, the participant is given a map of predictions of the location of enemy ships, along with the uncertainty in those predictions.
+The participant should place their missile strikes on locations with a high probability of containing a ship, and with high certainty in this probability.
+The ``defend'' framing is the opposite task: the participant has a list of likely missile locations, and ought to place their ships one locations with low probability of attack, and high certainty in this probability.}
+		\label{fig:taskTwoConditions}
 	\end{figure*}

### 2017-03-30 paper/template.tex
Filled out methods section

@@
 	\item [T2] a \textbf{prediction} task, similar to Battleship, where we gave participants charts with both \emph{forecast} and \emph{forecast uncertainty} information, and asked the participants to place tokens on the board in order to optimize expected value.
-E.g., ``place your $4$ ships on the safest locations on the board.''
+E.g., ``place your $4$ ships on safe locations on the board.''
 \end{enumerate}
@@
 Figure \ref{fig:stimuli} shows examples of these factor levels.
+Other factors included the size of the map (either a 4x4 or 8x8 grid), and (for the prediction task), the question framing (attack or defense).
+These secondary factors were evenly blocked across stimuli.
+Fatigue effects were significant in piloting, and so we limited the number of replicates: participants saw 18 stimuli (6 for each map type) for the identification task, and 12 (4 for each map type) for the prediction task.
+
+In all cases, we generated color maps based on the requirement that no two colors in the resulting color scheme were closer than 18 units of Eulcidean distance in CIELAB space.
+This threshold results in a $9$ color bivariate map, $15$ color VSUM for all color scales used in this study.
+Given the size of the marks used in our study, this threshold distance is sufficient for colors to be correctly distinguished in over 95\% of trials~\cite{stone2014engineering}.
+Other choices of threshold, and differing univariate scales, would result in different maps (see Fig.~\ref{fig:performance}).
 
@@
 For the identification task, we gave participants bivariate representations of uncertainty, and asked them in a written prompt to identify a location meeting a given criteria, e.g., ``Click on the location with the greatest uncertainty.'' There were three potential criteria, mapping to the three vertices of a VSUM: 1) high value, low uncertainty, 2) low value, low uncertainty, and 3) high uncertainty.
-In each map,
+Each map contained four locations, one of each combination of maximum and minimum value and uncertainty.
+The correct answer was always one of these locations.
+Other locations in the grid were created pseudo-randomly, with less extreme ranges of uncertainty and probability.
+We measured performance both in terms of accuracy (did the participant select the correct point) as well as response time.
 
 \subsubsection{Prediction Task}
+
+\taskTwoFig
 
@@
 Without direct access to the location of the opponent's ships, the player then has to place tokens representing missile strikes on their opponent's grid.
+Players had to place all of their tokens before continuing.
 If a player fires a missile at a square containing a ship, then that ship is sunk.
@@
 
-We included both a \textbf{gain} framing (given predictions of where your opponent's ships are, where should you launch your missiles?), and a \textbf{loss} framing (given predictions of where your opponent's missiles are, where should you place your ships?) in order to account for systematic differences in human judgments under uncertainty.
+We included both a \textbf{gain} framing (given predictions of where your opponent's ships are, where should you launch your missiles?), and a \textbf{loss} framing (given predictions of where your opponent's missiles will fall, where should you place your ships?) in order to account for systematic differences in human judgments under uncertainty.
 Tversky \& Kahneman~\cite{tversky1985framing} illustrate that framings in terms of gains or losses produce reliably different outcomes.
@@
 That is, avoiding a large loss is often more valuable than striving for a similarly large gain.
+To avoid potential (in this case, incorrect) transfer learning, the attacking and defending used different value color maps (Plasma and Cool) than the identification tasks, and from each other.
+Figure \ref{fig:taskTwoConditions} shows these differing color maps and framings.
 
-To avoid potential (in this case, incorrect) transfer learning, the attacking and defending used different heatmaps than the identification tasks, and from each other.
+The ideal strategy from a value-maximizing standpoint would be to place tokens on areas with the highest expected value, ignoring the uncertainty information.
+However, as with Roulette and other similar games of chance, the variability in expected value is relevant when considering where to place bets~\cite{mlodinow2009drunkard}.
+Over the short term, a player may seek out high expected value but risky locations, or, if risk averse, accept lower expected value but lower variability locations.
+We therefore measured the distribution of both \emph{value} and \emph{uncertainty} of the tokens placed by the participants.
+Each map contained a minimum number of ``risky'' locations: cells with high expected value, but also high uncertainty.
+There were also an equal number of ``safer'' locations, with slightly less expected value, but also less uncertainty.
+The remaining locations of the map varied pseudo-randomly in the middle ranges.
 
@@
 
+We had several hypotheses, stemming from our believe that VSUMs promote better \emph{integration} between uncertainty and value information, and also encourage \emph{caution} by highlighting the ambiguity or untrustworthiness introduced by uncertain data.
+In particular:
+\begin{enumerate}
+	\item Participants would be \textbf{faster and more accurate} when completing the identification task using a VSUM as compared to the other map types.
+Juxtaposed maps, by adding an additional search task to the lookup task, would have especially poor performance compared to VSUMs.
+	\item Participants would choose targets with \textbf{less uncertainty} in the prediction task when using a VSUM compared to the other maps types.
+This would result in a tradeoff where they would also choose targets with \textbf{less expected value} than the other map types.
+Again, juxtaposed maps, by increasing the difficulty of retrieving both variables simultaneously, would highlight this difference.
+\end{enumerate}
+
 \subsection{Results}
+
 \subsubsection{Participants}

### 2017-03-30 study/data/analysis.js
cleaning fix

@@ -11,9 +11,9 @@ function process(){
   // 6. save that sucker for the repo (in cleandata.csv)
   // 7. do analysis.
 
-  var demoTable = dl.csv("./demo.csv");
-  var dataOneTable = dl.csv("./dataOne.csv");
-  var dataTwoTable = dl.csv("./dataTwo.csv");
+  var demoTable = dl.csv("demo.csv");
+  var dataOneTable = dl.csv("dataOne.csv");
+  var dataTwoTable = dl.csv("dataTwo.csv");
 
   var cleanedOne = [];
   var cleanedTwo = [];
@@ -23,7 +23,7 @@ function process(){
 
   for(var i = 0;i<demoTable.length;i++){
     row = demoTable[i];
-    turkId = row.WorkerId;
+    turkId = row.workerId;
     valids = dataOneTable.filter(function(x){ return x.id == turkId;});
     valids = cleanRows(valids,i);
     cleanedOne = cleanedOne.concat(valids);
@@ -40,7 +40,7 @@ function process(){
   for(var i = 0;i<cleanedTwo.length;i++){
     writeRow(cleanedTwo[i]);
   }
-  
+
   dataOne = cleanedOne;
   dataTwo = cleanedTwo;
 }

### 2017-03-30 paper/figures.tex
Examples

@@
 \newcommand{\exampleFig}{
-\begin{figure}
+\begin{figure}[t]
 	\centering
@@
 \newcommand{\conditionFig}{
-	\begin{figure*}
+	\begin{figure*}[t]
 		\centering
@@
 \newcommand{\performanceFig}{
-	\begin{figure}
+	\begin{figure}[t]
 		\centering
@@
 \newcommand{\airlineFig}{
-\begin{figure}
+\begin{figure}[t]
+	\centering
+	\includegraphics[width=\columnwidth]{airline-2d.png}
+	\vspace{-20px}
+	\caption{Average departure delay for different times of the day and days of the week visualized with a 2D uncertainty map.}
+	\label{fig:airline2d}
+
+	\vspace{10px}
+
 	\centering
 	\includegraphics[width=\columnwidth]{airline-vsum.png}
-	\caption{Flight delay data visualized with VSUM.}
+	\vspace{-20px}
+	\caption{Flight delay data encoded with a VSUM.}
 	\label{fig:airlineVsum}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\columnwidth]{airline-2d.png}
-	\caption{Flight delay data visualized with 2D uncertainty map.}
-	\label{fig:airline2d}
 \end{figure}
@@
 \newcommand{\viralFig}{
-\begin{figure}
+\begin{figure}[t]
+	\centering
+	\includegraphics[width=\columnwidth]{viral-2d.png}
+	\vspace{-20px}
+	\caption{Virus genome data for reads between position $1309$ and $10124$ visualized with a 2D uncertainty map.}
+	\label{fig:viral2d}
+
 	\centering
 	\includegraphics[width=\columnwidth]{viral-vsum.png}
-	\caption{Virus genome data for reads between position $1309$ and $10124$ visualized with VSUM.}
+	\vspace{-20px}
+	\caption{Virus genome mutations and read error encoded with a VSUM.}
 	\label{fig:viralVsum}
 \end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\columnwidth]{viral-2d.png}
-	\caption{Virus genome data visualized with 2D uncertainty map.}
-	\label{fig:viral2d}
-\end{figure}
 }

### 2017-03-30 paper/template.tex

@@
 
+Every pair of colors quantized color maps should be easily distinguishable from any other color in the color map.
+VSUMs have a lower resolution for uncertain values and thus can have higher resolution for values with high certainty.
+We created 2D uncertainty maps and VSUMs and used them to visualize flight delay data and mutations in viral genomes to show that VSUMs have higher discriminability where it matters.
+
+The airline delay visualization shows average delays for different times of the day and days of the week.
+In this example we used data about flight delays of US carriers from the Bureau of Transportation Statistics~\cite{bts} for January 2017.
+In addition to the average delay for each pair of time of the day and day of the week, we computed the standard error of the mean as the standard deviation divided by the squareroot of the number of samples.
+
 \airlineFig
+
+\figref{fig:airline2d} shows the heatmap visualization where the delay and uncertainty are encoded with a 2D uncertainty map.
+We used a minimum discriminability  threshold of $18$ units in CIELAB space.
+We can see that the lowest delays are early in the day throughout the week.
+With the same threshold we can create a VSUM with much higher resolution for values with high certainty (\figref{fig:airlineVsum}).
+The VSUM also has more resolution for the uncertainty (4 different levels instead of 3 in the 2D heatmap).
+If the analysts is interested in delay patterns, VSUMs give them higher resolution.
+Since VSUMs blend uncertain values together, 2D maps are better if analysts are interested in high expected delays independent of the uncertainty.
+However, if analysts want to investigate uncertainty or values in isolation, separate channels (\eg juxtaposed heatmaps) are better as they prevent interaction effects.
+
 \viralFig
 
-\figref{fig:airlineVsum}
-\figref{fig:airline2d}
+In \figref{fig:viral2d} and \figref{fig:viralVsum}, we visualized the mutation rate of virus genomes.
+Biologists know that certain regions of the genome are ``coding regions'' that influence the the formation of proteins.
+The data you see here, however, has errors because the process of reading the genome is not perfect.
+Similarly to the flight delay example, the resolution for certain values with 2D uncertainty maps is much lower (\figref{fig:viral2d}).
+VSUMs enable biologists to see more fine grained differences as shown in \figref{fig:viralVsum}.
 

### 2017-03-31 paper/template.tex
Fix paper
prep for merge.

@@
 \newcommand{\figref}[1]{\hyperref[#1]{Figure~\ref*{#1}}}
-
+\usepackage{xspace}
 \newcommand{\ie}{{i.e.}\xspace}
@@
 Either two \textbf{juxtaposed} maps side-by-side, with separate encodings for value and uncertainty, a traditional \textbf{bivariate} map, where value and uncertainty are orthogonally encoded, or a \textbf{VSUM}.
-Figure \ref{fig:stimuli} shows examples of these factor levels.
+Figure \ref{fig:conditions} shows examples of these factor levels.
 Other factors included the size of the map (either a 4x4 or 8x8 grid), and (for the prediction task), the question framing (attack or defense).
@@
 Based on piloting, we paid participants \$2 dollars an hour, for a target rate of \$8/hour.
-After completion of the main tasks, we solicited demographic information, and performed a domain-agnostic general risk aversion assay~\cite{mandrik2005exploring}.
+After completion of the main tasks, we solicited demographic information.
+We recruited a total of 24 participants for this experiment: 4 female, 19 male, and 1 who declined to state ($M_{\text{age}}$= 32, $SD_{\text{age}}$ = 7.5).
+All reported having either college degrees, or some college experience.
+Average self-reported experience with interpreting graphs and charts was low (1.9 on a 5 point Likert scale, with 1 indicating ``No experience''), and no participant reported experience greater than 3 out of 5.
 
 \section{Discussion}
+
 \subsection{Limitations and Future Work}
@@
 
-\section{Conclusion}
+\subsection{Conclusion}
 

### 2017-03-31 study/data/analysis.js

@@ -1,7 +1,7 @@
 var dataOne;
 var dataTwo;
 
-process();
+analysis();
 
 function process(){
   //Steps:
@@ -23,12 +23,11 @@ function process(){
 
   for(var i = 0;i<demoTable.length;i++){
     row = demoTable[i];
-    turkId = row.workerId;
-    valids = dataOneTable.filter(function(x){ return x.id == turkId;});
-    valids = cleanRows(valids,i);
+    turkId = row.WorkerId;
+    valids = dataOneTable.filter(function(x){ return x.workerId == turkId;});
     cleanedOne = cleanedOne.concat(valids);
 
-    valids = dataTwoTable.filter(function(x){ return x.id == turkId;});
+    valids = dataTwoTable.filter(function(x){ return x.workerId == turkId;});
     valids = cleanRows(valids,i);
     cleanedOne = cleanedTwo.concat(valids);
   }
@@ -110,10 +109,8 @@ dl.bootstrap.midmeanci = function(values, a, b, c, d) {
 };
 
 function analysis(){
-  exp1 = dl.csv("./exp2clean.csv");
-  exp2 = dl.csv("./exp3clean.csv");
-  exp3 = dl.csv("./exp4clean.csv");
-
+  dataOne = dl.csv("taskOne.csv");
+  dataTwo = dl.csv("taskTwo.csv");
 }
 
 function bsci(data,facet,measure){

### 2017-03-31 paper/template.tex
Show flow
check in

@@
 
-The general VSUM algorithm is as follows:
+\flowFig
+
+The general VSUM algorithm demonstrated in \figref{fig:flow} is as follows:
 
@@
 Other factors included the size of the map (either a 4x4 or 8x8 grid), and (for the prediction task), the question framing (attack or defense).
-These secondary factors were evenly blocked across stimuli.
+All factors were presented within subjects.
+Secondary factors were evenly blocked across stimuli.
 Fatigue effects were significant in piloting, and so we limited the number of replicates: participants saw 18 stimuli (6 for each map type) for the identification task, and 12 (4 for each map type) for the prediction task.
@@
 
-\subsubsection{Identification Task}
+\subsection{Identification Task}
 
@@
 
-\subsubsection{Prediction Task}
+\subsection{Prediction Task}
 
@@
 
-\subsubsection{Hypotheses}
+\subsection{Hypotheses}
 
@@
 
+Rather than standard means, Cleveland \& McGill~\cite{cleveland1984graphical} use 25\% trimmed means to report effect sizes in graphical perception tasks.
+Trimmed means remove the first and last quartiles and then compute the mean from the remaining points.
+This filtering is important for graphical perception tasks (especially crowdsourced tasks) where error distributions may be long-tailed~\cite{heer2010crowdsourcing}.
+Trimmed means violate the sampling assumptions required for many null-hypothesis significance tests.
+Thus, while we report effect sizes in terms of trimmed means, we perform statistical tests on the standard means.
+
 \subsubsection{Participants}
@@
 Average self-reported experience with interpreting graphs and charts was low (1.9 on a 5 point Likert scale, with 1 indicating ``No experience''), and no participant reported experience greater than 3 out of 5.
+
+\subsubsection{Identification Task}
+
+We performed a one-way ANOVA investigating the effect of chart type (juxtaposed, full bivariate, or VSUM) on accuracy.
+We found only a marginal effect $F(2,859) = 2.4$, $p=0.09$, although more trails with VSUMs were ($66\%$ of trials) compared to the other conditions ()
+
+These results partially support our first hypothesis. While there
 

### 2017-03-31 paper/template.tex
Pdf changes

@@
 We performed a one-way ANOVA investigating the effect of chart type (juxtaposed, full bivariate, or VSUM) on accuracy.
-We found only a marginal effect $F(2,859) = 2.4$, $p=0.09$, although more trails with VSUMs were ($66\%$ of trials) compared to the other conditions ()
+We found only a marginal effect $F(2,859) = 2.4$, $p=0.09$, although VSUMs had the highest accuracy.
+Fig.
+X illustrates this result.
 
-These results partially support our first hypothesis. While there
+We also performed a one-way ANOVA on the effect of chart type on response time.
+In this case, we did find a significant main effect $(F,2,859) = 5.0$, $p<0.001$.
+A Tukey's test of Honest Significant Difference (HSD) found that participants were significantly slower at performing the task on juxtaposed maps, as opposed to the other map types.
+
+These results partially support our first hypothesis.
 

### 2017-03-31 paper/template.tex
Included analyses.

@@
 
+Each map contained a minimum number of extreme locations with each combination of very high and very low value and uncertainty.
+High expected value, high uncertainty locations were ``risky.'' As a balance to the risky locations, there were also an equal number of ``safer'' locations, with slightly less expected value, but also less uncertainty.
+The remaining locations of the map varied pseudo-randomly in the middle ranges.
+
 The ideal strategy from a value-maximizing standpoint would be to place tokens on areas with the highest expected value, ignoring the uncertainty information.
@@
 We therefore measured the distribution of both \emph{value} and \emph{uncertainty} of the tokens placed by the participants.
-Each map contained a minimum number of ``risky'' locations: cells with high expected value, but also high uncertainty.
-There were also an equal number of ``safer'' locations, with slightly less expected value, but also less uncertainty.
-The remaining locations of the map varied pseudo-randomly in the middle ranges.
 
@@
 Trimmed means violate the sampling assumptions required for many null-hypothesis significance tests.
-Thus, while we report effect sizes in terms of trimmed means, we perform statistical tests on the standard means.
+Thus, while we report effect sizes in terms of trimmed means, we perform inferential statistical tests on the standard means.
 
@@
 
+For the identification task, we were interested primarily in primary measures of performance, accuracy and response time.
+Our hypothesis was that VSUMs would have increased performance values compared to the other charts.
+
 We performed a one-way ANOVA investigating the effect of chart type (juxtaposed, full bivariate, or VSUM) on accuracy.
-We found only a marginal effect $F(2,859) = 2.4$, $p=0.09$, although VSUMs had the highest accuracy.
+We found only a marginal effect ($F(2,859) = 2.4$, $p=0.09$), although VSUMs had the highest accuracy.
 Fig.
@@
 We also performed a one-way ANOVA on the effect of chart type on response time.
-In this case, we did find a significant main effect $(F,2,859) = 5.0$, $p<0.001$.
+In this case, we did find a significant main effect ($(F,2,859) = 5.0$, $p<0.001$).
 A Tukey's test of Honest Significant Difference (HSD) found that participants were significantly slower at performing the task on juxtaposed maps, as opposed to the other map types.
+Fig.
+X illustrates this result.
 
-These results partially support our first hypothesis.
+These results only partially support our hypothesis.
+There was no significant difference in terms of accuracy between different maps, partially due to the simplicity of the identification task: participants only had to find a single target in a heatmap, with extreme values in both data and uncertainty axes.
+However, our results do provide evidence that juxtaposed maps, by introducing an additional search task to the problem of assessing the value and uncertainty of a location, require significantly more time to interpret.
+
+\subsubsection{Prediction Task}
+
+For the prediction task, where there is no dominating strategy, our primary measure was the distribution of guesses, in terms of both their average value, and their average uncertainty.
+A risky player would choose guesses with high value, regardless of the uncertainty of those points.
+A more conservative guesser might eschew high-risk, high-reward locations, resulting in a lower average value of guesses, but also lower uncertainty.
+Our hypothesis is that VSUMs, by communicating uncertainty information at the expense of value information, would promote more conservatives guesses than the other chart types.
+
+We performed a one-way ANOVA of chart type on average value of guesses.
+We found no significant effect ($F(2,861)=0.27$, $p=0.8$), indicating that participants were generating similar expected values from their guesses across conditions.
+However, an ANOVA of chart type on average \emph{uncertainty} of guesses did find a significant main effect ($F(2,861)=61$, $p<0.001$).
+A post-hoc Tukey's HSD confirmed that participants had guesses with significantly less uncertainty with VSUMs than with both other chart types.
+Additionally, participants' guesses had significantly more uncertainty with juxtaposed charts than the other two.
+
+As with the identification task, a one-way ANOVA also indicated that chart type had a significant effect on response time ($F(2,861)=13$, $p<0.001$).
+A Tukey's HSD confirmed that, as with the previous task, response time was significantly longer with juxtaposed maps (trimmed mean response time of $23$ seconds) compared to the other chart types (both with trimmed mean response time of $18$ seconds).
+Despite the documented effect of framing in prior work, a one-way ANOVA of task frame on expected value found no significant main effect ($F(1,862)=0.01$, $p=0.9$), and the trimmed means were quite similar (expected value of $0.76$ per guess for attackers, verus $0.75$ per guess for defenders).
+
+Our results indicates that VSUMs promote avoidance of uncertain values, while not noticeably negatively impacting the expected value.
+Juxtaposed maps, by contrast, make the retrieval of uncertainty information significantly more difficult, and so it is possible that participants ignore or downweight it when making predictions.
+By making the consideration of uncertainty values integral to the design of the chart, VSUMs promote caution and consideration of uncertainty when making decisions.
 

### 2017-03-31 paper/template.tex
Result figures
et al should never be separated

@@
 This is partially due to the complexity of uncertainty as a concept.
-In typologies from Thomson et al.~\cite{thomson2005typology} and Buttenfield \& Beard~\cite{buttenfield1994graphical}, the authors note that many, occasionally contradictory, concepts can fall under the category of ``uncertainty,'' including data quality, sampling error, credibility, and provenance.
+In typologies from Thomson \ea~\cite{thomson2005typology} and Buttenfield \& Beard~\cite{buttenfield1994graphical}, the authors note that many, occasionally contradictory, concepts can fall under the category of ``uncertainty,'' including data quality, sampling error, credibility, and provenance.
 
-Reviewing the state of the art in uncertainty visualization in the field, Greithe et al.~\cite{griethe2006visualization} and Brodlie et al.~\cite{brodlie2012review} present lists of potential techniques for conveying uncertainty in concert with data.
+Reviewing the state of the art in uncertainty visualization in the field, Greithe \ea~\cite{griethe2006visualization} and Brodlie \ea~\cite{brodlie2012review} present lists of potential techniques for conveying uncertainty in concert with data.
 Some of these options (interaction, animation, and sonification) are not applicable for static charts; even for dynamic charts, many users do not interact with charts in sufficient detail to recover uncertainty information~\cite{nyt2016}.
-While we acknowledge the potential utility of these techniques in uncertainty visualization (for instance, the animation used to convey sampling error in Hullman et al.'s \emph{HOPs}~\cite{hullman2015hypothetical}), we limit the scope of our discussion to techniques which center on static charts.
+While we acknowledge the potential utility of these techniques in uncertainty visualization (for instance, the animation used to convey sampling error in Hullman \ea's \emph{HOPs}~\cite{hullman2015hypothetical}), we limit the scope of our discussion to techniques which center on static charts.
 
 To support our goal of making uncertainty information well-integrated with data, we focus on the techniques of \emph{juxtaposition} (where uncertainty is presented alongside the data), \textbf{superposition} (where additional glyphs are overlaid on top of the data), and \emph{explicit encoding} (where an additional visual channel is used to encode the uncertainty).
-Gleicher et al.~\cite{gleicher2011visual} explore these three strategies from the perspective of affording \emph{comparisons}; here, we examine them from the perspective of uncertainty visualization.
+Gleicher \ea~\cite{gleicher2011visual} explore these three strategies from the perspective of affording \emph{comparisons}; here, we examine them from the perspective of uncertainty visualization.
 
@@
 Ware's~\cite{ware2009quantitative} \emph{textons} are an example of this strategy applied to thematic maps: monochrome glyphs that can encode an ordinal variable, or a binned quantitative variable.
-Greithe et al.~\cite{griethe2006visualization} and MacEachren~\cite{maceachren1992visualizing,maceachren1998visualizing} present other examples of superimposed uncertainty in thematic maps, including overlaid grids with ``fuzzy'' lines to indicate uncertainty, or overlaid treemaps with coarser and coarser resolution in uncertain regions.
+Greithe \ea~\cite{griethe2006visualization} and MacEachren~\cite{maceachren1992visualizing,maceachren1998visualizing} present other examples of superimposed uncertainty in thematic maps, including overlaid grids with ``fuzzy'' lines to indicate uncertainty, or overlaid treemaps with coarser and coarser resolution in uncertain regions.
 Costs to this approach include obscuring the underlying data map (which can cause uncertainty information to dominate the display~\cite{brodlie2012review}), potential visual interactions that can impair the legibility of glyphs (for instance, simultaneous contrast effects), and the necessity of performing spatial binning in order to ensure that there are a discrete number of uncertainty glyphs.
@@
 A further hurdle is that not all visual variables are well-suited for conveying uncertainty.
-MacEachren et al.~\cite{maceachren2012visual} evaluate a number of visual variables with respect to their \emph{semiotic fit} for representing uncertainty.
+MacEachren \ea~\cite{maceachren2012visual} evaluate a number of visual variables with respect to their \emph{semiotic fit} for representing uncertainty.
 They observe that certain visual variables such as blurriness and transparency seem to have a more intuitive connection to uncertainty than other variables such as shape or hue.
-Unfortunately, Boukhelifa et al.~\cite{boukhelifa2012evaluating} find that many visual variables habitually used for conveying uncertainty, such as blur and value, are also difficult to estimate.
+Unfortunately, Boukhelifa \ea~\cite{boukhelifa2012evaluating} find that many visual variables habitually used for conveying uncertainty, such as blur and value, are also difficult to estimate.
 This results in a preference/performance gap where designers must choose between encoding uncertainty in a way that is intuitive but error prone, or use higher fidelity channels that may be more difficult to interpret.
@@
 However, for many visualizations, the perceptual ambiguity caused by having marks that cannot be reliably disambiguated is higher than the gain from having a visual encoding that maps readily to the backing data.
-Padilla et al.~\cite{padilla2017evaluating} present one such example, where binned heatmaps result in higher task accuracy than continuous heatmaps.
+Padilla \ea~\cite{padilla2017evaluating} present one such example, where binned heatmaps result in higher task accuracy than continuous heatmaps.
 
@@
 For many visual variables (such as length or hue), the perceptual discriminability between marks has been assessed in prior literature, and so the only step is to choose a value in terms of the just noticeable distance (JND).
-For less common visual variables, a perceptual model may be built empirically (as in Demiralp et al.~\cite{demiralp2014learning}).
+For less common visual variables, a perceptual model may be built empirically (as in Demiralp \ea~\cite{demiralp2014learning}).
 	\item \textbf{Choose how quickly categories \emph{degrade}}.
@@
 
-Correll et al.~\cite{correll2015layercake,correll2011visualizing} use a pre-cursor of VSUMs in their LayerCake genomics visualization tool, where marks representing genomic data with increasing uncertainty are mapped to a smaller and smaller set of increasingly grey colors, creating the effect of uncertain values retreating into ``fog'' while highly certain values remain prominent.
+Correll \ea~\cite{correll2015layercake,correll2011visualizing} use a pre-cursor of VSUMs in their LayerCake genomics visualization tool, where marks representing genomic data with increasing uncertainty are mapped to a smaller and smaller set of increasingly grey colors, creating the effect of uncertain values retreating into ``fog'' while highly certain values remain prominent.
 Other bivariate visualizations implicitly alias together uncertain values.
@@
 \figref{fig:performance} shows how these benchmarks can be used to select reasonable thresholds.
-Other factors, such as nameability and aesthetic preference, can be added in as additional weights impacting the desirability of certain color palettes, as in Gramazio et al.~\cite{gramazio2017colorgorical}.
+Other factors, such as nameability and aesthetic preference, can be added in as additional weights impacting the desirability of certain color palettes, as in Gramazio \ea~\cite{gramazio2017colorgorical}.
 
@@
 For the prediction task, we gave participants the rules of a game similar to Battleship.
-Greis et al.~\cite{greis2016decision} employ these game-like experimental tasks to frame uncertainty information, which can be abstract or complex, to the general audience.
+Greis \ea~\cite{greis2016decision} employ these game-like experimental tasks to frame uncertainty information, which can be abstract or complex, to the general audience.
 In our task, the participant and a (fictional) adversary have to place tokens representing ships on a spatial grid.
@@
 Thus, while we report effect sizes in terms of trimmed means, we perform inferential statistical tests on the standard means.
+
+\responseTimeFig
+\accuracyFig
+\uncertaintyFig
+
+\figref{fig:rt1}
+\figref{fig:accuracy1}
+\figref{fig:uncertainty2}
 

### 2017-03-31 paper/figures.tex
Show flow
Result figures
Prep to receive figures.

@@
 
+\newcommand{\flowFig}{
+	\begin{figure}[t]
+		\centering
+		\includegraphics[width=0.95\columnwidth]{flow.pdf}
+		\caption{}
+		\label{fig:flow}
+	\end{figure}
+}
+
 \newcommand{\airlineFig}{
@@
 	\centering
-	\includegraphics[width=\columnwidth]{viral-2d.png}
+	\includegraphics[width=\columnwidth]{viral-2d.pdf}
 	\vspace{-20px}
-	\caption{Virus genome data for reads between position $1309$ and $10124$ visualized with a 2D uncertainty map.}
+	\caption{Variability for 25 different viral populations of the SIV virus, encoded with a traditional 2D bivariate map.
+Error in aligning short reads can lead to poor quality data, and so the amount of these errors is encoded as the uncertainty.
+}
 	\label{fig:viral2d}
@@
 	\centering
-	\includegraphics[width=\columnwidth]{viral-vsum.png}
+	\includegraphics[width=\columnwidth]{viral-vsum.pdf}
 	\vspace{-20px}
-	\caption{Virus genome mutations and read error encoded with a VSUM.}
+	\caption{Variability for 25 different viral populations of SIV, encoded with a VSUM.
+The magnitude of mutation rate, as well as the generally poorer quality of data towards the end of the sequence, is more readily visible than with the traditional bivariate color map.
+}
 	\label{fig:viralVsum}

### 2017-03-31 paper/template.tex
Prep to receive figures.

--- 2017-03-31 paper/template.tex
@@
 If the analysts is interested in delay patterns, VSUMs give them higher resolution.
-Since VSUMs blend uncertain values together, 2D maps are better if analysts are interested in high expected delays independent of the uncertainty.
-However, if analysts want to investigate uncertainty or values in isolation, separate channels (\eg juxtaposed heatmaps) are better as they prevent interaction effects.
 
@@
 
-In \figref{fig:viral2d} and \figref{fig:viralVsum}, we visualized the mutation rate of virus genomes.
-Biologists know that certain regions of the genome are ``coding regions'' that influence the the formation of proteins.
-The data you see here, however, has errors because the process of reading the genome is not perfect.
-Similarly to the flight delay example, the resolution for certain values with 2D uncertainty maps is much lower (\figref{fig:viral2d}).
-VSUMs enable biologists to see more fine grained differences as shown in \figref{fig:viralVsum}.
+In \figref{fig:viral2d} and \figref{fig:viralVsum}, we visualized the variability in viral genomes across 25 populations of the simian immunodeficiency virus (SIV)~\cite{o2012conditional}.
+Next generation sequencing (NGS) allows biologists to collect and analyze large amounts of genomics data.
+However, these techniques often create data quality issues, as they must align large numbers of potentially ambiguous ``reads'' of base pairs.
+In the dataset here, it is important not only to identify hotspots in the viral genome (locations with high rates of mutation or variability), but to not be distracted by false positives.
+Extending the ``confidence fog'' metaphor used in viral genome browers like LayerCake~\cite{correll2015layercake}, VSUMs enable biologists to explore interesting patterns while not being distracted by noise, as shown in \figref{fig:viralVsum}.
 
@@
 
-Our results indicates that VSUMs promote avoidance of uncertain values, while not noticeably negatively impacting the expected value.
+Our results indicate that VSUMs promote avoidance of uncertain values, while not noticeably negatively impacting the expected value.
 Juxtaposed maps, by contrast, make the retrieval of uncertainty information significantly more difficult, and so it is possible that participants ignore or downweight it when making predictions.
@@
 % When should you use these things?
+In general, VSUMs excel in scenarios where we wish to make firm guarantees about the discriminability of different marks in a chart, but must also represent, in as much detail as possible, two, potentially orthogonal variables of interest.
+VSUMs have improved data resolution over na\"ive bivariate maps by assuming that uncertain values ought to have less visual impact that highly certain values.
+VSUMs therefore act both as a kind of filtering device as well as a bivariate representation.
+This assumption holds in scenarios like the virology dataset (\figref{fig:viralVsum}), where the expectation is that highly uncertain data is ``junk'' or should otherwise be downweighted in analysis.
+If the analyst has a different interpretation of uncertainty, or wishes to quickly and orthogonally analyses the distributions of uncertainty and value, other strategies, such as juxtaposed maps, may be more appropriate.
+VSUMs are designed for the \emph{integration} of value and uncertainty.
+Designers should take care in considering how when and how this integration is desirable.
 
 \subsection{Conclusion}
+
+Often times, uncertainty, data quality, or confidence are considered separate information from the data itself, relegated to tooltips or visually distant supplemental charts. This
 

### 2017-03-31 paper/template.tex
Talk about 1d uncertainty viz

@@
 
+There are a number of techniques for communicating values and uncertainty in one-dimensional statistical graphics~\cite{olston2002visualizing, wickham201140, neyman1937outline}.
+Techniques such as error bars, confidence bands, and violin plots use a a separate position channel to encode uncertainty.
+The techniques in this paper also apply to 2D visualizations such as heatmaps.
+Today, there are no established standards of visualizing uncertainty in heatmaps.
+
 To support our goal of making uncertainty information well-integrated with data, we focus on the techniques of \emph{juxtaposition} (where uncertainty is presented alongside the data), \textbf{superposition} (where additional glyphs are overlaid on top of the data), and \emph{explicit encoding} (where an additional visual channel is used to encode the uncertainty).
@@
 Juxtaposition also creates an additional search task for viewers, who must register hot spots in one map with corresponding features of the other map.
-Nevertheless, especially in cases where the quantification of uncertainty is a separate analysis process from the measuring of data values, juxtaposition is a common design tactic for conveying uncertainty.
+Nevertheless, especially in cases where the quantification of uncertainty is a separate analysis process from the measuring of data values, juxtaposition is a common design tactic for conveying uncertainty~\cite{moritz2017trust}.
 

### 2017-03-31 paper/figures.tex
Full draft

@@
 		\includegraphics[width=0.95\columnwidth]{flow.pdf}
-		\caption{}
+		\caption{The process for creating a VSUM.
+VSUMs guarantee that no two marks in the resulting bivariate mapping are perceptually closer than some pre-specified threshold.
+This guarantee removes some of the issues of non-separability and ambiguity that are otherwise a concern when designing bivariate maps.}
 		\label{fig:flow}
@@
 }
+
+\newcommand{\responseTimeFig}{
+	\begin{figure}[t]
+		\centering
+		\includegraphics[width=.8\columnwidth]{rt1.png}
+		\caption{The effect of chart type on response time for the identification task.
+Juxtaposed maps of value and uncertainty introduce a visual search task, increasing the time to complete even simple tasks involving both value and uncertainty information.
+The confidence intervals are bootstrapped 95\% CIs.}
+		\label{fig:rt1}
+	\end{figure}
+}
+
+\newcommand{\accuracyFig}{
+	\begin{figure}[t]
+		\centering
+		\includegraphics[width=.8\columnwidth]{accuracy1.png}
+		\caption{The effect of chart type on accuracy for the identification task.
+VSUMS are similarly accurate for simple search tasks as more traditional bivariate visualizations, despite containing additional colors categories, and having a more complex relationship between uncertainty, value, and color than traditional charts.
+The confidence intervals are bootstrapped 95\% CIs.}
+		\label{fig:accuracy1}
+	\end{figure}
+}
+
+\newcommand{\uncertaintyFig}{
+	\begin{figure}[t]
+		\centering
+		\includegraphics[width=.8\columnwidth]{uncertainty2.png}
+		\caption{The effect of chart type on the average uncertainty in predictions for the prediction task.
+By reducing the number of color categories as uncertainty increases, VSUMs encourage more caution in predictions, making people less likely to consider data with strong, but spurious patterns.
+The confidence intervals are bootstrapped 95\% CIs.}
+		\label{fig:uncertainty2}
+	\end{figure}
+}

### 2017-03-31 paper/template.tex
Full draft

--- 2017-03-31 paper/template.tex
@@
 We describe the motivations behind the use of VSUMs, examples of their utility for decision-making under uncertainty, and assess VSUMs in a crowdsourced experiment.
-Our experimental results indicate that VSUMs promote better integration between uncertainty and data: X.
+Our experimental results indicate that VSUMs promote better integration between uncertainty and data, promoting rapid but cautious decision-making.
 
@@
 
-Figure X shows this process in detail, and how it can be used to generate a bivariate map for presenting uncertainty.
+Figure \ref{fig:flow} shows this process in detail, and how it can be used to generate a bivariate map for presenting uncertainty.
 JavaScript code for generating VSUMs given arbitrary d3 quantitative scale functions is available at \url{URL REMOVED FOR REVIEW}.
@@
 
-We discuss the two experimental tasks in more detail in the following sections.
+We discuss the two experimental tasks in more detail in the following sections. Experimental materials, including data tables, are available at \url{URL REMOVED FOR REVIEW}.
 
@@
 
-\responseTimeFig
-\accuracyFig
-\uncertaintyFig
-
-\figref{fig:rt1}
-\figref{fig:accuracy1}
-\figref{fig:uncertainty2}
-
 \subsubsection{Participants}
@@
 \subsubsection{Identification Task}
+\accuracyFig
+\responseTimeFig
 
@@
 We found only a marginal effect ($F(2,859) = 2.4$, $p=0.09$), although VSUMs had the highest accuracy.
-Fig.
-X illustrates this result.
+\figref{fig:accuracy1} illustrates this result.
 
@@
 A Tukey's test of Honest Significant Difference (HSD) found that participants were significantly slower at performing the task on juxtaposed maps, as opposed to the other map types.
-Fig.
-X illustrates this result.
+\figref{fig:rt1} illustrates this result.
 
@@
 \subsubsection{Prediction Task}
+
+\uncertaintyFig
 
@@
 Additionally, participants' guesses had significantly more uncertainty with juxtaposed charts than the other two.
+\figref{fig:uncertainty2} illustrates this result.
 
@@
 
+The results from our initial evaluation of VSUMs are promising, showing that even a general audience can make use of uncertainty information in a reasonable way.
+The way that uncertainty information is presented can have a measurable impact on decision-making.
+The VSUM strategy of only ``showing'' differences when uncertainty is sufficiently low is somewhat isomorphic to the inferential statistics such as effects tests.
+VSUMs can therefore be used to promote more caution in judgments, and lead analysts away from spurious signals in data.
+
+Bivariate maps have an inherent limitation in that visual channels are often difficult to attend to separately or orthogonally.
+Bivariate maps are also complex to interpret.
+Systems which simultaneously display large amounts of value and uncertainty data to general audiences (such as Pangloss~\cite{moritz2017trust}) avoid them for precisely these reasons, relying instead on juxtaposed maps.
+Our results indicate that juxtaposition produces additional problems, forcing users that wish to integrate uncertainty and data to locate an item of interest, and then re-locate that same item in a visually distant map, which may have few visual landmarks in common.
+This additional step before information fusion is reflected in longer response times, and riskier patterns of decision-making.
+Given the deleterious effects of latency on data analysis~\cite{liu2014effects}, designers ought to consider carefully the tradeoffs between expressiveness of visual encoding, and usability.
+VSUMs represent a middle ground between these two approaches, and have some of the benefits of both.
+
 \subsection{Limitations and Future Work}
@@
 
+Our experiments dealt with cases where both uncertainty and value were represented by color.
+The perceptual non-separability of color channels is well-known~\cite{ware2012information}, and so the concept of a limited ``budget'' of distinguishable marks easier to quantify and illustrate.
+In principle, a VSUM can be created for any combination of visual variables.
+All that is required is a perceptual model of the interaction between these two variables.
+Where these models exist, as with the interaction between size and color~\cite{stone2014engineering}, the creation of VSUMs is straightforward.
+Where these models do not exist, or where the perceptual interaction is too complex to efficiently model, experimental work remains to be done before VSUMs can be considered a feasible design strategy.
+We are currently experimenting with the creation of VSUMs using these other channels.
+
 \subsection{Conclusion}
 
-Often times, uncertainty, data quality, or confidence are considered separate information from the data itself, relegated to tooltips or visually distant supplemental charts. This
+Often times, uncertainty, data quality, or confidence are considered separate information from the data itself, relegated to tooltips or visually distant supplemental charts.
+We believe, in contrast, that uncertainty information ought to be directly integrated with charts of data itself.
+This integration introduces additional complexity in the design and presentation of data.
+Value Suppressing Uncertainty Maps represent one strategy for dealing with this complexity, by assigning marks properties in a way that supports the disambiguation of values in data where uncertainty is low, but suppresses these judgments when uncertainty is high.
 

### 2017-03-31 heatmap/heatmap.js
Virus heatmap changes

@@ -643,7 +643,8 @@ function makeViralExample(svg, colorScale, map, data, type) {
 
   var xAxis = d3.scaleLinear().range([0, w]).domain(d3.extent(data.map(function(d) { return d.Position; })));
 
-  var uScale = d3.scaleLinear().domain(d3.extent(data.map(function(d) { return d.Mutation; }))).range([0,1]).nice(map.length);
+  //var uScale = d3.scaleLinear().domain(d3.extent(data.map(function(d) { return d.Mutation; }))).range([0,1]).nice(map.length);
+  var uScale = d3.scaleLinear().domain([0,2]).range([0,1]).nice(map.length);
   var vScale = d3.scaleLinear().domain(d3.extent(data.map(function(d) { return d.BadReads; }))).range([0,1]).nice(map[0].length);
 
   var heatmap = svg.attr("width", w + 100).attr("height", h + 60).append("g")
@@ -654,9 +655,9 @@ function makeViralExample(svg, colorScale, map, data, type) {
     .enter()
     .append("rect")
     .attr("x", function(d) { return x(d.Position); })
-    .attr("y", function(d) { return y(d.Individual); })
+    .attr("y", function(d) { return y(d.Individual) + 2; })
     .attr("width", x.bandwidth())
-    .attr("height", y.bandwidth())
+    .attr("height", y.bandwidth() - 4)
     .attr("title", JSON.stringify)
     .attr("fill", function(d) { return colorScale({
       u: uScale(d.Mutation),
@@ -690,12 +691,12 @@ function makeViralExample(svg, colorScale, map, data, type) {
     var legendY = 60;
     var legendSize = 160;
     makeArcmap(heatmap, legendX, legendY, legendSize,map,colorScale);
-    makeArcLegend(heatmap, legendX, legendY, legendSize, map, vScale.ticks(6), uScale.domain(), "% Mutations", "% Bad Reads");
+    makeArcLegend(heatmap, legendX, legendY, legendSize, map, vScale.ticks(6), uScale.domain(), "Variant Reads (%)", "Alignment Error Rate (%)");
   } else {
     var legendY = 70;
     var legendSize = 140;
     makeHeatmap(heatmap, legendX, legendY, legendSize,map,colorScale);
-    makeHeatmapLegend(heatmap, legendX, legendY, legendSize, map, vScale.domain(), uScale.domain(), "% Mutations", "% Bad Reads");
[…]

### 2017-03-31 paper/figures.tex
Flow caption, viridis citation

@@
 	\caption{A standard bivariate map and a VSUM.
-Both use the same visual channels to encode value (position along the Viridis color map) and uncertainty (lightness and saturation), and have an equal standard of perceptual discriminability (at least 18 units of distance in CIELAB color space between colors).
+Both use the same visual channels to encode value (position along the Viridis~\cite{viridis} color map) and uncertainty (lightness and saturation), and have an equal standard of perceptual discriminability (at least 18 units of distance in CIELAB color space between colors).
 However, highly uncertain values result in colors that are very close together in the bivariate case, meaning the full bivariate map is only 9 bins under these constraints\,---\,a 4x4 map would result in colors that are perceptually too close together.
@@
 		\caption{
-			The number of discrete color categories in traditional bivariate color maps versus VSUMs, assuming value is encoded by the Viridis color map~\protect\cite{garnier2015viridis}, and uncertainty is encoded by saturation/value (whiter values are more uncertain).
+			The number of discrete color categories in traditional bivariate color maps versus VSUMs, assuming value is encoded by the Viridis color map~\protect\cite{viridis}, and uncertainty is encoded by saturation/value (whiter values are more uncertain).
 Since these whiter uncertain colors are closer together in CIELAB space, and VSUMs intentionally require fewer color bins in those regions, VSUMs always have more color categories than the corresponding 2D bivariate map.
@@
 VSUMs guarantee that no two marks in the resulting bivariate mapping are perceptually closer than some pre-specified threshold.
-This guarantee removes some of the issues of non-separability and ambiguity that are otherwise a concern when designing bivariate maps.}
+This guarantee removes some of the issues of non-separability and ambiguity that are otherwise a concern when designing bivariate maps.
+
+		In this example two colors from the map for $n=9$ are too close in CIELAB space.
+Thus, we create a VSUM with $n=8$ bins that encodes value with Viridis and uncertainty with increasingly white colors.}
 		\label{fig:flow}
@@
 	\includegraphics[width=\columnwidth]{airline-2d.png}
-	\vspace{-20px}
+	\vspace{-15px}
 	\caption{Average departure delay for different times of the day and days of the week visualized with a 2D uncertainty map.}
@@
 	\includegraphics[width=\columnwidth]{airline-vsum.png}
-	\vspace{-20px}
+	\vspace{-15px}
 	\caption{Flight delay data encoded with a VSUM.}
@@
 	\includegraphics[width=\columnwidth]{viral-2d.pdf}
-	\vspace{-20px}
+	\vspace{-15px}
 	\caption{Variability for 25 different viral populations of the SIV virus, encoded with a traditional 2D bivariate map.
@@
 
+	\vspace{10px}
+
 	\centering
 	\includegraphics[width=\columnwidth]{viral-vsum.pdf}
-	\vspace{-20px}
+	\vspace{-15px}
 	\caption{Variability for 25 different viral populations of SIV, encoded with a VSUM.

### 2017-03-31 paper/template.tex
Anonymizing pas
Flow caption, viridis citation

@@
 %% Author and Affiliation (multiple authors with multiple affiliations)
-\author{Michael Correll\\ %
-        \scriptsize University of Washington %
-\and Dominik Moritz\\ %
-     \scriptsize University of Washington %
-\and Jeffrey Heer\\ %
-     \scriptsize University of Washington}
+
+%\author{Michael Correll\\ %
+%        \scriptsize University of Washington %
+%\and Dominik Moritz\\ %
+%     \scriptsize University of Washington %
+%\and Jeffrey Heer\\ %
+%     \scriptsize University of Washington}
 
@@
 
-The general VSUM algorithm demonstrated in \figref{fig:flow} is as follows:
+The general VSUM algorithm is as follows:
 
@@
 
-Figure \ref{fig:flow} shows this process in detail, and how it can be used to generate a bivariate map for presenting uncertainty.
+\figref{fig:flow} shows this process in detail, and how it can be used to generate a bivariate map for presenting uncertainty.
 JavaScript code for generating VSUMs given arbitrary d3 quantitative scale functions is available at \url{URL REMOVED FOR REVIEW}.

### 2017-03-31 paper/template.tex
Use figrefs
Typos and violin plot citation

@@
 There are a number of techniques for communicating values and uncertainty in one-dimensional statistical graphics~\cite{olston2002visualizing, wickham201140, neyman1937outline}.
-Techniques such as error bars, confidence bands, and violin plots use a a separate position channel to encode uncertainty.
+Techniques such as error bars, confidence bands, and violin plots~\cite{hintze1998violin} use a separate position channel to encode uncertainty.
 The techniques in this paper also apply to 2D visualizations such as heatmaps.
@@
 \emph{Juxtaposition} creates two explicit visualizations: an ``uncertainty map'' and a ``data map.'' MacEachren~\cite{maceachren1992visualizing} portrays uncertainty visualization as the design problem of how to \emph{unify} these two maps, and so this explicit separation can be counterproductive to the goal of information fusion.
-Juxtaposition also creates an additional search task for viewers, who must register hot spots in one map with corresponding features of the other map.
+Juxtaposition also creates an additional search task for viewers, who must register hotspots in one map with corresponding features of the other map.
 Nevertheless, especially in cases where the quantification of uncertainty is a separate analysis process from the measuring of data values, juxtaposition is a common design tactic for conveying uncertainty~\cite{moritz2017trust}.
@@
 In the dataset here, it is important not only to identify hotspots in the viral genome (locations with high rates of mutation or variability), but to not be distracted by false positives.
-Extending the ``confidence fog'' metaphor used in viral genome browers like LayerCake~\cite{correll2015layercake}, VSUMs enable biologists to explore interesting patterns while not being distracted by noise, as shown in \figref{fig:viralVsum}.
+Extending the ``confidence fog'' metaphor used in viral genome browsers like LayerCake~\cite{correll2015layercake}, VSUMs enable biologists to explore interesting patterns while not being distracted by noise, as shown in \figref{fig:viralVsum}.
 
@@
 Either two \textbf{juxtaposed} maps side-by-side, with separate encodings for value and uncertainty, a traditional \textbf{bivariate} map, where value and uncertainty are orthogonally encoded, or a \textbf{VSUM}.
-Figure \ref{fig:conditions} shows examples of these factor levels.
+\figref{fig:conditions} shows examples of these factor levels.
 Other factors included the size of the map (either a 4x4 or 8x8 grid), and (for the prediction task), the question framing (attack or defense).
@@
 
-In all cases, we generated color maps based on the requirement that no two colors in the resulting color scheme were closer than 18 units of Eulcidean distance in CIELAB space.
+In all cases, we generated color maps based on the requirement that no two colors in the resulting color scheme were closer than 18 units of Euclidean distance in CIELAB space.
 This threshold results in a $9$ color bivariate map, $15$ color VSUM for all color scales used in this study.
@@
 To avoid potential (in this case, incorrect) transfer learning, the attacking and defending used different value color maps (Plasma and Cool) than the identification tasks, and from each other.
-Figure \ref{fig:taskTwoConditions} shows these differing color maps and framings.
+\figref{fig:taskTwoConditions} shows these differing color maps and framings.
 
@@
 A Tukey's HSD confirmed that, as with the previous task, response time was significantly longer with juxtaposed maps (trimmed mean response time of $23$ seconds) compared to the other chart types (both with trimmed mean response time of $18$ seconds).
-Despite the documented effect of framing in prior work, a one-way ANOVA of task frame on expected value found no significant main effect ($F(1,862)=0.01$, $p=0.9$), and the trimmed means were quite similar (expected value of $0.76$ per guess for attackers, verus $0.75$ per guess for defenders).
+Despite the documented effect of framing in prior work, a one-way ANOVA of task frame on expected value found no significant main effect ($F(1,862)=0.01$, $p=0.9$), and the trimmed means were quite similar (expected value of $0.76$ per guess for attackers, versus $0.75$ per guess for defenders).
 

### 2017-03-31 paper/template.tex
More minor fixes

@@
 Greithe \ea~\cite{griethe2006visualization} and MacEachren~\cite{maceachren1992visualizing,maceachren1998visualizing} present other examples of superimposed uncertainty in thematic maps, including overlaid grids with ``fuzzy'' lines to indicate uncertainty, or overlaid treemaps with coarser and coarser resolution in uncertain regions.
-Costs to this approach include obscuring the underlying data map (which can cause uncertainty information to dominate the display~\cite{brodlie2012review}), potential visual interactions that can impair the legibility of glyphs (for instance, simultaneous contrast effects), and the necessity of performing spatial binning in order to ensure that there are a discrete number of uncertainty glyphs.
+Costs to this approach include obscuring the underlying data map (which can cause uncertainty information to dominate the display~\cite{brodlie2012review}), potential visual interactions that can impair the legibility of glyphs (for instance, simultaneous contrast effects), and the necessity of performing spatial binning to ensure that there are a discrete number of uncertainty glyphs.
 How this binning is performed can introduce undesirable artifacts on the resulting visualization, suppressing important signals or highlighting spurious ones~\cite{battersby2016shapes}.
@@
 In these situations, data value is often encoded using color.
-Colors in such univariate quantitative color maps should be sufficiently far apart as to be perceptually distinguishable~\cite{ware1988color}, and vary in lightness as well as hue in order to afford an implicit ordering of value~\cite{borland2007rainbow,rogowitz2001blair}.
+Colors in such univariate quantitative color maps should be sufficiently far apart as to be perceptually distinguishable~\cite{ware1988color}, and vary in lightness as well as hue to afford an implicit ordering of value~\cite{borland2007rainbow,rogowitz2001blair}.
 Different choices of color maps can highlight different features of the data, and should be chosen with care based on the data types and important features to be visualized~\cite{rogowitz1996not}.
@@
 In some cases, this second assumption is violated.
-For instance, an analyst might be interested in ``long tail risks'' or other ``black swan events,'' where the impact of an value, no matter how uncertain, must be considered and planned for~\cite{taleb2011black}.
+For instance, an analyst might be interested in ``long tail risks'' or other ``black swan events,'' where the impact of a value, no matter how uncertain, must be considered and planned for~\cite{taleb2011black}.
 Other analysis tasks (such as filtering out outliers), require increased, rather than decreased, discriminability when uncertainty is high.
-However, for many information fusion tasks, the assumption is that uncertainty is related to data quality, or the uncertainty of particular data values~\cite{riveiro2007evaluation}.
+However, for many information fusion tasks, the assumption is that uncertainty is related to data quality, or the uncertainty of data values~\cite{riveiro2007evaluation}.
 As with null hypothesis significance testing, the aim would then be to avoid Type I errors, where patterns arising from statistical noise or measurement error are interpreted as significant signals.
@@
 Other channels with a semiotic connection to uncertainty, such as saturation, value, blur, or size, also have deleterious effects on the disambiguation of colors and shapes.
-In both of these cases, the property of aliasing is \emph{ad hoc}, and places no guarantees on the discriminability of colors.
+In both cases, the property of aliasing is \emph{ad hoc}, and places no guarantees on the discriminability of colors.
 The binning and degradation approach of VSUMs makes the choice to alias values explicit to both the designer and the viewer, and results in a bivariate mapping with known perceptual properties.
@@
 In this paper, we primarily deal with the case where value is encoded using color.
-While there are some similar conventions for representing uncertainty, generally designers have a greater degree of freedom in choosing a visual channel for uncertainty~\cite{maceachren1992visualizing}.
+While there are some similar conventions for representing uncertainty, generally designers have more freedom in choosing a visual channel for uncertainty~\cite{maceachren1992visualizing}.
 We recommend channels with both a strong semiotic connection to uncertainty as well as a relatively large number of perceptually distinguishable levels.
@@
 The airline delay visualization shows average delays for different times of the day and days of the week.
-In this example we used data about flight delays of US carriers from the Bureau of Transportation Statistics~\cite{bts} for January 2017.
+In this example, we use data about flight delays of US carriers from the Bureau of Transportation Statistics~\cite{bts} for January 2017.
 In addition to the average delay for each pair of time of the day and day of the week, we computed the standard error of the mean as the standard deviation divided by the squareroot of the number of samples.
@@
 The VSUM also has more resolution for the uncertainty (4 different levels instead of 3 in the 2D heatmap).
-If the analysts is interested in delay patterns, VSUMs give them higher resolution.
+If the analyst is interested in delay patterns, VSUMs give them higher resolution.
 
@@
 
-We performed a crowdsourced experiment on Amazon's Mechanical Turk in order to evaluate the effectiveness of VSUMs for integrating uncertainty and value information in thematic maps.
+We performed a crowdsourced experiment on Amazon's Mechanical Turk to evaluate the effectiveness of VSUMs for integrating uncertainty and value information in thematic maps.
 This focus on integration meant that we limited our experimental tasks to scenarios where the participants needed to consider both value and uncertainty before making a decision.
@@
 \begin{enumerate}
-	\item [T1] an \textbf{identification} task, where we gaves participants charts with value and uncertainty information, and asked them to locate specific regions.
+	\item [T1] an \textbf{identification} task, where we gave participants charts with value and uncertainty information, and asked them to locate specific regions.
 E.g., ``click on the region of the chart with highest uncertainty \emph{and} lowest value.''
@@
 
-For the prediction task, we gave participants the rules of a game similar to Battleship.
+For the prediction task, we gave participants the rules of a game like Battleship.
 Greis \ea~\cite{greis2016decision} employ these game-like experimental tasks to frame uncertainty information, which can be abstract or complex, to the general audience.
@@
 Without direct access to the location of the opponent's ships, the player then has to place tokens representing missile strikes on their opponent's grid.
-Players had to place all of their tokens before continuing.
+Players had to place all their tokens before continuing.
 If a player fires a missile at a square containing a ship, then that ship is sunk.
@@
 
-We included both a \textbf{gain} framing (given predictions of where your opponent's ships are, where should you launch your missiles?), and a \textbf{loss} framing (given predictions of where your opponent's missiles will fall, where should you place your ships?) in order to account for systematic differences in human judgments under uncertainty.
+We included both a \textbf{gain} framing (given predictions of where your opponent's ships are, where should you launch your missiles?), and a \textbf{loss} framing (given predictions of where your opponent's missiles will fall, where should you place your ships?) to account for systematic differences in human judgments under uncertainty.
 Tversky \& Kahneman~\cite{tversky1985framing} illustrate that framings in terms of gains or losses produce reliably different outcomes.
@@
 
-Often times, uncertainty, data quality, or confidence are considered separate information from the data itself, relegated to tooltips or visually distant supplemental charts.
+Often, uncertainty, data quality, or confidence are considered separate information from the data itself, relegated to tooltips or visually distant supplemental charts.
 We believe, in contrast, that uncertainty information ought to be directly integrated with charts of data itself.

### 2017-03-31 paper/figures.tex
Fix build
Typo pass

@@
 This decision affords more distinct colors in other regions of the map, and so an increase in overall bins to 15.
-The resulting map suppresses the value of uncertain data, but increases the discriminability of data with low uncertainty.}
+The resulting map suppresses the value of highly uncertain data, but increases the discriminability of data with low uncertainty.}
 	\label{fig:example}
@@
 The participant should place their missile strikes on locations with a high probability of containing a ship, and with high certainty in this probability.
-The ``defend'' framing is the opposite task: the participant has a list of likely missile locations, and ought to place their ships one locations with low probability of attack, and high certainty in this probability.}
+The ``defend'' framing is the opposite task: the participant has a list of likely missile locations, and ought to place their ships on locations with low probability of attack, and high certainty in this probability.}
 		\label{fig:taskTwoConditions}
@@
 This guarantee removes some of the issues of non-separability and ambiguity that are otherwise a concern when designing bivariate maps.
-
-		In this example two colors from the map for $n=9$ are too close in CIELAB space.
-Thus, we create a VSUM with $n=8$ bins that encodes value with Viridis and uncertainty with increasingly white colors.}
+\\
+		In this example two colors from the map for $n=9$ are too close in CIELAB space. Thus, we create a VSUM with $n=8$ color bins at the lowest level of uncertainty.}
 		\label{fig:flow}
@@
 	\vspace{-15px}
-	\caption{Average departure delay for different times of the day and days of the week visualized with a 2D uncertainty map.}
+	\caption{Average departure delay for different times of the day and days of the week visualized with a 2D uncertainty map.
+Horizontal position is the hour of scheduled departure, and vertical position is the day of the week.}
 	\label{fig:airline2d}
@@
 	\vspace{-15px}
-	\caption{Variability for 25 different viral populations of the SIV virus, encoded with a traditional 2D bivariate map.
+	\caption{Variability for 12 different viral populations of the SIV virus, encoded with a traditional 2D bivariate map.
+Horizontal position denotes location along the SIV genome.
+Each row is the viral population of a different infected animal.
 Error in aligning short reads can lead to poor quality data, and so the amount of these errors is encoded as the uncertainty.

### 2017-03-31 paper/template.tex
Typo pass

@@
 We describe the motivations behind the use of VSUMs, examples of their utility for decision-making under uncertainty, and assess VSUMs in a crowdsourced experiment.
-Our experimental results indicate that VSUMs promote better integration between uncertainty and data, promoting rapid but cautious decision-making.
+Our experimental results indicate that VSUMs create close integration between uncertainty and data, promoting rapid but cautious decision-making.
 
@@
 The techniques in this paper also apply to 2D visualizations such as heatmaps.
-Today, there are no established standards of visualizing uncertainty in heatmaps.
+To our knowledge, there are no established standards of visualizing uncertainty in thematic maps (but see MachEachren~\cite{maceachren1992visualizing} for a glossary of potential techniques).
 
-To support our goal of making uncertainty information well-integrated with data, we focus on the techniques of \emph{juxtaposition} (where uncertainty is presented alongside the data), \textbf{superposition} (where additional glyphs are overlaid on top of the data), and \emph{explicit encoding} (where an additional visual channel is used to encode the uncertainty).
+To support our goal of making uncertainty information well-integrated with data, we focus on the techniques of \emph{juxtaposition} (where uncertainty is presented alongside the data), \emph{superposition} (where additional glyphs are overlaid on top of the data), and \emph{explicit encoding} (where an additional visual channel is used to encode the uncertainty).
 Gleicher \ea~\cite{gleicher2011visual} explore these three strategies from the perspective of affording \emph{comparisons}; here, we examine them from the perspective of uncertainty visualization.
@@
 In addition to overcoming the drawbacks of the other techniques mentioned in this section, we believe that explicit encoding represents the most general form of the solution of how to fuse uncertainty information with data.
-The most pressing design decisions for explicit encoding are 1) which visual variable to use to encode uncertainty and 2) how to design bivariate maps which afford the encoding of both value and uncertainty.
+The most pressing design decisions for explicit encoding are 1) which visual variables to use to encode uncertainty and 2) how to design bivariate maps which afford the encoding of both value and uncertainty.
 We deal with these two questions in greater detail below.
@@
 
-For visualizations such as choropleth maps, heatmaps, or treemaps, visual channels such as position and length are reserved for data variables other than value (such as geographic location or relative size).
+For spatial visualizations like choropleth maps, heatmaps, and treemaps, visual channels such as position and length are reserved for data variables other than value (such as geographic location or relative size).
 In these situations, data value is often encoded using color.
@@
 
-Given these considerations, and the perceptual integrality of color channels, the construction of \emph{bivariate} color maps is difficult.
+Given these considerations, and the perceptual non-orthogonality of color channels, the construction of \emph{bivariate} color maps is difficult.
 Viewers need to be able to independently assess the variables of interest, but the resulting bivariate map should also be expressive and easy to interpret.
@@
 
-Dunn~\cite{dunn1989dynamic} attempts to circumvent these issues by dynamically allocating bins to the bivariate map.
+Dunn~\cite{dunn1989dynamic} attempts to circumvent these issues by dynamically allocating bins to bivariate maps.
 As the number of distinct, separable colors is limited, the dynamic approach allocates bins to regions of the highest utility.
@@
 
-Correll \ea~\cite{correll2015layercake,correll2011visualizing} use a pre-cursor of VSUMs in their LayerCake genomics visualization tool, where marks representing genomic data with increasing uncertainty are mapped to a smaller and smaller set of increasingly grey colors, creating the effect of uncertain values retreating into ``fog'' while highly certain values remain prominent.
+Correll \ea~\cite{correll2015layercake,correll2011visualizing} use a precursor of VSUMs in their LayerCake genomics visualization tool, where marks representing genomic data with increasing uncertainty are mapped to a smaller and smaller set of increasingly grey colors, creating the effect of uncertain values retreating into ``confidence fog'' while highly certain values remain prominent.
 Other bivariate visualizations implicitly alias together uncertain values.
@@
 How quickly the number of bins decreases is another parameter of VSUMs.
-If this degradation is too large, then we may only have a few discriminable levels of uncertainty.
-If this degradation is too small, then we can be overwhelmed with colors.
-For instance, if we only decrement by one bin per level of uncertainty, there would be \emph{1,485} different colors in a VSUM with a tolerance threshold of $1$ unit in CIELAB space.
+If this degradation is too fast, then we may only have a few discriminable levels of uncertainty.
+If this degradation is too slow, then we can be overwhelmed with colors.
+For instance, if we only decrement by one bin per level of uncertainty, there would be $1,485$ different colors in a VSUM with a tolerance threshold of $1$ unit in CIELAB space, far too much complexity for a discrete color mapping.
 In the examples in this paper, we halve the number of bins at each increasing level of uncertainty.
-When the initial bins at the highly certain region is a power of $2$ (as in Fig.
+When the number of initial bins at the highly certain region is a power of $2$ (as in Fig.
 \ref{fig:example}), this imparts a tree structure to the resulting VSUM.
@@
 
-Every pair of colors quantized color maps should be easily distinguishable from any other color in the color map.
-VSUMs have a lower resolution for uncertain values and thus can have higher resolution for values with high certainty.
-We created 2D uncertainty maps and VSUMs and used them to visualize flight delay data and mutations in viral genomes to show that VSUMs have higher discriminability where it matters.
+We present a set of examples showing how VSUMs can be applied to realworld datasets.
+The decision of VSUMs to allocate color bins assymmetrically amongst value and uncertainty regions affords greater discriminability in regions of the chart that warrant closer scrutiny, while discouraging exploration of regions with noisy or unreliable signals.
 
-The airline delay visualization shows average delays for different times of the day and days of the week.
-In this example, we use data about flight delays of US carriers from the Bureau of Transportation Statistics~\cite{bts} for January 2017.
-In addition to the average delay for each pair of time of the day and day of the week, we computed the standard error of the mean as the standard deviation divided by the squareroot of the number of samples.
+\subsubsection{Air Travel Delay}
 
@@
 
-\figref{fig:airline2d} shows the heatmap visualization where the delay and uncertainty are encoded with a 2D uncertainty map.
-We used a minimum discriminability threshold of $18$ units in CIELAB space.
-We can see that the lowest delays are early in the day throughout the week.
-With the same threshold we can create a VSUM with much higher resolution for values with high certainty (\figref{fig:airlineVsum}).
-The VSUM also has more resolution for the uncertainty (4 different levels instead of 3 in the 2D heatmap).
-If the analyst is interested in delay patterns, VSUMs give them higher resolution.
+Using a public data set from the U.S.
+Bureau of Transportation Statistics~\cite{bts}, we created a heatmap showing average flight delay for different times of the day and days of the week for U.S.
+carriers in January 2017.
+We use standard error ($\frac{\sigma}{\sqrt{n}}$) as our measure of uncertainty.
+Starting from an minimum discriminability threshold of $18$ units in CIELAB space, we created two alternate heatmaps of this delay information.
+\figref{fig:airline2d} shows a traditional bivariate map, whereas \figref{fig:airlineVsum} shows a VSUM generated under the same constraints.
+
+While both maps illustrate a similar temporal trend: flight delays are smaller towards the beginning of the day, increasing on average over time.
+The traditional bivariate map affords only a coarse examination of this trend: two visible ``blocks'' of color, early in the day, and later in the afternoon.
+The VSUM, in contrast, has sufficient color resolution to show a quasi-linear trend.
+The additional color resolution for uncertainty information in VSUMs also promotes caution when considering outliers: the small number of flights on Saturdays, or departing late at night, make standard error for these delays relatively high, and the connected values more suspect than the rest of the chart.
 
@@
 
-In \figref{fig:viral2d} and \figref{fig:viralVsum}, we visualized the variability in viral genomes across 25 populations of the simian immunodeficiency virus (SIV)~\cite{o2012conditional}.
+Using a dataset from a group of virologists interested in the population dynamics of viral mutation~\cite{o2012conditional}, we visualized the variability in viral genomes across 25 populations of the simian immunodeficiency virus (SIV).
 Next generation sequencing (NGS) allows biologists to collect and analyze large amounts of genomics data.
-However, these techniques often create data quality issues, as they must align large numbers of potentially ambiguous ``reads'' of base pairs.
+However, these techniques often create data quality issues, as they must align large numbers of potentially ambiguous ``reads'' of relatively small numbers of base pairs.
 In the dataset here, it is important not only to identify hotspots in the viral genome (locations with high rates of mutation or variability), but to not be distracted by false positives.
-Extending the ``confidence fog'' metaphor used in viral genome browsers like LayerCake~\cite{correll2015layercake}, VSUMs enable biologists to explore interesting patterns while not being distracted by noise, as shown in \figref{fig:viralVsum}.
+The traditional bivariate map (\figref{fig:viral2d}) affords the analysis of only a few, large hot spots.
+aExtending the ``confidence fog'' metaphor used in viral genome browsers like LayerCake~\cite{correll2015layercake}, VSUMs (\figref{fig:viralVsum}) enable biologists to explore interesting patterns while not being distracted by noise.
+Smaller scale ``hot spots'' that all be mapped to the same magenta color in the traditional map are visible as bright orange lines.
+Pink hot spots towards the end of the genome (past reference nucleotide $9,000$) that could be plausibly considered potential regions of interest in the 2D map, are encoded as highly uncertain and of ambiguous value in the VSUMs.
+In fact, these regions have systematically low coverage; the lack of reliable data for these regions should discourage analysts from giving their apparently pattern equal consideration from the  stronger signals earlier in the genome.
 
@@
 \begin{enumerate}
-	\item [T1] an \textbf{identification} task, where we gave participants charts with value and uncertainty information, and asked them to locate specific regions.
+	\item an \textbf{identification} task, where we gave participants charts with value and uncertainty information, and asked them to locate specific regions.
 E.g., ``click on the region of the chart with highest uncertainty \emph{and} lowest value.''
-	\item [T2] a \textbf{prediction} task, similar to Battleship, where we gave participants charts with both \emph{forecast} and \emph{forecast uncertainty} information, and asked the participants to place tokens on the board in order to optimize expected value.
+	\item a \textbf{prediction} task, similar to Battleship, where we gave participants charts with both \emph{forecast} and \emph{forecast uncertainty} information, and asked the participants to place tokens on the board in order to optimize expected value.
 E.g., ``place your $4$ ships on safe locations on the board.''
@@
 For each task, participants saw one of three types of thematic maps.
-Either two \textbf{juxtaposed} maps side-by-side, with separate encodings for value and uncertainty, a traditional \textbf{bivariate} map, where value and uncertainty are orthogonally encoded, or a \textbf{VSUM}.
+Either two \textbf{juxtaposed} maps side-by-side, with separate encodings for value and uncertainty, a traditional \textbf{2D bivariate} map, where value and uncertainty are orthogonally encoded, or a \textbf{VSUM}.
 \figref{fig:conditions} shows examples of these factor levels.
@@
 Secondary factors were evenly blocked across stimuli.
-Fatigue effects were significant in piloting, and so we limited the number of replicates: participants saw 18 stimuli (6 for each map type) for the identification task, and 12 (4 for each map type) for the prediction task.
+Participants reported fatigue effects in piloting, and so we limited the number of replicates in this study: participants saw 18 stimuli (6 for each map type) for the identification task, and 12 (4 for each map type) for the prediction task.
 
@@
 This threshold results in a $9$ color bivariate map, $15$ color VSUM for all color scales used in this study.
-Given the size of the marks used in our study, this threshold distance is sufficient for colors to be correctly distinguished in over 95\% of trials~\cite{stone2014engineering}.
+Given the size of the marks used in our study, this threshold distance is sufficient for a predicted discriminability rate of over 95\%~\cite{stone2014engineering}.
 Other choices of threshold, and differing univariate scales, would result in different maps (see Fig.~\ref{fig:performance}).
@@
 
-For the identification task, we gave participants bivariate representations of uncertainty, and asked them in a written prompt to identify a location meeting a given criteria, e.g., ``Click on the location with the greatest uncertainty.'' There were three potential criteria, mapping to the three vertices of a VSUM: 1) high value, low uncertainty, 2) low value, low uncertainty, and 3) high uncertainty.
+For the identification task, we gave participants bivariate representations of uncertainty, and asked to identify a location meeting a given criteria, e.g., ``Click on the location with the greatest uncertainty.'' There were three potential criteria, mapping to the three vertices of a VSUM: 1) high value, low uncertainty, 2) low value, low uncertainty, and 3) high uncertainty.
 Each map contained four locations, one of each combination of maximum and minimum value and uncertainty.
@@
 For the prediction task, we gave participants the rules of a game like Battleship.
-Greis \ea~\cite{greis2016decision} employ these game-like experimental tasks to frame uncertainty information, which can be abstract or complex, to the general audience.
+Greis \ea~\cite{greis2016decision} employ these game-like experimental tasks to assess how different visual designs communicate uncertainty information, which can be abstract or complex, to the general audience.
 In our task, the participant and a (fictional) adversary have to place tokens representing ships on a spatial grid.
 Without direct access to the location of the opponent's ships, the player then has to place tokens representing missile strikes on their opponent's grid.
-Players had to place all their tokens before continuing.
+Players have to place all their tokens before continuing.
 If a player fires a missile at a square containing a ship, then that ship is sunk.
 The objective is the maximize the number of enemy ships hit, while minimizing the number of your own ships that are hit.
-In our task, participants were given a chart encoding predictions on the location of enemy ships (in the attacking case) or potential missile strikes (in the defending case).
+In our task, participants were given a chart encoding predictions of the location of enemy ships (in the attacking case) or potential missile strikes (in the defending case).
 The \emph{value} component of the prediction was the likelihood of a item being present in a particular region.
@@
 
-We included both a \textbf{gain} framing (given predictions of where your opponent's ships are, where should you launch your missiles?), and a \textbf{loss} framing (given predictions of where your opponent's missiles will fall, where should you place your ships?) to account for systematic differences in human judgments under uncertainty.
+We included both a \textbf{gain} framing (the attacker's task: given predictions of where your opponent's ships are, where should you launch your missiles?), and a \textbf{loss} framing (the defender's task: given predictions of where your opponent's missiles will fall, where should you place your ships?) to account for systematic differences in human judgments under uncertainty.
 Tversky \& Kahneman~\cite{tversky1985framing} illustrate that framings in terms of gains or losses produce reliably different outcomes.
@@
 That is, avoiding a large loss is often more valuable than striving for a similarly large gain.
-To avoid potential (in this case, incorrect) transfer learning, the attacking and defending used different value color maps (Plasma and Cool) than the identification tasks, and from each other.
+To avoid potential (in this case, incorrect) transfer learning, the attacking and defending used different value color maps (Plasma and Cool) than in the identification task, and from the opposite framing in this task.
 \figref{fig:taskTwoConditions} shows these differing color maps and framings.
@@
 The ideal strategy from a value-maximizing standpoint would be to place tokens on areas with the highest expected value, ignoring the uncertainty information.
-However, as with Roulette and other similar games of chance, the variability in expected value is relevant when considering where to place bets~\cite{mlodinow2009drunkard}.
+However, as with roulette and other similar games of chance, the variability in expected value is relevant when considering where to place bets~\cite{mlodinow2009drunkard}.
 Over the short term, a player may seek out high expected value but risky locations, or, if risk averse, accept lower expected value but lower variability locations.
@@
 
-We had several hypotheses, stemming from our believe that VSUMs promote better \emph{integration} between uncertainty and value information, and also encourage \emph{caution} by highlighting the ambiguity or untrustworthiness introduced by uncertain data.
+Across both the identification and prediction tasks, we had several hypotheses, stemming from our belief that VSUMs promote better \emph{integration} between uncertainty and value information, and also encourage \emph{caution} by highlighting the ambiguity or untrustworthiness introduced by uncertain data.
 In particular:
@@
 
-For the identification task, we were interested primarily in primary measures of performance, accuracy and response time.
-Our hypothesis was that VSUMs would have increased performance values compared to the other charts.
+For the identification task, our primary measures of performance were accuracy and response time. Our hypothesis was that VSUMs would have increased performance values compared to the other charts.
 
 We performed a one-way ANOVA investigating the effect of chart type (juxtaposed, full bivariate, or VSUM) on accuracy.
-We found only a marginal effect ($F(2,859) = 2.4$, $p=0.09$), although VSUMs had the highest accuracy.
+We found only a marginal effect ($F(2,859) = 2.4$, $p=0.09$), although VSUMs had the highest accuracy among the three.
 \figref{fig:accuracy1} illustrates this result.
@@
 We found no significant effect ($F(2,861)=0.27$, $p=0.8$), indicating that participants were generating similar expected values from their guesses across conditions.
+
 However, an ANOVA of chart type on average \emph{uncertainty} of guesses did find a significant main effect ($F(2,861)=61$, $p<0.001$).
 A post-hoc Tukey's HSD confirmed that participants had guesses with significantly less uncertainty with VSUMs than with both other chart types.
-Additionally, participants' guesses had significantly more uncertainty with juxtaposed charts than the other two.
+Additionally, participants' guesses on juxtaposed charts were significantly riskier than with the other chart types.
 \figref{fig:uncertainty2} illustrates this result.
@@
 
-Our results indicate that VSUMs promote avoidance of uncertain values, while not noticeably negatively impacting the expected value.
-Juxtaposed maps, by contrast, make the retrieval of uncertainty information significantly more difficult, and so it is possible that participants ignore or downweight it when making predictions.
-By making the consideration of uncertainty values integral to the design of the chart, VSUMs promote caution and consideration of uncertainty when making decisions.
+While our results only partially support our results, in that we did not observe statistically different patterns of \emph{expected value}, our did indicate that VSUMs produce different patterns of \emph{acceptable uncertainty}.
+That is, VSUMs promote avoidance of uncertain values, while not noticeably negatively impacting the expected value.
+Juxtaposed maps, by contrast, make the retrieval of uncertainty information significantly more difficult, and so it is possible that participants ignore or downweight it when making predictions, resulting in riskier predictions overall.
+By making the consideration of uncertainty values integral to the design of the chart, VSUMs promote caution when making decisions.
 
@@
 Our results indicate that juxtaposition produces additional problems, forcing users that wish to integrate uncertainty and data to locate an item of interest, and then re-locate that same item in a visually distant map, which may have few visual landmarks in common.
-This additional step before information fusion is reflected in longer response times, and riskier patterns of decision-making.
-Given the deleterious effects of latency on data analysis~\cite{liu2014effects}, designers ought to consider carefully the tradeoffs between expressiveness of visual encoding, and usability.
+This additional step before information fusion is reflected in longer response times, and patterns of decision-making that are less mindful of risk.
+Given the deleterious effects of latency on data analysis~\cite{liu2014effects}, designers ought to carefully consider the tradeoffs between expressiveness of visual encoding, and usability.
 VSUMs represent a middle ground between these two approaches, and have some of the benefits of both.
@@
 VSUMs therefore act both as a kind of filtering device as well as a bivariate representation.
-This assumption holds in scenarios like the virology dataset (\figref{fig:viralVsum}), where the expectation is that highly uncertain data is ``junk'' or should otherwise be downweighted in analysis.
-If the analyst has a different interpretation of uncertainty, or wishes to quickly and orthogonally analyses the distributions of uncertainty and value, other strategies, such as juxtaposed maps, may be more appropriate.
+This assumption holds in scenarios like the virology dataset (\figref{fig:viralVsum}), where the expectation is that highly uncertain data is unreliable or should otherwise be downweighted in analysis.
+If the analyst has a different interpretation of uncertainty, or wishes to quickly and orthogonally analyze the distributions of uncertainty and value, other strategies, such as juxtaposed maps, may be more appropriate.
 VSUMs are designed for the \emph{integration} of value and uncertainty.
@@
 Value Suppressing Uncertainty Maps represent one strategy for dealing with this complexity, by assigning marks properties in a way that supports the disambiguation of values in data where uncertainty is low, but suppresses these judgments when uncertainty is high.
+This decision of how to allocate visual variables promotes patterns of decision-making that make responsible use of uncertainty information, discouraging comparison of values in unreliable regions of the data, and promoting comparison in regions of high certainty.
 

### 2017-03-31 paper/figures.tex
Added color/size fig

@@
 	\includegraphics[width=0.9\columnwidth]{example.pdf}
-	\caption{A standard bivariate map and a VSUM.
+	\caption{A standard bivariate map and a VSUM, used to encode an identical 5x5 grid of random data.
 Both use the same visual channels to encode value (position along the Viridis~\cite{viridis} color map) and uncertainty (lightness and saturation), and have an equal standard of perceptual discriminability (at least 18 units of distance in CIELAB color space between colors).
@@
 	\label{fig:example}
+\end{figure}
+}
+
+\newcommand{\sizeFig}{
+	\begin{figure}
+		\centering
+		\includegraphics[width=0.9\columnwidth]{size.pdf}
+		\caption{A VSUM where value is mapped to color, and uncertainty is mapped to glyph area.
+As glyphs become smaller, it becomes more difficult to distinguish their colors~\protect\cite{stone2014engineering}.
+VSUMs, by reducing the number of color categories as uncertainty increases, account for this ambiguity.
+On the right, the VSUM has been used to encode a 5x5 grid of random data.}
+		\label{fig:size}
 	\end{figure}

### 2017-03-31 paper/template.tex
Added color/size fig

@@
 
-While our results only partially support our results, in that we did not observe statistically different patterns of \emph{expected value}, our did indicate that VSUMs produce different patterns of \emph{acceptable uncertainty}.
+While our results only partially support our hypothesis, in that we did not observe statistically different patterns of \emph{expected value}, our did indicate that VSUMs produce different patterns of \emph{acceptable uncertainty}.
 That is, VSUMs promote avoidance of uncertain values, while not noticeably negatively impacting the expected value.
@@
 \section{Discussion}
+
+\sizeFig
 
@@
 Where these models do not exist, or where the perceptual interaction is too complex to efficiently model, experimental work remains to be done before VSUMs can be considered a feasible design strategy.
-We are currently experimenting with the creation of VSUMs using these other channels.
+We are currently experimenting with the creation of VSUMs using these other channels, as in \figref{fig:size}.
 

### 2017-03-31 paper/template.tex
Fixed typo

@@
 The traditional bivariate map (\figref{fig:viral2d}) affords the analysis of only a few, large hot spots.
-aExtending the ``confidence fog'' metaphor used in viral genome browsers like LayerCake~\cite{correll2015layercake}, VSUMs (\figref{fig:viralVsum}) enable biologists to explore interesting patterns while not being distracted by noise.
+Extending the ``confidence fog'' metaphor used in viral genome browsers like LayerCake~\cite{correll2015layercake}, VSUMs (\figref{fig:viralVsum}) enable biologists to explore interesting patterns while not being distracted by noise.
 Smaller scale ``hot spots'' that all be mapped to the same magenta color in the traditional map are visible as bright orange lines.

### 2017-03-31 paper/template.tex
Some small edits after the meeting

@@
 VSUMs intentionally alias together data values with high uncertainty, affording greater discriminability as uncertainty decreases.
-Rather than a traditional two-dimensional bivariate map, with differing outputs for each combination of value and uncertainty, VSUMs can be thought of as arcs: as uncertainty increases, values are mapped to smaller and smaller sets of outputs, culminating in a singularity where all inputs are mapped to an identical, highly uncertain mark regardless of their data value.
-\figref{fig:example} shows examples of both a VSUM and a more traditional bivariate map.
+Compared to a traditional two-dimensional bivariate map, with differing outputs for each combination of value and uncertainty, VSUMs can be thought of as arcs: as uncertainty increases, values are mapped to smaller and smaller sets of outputs, culminating in a singularity where all inputs are mapped to an identical, highly uncertain mark regardless of their data value.
+\figref{fig:example} shows examples of both a more traditional bivariate map and a VSUM.
 
@@
 \begin{enumerate}
-	\item \textbf{Select a visual variable to encode value}. In most cases in this paper, this variable is position along some established color scale.
+	\item \textbf{Select a visual variable to encode value}. In this paper, this variable is position along some established color scale.
 	\item \textbf{Select a visual variable to encode uncertainty}.
 We prefer variables that have a somewhat high semiotic connection with uncertainty (such as lightness and size), but still have a reasonable number of perceptually discriminable layers (as opposed to other, more intuitive channels such as blur or sketchiness).
+In most cases in this paper, we encode uncertainty as interpolation to white or as glyph size.
 	\item \textbf{Select a discriminability threshold}.

### 2017-03-31 paper/template.tex
Split up conditions figure, scoped back up from thematic maps to vis generally

@@
 
-In this paper, we contribute \textbf{Value-Suppressing Uncertainty Maps} (VSUMs) for integrating data and uncertainty information in thematic maps.
+In this paper, we contribute \textbf{Value-Suppressing Uncertainty Maps} (VSUMs) for integrating data and uncertainty information in visualizations.
 VSUMs intentionally alias together data values with high uncertainty, affording greater discriminability as uncertainty decreases.
@@
 The techniques in this paper also apply to 2D visualizations such as heatmaps.
-To our knowledge, there are no established standards of visualizing uncertainty in thematic maps (but see MachEachren~\cite{maceachren1992visualizing} for a glossary of potential techniques).
+To our knowledge, there are no established standards of visualizing uncertainty in heatmaps.
 
@@
 Often, the graphical conventions of a chart reserve variables such as position and area for other uses, and determine the visual channel to be used for value.
-In thematic maps, for instance, value is regularly encoded using color, while vertical and horizontal position are reserved for location.
+In heatmaps, for instance, value is regularly encoded using color, while vertical and horizontal position are reserved for location.
 In this paper, we primarily deal with the case where value is encoded using color.
@@
 
-We performed a crowdsourced experiment on Amazon's Mechanical Turk to evaluate the effectiveness of VSUMs for integrating uncertainty and value information in thematic maps.
+We performed a crowdsourced experiment on Amazon's Mechanical Turk to evaluate the effectiveness of VSUMs for integrating uncertainty and value information in visualizations.
 This focus on integration meant that we limited our experimental tasks to scenarios where the participants needed to consider both value and uncertainty before making a decision.
@@
 
-For each task, participants saw one of three types of thematic maps.
+For each task, participants saw one of three types of heatmaps.
 Either two \textbf{juxtaposed} maps side-by-side, with separate encodings for value and uncertainty, a traditional \textbf{2D bivariate} map, where value and uncertainty are orthogonally encoded, or a \textbf{VSUM}.

### 2017-03-31 paper/template.tex
post meeting edits

@@
 Juxtaposition also creates an additional search task for viewers, who must register hotspots in one map with corresponding features of the other map.
+Rather than complete this task, viewers may choose to ignore the uncertainty information altogether.
+This ignoreability is a key deficiency for information fusion tasks.
 Nevertheless, especially in cases where the quantification of uncertainty is a separate analysis process from the measuring of data values, juxtaposition is a common design tactic for conveying uncertainty~\cite{moritz2017trust}.
@@
 This results in a preference/performance gap where designers must choose between encoding uncertainty in a way that is intuitive but error prone, or use higher fidelity channels that may be more difficult to interpret.
+
+Foundational work in psychophysics and graphical perception make use of the concept of a ``just noticeable difference'' (JND)\cite{cleveland1984graphical}.
+A 50\% JND is a difference between two stimuli that is detectable 50\% of the time.
+For scales and legends, we would want a significantly higher standard of unambiguity.
+That is, the different levels of visual variables used in visualizations should be readily and reliably distinguishable.
+Complicating this goal is that a JND as measured in a controlled environment with cooperative participants can be considerably lower than JNDs measured in the sorts of heterogeneous conditions found in real environments~\cite{szafir2014adapting}.
 
@@
 We use values from these works as benchmarks for selecting thresholds of discriminability.
-The ``just noticeable difference'' (JND) of these channels provides a useful minimum; in practice, we would wish to create maps with marks that are far more reliably distinguishable than the 50\% or 75\% discriminability rate of a JND.
+The JNDs of these channels provides a useful minimum; in practice, we would wish to create maps with marks that are far more reliably distinguishable than the 50\% or 75\% discriminability rate of a JND.
 \figref{fig:performance} shows how these benchmarks can be used to select reasonable thresholds.
@@
 We limited our population to Turkers from the United States, with a prior task approval rate of at least 90\%.
-As the experimental tasks required multi-hue color perception, we presented participants with a set of Ishihara plates as a pretest, and excluded participants who either misidentified the values in the plates, or who self-reported as having a color vision deficiency (CVD) in the post-test.
+As the experimental tasks required multi-hue color perception, we presented participants with a set of Ishihara plates~\cite{hardy1945tests} as a pretest, and excluded participants who either misidentified the values in the plates, or who self-reported as having a color vision deficiency (CVD) in the post-test.
 Based on piloting, we paid participants \$2 dollars an hour, for a target rate of \$8/hour.

### 2017-03-31 paper/figures.tex
Some small edits after the meeting
Split up conditions figure, scoped back up from thematic maps to vis generally
Jeff revisions.

@@
 	\includegraphics[width=0.9\columnwidth]{example.pdf}
-	\caption{A standard bivariate map and a VSUM, used to encode an identical 5x5 grid of random data.
+	\caption{A standard bivariate map (left) and a VSUM (right), used to encode an identical 5x5 grid of random data.
 Both use the same visual channels to encode value (position along the Viridis~\cite{viridis} color map) and uncertainty (lightness and saturation), and have an equal standard of perceptual discriminability (at least 18 units of distance in CIELAB color space between colors).
@@
 Juxtaposed maps of value and uncertainty introduce a visual search task, increasing the time to complete even simple tasks involving both value and uncertainty information.
-The confidence intervals are bootstrapped 95\% CIs.}
+The confidence intervals are bootstrapped 95\% CIs of trimmed means.}
 		\label{fig:rt1}
@@
 VSUMS are similarly accurate for simple search tasks as more traditional bivariate visualizations, despite containing additional colors categories, and having a more complex relationship between uncertainty, value, and color than traditional charts.
-The confidence intervals are bootstrapped 95\% CIs.}
+The confidence intervals are bootstrapped 95\% CIs of trimmed means.}
 		\label{fig:accuracy1}
@@
 By reducing the number of color categories as uncertainty increases, VSUMs encourage more caution in predictions, making people less likely to consider data with strong, but spurious patterns.
-The confidence intervals are bootstrapped 95\% CIs.}
+The confidence intervals are bootstrapped 95\% CIs of trimmed means.}
 		\label{fig:uncertainty2}

### 2017-03-31 paper/template.tex
Made experimental design less mysterious.
Jeff revisions.

--- 2017-03-31 paper/template.tex
@@
 Some of these options (interaction, animation, and sonification) are not applicable for static charts; even for dynamic charts, many users do not interact with charts in sufficient detail to recover uncertainty information~\cite{nyt2016}.
-While we acknowledge the potential utility of these techniques in uncertainty visualization (for instance, the animation used to convey sampling error in Hullman \ea's \emph{HOPs}~\cite{hullman2015hypothetical}), we limit the scope of our discussion to techniques which center on static charts.
+While we acknowledge the potential utility of these techniques in uncertainty visualization (for instance, the animation used to convey sampling error in Hullman \ea's Hypothetical Outcome Plots~\cite{hullman2015hypothetical}), we limit the scope of our discussion to techniques which center on static charts.
 
@@
 To support our goal of making uncertainty information well-integrated with data, we focus on the techniques of \emph{juxtaposition} (where uncertainty is presented alongside the data), \emph{superposition} (where additional glyphs are overlaid on top of the data), and \emph{explicit encoding} (where an additional visual channel is used to encode the uncertainty).
-Gleicher \ea~\cite{gleicher2011visual} explore these three strategies from the perspective of affording \emph{comparisons}; here, we examine them from the perspective of uncertainty visualization.
+Gleicher \ea~\cite{gleicher2011visual} explore these three strategies from the perspective of affording \emph{comparisons}; here, we examine them from the perspective of incorporating measures of uncertainty.
 
@@
 
-Given these considerations, and the perceptual non-orthogonality of color channels, the construction of \emph{bivariate} color maps is difficult.
+Given these considerations, and the perceptual integrality of color channels~\cite{garner1970integrality, ware2012information}, the construction of \emph{bivariate} color maps is difficult.
 Viewers need to be able to independently assess the variables of interest, but the resulting bivariate map should also be expressive and easy to interpret.
@@
 For instance, a bar can have an arbitrary, non-integer height.
-However, for many visualizations, the perceptual ambiguity caused by having marks that cannot be reliably disambiguated is higher than the gain from having a visual encoding that maps readily to the backing data.
+However, for many visualizations, the perceptual ambiguity caused by having marks that cannot be reliably disambiguated dominates the gains from having a visual encoding that maps readily to the backing data.
 Padilla \ea~\cite{padilla2017evaluating} present one such example, where binned heatmaps result in higher task accuracy than continuous heatmaps.
@@
 
-The general VSUM algorithm is as follows:
+The general VSUM construction algorithm is as follows:
 
@@
 \figref{fig:flow} shows this process in detail, and how it can be used to generate a bivariate map for presenting uncertainty.
-JavaScript code for generating VSUMs given arbitrary d3 quantitative scale functions is available at \url{URL REMOVED FOR REVIEW}.
+JavaScript code for generating VSUMs given arbitrary D3~\cite{bostock2011d3} quantitative scale functions is available at \url{URL-REMOVED-FOR-REVIEW}.
 
@@
 We present a set of examples showing how VSUMs can be applied to realworld datasets.
-The decision of VSUMs to allocate color bins assymmetrically amongst value and uncertainty regions affords greater discriminability in regions of the chart that warrant closer scrutiny, while discouraging exploration of regions with noisy or unreliable signals.
+The decision of VSUMs to allocate color bins asymmetrically amongst value and uncertainty regions affords greater discriminability in regions of the chart that warrant closer scrutiny, while discouraging exploration of regions with noisy or unreliable signals.
 
@@
 Extending the ``confidence fog'' metaphor used in viral genome browsers like LayerCake~\cite{correll2015layercake}, VSUMs (\figref{fig:viralVsum}) enable biologists to explore interesting patterns while not being distracted by noise.
-Smaller scale ``hot spots'' that all be mapped to the same magenta color in the traditional map are visible as bright orange lines.
+Smaller scale ``hot spots'' that are mapped to the same magenta color in the traditional map are visible as bright orange lines.
 Pink hot spots towards the end of the genome (past reference nucleotide $9,000$) that could be plausibly considered potential regions of interest in the 2D map, are encoded as highly uncertain and of ambiguous value in the VSUMs.
-In fact, these regions have systematically low coverage; the lack of reliable data for these regions should discourage analysts from giving their apparently pattern equal consideration from the  stronger signals earlier in the genome.
+In fact, these regions have systematically low coverage; the lack of reliable data for these regions should discourage analysts from giving their apparent pattern equal consideration from the  stronger signals earlier in the genome.
 
@@
 E.g., ``click on the region of the chart with highest uncertainty \emph{and} lowest value.''
-	\item a \textbf{prediction} task, similar to Battleship, where we gave participants charts with both \emph{forecast} and \emph{forecast uncertainty} information, and asked the participants to place tokens on the board in order to optimize expected value.
+	\item a \textbf{prediction} task, inspired by the game Battleship, where we gave participants charts with both \emph{forecast} and \emph{forecast uncertainty} information, and asked the participants to place tokens on the board in order to optimize expected value.
 E.g., ``place your $4$ ships on safe locations on the board.''
@@
 
-We discuss the two experimental tasks in more detail in the following sections. Experimental materials, including data tables, are available at \url{URL REMOVED FOR REVIEW}.
+We discuss the two experimental tasks in more detail in the following sections. Experimental materials, including data tables, are available at \url{URL-REMOVED-FOR-REVIEW}.
 
@@
 If a player fires a missile at a square containing a ship, then that ship is sunk.
-The objective is the maximize the number of enemy ships hit, while minimizing the number of your own ships that are hit.
+The objective is to maximize the number of enemy ships hit, while minimizing the number of your own ships that are hit.
 In our task, participants were given a chart encoding predictions of the location of enemy ships (in the attacking case) or potential missile strikes (in the defending case).
-The \emph{value} component of the prediction was the likelihood of a item being present in a particular region.
+The \emph{value} component of the prediction was the likelihood of an item being present in a particular region.
 The \emph{uncertainty} component was the confidence in this prediction.
@@
 That is, avoiding a large loss is often more valuable than striving for a similarly large gain.
-To avoid potential (in this case, incorrect) transfer learning, the attacking and defending used different value color maps (Plasma and Cool) than in the identification task, and from the opposite framing in this task.
+To avoid the potential for (incorrect) transfer learning, the attacking and defending tasks used different value color maps (Plasma and Cool) than the identification task, and from the opposite framing in this task.
 \figref{fig:taskTwoConditions} shows these differing color maps and framings.
@@
 
-For the identification task, our primary measures of performance were accuracy and response time. Our hypothesis was that VSUMs would have increased performance values compared to the other charts.
+For the identification task, our primary measures of performance were accuracy and response time. Our hypothesis was that VSUMs would exhibit better performance than the other charts.
 
 We performed a one-way ANOVA investigating the effect of chart type (juxtaposed, full bivariate, or VSUM) on accuracy.
-We found only a marginal effect ($F(2,859) = 2.4$, $p=0.09$), although VSUMs had the highest accuracy among the three.
+We did not find a significant effect ($F(2,859) = 2.4$, $p=0.09$), although VSUMs had the highest accuracy.
 \figref{fig:accuracy1} illustrates this result.
@@
 We also performed a one-way ANOVA on the effect of chart type on response time.
-In this case, we did find a significant main effect ($(F,2,859) = 5.0$, $p<0.001$).
+In this case, we found a significant main effect ($(F,2,859) = 5.0$, $p<0.001$).
 A Tukey's test of Honest Significant Difference (HSD) found that participants were significantly slower at performing the task on juxtaposed maps, as opposed to the other map types.
@@
 A Tukey's HSD confirmed that, as with the previous task, response time was significantly longer with juxtaposed maps (trimmed mean response time of $23$ seconds) compared to the other chart types (both with trimmed mean response time of $18$ seconds).
+
 Despite the documented effect of framing in prior work, a one-way ANOVA of task frame on expected value found no significant main effect ($F(1,862)=0.01$, $p=0.9$), and the trimmed means were quite similar (expected value of $0.76$ per guess for attackers, versus $0.75$ per guess for defenders).
 
-While our results only partially support our hypothesis, in that we did not observe statistically different patterns of \emph{expected value}, our did indicate that VSUMs produce different patterns of \emph{acceptable uncertainty}.
-That is, VSUMs promote avoidance of uncertain values, while not noticeably negatively impacting the expected value.
-Juxtaposed maps, by contrast, make the retrieval of uncertainty information significantly more difficult, and so it is possible that participants ignore or downweight it when making predictions, resulting in riskier predictions overall.
-By making the consideration of uncertainty values integral to the design of the chart, VSUMs promote caution when making decisions.
+Our results again partially support our hypotheses.
+We did not observe statistically different patterns of \emph{expected value}, but did find that VSUMs produce different patterns of \emph{acceptable uncertainty}.
+That is, VSUMs promote avoidance of uncertain values, without notably impacting the expected value.
+Juxtaposed maps, by contrast, make the retrieval of uncertainty information significantly more difficult; it is possible that participants ignore or downweight it when making predictions, resulting in riskier predictions overall.
+By making the consideration of uncertainty values integral to the design of the chart, VSUMs appear to promote more cautious decisions.
 
@@
 The way that uncertainty information is presented can have a measurable impact on decision-making.
-The VSUM strategy of only ``showing'' differences when uncertainty is sufficiently low is somewhat isomorphic to the inferential statistics such as effects tests.
-VSUMs can therefore be used to promote more caution in judgments, and lead analysts away from spurious signals in data.
+The VSUM strategy of only ``showing'' differences when uncertainty is sufficiently low is somewhat analogous to the inferential statistics such as effects tests.
+VSUMs might therefore be used to promote more caution in judgments, and lead analysts away from spurious signals in data.
 
@@
 Our experiments dealt with cases where both uncertainty and value were represented by color.
-The perceptual non-separability of color channels is well-known~\cite{ware2012information}, and so the concept of a limited ``budget'' of distinguishable marks easier to quantify and illustrate.
+The perceptual non-separability of color channels is well-known~\cite{garner1970integrality, ware2012information}, and so the concept of a limited ``budget'' of distinguishable marks easier to quantify and illustrate.
 In principle, a VSUM can be created for any combination of visual variables.

### 2017-03-31 paper/template.tex
Inline formulas

--- 2017-03-31 paper/template.tex
@@
 That is, if there are $n$ bins at some uncertainty level $m$, how many bins should there be when uncertainty increases to level $m+1$?
-In \figref{fig:example}, this function is $\left \lfloor {\frac{n}{2}}\right \rfloor$, meaning that the number of value bins decreases by half for each level of uncertainty bins.
+In \figref{fig:example}, this function is $\left \lfloor n / 2 \right \rfloor$, meaning that the number of value bins decreases by half for each level of uncertainty bins.
 If there is too much degradation, then there are fewer levels of uncertainty.
@@
 carriers in January 2017.
-We use standard error ($\frac{\sigma}{\sqrt{n}}$) as our measure of uncertainty.
+We use standard error $\left(\sigma / \sqrt{n}\right)$ as our measure of uncertainty.
 Starting from an minimum discriminability threshold of $18$ units in CIELAB space, we created two alternate heatmaps of this delay information.
@@
 Other factors included the size of the map (either a 4x4 or 8x8 grid), and (for the prediction task), the question framing (attack or defense).
-All factors were presented within subjects.
-Secondary factors were evenly blocked across stimuli.
+Secondary factors were evenly blocked across stimuli, resulting in a 3x2 within subjects design for the identification task, and a 3x2x2 within subjects design for the prediction task.
 Participants reported fatigue effects in piloting, and so we limited the number of replicates in this study: participants saw 18 stimuli (6 for each map type) for the identification task, and 12 (4 for each map type) for the prediction task.

### 2017-03-31 paper/template.tex
Typos

@@
 
-While both maps illustrate a similar temporal trend: flight delays are smaller towards the beginning of the day, increasing on average over time.
+Both maps illustrate a similar temporal trend: flight delays are shorter towards the beginning of the day, increasing on average over time.
 The traditional bivariate map affords only a coarse examination of this trend: two visible ``blocks'' of color, early in the day, and later in the afternoon.
@@
 Smaller scale ``hot spots'' that are mapped to the same magenta color in the traditional map are visible as bright orange lines.
-Pink hot spots towards the end of the genome (past reference nucleotide $9,000$) that could be plausibly considered potential regions of interest in the 2D map, are encoded as highly uncertain and of ambiguous value in the VSUMs.
+Pink hot spots towards the end of the genome (around reference nucleotide $9,500$) that could be plausibly considered potential regions of interest in the 2D map, are encoded as highly uncertain and of ambiguous value in the VSUMs.
 In fact, these regions have systematically low coverage; the lack of reliable data for these regions should discourage analysts from giving their apparent pattern equal consideration from the  stronger signals earlier in the genome.
@@
 \begin{enumerate}
-	\item an \textbf{identification} task, where we gave participants charts with value and uncertainty information, and asked them to locate specific regions.
+	\item An \textbf{identification} task, where we gave participants charts with value and uncertainty information, and asked them to locate specific regions.
 E.g., ``click on the region of the chart with highest uncertainty \emph{and} lowest value.''
-	\item a \textbf{prediction} task, inspired by the game Battleship, where we gave participants charts with both \emph{forecast} and \emph{forecast uncertainty} information, and asked the participants to place tokens on the board in order to optimize expected value.
+	\item A \textbf{prediction} task, inspired by the game Battleship, where we gave participants charts with both \emph{forecast} and \emph{forecast uncertainty} information, and asked the participants to place tokens on the board in order to optimize expected value.
 E.g., ``place your $4$ ships on safe locations on the board.''
@@
 In all cases, we generated color maps based on the requirement that no two colors in the resulting color scheme were closer than 18 units of Euclidean distance in CIELAB space.
-This threshold results in a $9$ color bivariate map, $15$ color VSUM for all color scales used in this study.
+This threshold results in a $9$-color bivariate map and $15$-color VSUM for all color scales used in this study.
 Given the size of the marks used in our study, this threshold distance is sufficient for a predicted discriminability rate of over 95\%~\cite{stone2014engineering}.
@@
 
-For the identification task, we gave participants bivariate representations of uncertainty, and asked to identify a location meeting a given criteria, e.g., ``Click on the location with the greatest uncertainty.'' There were three potential criteria, mapping to the three vertices of a VSUM: 1) high value, low uncertainty, 2) low value, low uncertainty, and 3) high uncertainty.
+For the identification task, we gave participants bivariate representations of uncertainty, and asked them to identify a location meeting a given criteria, e.g., ``Click on the location with the greatest uncertainty.'' There were three potential criteria, mapping to the three vertices of a VSUM: 1) high value, low uncertainty, 2) low value, low uncertainty, and 3) high uncertainty.
 Each map contained four locations, one of each combination of maximum and minimum value and uncertainty.
@@
 VSUMs are designed for the \emph{integration} of value and uncertainty.
-Designers should take care in considering how when and how this integration is desirable.
+Designers should take care in considering when and how this integration is desirable.
 

### 2017-03-31 paper/template.tex
typo fix
Petty hyphenation fix.
dropped bad sentence rather than fix it
Re-hyphenating pass

@@
 \newcommand{\etc}{{etc.}\xspace}
+\hyphenation{LayerCake}
 
@@
 As the experimental tasks required multi-hue color perception, we presented participants with a set of Ishihara plates~\cite{hardy1945tests} as a pretest, and excluded participants who either misidentified the values in the plates, or who self-reported as having a color vision deficiency (CVD) in the post-test.
-Based on piloting, we paid participants \$2 dollars an hour, for a target rate of \$8/hour.
+Based on piloting, we paid participants \$2 dollars for participation, for a target rate of \$8/hour.
 After completion of the main tasks, we solicited demographic information.
@@
 Given the deleterious effects of latency on data analysis~\cite{liu2014effects}, designers ought to carefully consider the tradeoffs between expressiveness of visual encoding, and usability.
-VSUMs represent a middle ground between these two approaches, and have some of the benefits of both.
 
@@
 This integration introduces additional complexity in the design and presentation of data.
-Value Suppressing Uncertainty Maps represent one strategy for dealing with this complexity, by assigning marks properties in a way that supports the disambiguation of values in data where uncertainty is low, but suppresses these judgments when uncertainty is high.
+Value-Suppressing Uncertainty Maps represent one strategy for dealing with this complexity, by assigning marks properties in a way that supports the disambiguation of values in data where uncertainty is low, but suppresses these judgments when uncertainty is high.
 This decision of how to allocate visual variables promotes patterns of decision-making that make responsible use of uncertainty information, discouraging comparison of values in unreliable regions of the data, and promoting comparison in regions of high certainty.

### 2017-03-31 paper/template.tex
Strategies
Added data tables
Submitted version.

@@
 
+\strategyFig
+
 To support our goal of making uncertainty information well-integrated with data, we focus on the techniques of \emph{juxtaposition} (where uncertainty is presented alongside the data), \emph{superposition} (where additional glyphs are overlaid on top of the data), and \emph{explicit encoding} (where an additional visual channel is used to encode the uncertainty).
 Gleicher \ea~\cite{gleicher2011visual} explore these three strategies from the perspective of affording \emph{comparisons}; here, we examine them from the perspective of incorporating measures of uncertainty.
+\figref{fig:strategies} shows simple examples of each of these three techniques for presenting both uncertainty and value.
 
@@
 Thus, while we report effect sizes in terms of trimmed means, we perform inferential statistical tests on the standard means.
+Asterisks in our charts denote significant differences, as the reported by a post-hoc Tukey's test of Honest Significant Difference (HSD).
 
@@
 In this case, we found a significant main effect ($(F,2,859) = 5.0$, $p<0.001$).
-A Tukey's test of Honest Significant Difference (HSD) found that participants were significantly slower at performing the task on juxtaposed maps, as opposed to the other map types.
+A Tukey's HSD found that participants were significantly slower at performing the task on juxtaposed maps, as opposed to the other map types.
 \figref{fig:rt1} illustrates this result.

### 2017-08-22 paper/figures.tex
Re-hyphenating pass
Strategies
Annotate significance
Pretty figures
Fixed minor typo in submitted version.
Moved paper to CHI format

@@
 	\caption{A standard bivariate map (left) and a VSUM (right), used to encode an identical 5x5 grid of random data.
-Both use the same visual channels to encode value (position along the Viridis~\cite{viridis} color map) and uncertainty (lightness and saturation), and have an equal standard of perceptual discriminability (at least 18 units of distance in CIELAB color space between colors).
+Both use the same visual channels to encode value (position along the Viridis~\protect\cite{viridis} color map) and uncertainty (lightness and saturation), and have an equal standard of perceptual discriminability (at least 18 units of distance in CIELAB color space between colors).
 However, highly uncertain values result in colors that are very close together in the bivariate case, meaning the full bivariate map is only 9 bins under these constraints\,---\,a 4x4 map would result in colors that are perceptually too close together.
@@
 		\centering
-		\includegraphics[width=0.8\textwidth]{conditions.pdf}
+		\begin{subfigure}{0.3\textwidth}
+			\centering
+			\includegraphics[width=.45\textwidth]{conditions1.pdf}
+			\caption{Traditional Bivariate Map}
+		\end{subfigure}
+		~
+		\begin{subfigure}{0.3\textwidth}
+			\centering
+			\includegraphics[width=\textwidth]{conditions2.pdf}
+			\caption{Juxtaposed Univariate Maps}
+		\end{subfigure}
+		~
+		\begin{subfigure}{0.3\textwidth}
+			\centering
+			\includegraphics[width=0.45\textwidth]{conditions3.pdf}
+			\caption{Value-Suppressing Uncertainty Map}
+		\end{subfigure}
 		\caption{The three graph types in our evaluation.
@@
 		\centering
-		\includegraphics[width=.8\columnwidth]{rt1.png}
+		\includegraphics[height=2.4cm]{rt1.pdf}
 		\caption{The effect of chart type on response time for the identification task.
@@
 		\centering
-		\includegraphics[width=.8\columnwidth]{accuracy1.png}
+		\includegraphics[height=2.4cm]{accuracy1.pdf}
 		\caption{The effect of chart type on accuracy for the identification task.
@@
 		\centering
-		\includegraphics[width=.8\columnwidth]{uncertainty2.png}
+		\includegraphics[height=2.4cm]{uncertainty2.pdf}
 		\caption{The effect of chart type on the average uncertainty in predictions for the prediction task.
@@
 }
+
+\newcommand{\strategyFig}{
+	\begin{figure}[t]
+		\centering
+		\includegraphics[width=.9\columnwidth]{strategies.pdf}
+		\caption{Strategies for encoding uncertainty information along with values.
+From left to right: encode uncertainty as a separate visualization, overlay uncertainty information with separate glyphs in the same visualization, or encode uncertainty with a separate visual channel on the same glyphs.}
+		\label{fig:strategies}
+	\end{figure}
+}

### 2017-08-22 paper/main.tex
Moved paper to CHI format

@@
-% ---------------------------------------------------------------------
-% EG author guidelines plus sample file for EG publication using LaTeX2e input
-% D.Fellner, v1.17, Sep 23, 2010
+\section{Introduction}
+certainty is an inescapable component of collecting, presenting, and using data.
+A common goal in the communication of uncertainty is promoting \emph{uncertainty-aware decisions}: the audience should be aware of the risks and rewards of certain decisions, modulate their confidence in their conclusions, and perhaps refrain from making a decision at all if there is too much uncertainty.
+A way that designers can contribute to this goal is by ensuring that uncertainty information is \emph{well-integrated} with the rest of the data.
+That is, it should be difficult to discount or ignore the uncertainty in a dataset.
 
-\title[EG \LaTeX\ Author Guidelines]%
-      {\LaTeX\ Author Guidelines for EUROGRAPHICS Proceedings Manuscripts}
+Simultaneous presentation of uncertainty and value necessitates the construction of a bivariate map\,---\,a relation, in terms of visual variables, between 2-tuples $(\text{value}, \text{uncertainty})$ and mark properties.
+The design of bivariate maps is difficult and, due to the interference and interplay between different visual variables, often leads to maps with few discriminable categories.
 
-% for anonymous conference submission please enter your SUBMISSION ID
-% instead of the author's name (and leave the affiliation blank) !!
-\author[D. Fellner \& S. Behnke]
-       {D.\,W. Fellner\thanks{Chairman Eurographics Publications Board}$^{1,2}$
-        and S. Behnke$^{2}$
-%        S. Spencer$^2$\thanks{Chairman Siggraph Publications Board}
-        \\
-% For Computer Graphics Forum: Please use the abbreviation of your first name.
-         $^1$TU Darmstadt \& Fraunhofer IGD, Germany\\
-         $^2$Institut f{\"u}r ComputerGraphik \& Wissensvisualisierung, TU Graz, Austria
-%        $^2$ Another Department to illustrate the use in papers from authors
-%             with different affiliations
-       }
+In this paper, we contribute \textbf{Value-Suppressing Uncertainty Maps} (VSUMs) for integrating data and uncertainty information in visualizations.
+VSUMs intentionally alias together data values with high uncertainty, affording greater discriminability as uncertainty decreases.
+Compared to a traditional two-dimensional bivariate map, with differing outputs for each combination of value and uncertainty, VSUMs can be thought of as arcs: as uncertainty increases, values are mapped to smaller and smaller sets of outputs, culminating in a singularity where all inputs are mapped to an identical, highly uncertain mark regardless of their data value.
+\figref{fig:example} shows examples of both a more traditional bivariate map and a VSUM.
 
-% ------------------------------------------------------------------------
+We describe the motivations behind the use of VSUMs, examples of their utility for decision-making under uncertainty, and assess VSUMs in a crowdsourced experiment.
+Our experimental results indicate that VSUMs create close integration between uncertainty and data, promoting rapid but cautious decision-making.
 
-% if the Editors-in-Chief have given you the data, you may uncomment
-% the following five lines and insert it here
-%
-% \volume{27}   % the volume in which the issue will be published;
-% \issue{1}     % the issue number of the publication
-% \pStartPage{1}      % set starting page
+\exampleFig
+\section{Related Work}
 
-%-------------------------------------------------------------------------
-\begin{document}
+% What this section needs to do:
+% State of the art in uncertainty vis, with heavy emphasis on MacEachren & co.
+% Bivariate maps are hard!
+% Binning colormaps is de rigueur
+% The variables we'd want to use for uncertainty (like value or alpha or size) mess with color discriminability
 
-% \teaser{
-%  \includegraphics[width=\linewidth]{eg_new}
-%  \centering
-%   \caption{New EG Logo}
-% \label{fig:teaser}
-% }
+Despite the acknowledged importance of uncertainty in understanding data, explicit representations of uncertainty are often missing from visualizations~\cite{boukhelifa2009uncertainty}.
+This is partially due to the complexity of uncertainty as a concept.
+In typologies from Thomson \ea~\cite{thomson2005typology} and Buttenfield \& Beard~\cite{buttenfield1994graphical}, the authors note that many, occasionally contradictory, concepts can fall under the category of ``uncertainty,'' including data quality, sampling error, credibility, and provenance.
 
-\maketitle
+Reviewing the state of the art in uncertainty visualization in the field, Greithe \ea~\cite{griethe2006visualization} and Brodlie \ea~\cite{brodlie2012review} present lists of potential techniques for conveying uncertainty in concert with data.
+Some of these options (interaction, animation, and sonification) are not applicable for static charts; even for dynamic charts, many users do not interact with charts in sufficient detail to recover uncertainty information~\cite{nyt2016}.
+While we acknowledge the potential utility of these techniques in uncertainty visualization (for instance, the animation used to convey sampling error in Hullman \ea's Hypothetical Outcome Plots~\cite{hullman2015hypothetical}), we limit the scope of our discussion to techniques which center on static charts.
 
-\begin{abstract}
-   The ABSTRACT is to be in fully-justified italicized text, 
-   between two horizontal lines,
-   in one-column format, 
-   below the author and affiliation information. 
-   Use the word ``Abstract'' as the title, in 9-point Times, boldface type, 
-   left-aligned to the text, initially capitalized. 
-   The abstract is to be in 9-point, single-spaced type.
-   The abstract may be up to 3 inches (7.62 cm) long. \\
-   Leave one blank line after the abstract, 
-   then add the subject categories according to the ACM Classification Index 
-   (see http://www.acm.org/class/1998/).
+There are a number of techniques for communicating values and uncertainty in one-dimensional statistical graphics~\cite{olston2002visualizing, wickham201140, neyman1937outline}.
+Techniques such as error bars, confidence bands, and violin plots~\cite{hintze1998violin} use a separate position channel to encode uncertainty.
+The techniques in this paper also apply to 2D visualizations such as heatmaps.
+To our knowledge, there are no established standards of visualizing uncertainty in heatmaps.
 
-\begin{classification} % according to http://www.acm.org/class/1998/
-\CCScat{Computer Graphics}{I.3.3}{Picture/Image Generation}{Line and curve generation}
-\end{classification}
+\strategyFig
 
-\end{abstract}
+To support our goal of making uncertainty information well-integrated with data, we focus on the techniques of \emph{juxtaposition} (where uncertainty is presented alongside the data), \emph{superposition} (where additional glyphs are overlaid on top of the data), and \emph{explicit encoding} (where an additional visual channel is used to encode the uncertainty).
+Gleicher \ea~\cite{gleicher2011visual} explore these three strategies from the perspective of affording \emph{comparisons}; here, we examine them from the perspective of incorporating measures of uncertainty.
+\figref{fig:strategies} shows simple examples of each of these three techniques for presenting both uncertainty and value.
 
-%-------------------------------------------------------------------------
-\section{Introduction}
+\subsection{Strategies for Encoding Uncertainty}
 
-Please follow the steps outlined in this document very carefully when
-submitting your manuscript to Eurographics.
+\emph{Juxtaposition} creates two explicit visualizations: an ``uncertainty map'' and a ``data map.'' MacEachren~\cite{maceachren1992visualizing} portrays uncertainty visualization as the design problem of how to \emph{unify} these two maps, and so this explicit separation can be counterproductive to the goal of information fusion.
+Juxtaposition also creates an additional search task for viewers, who must register hotspots in one map with corresponding features of the other map.
+Rather than complete this task, viewers may choose to ignore the uncertainty information altogether.
+This ignoreability is a key deficiency for information fusion tasks.
+Nevertheless, especially in cases where the quantification of uncertainty is a separate analysis process from the measuring of data values, juxtaposition is a common design tactic for conveying uncertainty~\cite{moritz2017trust}.
 
-You may as well use the \LaTeX\ source as a template to typeset your own
-paper. In this case we encourage you to also read the \LaTeX\ comments
-embedded in the document.
+\emph{Superposition}, or the overlay of glyphs representing uncertainty on top of a data map, is another common strategy.
+Error bars are a common example of this strategy: a mark indicating measurement error is superimposed atop a mark representing a point value.
+Ware's~\cite{ware2009quantitative} \emph{textons} are an example of this strategy applied to thematic maps: monochrome glyphs that can encode an ordinal variable, or a binned quantitative variable.
+Greithe \ea~\cite{griethe2006visualization} and MacEachren~\cite{maceachren1992visualizing,maceachren1998visualizing} present other examples of superimposed uncertainty in thematic maps, including overlaid grids with ``fuzzy'' lines to indicate uncertainty, or overlaid treemaps with coarser and coarser resolution in uncertain regions.
+Costs to this approach include obscuring the underlying data map (which can cause uncertainty information to dominate the display~\cite{brodlie2012review}), potential visual interactions that can impair the legibility of glyphs (for instance, simultaneous contrast effects), and the necessity of performing spatial binning to ensure that there are a discrete number of uncertainty glyphs.
+How this binning is performed can introduce undesirable artifacts on the resulting visualization, suppressing important signals or highlighting spurious ones~\cite{battersby2016shapes}.
 
-%-------------------------------------------------------------------------
-\section{Instructions}
+\emph{Explicit encoding} directly represents uncertainty with a visual channel, simultaneously with data.
+VSUMs employ this strategy.
+In addition to overcoming the drawbacks of the other techniques mentioned in this section, we believe that explicit encoding represents the most general form of the solution of how to fuse uncertainty information with data.
+The most pressing design decisions for explicit encoding are 1) which visual variables to use to encode uncertainty and 2) how to design bivariate maps which afford the encoding of both value and uncertainty.
+We deal with these two questions in greater detail below.
 
-Please read the following carefully.
+\subsection{Visual Variables for Representing Uncertainty}
 
-%-------------------------------------------------------------------------
-\subsection{Language}
+The decision to explicitly encode uncertainty increases the dimensionality of the data, and so requires the use of (at least one) additional visual channel ~\cite{brodlie2012review}.
+Uncertainty therefore inherently increases the visual complexity of a visualization.
+When the data are already complex to convey, and many of the more common or accurate visual variables are in use, allocating an additional dimension is non-trivial.
+As the number of dimensions increases, finding visual variables that are both perceptually accurate (in either estimation of quantity or discrimination of category) as well as perceptually separable from all the other encoding channels, becomes more and more difficult.
 
-All manuscripts must be in English.
+A further hurdle is that not all visual variables are well-suited for conveying uncertainty.
+MacEachren \ea~\cite{maceachren2012visual} evaluate a number of visual variables with respect to their \emph{semiotic fit} for representing uncertainty.
+They observe that certain visual variables such as blurriness and transparency seem to have a more intuitive connection to uncertainty than other variables such as shape or hue.
+Unfortunately, Boukhelifa \ea~\cite{boukhelifa2012evaluating} find that many visual variables habitually used for conveying uncertainty, such as blur and value, are also difficult to estimate.
+This results in a preference/performance gap where designers must choose between encoding uncertainty in a way that is intuitive but error prone, or use higher fidelity channels that may be more difficult to interpret.
 
-%-------------------------------------------------------------------------
-\subsection{Margins and page numbering}
+Foundational work in psychophysics and graphical perception make use of the concept of a ``just noticeable difference'' (JND)\cite{cleveland1984graphical}.
+A 50\% JND is a difference between two stimuli that is detectable 50\% of the time.
+For scales and legends, we would want a significantly higher standard of unambiguity.
+That is, the different levels of visual variables used in visualizations should be readily and reliably distinguishable.
+Complicating this goal is that a JND as measured in a controlled environment with cooperative participants can be considerably lower than JNDs measured in the sorts of heterogeneous conditions found in real environments~\cite{szafir2014adapting}.
 
-All printed material, including text, illustrations, and charts,
-must be kept within a print area %6.31 inches (16.03 cm) 
-7 inches (17.7 cm) wide by
-%9.10 inches (23.13 cm) 
-9.44 inches (24 cm) high. Do not write or print anything
-outside the print area. Number your pages on odd sites right
-above, on even sites left above, no page number on the first site.
-Do not use page numbering within the final version of your paper.
+\subsection{Bivariate Maps}
 
-%------------------------------------------------------------------------
-\subsection{Formatting your paper}
+For spatial visualizations like choropleth maps, heatmaps, and treemaps, visual channels such as position and length are reserved for data variables other than value (such as geographic location or relative size).
+In these situations, data value is often encoded using color.
+Colors in such univariate quantitative color maps should be sufficiently far apart as to be perceptually distinguishable~\cite{ware1988color}, and vary in lightness as well as hue to afford an implicit ordering of value~\cite{borland2007rainbow,rogowitz2001blair}.
+Different choices of color maps can highlight different features of the data, and should be chosen with care based on the data types and important features to be visualized~\cite{rogowitz1996not}.
 
-% All text with the exception of the abstract must be in a two-column format.
-% The total allowable width of the text area -- including header and footer
-% lines -- is 161\,mm (6.34 inch) wide by 231\,mm (9.10 inch) high.
-% 
-% Columns are to be 76\,mm (3.0 inch) wide, with a 8\,mm (0.315 inch) space 
-% between them.
+Given these considerations, and the perceptual integrality of color channels~\cite{garner1970integrality, ware2012information}, the construction of \emph{bivariate} color maps is difficult.
+Viewers need to be able to independently assess the variables of interest, but the resulting bivariate map should also be expressive and easy to interpret.
+Bivariate maps with these constraints are often limited to relatively few categories (say, a 3x3 matrix as in \figref{fig:example}) \cite{robertson1986generation,trumbo1981theory}.
+In general, the quality of bivariate color maps is a multivariate measure involving consideration of not just the component color channels, but also the interpolation scheme and the color of the surround~\cite{bernard2015survey}.
+However, even simple bivariate maps can be difficult for a general audience to interpret \cite{wainer1980empirical}.
 
-All text with the exception of the abstract must be in a two-column format.
-The total allowable width of the text area -- including header and footer
-lines -- is 177\,mm (7 inch) wide by 245\,mm (9.64 inch) high.
+Dunn~\cite{dunn1989dynamic} attempts to circumvent these issues by dynamically allocating bins to bivariate maps.
+As the number of distinct, separable colors is limited, the dynamic approach allocates bins to regions of the highest utility.
+For instance, placing the dividing line between bins in areas aligned with a trend line allows the color to encode information about the sign of the residual with respect to each variable.
+This insight, that the bins in a bivariate map are a limited resource and so should be ``spent'' wisely, drives the creation of VSUMs, which allocate more bins in a bivariate map to regions with less uncertainty.
 
-Columns are to be 84\,mm (3.3 inch) wide, with a 8\,mm (0.315 inch) space 
-between them.
+\section{Value-Suppressing Uncertainty Maps}
 
-The space between the header line and the first line of the text body and
-between the last line of the text body and the footer line is 5\,mm
-(0.196 inch) each.
+Value-Suppressing Uncertainty Maps (VSUMs) are a technique for creating bivariate maps of data \emph{value} and \emph{uncertainty}.
+These maps circumvent the issues in the creation of traditional bivariate maps by making the following assumptions:
 
-%-------------------------------------------------------------------------
-\subsection{Type-style and fonts}
+\begin{enumerate}
+	\item \textbf{Designers have a limited budget of discriminable marks.} That is, viewers cannot easily or reliably discriminate between an infinite number of categories.
+Rather than accept this perceptual ambiguity, designers ought to \emph{alias} together marks into a finite, but sufficiently discriminable, set.
+	\item \textbf{Uncertain values should have less weight than certain values.} Weight, in this case, might mean literal visual weight, or impact on the decision-making process.
+\end{enumerate}
 
-Wherever Times is specified, Times Roman may also be used. If
-neither is available on your word processor, please use the font
-closest in appearance to Times that you have access to. Only
-Type-1 fonts will be accepted.
+%MC this paragraph is a little iffy, and sort of "the lady doth protest too much" territory.
+we might be able to get away with "people bin colors all the time.
+and in fact, you really should be binning your colors anyway."
+The first assumption seems to contradict Mackinlay's \emph{expressiveness principle}~\cite{mackinlay1986automating}, where the visual encoding of the data should reflect the composition of the data.
+For instance, if my data are a collection of measurements of some continuous variable, then it would make sense to encode these data using a continuous \emph{visual} variable, which would then mean an arbitrary number of potential marks.
+For instance, a bar can have an arbitrary, non-integer height.
+However, for many visualizations, the perceptual ambiguity caused by having marks that cannot be reliably disambiguated dominates the gains from having a visual encoding that maps readily to the backing data.
+Padilla \ea~\cite{padilla2017evaluating} present one such example, where binned heatmaps result in higher task accuracy than continuous heatmaps.
 
-MAIN TITLE. The title should be in Times 17-point, boldface type and
-centered. Capitalize the first letter of nouns, pronouns, verbs, adjectives,
-and adverbs; do not capitalize articles, coordinate conjunctions, or
-prepositions (unless the title begins with such a word). Leave two blank
-lines after the title.
+In some cases, this second assumption is violated.
+For instance, an analyst might be interested in ``long tail risks'' or other ``black swan events,'' where the impact of a value, no matter how uncertain, must be considered and planned for~\cite{taleb2011black}.
+Other analysis tasks (such as filtering out outliers), require increased, rather than decreased, discriminability when uncertainty is high.
+However, for many information fusion tasks, the assumption is that uncertainty is related to data quality, or the uncertainty of data values~\cite{riveiro2007evaluation}.
+As with null hypothesis significance testing, the aim would then be to avoid Type I errors, where patterns arising from statistical noise or measurement error are interpreted as significant signals.
 
-AUTHOR NAME(s) and AFFILIATION(s) are to be centered beneath the title and
-printed in Times 9-point, non-boldface type. This information is to be
-followed by two blank lines.
+If both of these assumptions hold, then it follows that the designer of a bivariate map should allocate more mark types to certain values, and fewer mark types to uncertain values.
+VSUMs codify this decision by \emph{reducing the number of mark categories for representing value as uncertainty increases.} This means that data encoded using a VSUM will make visible only the largest of differences in uncertain data, but highlight comparatively small differences in value when uncertainty is low.
+VSUMs therefore act as both a filtering mechanism (in that values with too much uncertainty are all mapped to the same glyph), as well as an implicit test of effect size (in that smaller and smaller changes in value are visible as uncertainty decreases).
+This strategy of dampening the low quality or high uncertainty values in maps in order to focus on more informative regions has measurable benefits, including the removal of statistically spurious visual patterns, and the highlighting of regions of interest~\cite{correll2017surprise}.
 
-The ABSTRACT is to be in a one-column format. The MAIN TEXT is to be in a
-two-column format.
+\subsection{Method}
 
-MAIN TEXT. Type main text in 9-point Times, single-spaced. Do \emph{not} use
-double-spacing. All paragraphs should be indented 1 em (the length of the
-dash in the actual font). Make sure your text is fully justified -- that is,
-flush left and flush right. Please do not place any additional blank lines
-between paragraphs. Figure and table captions should be 9-point Times
-boldface type as in Figure~\ref{fig:firstExample}.
+\flowFig
 
-\noindent Long captions should be set as in Figure~\ref{fig:ex1} or
-Figure~\ref{fig:ex3}.
+The general VSUM construction algorithm is as follows:
 
-\begin{figure}[htb]
-   % an empty figure just consisting of the caption lines
-   \caption{\label{fig:ex1}
-     'Empty' figure only to serve as an example of long caption requiring 
-     more than one line. It is not typed centered but aligned on both sides.}
-\end{figure}
+\begin{enumerate}
+	\item \textbf{Select a visual variable to encode value}. In this paper, this variable is position along some established color scale.
+	\item \textbf{Select a visual variable to encode uncertainty}.
+We prefer variables that have a somewhat high semiotic connection with uncertainty (such as lightness and size), but still have a reasonable number of perceptually discriminable layers (as opposed to other, more intuitive channels such as blur or sketchiness).
+In most cases in this paper, we encode uncertainty as interpolation to white or as glyph size.
+	\item \textbf{Select a discriminability threshold}.
+Choose some minimum distance, in a perceptual space, between marks.
+For many visual variables (such as length or hue), the perceptual discriminability between marks has been assessed in prior literature, and so the only step is to choose a value in terms of the just noticeable distance (JND).
+For less common visual variables, a perceptual model may be built empirically (as in Demiralp \ea~\cite{demiralp2014learning}).
+	\item \textbf{Choose how quickly categories \emph{degrade}}.
+That is, if there are $n$ bins at some uncertainty level $m$, how many bins should there be when uncertainty increases to level $m+1$?
+In \figref{fig:example}, this function is $\left \lfloor n / 2 \right \rfloor$, meaning that the number of value bins decreases by half for each level of uncertainty bins.
+If there is too much degradation, then there are fewer levels of uncertainty.
+If there is no degradation, the resulting map is just a traditional bivariate map.
+	\item \textbf{Choose the largest number of initial categories $n$ such that the discriminability threshold is met.} For each pair of marks in the map, perform a pairwise comparison of discriminability.
+If no marks are perceptually ambiguous by the standards of your threshold, the map is complete.
+Otherwise, decrement $n$.
+\end{enumerate}
 
-\noindent
-Figures which need the full textwidth can be typeset as Figure~\ref{fig:ex3}.
+\figref{fig:flow} shows this process in detail, and how it can be used to generate a bivariate map for presenting uncertainty.
+JavaScript code for generating VSUMs given arbitrary D3~\cite{bostock2011d3} quantitative scale functions is available at \url{URL-REMOVED-FOR-REVIEW}.
 
-\noindent Callouts should be 9-point Times, non-boldface type. Initially
-capitalize only the first word of section titles and first-, second-, and
-third-order headings.
+Correll \ea~\cite{correll2015layercake,correll2011visualizing} use a precursor of VSUMs in their LayerCake genomics visualization tool, where marks representing genomic data with increasing uncertainty are mapped to a smaller and smaller set of increasingly grey colors, creating the effect of uncertain values retreating into ``confidence fog'' while highly certain values remain prominent.
+Other bivariate visualizations implicitly alias together uncertain values.
+For instance, if uncertainty is encoded by transparency, a maximally uncertain glyph may be entirely transparent, and so impossible to distinguish from any other maximally uncertain glyph.
+Other channels with a semiotic connection to uncertainty, such as saturation, value, blur, or size, also have deleterious effects on the disambiguation of colors and shapes.
+In both cases, the property of aliasing is \emph{ad hoc}, and places no guarantees on the discriminability of colors.
+The binning and degradation approach of VSUMs makes the choice to alias values explicit to both the designer and the viewer, and results in a bivariate mapping with known perceptual properties.
 
-FIRST-ORDER HEADINGS. (For example, \textbf{1. Introduction}) should be Times
-9-point boldface, initially capitalized, flush left, with one blank line
-before, and one blank line after.
+\subsection{Design Considerations}
 
-SECOND-ORDER HEADINGS. (For example, \textbf{2.1. Language}) should be Times
-9-point boldface, initially capitalized, flush left, with one blank line
-before, and one after. If you require a third-order heading (we discourage
-it), use 9-point Times, boldface, initially capitalized, flush left, preceded
-by one blank line, followed by a period and your text on the same line.
+\performanceFig
 
-The headline \emph{(authors / title)} must be shortened if it uses the full 
-two column width of the main text.
-There must be enough space for the page numbers. Please use ``et al.'' if 
-there are more than three authors and specify a shortened version for your title.
-%-------------------------------------------------------------------------
-\subsection{Footnotes}
+There are multiple choices that designers must make before creating a VSUM.
+In particular, they must choose 1) which visual channels map to value and uncertainty, 2) a threshold of perceptual distinguishability (and, implicitly, a corresponding metric), and 3) a degradation function to determine how quickly uncertain values are aliased.
 
-Please do \emph{not} use footnotes at all!
+Often, the graphical conventions of a chart reserve variables such as position and area for other uses, and determine the visual channel to be used for value.
+In heatmaps, for instance, value is regularly encoded using color, while vertical and horizontal position are reserved for location.
+In this paper, we primarily deal with the case where value is encoded using color.
+While there are some similar conventions for representing uncertainty, generally designers have more freedom in choosing a visual channel for uncertainty~\cite{maceachren1992visualizing}.
+We recommend channels with both a strong semiotic connection to uncertainty as well as a relatively large number of perceptually distinguishable levels.
+Here, we focus on value/saturation (or, equivalently, transparency against a white background) and size.
+Correll \& Gleicher~\cite{correll2013error} show that even audiences without statistical backgrounds can interpret uncertainty information encoded in these channels.
 
-%-------------------------------------------------------------------------
-\subsection{References}
+Another reason for limiting our discussion of bivariate mappings to color and size channels is that the perceptual distinguishability of these channels has been well-studied by prior work.
+CIELAB and other color spaces are designed to approximate perceptual uniformity\,---\,that is, distance in the color space should universally correspond with perceptual distance.
+Recent work~\cite{stone2014engineering,szafir2014adapting} has investigated the discriminability of colors in such spaces not only under laboratory conditions, but also among the disparate display conditions and color vision proficiency of crowdworkers.
+Similarly, formative work by Cleveland \& McGill~\cite{cleveland1984graphical} assesses the discriminability of marks of different height, and this too has been extended both to a larger population of crowdworkers, and to the area channel specifically~\cite{heer2010crowdsourcing,talbot2014four}.
 
-List all bibliographical references in 9-point Times, single-spaced, at the
-end of your paper in alphabetical order. When referenced in the text, enclose
-the citation index in square brackets, for example~\cite{Lous90}. Where
-appropriate, include the name(s) of editors of referenced books.
+We use values from these works as benchmarks for selecting thresholds of discriminability.
+The JNDs of these channels provides a useful minimum; in practice, we would wish to create maps with marks that are far more reliably distinguishable than the 50\% or 75\% discriminability rate of a JND.
+\figref{fig:performance} shows how these benchmarks can be used to select reasonable thresholds.
+Other factors, such as nameability and aesthetic preference, can be added in as additional weights impacting the desirability of certain color palettes, as in Gramazio \ea~\cite{gramazio2017colorgorical}.
 
-For your references please use the following algorithm:
-\begin{itemize} 
-\item \textbf{one} author: first 3 chars plus year -- 
-      e.g.\ \cite{Lous90}
-\item \textbf{two}, \textbf{three} or \textbf{four} authors: first char
-      of each family name plus year --  e.g.\ \cite{Fellner-Helmberg93} 
-      or \cite{Kobbelt97-USHDR} or \cite{Lafortune97-NARF}
-\item \textbf{more than 4} authors: first char of family name from 
-      first 3 authors followed by a '*' followed by the year -- 
-      e.g.\ \cite{Buhmann:1998:DCQ} or \cite{FolDamFeiHug.etal93} 
-\end{itemize}
+How quickly the number of bins decreases is another parameter of VSUMs.
+If this degradation is too fast, then we may only have a few discriminable levels of uncertainty.
+If this degradation is too slow, then we can be overwhelmed with colors.
+For instance, if we only decrement by one bin per level of uncertainty, there would be $1,485$ different colors in a VSUM with a tolerance threshold of $1$ unit in CIELAB space, far too much complexity for a discrete color mapping.
+In the examples in this paper, we halve the number of bins at each increasing level of uncertainty.
+When the number of initial bins at the highly certain region is a power of $2$ (as in Fig.
+\ref{fig:example}), this imparts a tree structure to the resulting VSUM.
+As uncertainty increases, bins in the previous level map to one and only one ``parent'' in the next uncertainty level, resulting in fewer crossings of color boundaries among rows.
 
-For BibTeX users a style file \ \texttt{eg-alpha.bst} \ is available which
-uses the above algorithm.
+\subsection{Examples}
 
-%-------------------------------------------------------------------------
-\subsection{Illustrations, graphs, and photographs}
+We present a set of examples showing how VSUMs can be applied to realworld datasets.
+The decision of VSUMs to allocate color bins asymmetrically amongst value and uncertainty regions affords greater discriminability in regions of the chart that warrant closer scrutiny, while discouraging exploration of regions with noisy or unreliable signals.
 
-All graphics should be centered.
+\subsubsection{Air Travel Delay}
 
-%%%
-%%% Figure 1
-%%%
-\begin{figure}[htb]
-  \centering
-  % the following command controls the width of the embedded PS file
-  % (relative to the width of the current column)
-  \includegraphics[width=.8\linewidth]{sampleFig}
-  % replacing the above command with the one below will explicitly set
-  % the bounding box of the PS figure to the rectangle (xl,yl),(xh,yh).
-  % It will also prevent LaTeX from reading the PS file to determine
-  % the bounding box (i.e., it will speed up the compilation process)
-  % \includegraphics[width=.95\linewidth, bb=39 696 126 756]{sampleFig}
-  %
-  \parbox[t]{.9\columnwidth}{\relax
-           For all figures please keep in mind that you \textbf{must not}
-           use images with transparent background! 
-           }
-  %
-  \caption{\label{fig:firstExample}
-           Here is a sample figure.}
-\end{figure}
+\airlineFig
 
-If your paper includes images, it is very important that they are of
-sufficient resolution to be faithfully reproduced.
+Using a public data set from the U.S.
+Bureau of Transportation Statistics~\cite{bts}, we created a heatmap showing average flight delay for different times of the day and days of the week for U.S.
+carriers in January 2017.
+We use standard error $\left(\sigma / \sqrt{n}\right)$ as our measure of uncertainty.
+Starting from an minimum discriminability threshold of $18$ units in CIELAB space, we created two alternate heatmaps of this delay information.
+\figref{fig:airline2d} shows a traditional bivariate map, whereas \figref{fig:airlineVsum} shows a VSUM generated under the same constraints.
 
-To determine the optimum size (width and height) of an image, measure
-the image's size as it appears in your document (in millimeters), and
-then multiply those two values by 12. The resulting values are the
-optimum $x$ and $y$ resolution, in pixels, of the image. Image quality
-will suffer if these guidelines are not followed.
+Both maps illustrate a similar temporal trend: flight delays are shorter towards the beginning of the day, increasing on average over time.
+The traditional bivariate map affords only a coarse examination of this trend: two visible ``blocks'' of color, early in the day, and later in the afternoon.
+The VSUM, in contrast, has sufficient color resolution to show a quasi-linear trend.
+The additional color resolution for uncertainty information in VSUMs also promotes caution when considering outliers: the small number of flights on Saturdays, or departing late at night, make standard error for these delays relatively high, and the connected values more suspect than the rest of the chart.
 
-Example 1: 
-%
-An image measures 50\,mm by 75\,mm when placed in a document. This
-image should have a resolution of no less than 600 pixels by 900
-pixels in order to be reproduced faithfully.
+\viralFig
 
-Example 2: 
-%
-Capturing a screenshot of your entire $1024 \times 768$ pixel display
-monitor may be useful in illustrating a concept from your research. In
-order to be reproduced faithfully, that $1024 \times 768$ image should
-be no larger than 85 mm by 64 mm (approximately) when placed in your
-document.
+Using a dataset from a group of virologists interested in the population dynamics of viral mutation~\cite{o2012conditional}, we visualized the variability in viral genomes across 25 populations of the simian immunodeficiency virus (SIV).
+Next generation sequencing (NGS) allows biologists to collect and analyze large amounts of genomics data.
+However, these techniques often create data quality issues, as they must align large numbers of potentially ambiguous ``reads'' of relatively small numbers of base pairs.
+In the dataset here, it is important not only to identify hotspots in the viral genome (locations with high rates of mutation or variability), but to not be distracted by false positives.
+The traditional bivariate map (\figref{fig:viral2d}) affords the analysis of only a few, large hot spots.
+Extending the ``confidence fog'' metaphor used in viral genome browsers like LayerCake~\cite{correll2015layercake}, VSUMs (\figref{fig:viralVsum}) enable biologists to explore interesting patterns while not being distracted by noise.
+Smaller scale ``hot spots'' that are mapped to the same magenta color in the traditional map are visible as bright orange lines.
+Pink hot spots towards the end of the genome (around reference nucleotide $9,500$) that could be plausibly considered potential regions of interest in the 2D map, are encoded as highly uncertain and of ambiguous value in the VSUMs.
+In fact, these regions have systematically low coverage; the lack of reliable data for these regions should discourage analysts from giving their apparent pattern equal consideration from the  stronger signals earlier in the genome.
 
-%-------------------------------------------------------------------------
-\subsection{Color}
+\section{Evaluation}
 
-\textbf{Please observe:} as of 2003 publications in the proceedings of the
-Eurographics Conference can use color images throughout the paper. No
-separate color tables are necessary.
+\conditionFig
 
-However, workshop proceedings might have different agreements! 
-Figure~\ref{fig:ex3} is an example for creating color plates.
+We performed a crowdsourced experiment on Amazon's Mechanical Turk to evaluate the effectiveness of VSUMs for integrating uncertainty and value information in visualizations.
+This focus on integration meant that we limited our experimental tasks to scenarios where the participants needed to consider both value and uncertainty before making a decision.
+We gave participants two main tasks:
 
-%------------------------------------------------------------------------
-\subsection{Embedding of Hyperlinks / Typesetting of URLs}
+\begin{enumerate}
+	\item An \textbf{identification} task, where we gave participants charts with value and uncertainty information, and asked them to locate specific regions.
+E.g., ``click on the region of the chart with highest uncertainty \emph{and} lowest value.''
+	\item A \textbf{prediction} task, inspired by the game Battleship, where we gave participants charts with both \emph{forecast} and \emph{forecast uncertainty} information, and asked the participants to place tokens on the board in order to optimize expected value.
+E.g., ``place your $4$ ships on safe locations on the board.''
+\end{enumerate}
 
-Due to the use of the package \texttt{hyperref} the original behavior
-of the command $\backslash$\texttt{url} from the package \texttt{url}
-is not available. To circumvent this problem we either recommend to
-use the command $\backslash$\texttt{httpAddr} from the 
-included package \texttt{egweblnk} (see below) or to replace the
-command $\backslash$\texttt{url} by the command $\backslash$\texttt{webLink} 
--- e.g. in cases where $\backslash$\texttt{url} has been used
-widely in BibTeX-References. In the latter case we suggest to run
-BibTeX as usual and then replace all occurences of $\backslash$\texttt{url}  by
-$\backslash$\texttt{webLink}
+We discuss the two experimental tasks in more detail in the following sections. Experimental materials, including data tables, are available at \url{URL-REMOVED-FOR-REVIEW}.
 
-\noindent
-The provided commands for hyperlinks, in a nutshell, are:
+For each task, participants saw one of three types of heatmaps.
+Either two \textbf{juxtaposed} maps side-by-side, with separate encodings for value and uncertainty, a traditional \textbf{2D bivariate} map, where value and uncertainty are orthogonally encoded, or a \textbf{VSUM}.
+\figref{fig:conditions} shows examples of these factor levels.
+Other factors included the size of the map (either a 4x4 or 8x8 grid), and (for the prediction task), the question framing (attack or defense).
+Secondary factors were evenly blocked across stimuli, resulting in a 3x2 within subjects design for the identification task, and a 3x2x2 within subjects design for the prediction task.
+Participants reported fatigue effects in piloting, and so we limited the number of replicates in this study: participants saw 18 stimuli (6 for each map type) for the identification task, and 12 (4 for each map type) for the prediction task.
 
-\begin{description} \itemsep 1ex
-\item [\webLinkFont $\backslash$httpAddr \{URL without leading 'http:'\}]
-      \mbox{}\\
-      e.g. \  \httpAddr{//diglib.eg.org/EG/DL/WS}
+In all cases, we generated color maps based on the requirement that no two colors in the resulting color scheme were closer than 18 units of Euclidean distance in CIELAB space.
+This threshold results in a $9$-color bivariate map and $15$-color VSUM for all color scales used in this study.
+Given the size of the marks used in our study, this threshold distance is sufficient for a predicted discriminability rate of over 95\%~\cite{stone2014engineering}.
+Other choices of threshold, and differing univariate scales, would result in different maps (see Fig.~\ref{fig:performance}).
 
-\item[\webLinkFont $\backslash$ftpAddr \{URL without leading 'ftp:'\}]
-      \mbox{}\\
-      e.g. \  \ftpAddr{//www.eg.org/EG/DL/ftpupload}
+\subsection{Identification Task}
 
-\item[\webLinkFont $\backslash$URL \{url\}]
-      \mbox{}\\
-      e.g. \  \URL{http://www.eg.org/EG/DL/WS}
+For the identification task, we gave participants bivariate representations of uncertainty, and asked them to identify a location meeting a given criteria, e.g., ``Click on the location with the greatest uncertainty.'' There were three potential criteria, mapping to the three vertices of a VSUM: 1) high value, low uncertainty, 2) low value, low uncertainty, and 3) high uncertainty.
+Each map contained four locations, one of each combination of maximum and minimum value and uncertainty.
+The correct answer was always one of these locations.
+Other locations in the grid were created pseudo-randomly, with less extreme ranges of uncertainty and probability.
+We measured performance both in terms of accuracy (did the participant select the correct point) as well as response time.
 
-\item[\webLinkFont $\backslash$MailTo \{Email addr\}]
-      \mbox{}\\
-      e.g. \  \MailTo{publishing@eg.org}
+\subsection{Prediction Task}
 
-\item[\webLinkFont $\backslash$MailToNA \{emailName\}\{@emailSiteAddress\}]
-      \mbox{}\\
-      e.g. \  \MailToNA{publishing}{@eg.org}
+\taskTwoFig
 
-\item[\webLinkFont $\backslash$webLink\{URL without hyperlink creation\}]
-      \mbox{}\\
-      e.g. \  \webLink{http://www.eg.org/some_arbitrary_long/but_useless/URL}
+For the prediction task, we gave participants the rules of a game like Battleship.
+Greis \ea~\cite{greis2016decision} employ these game-like experimental tasks to assess how different visual designs communicate uncertainty information, which can be abstract or complex, to the general audience.
+In our task, the participant and a (fictional) adversary have to place tokens representing ships on a spatial grid.
+Without direct access to the location of the opponent's ships, the player then has to place tokens representing missile strikes on their opponent's grid.
+Players have to place all their tokens before continuing.
+If a player fires a missile at a square containing a ship, then that ship is sunk.
+The objective is to maximize the number of enemy ships hit, while minimizing the number of your own ships that are hit.
+In our task, participants were given a chart encoding predictions of the location of enemy ships (in the attacking case) or potential missile strikes (in the defending case).
+The \emph{value} component of the prediction was the likelihood of an item being present in a particular region.
+The \emph{uncertainty} component was the confidence in this prediction.
 
-\end{description}
+We included both a \textbf{gain} framing (the attacker's task: given predictions of where your opponent's ships are, where should you launch your missiles?), and a \textbf{loss} framing (the defender's task: given predictions of where your opponent's missiles will fall, where should you place your ships?) to account for systematic differences in human judgments under uncertainty.
+Tversky \& Kahneman~\cite{tversky1985framing} illustrate that framings in terms of gains or losses produce reliably different outcomes.
+In general, the function of the value of gains is concave, but is convex for losses~\cite{kahneman1979prospect}.
+That is, avoiding a large loss is often more valuable than striving for a similarly large gain.
+To avoid the potential for (incorrect) transfer learning, the attacking and defending tasks used different value color maps (Plasma and Cool) than the identification task, and from the opposite framing in this task.
+\figref{fig:taskTwoConditions} shows these differing color maps and framings.
 
-%------------------------------------------------------------------------
-\subsection{PDF Generation}
+Each map contained a minimum number of extreme locations with each combination of very high and very low value and uncertainty.
+High expected value, high uncertainty locations were ``risky.'' As a balance to the risky locations, there were also an equal number of ``safer'' locations, with slightly less expected value, but also less uncertainty.
+The remaining locations of the map varied pseudo-randomly in the middle ranges.
 
-Your final paper should be delivered as a PDF document with all typefaces
-embedded. \LaTeX{} users should use \texttt{dvips} and \texttt{ps2pdf} to
-create this PDF document. Adobe Acrobat Distiller may be used in place of
-\texttt{ps2pdf}.
+The ideal strategy from a value-maximizing standpoint would be to place tokens on areas with the highest expected value, ignoring the uncertainty information.
+However, as with roulette and other similar games of chance, the variability in expected value is relevant when considering where to place bets~\cite{mlodinow2009drunkard}.
+Over the short term, a player may seek out high expected value but risky locations, or, if risk averse, accept lower expected value but lower variability locations.
+We therefore measured the distribution of both \emph{value} and \emph{uncertainty} of the tokens placed by the participants.
 
-Adobe PDFWriter is \emph{not} acceptable for use. Documents created with
-PDFWriter will be returned to the author for revision. \texttt{pdftex} and
-\texttt{pdflatex} (and its variants) can be used only if the author can
-make certain that all typefaces are embedded and images are not downsampled
-or subsampled during the PDF creation process.
+%Show:
+% 1) Juxtaposition makes people ignore/underweight uncertainty (even when they don't screw up the search task)
+% 2) VSUMs make people, surprise surprise, bad at distinguishing highly uncertain values
+% 3) But, they make uncertainty better-integrated
+%    i) lower weights on uncertain values in decision-making
+%    ii) a priori better discrimination
 
-Users with no access to these PDF creation tools should make available a
-PostScript file and we will make a PDF document from it.
+%conditions:
+% 1) juxtaposed d map, u map
+% 2) binned d/u map as per a priori jnds
+% 3) vsm
+% 4) continuous d/u map?
 
-The PDF file \emph{must not} be change protected.
+\subsection{Hypotheses}
 
-%------------------------------------------------------------------------
-\subsubsection*{Configuration Notes: dvips / ps2pdf / etc.}
+Across both the identification and prediction tasks, we had several hypotheses, stemming from our belief that VSUMs promote better \emph{integration} between uncertainty and value information, and also encourage \emph{caution} by highlighting the ambiguity or untrustworthiness introduced by uncertain data.
+In particular:
+\begin{enumerate}
+	\item Participants would be \textbf{faster and more accurate} when completing the identification task using a VSUM as compared to the other map types.
+Juxtaposed maps, by adding an additional search task to the lookup task, would have especially poor performance compared to VSUMs.
+	\item Participants would choose targets with \textbf{less uncertainty} in the prediction task when using a VSUM compared to the other maps types.
+This would result in a tradeoff where they would also choose targets with \textbf{less expected value} than the other map types.
+Again, juxtaposed maps, by increasing the difficulty of retrieving both variables simultaneously, would highlight this difference.
+\end{enumerate}
 
-\noindent
-\texttt{dvips} should be invoked with the \texttt{-Ppdf} and \texttt{-G0}
-flags in order to use Type 1 PostScript typefaces:
+\subsection{Results}
 
-\begin{verbatim}
-    dvips -t a4 -Ppdf -G0 -o my.ps my.dvi
-\end{verbatim}
+Rather than standard means, Cleveland \& McGill~\cite{cleveland1984graphical} use 25\% trimmed means to report effect sizes in graphical perception tasks.
+Trimmed means remove the first and last quartiles and then compute the mean from the remaining points.
+This filtering is important for graphical perception tasks (especially crowdsourced tasks) where error distributions may be long-tailed~\cite{heer2010crowdsourcing}.
+Trimmed means violate the sampling assumptions required for many null-hypothesis significance tests.
+Thus, while we report effect sizes in terms of trimmed means, we perform inferential statistical tests on the standard means.
+Asterisks in our charts denote significant differences, as the reported by a post-hoc Tukey's test of Honest Significant Difference (HSD).
 
-\noindent
-If you are using version 7.x of GhostScript, please use the following method of invoking \texttt{ps2pdf}, in
-order to embed all typefaces and ensure that images are not downsampled or subsampled in the PDF
-creation process:
+\subsubsection{Participants}
 
-\begin{verbatim}
-  ps2pdf -dMaxSubsetPct=100 \
-         -dCompatibilityLevel=1.3 \
-         -dSubsetFonts=true \
-         -dEmbedAllFonts=true \
-         -dAutoFilterColorImages=false \
-         -dAutoFilterGrayImages=false \
-         -dColorImageFilter=/FlateEncode \
-         -dGrayImageFilter=/FlateEncode \
-         -dMonoImageFilter=/FlateEncode \
-         mypaper.ps mypaper.pdf
-\end{verbatim}
+We limited our population to Turkers from the United States, with a prior task approval rate of at least 90\%.
+As the experimental tasks required multi-hue color perception, we presented participants with a set of Ishihara plates~\cite{hardy1945tests} as a pretest, and excluded participants who either misidentified the values in the plates, or who self-reported as having a color vision deficiency (CVD) in the post-test.
+Based on piloting, we paid participants \$2 dollars for participation, for a target rate of \$8/hour.
+After completion of the main tasks, we solicited demographic information.
+We recruited a total of 24 participants for this experiment: 4 female, 19 male, and 1 who declined to state ($M_{\text{age}}$= 32, $SD_{\text{age}}$ = 7.5).
+All reported having either college degrees, or some college experience.
+Average self-reported experience with interpreting graphs and charts was low (1.9 on a 5 point Likert scale, with 1 indicating ``No experience''), and no participant reported experience greater than 3 out of 5.
 
-If you are using version 8.x of GhostScript, please use this method in place of the example above:
-\begin{verbatim}
-  ps2pdf -dPDFSETTINGS=/prepress \
-         -dCompatibilityLevel=1.3 \
-         -dAutoFilterColorImages=false \
-         -dAutoFilterGrayImages=false \
-         -dColorImageFilter=/FlateEncode \
-         -dGrayImageFilter=/FlateEncode \
-         -dMonoImageFilter=/FlateEncode \
-         -dDownsampleColorImages=false \
-         -dDownsampleGrayImages=false \
-         mypaper.ps mypaper.pdf
-\end{verbatim}
+\subsubsection{Identification Task}
+\accuracyFig
+\responseTimeFig
 
-%------------------------------------------------------------------------
-\subsubsection*{Configuration Notes: pdftex / pdflatex / etc.}
+For the identification task, our primary measures of performance were accuracy and response time. Our hypothesis was that VSUMs would exhibit better performance than the other charts.
 
-\noindent
-Configuration of these tools to embed all typefaces can be accomplished by editing the \texttt{updmap.cfg} file
-to enable inclusion of the standard (or base) 14 typefaces.
+We performed a one-way ANOVA investigating the effect of chart type (juxtaposed, full bivariate, or VSUM) on accuracy.
+We did not find a significant effect ($F(2,859) = 2.4$, $p=0.09$), although VSUMs had the highest accuracy.
+\figref{fig:accuracy1} illustrates this result.
 
-Linux users can run the \texttt{updmap} script to do this:
-\begin{verbatim}
-updmap --setoption pdftexDownloadBase14 true
-\end{verbatim}
+We also performed a one-way ANOVA on the effect of chart type on response time.
+In this case, we found a significant main effect ($(F,2,859) = 5.0$, $p<0.001$).
+A Tukey's HSD found that participants were significantly slower at performing the task on juxtaposed maps, as opposed to the other map types.
+\figref{fig:rt1} illustrates this result.
 
-Windows users should edit the \texttt{updmap.cfg} files found in their TeX installation directories (one or both
-of the following may be present):
-\begin{verbatim}
-  INSTALLDIR\texmf\web2c\updmap.cfg
-  INSTALLDIR\localtexmf\miktex\config\updmap.cfg
-\end{verbatim}
+These results only partially support our hypothesis.
+There was no significant difference in terms of accuracy between different maps, partially due to the simplicity of the identification task: participants only had to find a single target in a heatmap, with extreme values in both data and uncertainty axes.
+However, our results do provide evidence that juxtaposed maps, by introducing an additional search task to the problem of assessing the value and uncertainty of a location, require significantly more time to interpret.
 
-Ensure the value for \texttt{pdftexDownloadBase14} is "true," and then follow the instructions found here:
-\httpAddr{//docs.miktex.org/manual/} to update your MikTeX installation.
+\subsubsection{Prediction Task}
 
-%------------------------------------------------------------------------
-\subsubsection*{Configuration Notes: Acrobat Distiller}
+\uncertaintyFig
 
-We recommend to download and install the version of the ``CMW'' Adobe Acrobat Distiller job options file
-appropriate for your operating system and version of Acrobat from the following URL:
+For the prediction task, where there is no dominating strategy, our primary measure was the distribution of guesses, in terms of both their average value, and their average uncertainty.
+A risky player would choose guesses with high value, regardless of the uncertainty of those points.
+A more conservative guesser might eschew high-risk, high-reward locations, resulting in a lower average value of guesses, but also lower uncertainty.
+Our hypothesis is that VSUMs, by communicating uncertainty information at the expense of value information, would promote more conservatives guesses than the other chart types.
 
-\httpAddr{//www.cadmusmediaworks.com/index2.html}\\
-in the ``(Operating System)/Applications/Distiller Settings'' folder. The ``CMW'' job options file embeds
-all typefaces and does not downsample or subsample images when creating the PDF document.
-%------------------------------------------------------------------------
-\subsection{Copyright forms}
+We performed a one-way ANOVA of chart type on average value of guesses.
+We found no significant effect ($F(2,861)=0.27$, $p=0.8$), indicating that participants were generating similar expected values from their guesses across conditions.
 
-You must include your signed Eurographics copyright release form
-when you submit your finished paper. We MUST have this form before
-your paper can be published in the proceedings.
+However, an ANOVA of chart type on average \emph{uncertainty} of guesses did find a significant main effect ($F(2,861)=61$, $p<0.001$).
+A post-hoc Tukey's HSD confirmed that participants had guesses with significantly less uncertainty with VSUMs than with both other chart types.
+Additionally, participants' guesses on juxtaposed charts were significantly riskier than with the other chart types.
+\figref{fig:uncertainty2} illustrates this result.
 
-%-------------------------------------------------------------------------
-\subsection{Conclusions}
+As with the identification task, a one-way ANOVA also indicated that chart type had a significant effect on response time ($F(2,861)=13$, $p<0.001$).
+A Tukey's HSD confirmed that, as with the previous task, response time was significantly longer with juxtaposed maps (trimmed mean response time of $23$ seconds) compared to the other chart types (both with trimmed mean response time of $18$ seconds).
 
-Please direct any questions to the production editor in charge of
-these proceedings.
+Despite the documented effect of framing in prior work, a one-way ANOVA of task frame on expected value found no significant main effect ($F(1,862)=0.01$, $p=0.9$), and the trimmed means were quite similar (expected value of $0.76$ per guess for attackers, versus $0.75$ per guess for defenders).
 
-%-------------------------------------------------------------------------
+Our results again partially support our hypotheses.
+We did not observe statistically different patterns of \emph{expected value}, but did find that VSUMs produce different patterns of \emph{acceptable uncertainty}.
+That is, VSUMs promote avoidance of uncertain values, without notably impacting the expected value.
+Juxtaposed maps, by contrast, make the retrieval of uncertainty information significantly more difficult; it is possible that participants ignore or downweight it when making predictions, resulting in riskier predictions overall.
+By making the consideration of uncertainty values integral to the design of the chart, VSUMs appear to promote more cautious decisions.
 
-%\bibliographystyle{eg-alpha}
-\bibliographystyle{eg-alpha-doi}
+\section{Discussion}
 
-\bibliography{egbibsample}
+\sizeFig
 
-%-------------------------------------------------------------------------
-\newpage
+The results from our initial evaluation of VSUMs are promising, showing that even a general audience can make use of uncertainty information in a reasonable way.
+The way that uncertainty information is presented can have a measurable impact on decision-making.
+The VSUM strategy of only ``showing'' differences when uncertainty is sufficiently low is somewhat analogous to the inferential statistics such as effects tests.
+VSUMs might therefore be used to promote more caution in judgments, and lead analysts away from spurious signals in data.
 
-\begin{figure*}[tbp]
-  \centering
-  \mbox{} \hfill
-  % the following command controls the width of the embedded PS file
-  % (relative to the width of the current column)
-  \includegraphics[width=.3\linewidth]{sampleFig}
-  % replacing the above command with the one below will explicitly set
-  % the bounding box of the PS figure to the rectangle (xl,yl),(xh,yh).
-  % It will also prevent LaTeX from reading the PS file to determine
-  % the bounding box (i.e., it will speed up the compilation process)
-  % \includegraphics[width=.3\linewidth, bb=39 696 126 756]{sampleFig}
-  \hfill
-  \includegraphics[width=.3\linewidth]{sampleFig}
-  \hfill \mbox{}
-  \caption{\label{fig:ex3}%
-           For publications with color tables (i.e., publications not offering
-           color throughout the paper) please \textbf{observe}: 
-           for the printed version -- and ONLY for the printed
-           version -- color figures have to be placed in the last page.
-           \newline
-           For the electronic version, which will be converted to PDF before
-           making it available electronically, the color images should be
-           embedded within the document. Optionally, other multimedia
-           material may be attached to the electronic version. }
-\end{figure*}
+Bivariate maps have an inherent limitation in that visual channels are often difficult to attend to separately or orthogonally.
+Bivariate maps are also complex to interpret.
+Systems which simultaneously display large amounts of value and uncertainty data to general audiences (such as Pangloss~\cite{moritz2017trust}) avoid them for precisely these reasons, relying instead on juxtaposed maps.
+Our results indicate that juxtaposition produces additional problems, forcing users that wish to integrate uncertainty and data to locate an item of interest, and then re-locate that same item in a visually distant map, which may have few visual landmarks in common.
+This additional step before information fusion is reflected in longer response times, and patterns of decision-making that are less mindful of risk.
+Given the deleterious effects of latency on data analysis~\cite{liu2014effects}, designers ought to carefully consider the tradeoffs between expressiveness of visual encoding, and usability.
 
-\end{document}
+\subsection{Limitations and Future Work}
+% When should you use these things?
+In general, VSUMs excel in scenarios where we wish to make firm guarantees about the discriminability of different marks in a chart, but must also represent, in as much detail as possible, two, potentially orthogonal variables of interest.
+VSUMs have improved data resolution over na\"ive bivariate maps by assuming that uncertain values ought to have less visual impact that highly certain values.
+VSUMs therefore act both as a kind of filtering device as well as a bivariate representation.
+This assumption holds in scenarios like the virology dataset (\figref{fig:viralVsum}), where the expectation is that highly uncertain data is unreliable or should otherwise be downweighted in analysis.
+If the analyst has a different interpretation of uncertainty, or wishes to quickly and orthogonally analyze the distributions of uncertainty and value, other strategies, such as juxtaposed maps, may be more appropriate.
+VSUMs are designed for the \emph{integration} of value and uncertainty.
+Designers should take care in considering when and how this integration is desirable.
+
+Our experiments dealt with cases where both uncertainty and value were represented by color.
+The perceptual non-separability of color channels is well-known~\cite{garner1970integrality, ware2012information}, and so the concept of a limited ``budget'' of distinguishable marks easier to quantify and illustrate.
+In principle, a VSUM can be created for any combination of visual variables.
+All that is required is a perceptual model of the interaction between these two variables.
+Where these models exist, as with the interaction between size and color~\cite{stone2014engineering}, the creation of VSUMs is straightforward.
+Where these models do not exist, or where the perceptual interaction is too complex to efficiently model, experimental work remains to be done before VSUMs can be considered a feasible design strategy.
+We are currently experimenting with the creation of VSUMs using these other channels, as in \figref{fig:size}.
+
+\subsection{Conclusion}
+
+Often, uncertainty, data quality, or confidence are considered separate information from the data itself, relegated to tooltips or visually distant supplemental charts.
+We believe, in contrast, that uncertainty information ought to be directly integrated with charts of data itself.
+This integration introduces additional complexity in the design and presentation of data.
+Value-Suppressing Uncertainty Maps represent one strategy for dealing with this complexity, by assigning marks properties in a way that supports the disambiguation of values in data where uncertainty is low, but suppresses these judgments when uncertainty is high.
+This decision of how to allocate visual variables promotes patterns of decision-making that make responsible use of uncertainty information, discouraging comparison of values in unreliable regions of the data, and promoting comparison in regions of high certainty.

### 2017-08-31 modules/src/index.js
Added "square" scale, abstract for workshop tomorrow.

@@ -1,5 +1,5 @@
 export {simpleHeatmap as heatmap, simpleArcmap as arcmap} from "./heatmap";
-export {treeScale as scale} from "./scale";
+export {treeScale as scale, squareScale as iscale} from "./scale";
 export {simpleMap as map} from "./map";
 export {continuousSquare as csquare, continuousArc as carc} from "./continuous";
 import * as legend_ from "./legend";

### 2017-08-31 modules/src/scale.js

@@ -3,6 +3,69 @@ A tree-like bivariate quantization scheme.
 As uncertainty increases, the number of quantization bins decreases.
 */
 
+
+export function squareScale(m_n) {
+  var n = m_n,
+      uscale,
+      vscale,
+      matrix = makeMatrix();
+
+  function scale(value, uncertainty) {
+    var u = uncertainty!=undefined ? uncertainty : value.u,
+        v = uncertainty!=undefined ? value : value.v,
+        i = 0;
+
+    //find the right layer of the tree, based on uncertainty
+    while(i<matrix.length-1 && u < 1 - ((i+1)/n)){
+      i++;
+    }
+
+    //find right leaf of tree, based on value
+    var vgap = (matrix[i].length>1) ? (matrix[i][1].v - matrix[i][0].v) / 2 : 0,
+        j = 0;
+
+    while(j<(matrix[i].length-1) && v > matrix[i][j].v+vgap){
+      j++;
+    }
+
+    return matrix[i][j];
+  }
+
+  function makeMatrix() {
+    var matrix = [];
+
+    for(var i = 0;i<n;i++){
+      matrix[i] = [];
+      for(var j = 1;j<(2*n);j+=2){
+        matrix[i].push({ u: 1 - ( i/(n-1)), v: (j/(2*n))});
+      }
+    }
+
+    return matrix;
+  }
+
+  scale.range = function(newRange) {
+    return [].concat.apply([],matrix);
+  }
+
+  scale.n = function(newN) {
+      if(!arguments.length) {
+        return n;
+      }
+      else{
+        n = newN;
+        matrix = makeMatrix();
+        return scale;
+      }
+  }
+
+  scale.matrix = function() {
+    return matrix;
+  }
+
+  return scale;
+}
+
 export function treeScale(branchingFactor,treeLayers) {
   var branch = branchingFactor ? branchingFactor : 2,
       layers = treeLayers ? treeLayers : 2,

### 2017-09-12 paper/figures.tex
Check in for pilot fixes

@@
 	\includegraphics[width=0.9\columnwidth]{example.pdf}
-	\caption{A standard bivariate map (left) and a VSUM (right), used to encode an identical 5x5 grid of random data.
+	\caption{A standard bivariate map (left) and a VSUP (right), used to encode an identical 5x5 grid of random data.
 Both use the same visual channels to encode value (position along the Viridis~\protect\cite{viridis} color map) and uncertainty (lightness and saturation), and have an equal standard of perceptual discriminability (at least 18 units of distance in CIELAB color space between colors).
 However, highly uncertain values result in colors that are very close together in the bivariate case, meaning the full bivariate map is only 9 bins under these constraints\,---\,a 4x4 map would result in colors that are perceptually too close together.
-By contrast, the VSUM intentionally reduces bins when uncertainty is high, eventually aliasing all highly uncertain values to the same color.
+By contrast, the VSUP intentionally reduces bins when uncertainty is high, eventually aliasing all highly uncertain values to the same color.
 This decision affords more distinct colors in other regions of the map, and so an increase in overall bins to 15.
@@
 		\includegraphics[width=0.9\columnwidth]{size.pdf}
-		\caption{A VSUM where value is mapped to color, and uncertainty is mapped to glyph area.
+		\caption{A VSUP where value is mapped to color, and uncertainty is mapped to glyph area.
 As glyphs become smaller, it becomes more difficult to distinguish their colors~\protect\cite{stone2014engineering}.
-VSUMs, by reducing the number of color categories as uncertainty increases, account for this ambiguity.
-On the right, the VSUM has been used to encode a 5x5 grid of random data.}
+VSUPs, by reducing the number of color categories as uncertainty increases, account for this ambiguity.
+On the right, the VSUP has been used to encode a 5x5 grid of random data.}
 		\label{fig:size}
@@
 			\includegraphics[width=0.45\textwidth]{conditions3.pdf}
-			\caption{Value-Suppressing Uncertainty Map}
+			\caption{Value-Suppressing Uncertainty Palette}
 		\end{subfigure}
@@
 Juxtaposed maps, by keeping these channels separate, afford a larger range of colors, but require a visual search task to recover both value and uncertainty in a region.
-VSUMs have many of the advantages of bivariate maps, but assign more color categories to certain data, at the expense of ambiguity when uncertainty is high.}
+VSUPs have many of the advantages of bivariate maps, but assign more color categories to certain data, at the expense of ambiguity when uncertainty is high.}
 		\label{fig:conditions}
@@
 		\caption{
-			The number of discrete color categories in traditional bivariate color maps versus VSUMs, assuming value is encoded by the Viridis color map~\protect\cite{viridis}, and uncertainty is encoded by saturation/value (whiter values are more uncertain).
-Since these whiter uncertain colors are closer together in CIELAB space, and VSUMs intentionally require fewer color bins in those regions, VSUMs always have more color categories than the corresponding 2D bivariate map.
+			The number of discrete color categories in traditional bivariate color maps versus VSUPs, assuming value is encoded by the Viridis color map~\protect\cite{viridis}, and uncertainty is encoded by saturation/value (whiter values are more uncertain).
+Since these whiter uncertain colors are closer together in CIELAB space, and VSUPs intentionally require fewer color bins in those regions, VSUPs always have more color categories than the corresponding 2D bivariate map.
 The red lines denote:
@@
 		\includegraphics[width=0.95\columnwidth]{flow.pdf}
-		\caption{The process for creating a VSUM.
-VSUMs guarantee that no two marks in the resulting bivariate mapping are perceptually closer than some pre-specified threshold.
+		\caption{The process for creating a VSUP.
+VSUPs guarantee that no two marks in the resulting bivariate mapping are perceptually closer than some pre-specified threshold.
 This guarantee removes some of the issues of non-separability and ambiguity that are otherwise a concern when designing bivariate maps.
 \\
-		In this example two colors from the map for $n=9$ are too close in CIELAB space. Thus, we create a VSUM with $n=8$ color bins at the lowest level of uncertainty.}
+		In this example two colors from the map for $n=9$ are too close in CIELAB space. Thus, we create a VSUP with $n=8$ color bins at the lowest level of uncertainty.}
 		\label{fig:flow}
@@
 	\vspace{-15px}
-	\caption{Flight delay data encoded with a VSUM.}
+	\caption{Flight delay data encoded with a VSUP.}
 	\label{fig:airlineVsum}
@@
 	\vspace{-15px}
-	\caption{Variability for 25 different viral populations of SIV, encoded with a VSUM.
+	\caption{Variability for 25 different viral populations of SIV, encoded with a VSUP.
 The magnitude of mutation rate, as well as the generally poorer quality of data towards the end of the sequence, is more readily visible than with the traditional bivariate color map.
@@
 		\caption{The effect of chart type on accuracy for the identification task.
-VSUMS are similarly accurate for simple search tasks as more traditional bivariate visualizations, despite containing additional colors categories, and having a more complex relationship between uncertainty, value, and color than traditional charts.
+VSUPS are similarly accurate for simple search tasks as more traditional bivariate visualizations, despite containing additional colors categories, and having a more complex relationship between uncertainty, value, and color than traditional charts.
 The confidence intervals are bootstrapped 95\% CIs of trimmed means.}
@@
 		\caption{The effect of chart type on the average uncertainty in predictions for the prediction task.
-By reducing the number of color categories as uncertainty increases, VSUMs encourage more caution in predictions, making people less likely to consider data with strong, but spurious patterns.
+By reducing the number of color categories as uncertainty increases, VSUPs encourage more caution in predictions, making people less likely to consider data with strong, but spurious patterns.
 The confidence intervals are bootstrapped 95\% CIs of trimmed means.}

### 2017-09-12 paper/main.tex
Check in for pilot fixes

@@
 \section{Introduction}
-certainty is an inescapable component of collecting, presenting, and using data.
+Uncertainty is an inescapable component of collecting, presenting, and using data.
 A common goal in the communication of uncertainty is promoting \emph{uncertainty-aware decisions}: the audience should be aware of the risks and rewards of certain decisions, modulate their confidence in their conclusions, and perhaps refrain from making a decision at all if there is too much uncertainty.
@@
 
-In this paper, we contribute \textbf{Value-Suppressing Uncertainty Maps} (VSUMs) for integrating data and uncertainty information in visualizations.
-VSUMs intentionally alias together data values with high uncertainty, affording greater discriminability as uncertainty decreases.
-Compared to a traditional two-dimensional bivariate map, with differing outputs for each combination of value and uncertainty, VSUMs can be thought of as arcs: as uncertainty increases, values are mapped to smaller and smaller sets of outputs, culminating in a singularity where all inputs are mapped to an identical, highly uncertain mark regardless of their data value.
-\figref{fig:example} shows examples of both a more traditional bivariate map and a VSUM.
+In this paper, we contribute \textbf{Value-Suppressing Uncertainty Palettes} (VSUPs) for integrating data and uncertainty information in visualizations.
+VSUP intentionally alias together data values with high uncertainty, affording greater discriminability as uncertainty decreases.
+If traditional bivariate maps are a two-dimensional square, with differing outputs for each combination of value and uncertainty, VSUP can be thought of as arcs: as uncertainty increases, values are mapped to smaller and smaller sets of outputs, culminating in a singularity where all inputs are mapped to an identical, highly uncertain mark regardless of their data value.
+\figref{fig:example} shows examples of both a more traditional bivariate map and a VSUP.
 
-We describe the motivations behind the use of VSUMs, examples of their utility for decision-making under uncertainty, and assess VSUMs in a crowdsourced experiment.
-Our experimental results indicate that VSUMs create close integration between uncertainty and data, promoting rapid but cautious decision-making.
+We describe the motivations behind the use of VSUP, examples of their utility for decision-making under uncertainty, and assess VSUP in a crowdsourced experiment.
+Our experimental results indicate that VSUP create close integration between uncertainty and data, promoting rapid but cautious decision-making.
 
@@
 
+To our knowledge, there are no established standards for visualizing uncertainty in heatmaps.
 Reviewing the state of the art in uncertainty visualization in the field, Greithe \ea~\cite{griethe2006visualization} and Brodlie \ea~\cite{brodlie2012review} present lists of potential techniques for conveying uncertainty in concert with data.
@@
 
-There are a number of techniques for communicating values and uncertainty in one-dimensional statistical graphics~\cite{olston2002visualizing, wickham201140, neyman1937outline}.
+%There are a number of techniques for communicating values and uncertainty in one-dimensional statistical graphics~\cite{olston2002visualizing, wickham201140, neyman1937outline}.
 Techniques such as error bars, confidence bands, and violin plots~\cite{hintze1998violin} use a separate position channel to encode uncertainty.
 The techniques in this paper also apply to 2D visualizations such as heatmaps.
-To our knowledge, there are no established standards of visualizing uncertainty in heatmaps.
 
-\strategyFig
+%\strategyFig
 
-To support our goal of making uncertainty information well-integrated with data, we focus on the techniques of \emph{juxtaposition} (where uncertainty is presented alongside the data), \emph{superposition} (where additional glyphs are overlaid on top of the data), and \emph{explicit encoding} (where an additional visual channel is used to encode the uncertainty).
+%To support our goal of making uncertainty information well-integrated with data, we focus on the techniques of \emph{juxtaposition} (where uncertainty is presented alongside the data), \emph{superposition} (where additional glyphs are overlaid on top of the data), and \emph{explicit encoding} (where an additional visual channel is used to encode the uncertainty).
 Gleicher \ea~\cite{gleicher2011visual} explore these three strategies from the perspective of affording \emph{comparisons}; here, we examine them from the perspective of incorporating measures of uncertainty.
@@
 
-\subsection{Strategies for Encoding Uncertainty}
+%\subsection{Strategies for Encoding Uncertainty}
 
-\emph{Juxtaposition} creates two explicit visualizations: an ``uncertainty map'' and a ``data map.'' MacEachren~\cite{maceachren1992visualizing} portrays uncertainty visualization as the design problem of how to \emph{unify} these two maps, and so this explicit separation can be counterproductive to the goal of information fusion.
+%\emph{Juxtaposition} creates two explicit visualizations: an ``uncertainty map'' and a ``data map.'' MacEachren~\cite{maceachren1992visualizing} portrays uncertainty visualization as the design problem of how to \emph{unify} these two maps, and so this explicit separation can be counterproductive to the goal of information fusion.
 Juxtaposition also creates an additional search task for viewers, who must register hotspots in one map with corresponding features of the other map.
@@
 
-\emph{Superposition}, or the overlay of glyphs representing uncertainty on top of a data map, is another common strategy.
+%\emph{Superposition}, or the overlay of glyphs representing uncertainty on top of a data map, is another common strategy.
 Error bars are a common example of this strategy: a mark indicating measurement error is superimposed atop a mark representing a point value.
@@
 
-\emph{Explicit encoding} directly represents uncertainty with a visual channel, simultaneously with data.
-VSUMs employ this strategy.
+%\emph{Explicit encoding} directly represents uncertainty with a visual channel, simultaneously with data.
+VSUP employ this strategy.
 In addition to overcoming the drawbacks of the other techniques mentioned in this section, we believe that explicit encoding represents the most general form of the solution of how to fuse uncertainty information with data.
@@
 
-Foundational work in psychophysics and graphical perception make use of the concept of a ``just noticeable difference'' (JND)\cite{cleveland1984graphical}.
+%TODO stuff that needs to get cited here:
+
+% uncertainty vis surveys: kinkeldey2017evaluating,maceachren1992visualizing,brodlie2012review
+% 2d uvis: griethe2006visualization, maceachren1992visualizing,maceachren1998visualizing
+% 1d uvis: olston2002visualizing, wickham201140, neyman1937outline
+% ignoreability: moritz2017trust
+
+%Foundational work in psychophysics and graphical perception make use of the concept of a ``just noticeable difference'' (JND)\cite{cleveland1984graphical}.
 A 50\% JND is a difference between two stimuli that is detectable 50\% of the time.
@@
 \subsection{Bivariate Maps}
+
+%TODO cites, fill out section
+
+% examples: roth2010value,ware2009quantitative
+
+\subsection{Quantizing Color Scales}
+
+%TODO cites, fill out section
+% binning stuff: [data] battersby2016shapes, [legends] padilla2017evaluating,muller1979perception,dobson1973choropleth,dunn1989dynamic
 
@@
 
-Dunn~\cite{dunn1989dynamic} attempts to circumvent these issues by dynamically allocating bins to bivariate maps.
+%Dunn~\cite{dunn1989dynamic} attempts to circumvent these issues by dynamically allocating bins to bivariate maps.
 As the number of distinct, separable colors is limited, the dynamic approach allocates bins to regions of the highest utility.
 For instance, placing the dividing line between bins in areas aligned with a trend line allows the color to encode information about the sign of the residual with respect to each variable.
-This insight, that the bins in a bivariate map are a limited resource and so should be ``spent'' wisely, drives the creation of VSUMs, which allocate more bins in a bivariate map to regions with less uncertainty.
+This insight, that the bins in a bivariate map are a limited resource and so should be ``spent'' wisely, drives the creation of VSUPs, which allocate more bins in a bivariate map to regions with less uncertainty.
 
-\section{Value-Suppressing Uncertainty Maps}
+\section{Value-Suppressing Uncertainty Palettes}
 
-Value-Suppressing Uncertainty Maps (VSUMs) are a technique for creating bivariate maps of data \emph{value} and \emph{uncertainty}.
-These maps circumvent the issues in the creation of traditional bivariate maps by making the following assumptions:
-
-\begin{enumerate}
-	\item \textbf{Designers have a limited budget of discriminable marks.} That is, viewers cannot easily or reliably discriminate between an infinite number of categories.
-Rather than accept this perceptual ambiguity, designers ought to \emph{alias} together marks into a finite, but sufficiently discriminable, set.
-	\item \textbf{Uncertain values should have less weight than certain values.} Weight, in this case, might mean literal visual weight, or impact on the decision-making process.
-\end{enumerate}
-
-%MC this paragraph is a little iffy, and sort of "the lady doth protest too much" territory.
-we might be able to get away with "people bin colors all the time.
-and in fact, you really should be binning your colors anyway."
-The first assumption seems to contradict Mackinlay's \emph{expressiveness principle}~\cite{mackinlay1986automating}, where the visual encoding of the data should reflect the composition of the data.
-For instance, if my data are a collection of measurements of some continuous variable, then it would make sense to encode these data using a continuous \emph{visual} variable, which would then mean an arbitrary number of potential marks.
-For instance, a bar can have an arbitrary, non-integer height.
-However, for many visualizations, the perceptual ambiguity caused by having marks that cannot be reliably disambiguated dominates the gains from having a visual encoding that maps readily to the backing data.
-Padilla \ea~\cite{padilla2017evaluating} present one such example, where binned heatmaps result in higher task accuracy than continuous heatmaps.
+Value-Suppressing Uncertainty Palettes (VSUPs) are a technique for creating bivariate maps of data \emph{value} and \emph{uncertainty}.
 
@@
 If both of these assumptions hold, then it follows that the designer of a bivariate map should allocate more mark types to certain values, and fewer mark types to uncertain values.
-VSUMs codify this decision by \emph{reducing the number of mark categories for representing value as uncertainty increases.} This means that data encoded using a VSUM will make visible only the largest of differences in uncertain data, but highlight comparatively small differences in value when uncertainty is low.
-VSUMs therefore act as both a filtering mechanism (in that values with too much uncertainty are all mapped to the same glyph), as well as an implicit test of effect size (in that smaller and smaller changes in value are visible as uncertainty decreases).
+VSUPs codify this decision by \emph{reducing the number of mark categories for representing value as uncertainty increases.} This means that data encoded using a VSUP will make visible only the largest of differences in uncertain data, but highlight comparatively small differences in value when uncertainty is low.
+VSUPs therefore act as both a filtering mechanism (in that values with too much uncertainty are all mapped to the same glyph), as well as an implicit test of effect size (in that smaller and smaller changes in value are visible as uncertainty decreases).
 This strategy of dampening the low quality or high uncertainty values in maps in order to focus on more informative regions has measurable benefits, including the removal of statistically spurious visual patterns, and the highlighting of regions of interest~\cite{correll2017surprise}.
@@
 
-The general VSUM construction algorithm is as follows:
+The general VSUP construction algorithm is as follows:
 
@@
 	\item \textbf{Select a visual variable to encode value}. In this paper, this variable is position along some established color scale.
-	\item \textbf{Select a visual variable to encode uncertainty}.
-We prefer variables that have a somewhat high semiotic connection with uncertainty (such as lightness and size), but still have a reasonable number of perceptually discriminable layers (as opposed to other, more intuitive channels such as blur or sketchiness).
-In most cases in this paper, we encode uncertainty as interpolation to white or as glyph size.
-	\item \textbf{Select a discriminability threshold}.
-Choose some minimum distance, in a perceptual space, between marks.
-For many visual variables (such as length or hue), the perceptual discriminability between marks has been assessed in prior literature, and so the only step is to choose a value in terms of the just noticeable distance (JND).
-For less common visual variables, a perceptual model may be built empirically (as in Demiralp \ea~\cite{demiralp2014learning}).
-	\item \textbf{Choose how quickly categories \emph{degrade}}.
-That is, if there are $n$ bins at some uncertainty level $m$, how many bins should there be when uncertainty increases to level $m+1$?
-In \figref{fig:example}, this function is $\left \lfloor n / 2 \right \rfloor$, meaning that the number of value bins decreases by half for each level of uncertainty bins.
-If there is too much degradation, then there are fewer levels of uncertainty.
-If there is no degradation, the resulting map is just a traditional bivariate map.
-	\item \textbf{Choose the largest number of initial categories $n$ such that the discriminability threshold is met.} For each pair of marks in the map, perform a pairwise comparison of discriminability.
-If no marks are perceptually ambiguous by the standards of your threshold, the map is complete.
-Otherwise, decrement $n$.
 \end{enumerate}
@@
 \figref{fig:flow} shows this process in detail, and how it can be used to generate a bivariate map for presenting uncertainty.
-JavaScript code for generating VSUMs given arbitrary D3~\cite{bostock2011d3} quantitative scale functions is available at \url{URL-REMOVED-FOR-REVIEW}.
+JavaScript code for generating VSUPs given arbitrary D3~\cite{bostock2011d3} quantitative scale functions is available at \url{URL-REMOVED-FOR-REVIEW}.
 
-Correll \ea~\cite{correll2015layercake,correll2011visualizing} use a precursor of VSUMs in their LayerCake genomics visualization tool, where marks representing genomic data with increasing uncertainty are mapped to a smaller and smaller set of increasingly grey colors, creating the effect of uncertain values retreating into ``confidence fog'' while highly certain values remain prominent.
+Correll \ea~\cite{correll2015layercake,correll2011visualizing} use a precursor of VSUPs in their LayerCake genomics visualization tool, where marks representing genomic data with increasing uncertainty are mapped to a smaller and smaller set of increasingly grey colors, creating the effect of uncertain values retreating into ``confidence fog'' while highly certain values remain prominent.
 Other bivariate visualizations implicitly alias together uncertain values.
@@
 In both cases, the property of aliasing is \emph{ad hoc}, and places no guarantees on the discriminability of colors.
-The binning and degradation approach of VSUMs makes the choice to alias values explicit to both the designer and the viewer, and results in a bivariate mapping with known perceptual properties.
+The binning and degradation approach of VSUPs makes the choice to alias values explicit to both the designer and the viewer, and results in a bivariate mapping with known perceptual properties.
 
@@
 
-There are multiple choices that designers must make before creating a VSUM.
+There are multiple choices that designers must make before creating a VSUP.
 In particular, they must choose 1) which visual channels map to value and uncertainty, 2) a threshold of perceptual distinguishability (and, implicitly, a corresponding metric), and 3) a degradation function to determine how quickly uncertain values are aliased.
@@
 
-How quickly the number of bins decreases is another parameter of VSUMs.
+How quickly the number of bins decreases is another parameter of VSUPs.
 If this degradation is too fast, then we may only have a few discriminable levels of uncertainty.
 If this degradation is too slow, then we can be overwhelmed with colors.
-For instance, if we only decrement by one bin per level of uncertainty, there would be $1,485$ different colors in a VSUM with a tolerance threshold of $1$ unit in CIELAB space, far too much complexity for a discrete color mapping.
+For instance, if we only decrement by one bin per level of uncertainty, there would be $1,485$ different colors in a VSUP with a tolerance threshold of $1$ unit in CIELAB space, far too much complexity for a discrete color mapping.
 In the examples in this paper, we halve the number of bins at each increasing level of uncertainty.
 When the number of initial bins at the highly certain region is a power of $2$ (as in Fig.
-\ref{fig:example}), this imparts a tree structure to the resulting VSUM.
+\ref{fig:example}), this imparts a tree structure to the resulting VSUP.
 As uncertainty increases, bins in the previous level map to one and only one ``parent'' in the next uncertainty level, resulting in fewer crossings of color boundaries among rows.
@@
 
-We present a set of examples showing how VSUMs can be applied to realworld datasets.
-The decision of VSUMs to allocate color bins asymmetrically amongst value and uncertainty regions affords greater discriminability in regions of the chart that warrant closer scrutiny, while discouraging exploration of regions with noisy or unreliable signals.
+We present a set of examples showing how VSUPs can be applied to realworld datasets.
+The decision of VSUPs to allocate color bins asymmetrically amongst value and uncertainty regions affords greater discriminability in regions of the chart that warrant closer scrutiny, while discouraging exploration of regions with noisy or unreliable signals.
 
@@
 Starting from an minimum discriminability threshold of $18$ units in CIELAB space, we created two alternate heatmaps of this delay information.
-\figref{fig:airline2d} shows a traditional bivariate map, whereas \figref{fig:airlineVsum} shows a VSUM generated under the same constraints.
+\figref{fig:airline2d} shows a traditional bivariate map, whereas \figref{fig:airlineVsum} shows a VSUP generated under the same constraints.
 
@@
 The traditional bivariate map affords only a coarse examination of this trend: two visible ``blocks'' of color, early in the day, and later in the afternoon.
-The VSUM, in contrast, has sufficient color resolution to show a quasi-linear trend.
-The additional color resolution for uncertainty information in VSUMs also promotes caution when considering outliers: the small number of flights on Saturdays, or departing late at night, make standard error for these delays relatively high, and the connected values more suspect than the rest of the chart.
+The VSUP, in contrast, has sufficient color resolution to show a quasi-linear trend.
+The additional color resolution for uncertainty information in VSUPs also promotes caution when considering outliers: the small number of flights on Saturdays, or departing late at night, make standard error for these delays relatively high, and the connected values more suspect than the rest of the chart.
 
@@
 The traditional bivariate map (\figref{fig:viral2d}) affords the analysis of only a few, large hot spots.
-Extending the ``confidence fog'' metaphor used in viral genome browsers like LayerCake~\cite{correll2015layercake}, VSUMs (\figref{fig:viralVsum}) enable biologists to explore interesting patterns while not being distracted by noise.
+Extending the ``confidence fog'' metaphor used in viral genome browsers like LayerCake~\cite{correll2015layercake}, VSUPs (\figref{fig:viralVsum}) enable biologists to explore interesting patterns while not being distracted by noise.
 Smaller scale ``hot spots'' that are mapped to the same magenta color in the traditional map are visible as bright orange lines.
-Pink hot spots towards the end of the genome (around reference nucleotide $9,500$) that could be plausibly considered potential regions of interest in the 2D map, are encoded as highly uncertain and of ambiguous value in the VSUMs.
+Pink hot spots towards the end of the genome (around reference nucleotide $9,500$) that could be plausibly considered potential regions of interest in the 2D map, are encoded as highly uncertain and of ambiguous value in the VSUPs.
 In fact, these regions have systematically low coverage; the lack of reliable data for these regions should discourage analysts from giving their apparent pattern equal consideration from the  stronger signals earlier in the genome.
@@
 
-We performed a crowdsourced experiment on Amazon's Mechanical Turk to evaluate the effectiveness of VSUMs for integrating uncertainty and value information in visualizations.
+We performed a crowdsourced experiment on Amazon's Mechanical Turk to evaluate the effectiveness of VSUPs for integrating uncertainty and value information in visualizations.
 This focus on integration meant that we limited our experimental tasks to scenarios where the participants needed to consider both value and uncertainty before making a decision.
@@
 For each task, participants saw one of three types of heatmaps.
-Either two \textbf{juxtaposed} maps side-by-side, with separate encodings for value and uncertainty, a traditional \textbf{2D bivariate} map, where value and uncertainty are orthogonally encoded, or a \textbf{VSUM}.
+Either two \textbf{juxtaposed} maps side-by-side, with separate encodings for value and uncertainty, a traditional \textbf{2D bivariate} map, where value and uncertainty are orthogonally encoded, or a \textbf{VSUP}.
 \figref{fig:conditions} shows examples of these factor levels.
@@
 In all cases, we generated color maps based on the requirement that no two colors in the resulting color scheme were closer than 18 units of Euclidean distance in CIELAB space.
-This threshold results in a $9$-color bivariate map and $15$-color VSUM for all color scales used in this study.
+This threshold results in a $9$-color bivariate map and $15$-color VSUP for all color scales used in this study.
 Given the size of the marks used in our study, this threshold distance is sufficient for a predicted discriminability rate of over 95\%~\cite{stone2014engineering}.
@@
 
-For the identification task, we gave participants bivariate representations of uncertainty, and asked them to identify a location meeting a given criteria, e.g., ``Click on the location with the greatest uncertainty.'' There were three potential criteria, mapping to the three vertices of a VSUM: 1) high value, low uncertainty, 2) low value, low uncertainty, and 3) high uncertainty.
+For the identification task, we gave participants bivariate representations of uncertainty, and asked them to identify a location meeting a given criteria, e.g., ``Click on the location with the greatest uncertainty.'' There were three potential criteria, mapping to the three vertices of a VSUP: 1) high value, low uncertainty, 2) low value, low uncertainty, and 3) high uncertainty.
 Each map contained four locations, one of each combination of maximum and minimum value and uncertainty.
@@
 % 1) Juxtaposition makes people ignore/underweight uncertainty (even when they don't screw up the search task)
-% 2) VSUMs make people, surprise surprise, bad at distinguishing highly uncertain values
+% 2) VSUPs make people, surprise surprise, bad at distinguishing highly uncertain values
 % 3) But, they make uncertainty better-integrated
@@
 
-Across both the identification and prediction tasks, we had several hypotheses, stemming from our belief that VSUMs promote better \emph{integration} between uncertainty and value information, and also encourage \emph{caution} by highlighting the ambiguity or untrustworthiness introduced by uncertain data.
+Across both the identification and prediction tasks, we had several hypotheses, stemming from our belief that VSUPs promote better \emph{integration} between uncertainty and value information, and also encourage \emph{caution} by highlighting the ambiguity or untrustworthiness introduced by uncertain data.
 In particular:
 \begin{enumerate}
-	\item Participants would be \textbf{faster and more accurate} when completing the identification task using a VSUM as compared to the other map types.
-Juxtaposed maps, by adding an additional search task to the lookup task, would have especially poor performance compared to VSUMs.
-	\item Participants would choose targets with \textbf{less uncertainty} in the prediction task when using a VSUM compared to the other maps types.
+	\item Participants would be \textbf{faster and more accurate} when completing the identification task using a VSUP as compared to the other map types.
+Juxtaposed maps, by adding an additional search task to the lookup task, would have especially poor performance compared to VSUPs.
+	\item Participants would choose targets with \textbf{less uncertainty} in the prediction task when using a VSUP compared to the other maps types.
 This would result in a tradeoff where they would also choose targets with \textbf{less expected value} than the other map types.
@@
 
-For the identification task, our primary measures of performance were accuracy and response time. Our hypothesis was that VSUMs would exhibit better performance than the other charts.
+For the identification task, our primary measures of performance were accuracy and response time. Our hypothesis was that VSUPs would exhibit better performance than the other charts.
 
-We performed a one-way ANOVA investigating the effect of chart type (juxtaposed, full bivariate, or VSUM) on accuracy.
-We did not find a significant effect ($F(2,859) = 2.4$, $p=0.09$), although VSUMs had the highest accuracy.
+We performed a one-way ANOVA investigating the effect of chart type (juxtaposed, full bivariate, or VSUP) on accuracy.
+We did not find a significant effect ($F(2,859) = 2.4$, $p=0.09$), although VSUPs had the highest accuracy.
 \figref{fig:accuracy1} illustrates this result.
@@
 A more conservative guesser might eschew high-risk, high-reward locations, resulting in a lower average value of guesses, but also lower uncertainty.
-Our hypothesis is that VSUMs, by communicating uncertainty information at the expense of value information, would promote more conservatives guesses than the other chart types.
+Our hypothesis is that VSUPs, by communicating uncertainty information at the expense of value information, would promote more conservatives guesses than the other chart types.
 
@@
 However, an ANOVA of chart type on average \emph{uncertainty} of guesses did find a significant main effect ($F(2,861)=61$, $p<0.001$).
-A post-hoc Tukey's HSD confirmed that participants had guesses with significantly less uncertainty with VSUMs than with both other chart types.
+A post-hoc Tukey's HSD confirmed that participants had guesses with significantly less uncertainty with VSUPs than with both other chart types.
 Additionally, participants' guesses on juxtaposed charts were significantly riskier than with the other chart types.
@@
 Our results again partially support our hypotheses.
-We did not observe statistically different patterns of \emph{expected value}, but did find that VSUMs produce different patterns of \emph{acceptable uncertainty}.
-That is, VSUMs promote avoidance of uncertain values, without notably impacting the expected value.
+We did not observe statistically different patterns of \emph{expected value}, but did find that VSUPs produce different patterns of \emph{acceptable uncertainty}.
+That is, VSUPs promote avoidance of uncertain values, without notably impacting the expected value.
 Juxtaposed maps, by contrast, make the retrieval of uncertainty information significantly more difficult; it is possible that participants ignore or downweight it when making predictions, resulting in riskier predictions overall.
-By making the consideration of uncertainty values integral to the design of the chart, VSUMs appear to promote more cautious decisions.
+By making the consideration of uncertainty values integral to the design of the chart, VSUPs appear to promote more cautious decisions.
 
@@
 
-The results from our initial evaluation of VSUMs are promising, showing that even a general audience can make use of uncertainty information in a reasonable way.
+The results from our initial evaluation of VSUPs are promising, showing that even a general audience can make use of uncertainty information in a reasonable way.
 The way that uncertainty information is presented can have a measurable impact on decision-making.
-The VSUM strategy of only ``showing'' differences when uncertainty is sufficiently low is somewhat analogous to the inferential statistics such as effects tests.
-VSUMs might therefore be used to promote more caution in judgments, and lead analysts away from spurious signals in data.
+The VSUP strategy of only ``showing'' differences when uncertainty is sufficiently low is somewhat analogous to the inferential statistics such as effects tests.
+VSUPs might therefore be used to promote more caution in judgments, and lead analysts away from spurious signals in data.
 
@@
 % When should you use these things?
-In general, VSUMs excel in scenarios where we wish to make firm guarantees about the discriminability of different marks in a chart, but must also represent, in as much detail as possible, two, potentially orthogonal variables of interest.
-VSUMs have improved data resolution over na\"ive bivariate maps by assuming that uncertain values ought to have less visual impact that highly certain values.
-VSUMs therefore act both as a kind of filtering device as well as a bivariate representation.
+In general, VSUPs excel in scenarios where we wish to make firm guarantees about the discriminability of different marks in a chart, but must also represent, in as much detail as possible, two, potentially orthogonal variables of interest.
+VSUPs have improved data resolution over na\"ive bivariate maps by assuming that uncertain values ought to have less visual impact that highly certain values.
+VSUPs therefore act both as a kind of filtering device as well as a bivariate representation.
 This assumption holds in scenarios like the virology dataset (\figref{fig:viralVsum}), where the expectation is that highly uncertain data is unreliable or should otherwise be downweighted in analysis.
 If the analyst has a different interpretation of uncertainty, or wishes to quickly and orthogonally analyze the distributions of uncertainty and value, other strategies, such as juxtaposed maps, may be more appropriate.
-VSUMs are designed for the \emph{integration} of value and uncertainty.
+VSUPs are designed for the \emph{integration} of value and uncertainty.
 Designers should take care in considering when and how this integration is desirable.
@@
 The perceptual non-separability of color channels is well-known~\cite{garner1970integrality, ware2012information}, and so the concept of a limited ``budget'' of distinguishable marks easier to quantify and illustrate.
-In principle, a VSUM can be created for any combination of visual variables.
+In principle, a VSUP can be created for any combination of visual variables.
 All that is required is a perceptual model of the interaction between these two variables.
-Where these models exist, as with the interaction between size and color~\cite{stone2014engineering}, the creation of VSUMs is straightforward.
-Where these models do not exist, or where the perceptual interaction is too complex to efficiently model, experimental work remains to be done before VSUMs can be considered a feasible design strategy.
-We are currently experimenting with the creation of VSUMs using these other channels, as in \figref{fig:size}.
+Where these models exist, as with the interaction between size and color~\cite{stone2014engineering}, the creation of VSUPs is straightforward.
+Where these models do not exist, or where the perceptual interaction is too complex to efficiently model, experimental work remains to be done before VSUPs can be considered a feasible design strategy.
+We are currently experimenting with the creation of VSUPs using these other channels, as in \figref{fig:size}.
 
@@
 This integration introduces additional complexity in the design and presentation of data.
-Value-Suppressing Uncertainty Maps represent one strategy for dealing with this complexity, by assigning marks properties in a way that supports the disambiguation of values in data where uncertainty is low, but suppresses these judgments when uncertainty is high.
+Value-Suppressing Uncertainty Palettes represent one strategy for dealing with this complexity, by assigning marks properties in a way that supports the disambiguation of values in data where uncertainty is low, but suppresses these judgments when uncertainty is high.
 This decision of how to allocate visual variables promotes patterns of decision-making that make responsible use of uncertainty information, discouraging comparison of values in unreliable regions of the data, and promoting comparison in regions of high certainty.

### 2017-09-12 modules/src/continuous.js

@@ -128,3 +128,26 @@ export function continuousArc(m_size, m_scale) {
 
   return arc;
 }
+
+export function continuousLine(m_size, m_scale) {
+  var line = continuousSquare(m_size, m_scale);
+
+  line.makePixelData = function() {
+    var pixelData = [];
+    var c;
+    var size = line.size();
+
+    for (var i = 0;i < size;i++) {
+      for (var j = 0;j < size;j++) {
+        c = line.scale()(j / size);
+        pixelData.push(c.r);
+        pixelData.push(c.g);
+        pixelData.push(c.b);
+        pixelData.push(255);
+      }
+    }
+    return pixelData;
+  }
+
+  return line;
+}

### 2017-09-12 modules/src/index.js

@@ -1,6 +1,6 @@
 export {simpleHeatmap as heatmap, simpleArcmap as arcmap} from "./heatmap";
-export {treeQuantization as quantization, squareQuantization} from "./quantization";
+export {treeQuantization as quantization, squareQuantization, linearQuantization} from "./quantization";
 export {simpleScale as scale} from "./scale";
-export {continuousSquare as csquare, continuousArc as carc} from "./continuous";
+export {continuousSquare as csquare, continuousArc as carc, continuousLine as line} from "./continuous";
 import * as legend_ from "./legend";
 export const legend = legend_;

### 2017-09-12 modules/src/legend.js

@@ -27,7 +27,7 @@ export function simpleLegend(m_scale, m_size, m_height, m_format, m_title, m_x, 
       return;
     }
 
-    var domain = scale.domain(),
+    var domain = scale.domain ? scale.domain() : [0, 1],
         w = size / scale.range().length,
         step = (domain[1] - domain[0]) / scale.range().length,
         dom = d3.range(domain[0], domain[1] + step, step),
@@ -38,7 +38,7 @@ export function simpleLegend(m_scale, m_size, m_height, m_format, m_title, m_x, 
       .attr("transform", "translate(" + x + "," + y + ")");
 
     var rect = el.selectAll("rect").data(scale.range())
-    
+
     rect.enter()
         .append("rect")
       .merge(rect)
@@ -181,7 +181,7 @@ export function heatmapLegend(data, m_scale, m_size, m_format, m_utitle, m_vtitl
 
     var uStep = 1 / data.length;
     var uDom = d3.range(0, 1 + uStep - epsylon, uStep);
-  
+
     var vStep = 1 / data.length;
     var vDom = d3.range(0, 1 + vStep - epsylon, vStep);
 
@@ -356,9 +356,9 @@ export function arcmapLegend(data, m_scale, m_size, m_format, m_utitle, m_vtitle
     var angle = d3.scaleLinear()
       .domain([vTicks[0], vTicks[vTicks.length - 1]])
       .range([-30, 30]);
-  
+
     var offset = 3 * px;
-  
+
     var myArc = d3.arc()
       .innerRadius(size + offset)
       .outerRadius(size + offset + 1)
@@ -367,32 +367,32 @@ export function arcmapLegend(data, m_scale, m_size, m_format, m_utitle, m_vtitle
 
     var arcAxis = el.append("g")
       .attr("transform", "translate(" + (size / 2) + "," + (size - offset) + ")");
-  
+
     arcAxis.append("path")
       .attr("fill", "black")
       .attr("stroke", "transparent")
       .attr("d", myArc);
-  
+
     var labelEnter = arcAxis.selectAll(".arc-label").data(vTicks).enter()
       .append("g")
         .attr("class", "arc-label")
         .attr("transform", function(d) {
           return "rotate(" + angle(d) + ")translate(" + 0 + "," + (-size - offset) + ")";
         })
-  
+
     labelEnter.append("text")
       .style("font-size", "11")
       .style("text-anchor", "middle")
       .attr("y", -10)
[…]

### 2017-09-12 modules/src/quantization.js

@@ -1,8 +1,36 @@
-/*
-A tree-like bivariate quantization scheme.
-As uncertainty increases, the number of quantization bins decreases.
-*/
+import * as d3 from "d3";
 
+export function linearQuantization(m_n, m_range) {
+  var n = m_n,
+  range = m_range,
+  scale = makeScale();
+
+  function makeScale() {
+    return d3.scaleQuantize().range(d3.quantize(range, n));
+  }
+
+  function quantization(value) {
+    return scale(value);
+  }
+
+
+  quantization.range = function() {
+    return scale.range();
+  }
+
+  quantization.n = function(newN) {
+      if (!arguments.length) {
+        return n;
+      }
+      else {
+        n = newN;
+        scale = makeScale();
+        return quantization;
+      }
+  }
+
+  return quantization;
+}
 
 export function squareQuantization(m_n) {
   var n = m_n,

### 2017-09-12 modules/src/scale.js

@@ -11,7 +11,7 @@ export function simpleScale(m_mode, m_range, m_scale) {
       },
       mode = m_mode;
 
-  function CIEdist(color1, color2) {
+  function CIEDist(color1, color2) {
       var c1 = d3.lab(d3.color(color1)),
           c2 = d3.lab(d3.color(color2));
       return Math.sqrt(Math.pow( (c1.l - c2.l), 2) + Math.pow(c1.a - c2.a, 2) + Math.pow(c1.b - c2.b, 2));

### 2017-09-12 study/data/analysis.js

@@ -3,7 +3,7 @@ var dataTwo;
 
 analysis();
 
-function process(){
+function process() {
   //Steps:
   // 1. Load in csvs of both raw data, and turk completion information.
   // 2. Filter out rows from turkIDs who didn't complete the HIT.
@@ -21,7 +21,7 @@ function process(){
   var turkId;
   var row;
 
-  for(var i = 0;i<demoTable.length;i++){
+  for(var i = 0;i<demoTable.length;i++) {
     row = demoTable[i];
     turkId = row.WorkerId;
     valids = dataOneTable.filter(function(x){ return x.workerId == turkId;});
@@ -32,11 +32,11 @@ function process(){
     cleanedOne = cleanedTwo.concat(valids);
   }
 
-  for(var i = 0;i<cleanedOne.length;i++){
+  for(var i = 0;i<cleanedOne.length;i++) {
     writeRow(cleanedOne[i]);
   }
 
-  for(var i = 0;i<cleanedTwo.length;i++){
+  for(var i = 0;i<cleanedTwo.length;i++) {
     writeRow(cleanedTwo[i]);
   }
 
@@ -44,7 +44,7 @@ function process(){
   dataTwo = cleanedTwo;
 }
 
-function writeRow(row){
+function writeRow(row) {
   var writeRequest = new XMLHttpRequest();
   var writeString = "clean=true&answer="+JSON.stringify(row);
   //console.log(writeString);
@@ -53,10 +53,10 @@ function writeRow(row){
   writeRequest.send();
 }
 
-function cleanRows(rows,index){
+function cleanRows(rows, index) {
   var cleanedRows = [];
   var seen = dl.repeat(false,100);
-  for(let row of rows){
+  for(let row of rows) {
     row.id = index;
     if(!seen[row.index]){
       seen[row.index]=true;
@@ -108,12 +108,12 @@ dl.bootstrap.midmeanci = function(values, a, b, c, d) {
           ];
 };
 
-function analysis(){
+function analysis() {
   dataOne = dl.csv("taskOne.csv");
   dataTwo = dl.csv("taskTwo.csv");
 }
 
-function bsci(data,facet,measure){
+function bsci(data, facet, measure) {
   var values = d3.nest().key(function(d){ return d[facet];}).entries(data);
   var cis = values.map(function(d){ return {"key": d.key, "midmean": dl.mean.iqm(d.values,measure), "ci": dl.bootstrap.midmeanci(d.values,measure,1000)};});
   return cis;

### 2017-09-12 study/src/script.js

@@ -25,7 +25,7 @@ Include both gain and loss framings.
 5. Collect demographic information, including risk assay.
 */
 
-var experiment = "Pilot";
+var experiment = "Exp1";
 var startTime;
 var main = d3.select("#fcontainer");
 var done = false;
@@ -37,65 +37,59 @@ var taskTwoStimuli = [];
 var taskTwoGrid = [];
 var taskTwoTokens = 0;
 
-var maxSize = 200/16;;
-
-//var map = d3.interpolateViridis;
-
-function scaleGenerator(map){
-  var maps = makeMaps(map, 15,maxSize);
-  var uSL = makeuSL(map);
-  var uSize = makeuSize(map,maxSize);
-
-  var squareScale = makeScaleFunction(maps.square, uSL);
-  var arcScale = makeScaleFunction(maps.arc, uSL);
-  var arcSizeScale = makeScaleFunction(maps.arcSize, uSize);
-
-  var juxtaposedValueScale = function(d) {
-    return map(d3.scaleQuantize().domain([0,1]).range(maps.linearValue)(d.v));
-  }
-
-  var juxtaposedUncertaintyScale = function(d) {
-    return reverseGrey(d3.scaleQuantize().domain([0,1]).range(maps.linearUncertainty)(d.u));
-  }
-
-  var juxtaposedSizeScale = function(d){
-    return {"c":"black", "s":Math.max(0.1*maxSize,d3.scaleQuantize().domain([0,1]).range(maps.linearUncertainty)(d.u)*maxSize)};
-  }
+var maxSize = 200/16;
 
 
-  return {"square":maps.square, "arc":maps.arc, "arcSize":maps.arcSize,
-  "cs":uSize, "cl":uSL, "squareScale": squareScale, "arcScale": arcScale,
-   "arcSizeScale":arcSizeScale, "juxtaposedUncertaintyScale":juxtaposedUncertaintyScale,
-    "juxtaposedValueScale":juxtaposedValueScale, "juxtaposedSizeScale":juxtaposedSizeScale,
-  "linearValue":maps.linearValue, "linearUncertainty":maps.linearUncertainty};
+//Variables for our 8 conditions
+
+//8 conditions
+// 1. Discrete square
+// 2. Continuous square
+// 3. Discrete juxtaposed
+// 4. Continuous juxtaposed
+// 5. Discrete arc
+// 6. Continuous arc
+// 7. VSUM square
+// 8. VSUM arc
[…]

### 2017-09-13 paper/main.tex
Implemented Dom's feedback

@@
 
-We describe the motivations behind the use of VSUP, examples of their utility for decision-making under uncertainty, and assess VSUP in a crowdsourced experiment.
-Our experimental results indicate that VSUP create close integration between uncertainty and data, promoting rapid but cautious decision-making.
+We describe the motivations behind the use of VSUPs, examples of their utility for decision-making under uncertainty, and assess VSUPs in a crowdsourced experiment.
+Our experimental results indicate that VSUPs create close integration between uncertainty and data, promoting rapid but cautious decision-making.
 
@@
 While we acknowledge the potential utility of these techniques in uncertainty visualization (for instance, the animation used to convey sampling error in Hullman \ea's Hypothetical Outcome Plots~\cite{hullman2015hypothetical}), we limit the scope of our discussion to techniques which center on static charts.
-
-%There are a number of techniques for communicating values and uncertainty in one-dimensional statistical graphics~\cite{olston2002visualizing, wickham201140, neyman1937outline}.
-Techniques such as error bars, confidence bands, and violin plots~\cite{hintze1998violin} use a separate position channel to encode uncertainty.
-The techniques in this paper also apply to 2D visualizations such as heatmaps.
-
-%\strategyFig
-
-%To support our goal of making uncertainty information well-integrated with data, we focus on the techniques of \emph{juxtaposition} (where uncertainty is presented alongside the data), \emph{superposition} (where additional glyphs are overlaid on top of the data), and \emph{explicit encoding} (where an additional visual channel is used to encode the uncertainty).
-Gleicher \ea~\cite{gleicher2011visual} explore these three strategies from the perspective of affording \emph{comparisons}; here, we examine them from the perspective of incorporating measures of uncertainty.
-\figref{fig:strategies} shows simple examples of each of these three techniques for presenting both uncertainty and value.
-
-%\subsection{Strategies for Encoding Uncertainty}
-
-%\emph{Juxtaposition} creates two explicit visualizations: an ``uncertainty map'' and a ``data map.'' MacEachren~\cite{maceachren1992visualizing} portrays uncertainty visualization as the design problem of how to \emph{unify} these two maps, and so this explicit separation can be counterproductive to the goal of information fusion.
-Juxtaposition also creates an additional search task for viewers, who must register hotspots in one map with corresponding features of the other map.
-Rather than complete this task, viewers may choose to ignore the uncertainty information altogether.
-This ignoreability is a key deficiency for information fusion tasks.
-Nevertheless, especially in cases where the quantification of uncertainty is a separate analysis process from the measuring of data values, juxtaposition is a common design tactic for conveying uncertainty~\cite{moritz2017trust}.
-
-%\emph{Superposition}, or the overlay of glyphs representing uncertainty on top of a data map, is another common strategy.
-Error bars are a common example of this strategy: a mark indicating measurement error is superimposed atop a mark representing a point value.
-Ware's~\cite{ware2009quantitative} \emph{textons} are an example of this strategy applied to thematic maps: monochrome glyphs that can encode an ordinal variable, or a binned quantitative variable.
-Greithe \ea~\cite{griethe2006visualization} and MacEachren~\cite{maceachren1992visualizing,maceachren1998visualizing} present other examples of superimposed uncertainty in thematic maps, including overlaid grids with ``fuzzy'' lines to indicate uncertainty, or overlaid treemaps with coarser and coarser resolution in uncertain regions.
-Costs to this approach include obscuring the underlying data map (which can cause uncertainty information to dominate the display~\cite{brodlie2012review}), potential visual interactions that can impair the legibility of glyphs (for instance, simultaneous contrast effects), and the necessity of performing spatial binning to ensure that there are a discrete number of uncertainty glyphs.
-How this binning is performed can introduce undesirable artifacts on the resulting visualization, suppressing important signals or highlighting spurious ones~\cite{battersby2016shapes}.
-
-%\emph{Explicit encoding} directly represents uncertainty with a visual channel, simultaneously with data.
-VSUP employ this strategy.
-In addition to overcoming the drawbacks of the other techniques mentioned in this section, we believe that explicit encoding represents the most general form of the solution of how to fuse uncertainty information with data.
-The most pressing design decisions for explicit encoding are 1) which visual variables to use to encode uncertainty and 2) how to design bivariate maps which afford the encoding of both value and uncertainty.
-We deal with these two questions in greater detail below.
 
@@
 \subsection{Bivariate Maps}
-
 %TODO cites, fill out section
 
+While uncertainty can be visualized entirely separately from the rest of the data (for instance, in a juxtaposed chart), this runs the risk that uncertainty information becomes \emph{ignorable}\cite{moritz2017trust}.
+For our focus on \emph{integrating} uncertainty information into analysis, we focus instead on the construction of \emph{bivariate maps}, where value and uncertainty information is displayed simultaneously in the same visualization.
+
 % examples: roth2010value,ware2009quantitative
+% weaving: lucchesi2017visualizing
+% theory: trumbo1981theory,bernard2015survey,robertson1986generation,dunn1989dynamic,wainer1980empirical
+
+Given these considerations, and the perceptual integrality of color channels~\cite{garner1970integrality, ware2012information}, the construction of bivariate color maps is difficult.
+Viewers need to be able to independently assess the variables of interest, but the resulting bivariate map should also be expressive and easy to interpret.
+Bivariate maps with these constraints are often limited to relatively few categories (say, a 3x3 matrix as in \figref{fig:example}) \cite{robertson1986generation,trumbo1981theory}.
+In general, the quality of bivariate color maps is a multivariate measure involving consideration of not just the component color channels, but also the interpolation scheme and the color of the surround~\cite{bernard2015survey}.
+However, even simple bivariate maps can be difficult for a general audience to interpret \cite{wainer1980empirical}.
 
@@
 
-Given these considerations, and the perceptual integrality of color channels~\cite{garner1970integrality, ware2012information}, the construction of \emph{bivariate} color maps is difficult.
-Viewers need to be able to independently assess the variables of interest, but the resulting bivariate map should also be expressive and easy to interpret.
-Bivariate maps with these constraints are often limited to relatively few categories (say, a 3x3 matrix as in \figref{fig:example}) \cite{robertson1986generation,trumbo1981theory}.
-In general, the quality of bivariate color maps is a multivariate measure involving consideration of not just the component color channels, but also the interpolation scheme and the color of the surround~\cite{bernard2015survey}.
-However, even simple bivariate maps can be difficult for a general audience to interpret \cite{wainer1980empirical}.
-
 %Dunn~\cite{dunn1989dynamic} attempts to circumvent these issues by dynamically allocating bins to bivariate maps.
@@
 
+\flowFig
+
 Value-Suppressing Uncertainty Palettes (VSUPs) are a technique for creating bivariate maps of data \emph{value} and \emph{uncertainty}.
 
-In some cases, this second assumption is violated.
+%In some cases, this second assumption is violated.
 For instance, an analyst might be interested in ``long tail risks'' or other ``black swan events,'' where the impact of a value, no matter how uncertain, must be considered and planned for~\cite{taleb2011black}.
@@
 
-If both of these assumptions hold, then it follows that the designer of a bivariate map should allocate more mark types to certain values, and fewer mark types to uncertain values.
+%If both of these assumptions hold, then it follows that the designer of a bivariate map should allocate more mark types to certain values, and fewer mark types to uncertain values.
 VSUPs codify this decision by \emph{reducing the number of mark categories for representing value as uncertainty increases.} This means that data encoded using a VSUP will make visible only the largest of differences in uncertain data, but highlight comparatively small differences in value when uncertainty is low.

### 2017-09-13 study/src/script.js

@@ -145,6 +145,8 @@ function consent() {
     readyBtn.text("PREVIEW");
   }
 
+  window.scrollTo(0,0);
+
 }
 
 function finishConsent() {
@@ -223,6 +225,7 @@ function tutorial() {
   main.selectAll("*").remove();
 
   main.append("iframe")
+  .attr("height","100%")
   .attr("src", "tutorial.html");
 
   main.append("button")
@@ -232,6 +235,8 @@ function tutorial() {
   .attr("name", "answer")
   .text("Ready")
   .on("click", finishTutorial);
+
+  window.scrollTo(0,0);
 }
 
 function finishTutorial() {
@@ -304,7 +309,7 @@ function makeTaskTwoStimuli() {
     {binned: "discrete", shape:"arc", vsum:true}
   ];
 
-  var replicates = 8;
+  var replicates = 4;
 
   for (var type of types) {
     for (var i = 0;i < replicates;i++) {
@@ -357,7 +362,7 @@ function taskOne() {
 
   main.append("svg")
   .attr("id", "map")
-  .attr("style", "width: 700px; height: 250px;");
+  .attr("style", "width: 700px; height: 250px; padding-bottom: 10px");
 
   main.append("div")
   .attr("id", "legend")
@@ -446,7 +451,7 @@ function drawMap(stim, data, task) {
 
   if (stim.binned == "continuous") {
     var canvasDiv = d3.select("#legend").append("div")
-      .attr("style", "position: relative; margin-top: -175px; margin-left: 275px;")
+      .attr("style", "position: relative; margin-top: -175px; margin-left: 275px; padding-bottom: 25px;")
     if (stim.shape == "square") {
       var square = bvu.csquare(150, d3.interpolateViridis);
       canvasDiv.call(square);
@@ -636,7 +641,7 @@ function taskTwo() {
 
   main.append("svg")
   .attr("id", "map")
-  .attr("style", "width: 700px; height: 250px;");
+  .attr("style", "width: 700px; height: 250px; padding-bottom: 10px");
 
   main.append("div")
   .attr("id", "legend")

### 2017-09-14 paper/figures.tex
Updated easy figures

@@
 
-\newcommand{\teaserFig}{
-  \teaser{
-		\centering
-		\includegraphics[width=0.9\textwidth]{example.pdf}
-		\caption{Lookit! Lookit!}
-		\label{fig:teaser}
-	}
-}
-
 \newcommand{\exampleFig}{
@@
 	\includegraphics[width=0.9\columnwidth]{example.pdf}
-	\caption{A standard bivariate map (left) and a VSUP (right), used to encode an identical 5x5 grid of random data.
-Both use the same visual channels to encode value (position along the Viridis~\protect\cite{viridis} color map) and uncertainty (lightness and saturation), and have an equal standard of perceptual discriminability (at least 18 units of distance in CIELAB color space between colors).
-However, highly uncertain values result in colors that are very close together in the bivariate case, meaning the full bivariate map is only 9 bins under these constraints\,---\,a 4x4 map would result in colors that are perceptually too close together.
-By contrast, the VSUP intentionally reduces bins when uncertainty is high, eventually aliasing all highly uncertain values to the same color.
-This decision affords more distinct colors in other regions of the map, and so an increase in overall bins to 15.
-The resulting map suppresses the value of highly uncertain data, but increases the discriminability of data with low uncertainty.}
+	\caption{A standard bivariate map (left) and a VSUP (right), used to encode an identical 10x10 grid of random data.
+Both use the same visual channels to encode value (position along the Viridis~\protect\cite{viridis} color map) and uncertainty (lightness and saturation).
+However, the VSUP uses a tree-like structure to allocate colors, defining more bins when uncertainty is low.
+This non-uniform budgeting affords better discrimination between values when uncertainty is low, even though the VSUP has fewer color bins (15 to the bivariate map's 16).
+This tree-like structure also discourages analysis in regions where uncertainty may be unacceptably high.}
 	\label{fig:example}
@@
 	\centering
-	\includegraphics[width=\columnwidth]{airline-2d.png}
+	\includegraphics[width=\columnwidth]{airline-2d}
 	\vspace{-15px}
@@
 	\centering
-	\includegraphics[width=\columnwidth]{airline-vsum.png}
+	\includegraphics[width=\columnwidth]{airline-vsum}
 	\vspace{-15px}

### 2017-09-14 paper/figures.tex
Finalized todo list

@@
 However, the VSUP uses a tree-like structure to allocate colors, defining more bins when uncertainty is low.
-This non-uniform budgeting affords better discrimination between values when uncertainty is low, even though the VSUP has fewer color bins (15 to the bivariate map's 16).
+This non-uniform budgeting affords better discrimination between values when uncertainty is low, even though the VSUP has fewer color bins (in this case, 15 to the bivariate map's 16).
 This tree-like structure also discourages analysis in regions where uncertainty may be unacceptably high.}
@@
 \newcommand{\taskTwoFig}{
-	\begin{figure*}
+	\begin{figure}
 		\centering
-		\begin{subfigure}{0.45\textwidth}
-			\includegraphics[width=\textwidth]{attack.png}
-			\caption{Attacking}
-			\label{fig:taskTwoAttack}
-		\end{subfigure}
-		~
-		\begin{subfigure}{0.45\textwidth}
-			\includegraphics[width=\textwidth]{defend.png}
-			\caption{Defending}
-			\label{fig:taskTwoDefend}
-		\end{subfigure}
-		\caption{The two framings of the Prediction task.
-In the ``attack'' framing, the participant is given a map of predictions of the location of enemy ships, along with the uncertainty in those predictions.
-The participant should place their missile strikes on locations with a high probability of containing a ship, and with high certainty in this probability.
-The ``defend'' framing is the opposite task: the participant has a list of likely missile locations, and ought to place their ships on locations with low probability of attack, and high certainty in this probability.}
+		\includegraphics[width=0.45\columnwidth]{defend.png}
+		\caption{The prediction task. The participant has a list of locations, and ought to place their ships on locations with low probability of attack, and high certainty in this probability.}
 		\label{fig:taskTwoConditions}
-	\end{figure*}
-}
-
-\newcommand{\performanceFig}{
-	\begin{figure}[t]
-		\centering
-		\includegraphics[width=0.9\columnwidth]{performance.pdf}
-		\caption{
-			The number of discrete color categories in traditional bivariate color maps versus VSUPs, assuming value is encoded by the Viridis color map~\protect\cite{viridis}, and uncertainty is encoded by saturation/value (whiter values are more uncertain).
-Since these whiter uncertain colors are closer together in CIELAB space, and VSUPs intentionally require fewer color bins in those regions, VSUPs always have more color categories than the corresponding 2D bivariate map.
-The red lines denote:
-		  \textbf{1)} Theoretical 50\% JND from CIELAB specification.
-			\textbf{2)} 50\% JND from controlled lab study of Mahy et al.~\protect\cite{mahy1994evaluation}.
-			\textbf{3)} 50\% JND from Szafir et al.\protect\cite{szafir2014adapting}.
-			\textbf{4)} 50\% JND for small visual targets from Stone et al.~\protect\cite{stone2014engineering}.
-			\textbf{5)} The threshold resulting in the simplest possible (2x2) bivariate map.
-	    }
-		\label{fig:performance}
 	\end{figure}
@@
 
-\newcommand{\flowFig}{
-	\begin{figure}[t]
-		\centering
-		\includegraphics[width=0.95\columnwidth]{flow.pdf}
-		\caption{The process for creating a VSUP.
+%\newcommand{\performanceFig}{
+%	\begin{figure}[t]
+%		\centering
+%		\includegraphics[width=0.9\columnwidth]{performance.pdf}
+%		\caption{
+%			The number of discrete color categories in traditional bivariate color maps versus VSUPs, assuming value is encoded by the Viridis color map~\protect\cite{viridis}, and uncertainty is encoded by saturation/value (whiter values are more uncertain).
+Since these whiter uncertain colors are closer together in CIELAB space, and VSUPs intentionally require fewer color bins in those regions, VSUPs always have more color categories than the corresponding 2D bivariate map.
+The red lines denote:
+%		  \textbf{1)} Theoretical 50\% JND from CIELAB specification.
+%			\textbf{2)} 50\% JND from controlled lab study of Mahy et al.~\protect\cite{mahy1994evaluation}.
+%			\textbf{3)} 50\% JND from Szafir et al.\protect\cite{szafir2014adapting}.
+%			\textbf{4)} 50\% JND for small visual targets from Stone et al.~\protect\cite{stone2014engineering}.
+%			\textbf{5)} The threshold resulting in the simplest possible (2x2) bivariate map.
+%	    }
+%		\label{fig:performance}
+%	\end{figure}
+%}
+
+%\newcommand{\flowFig}{
+%	\begin{figure}[t]
+%		\centering
+%		\includegraphics[width=0.95\columnwidth]{flow.pdf}
+%		\caption{The process for creating a VSUP.
 VSUPs guarantee that no two marks in the resulting bivariate mapping are perceptually closer than some pre-specified threshold.
@@
 \\
-		In this example two colors from the map for $n=9$ are too close in CIELAB space. Thus, we create a VSUP with $n=8$ color bins at the lowest level of uncertainty.}
-		\label{fig:flow}
-	\end{figure}
-}
+%		In this example two colors from the map for $n=9$ are too close in CIELAB space. Thus, we create a VSUP with $n=8$ color bins at the lowest level of uncertainty.}
+%		\label{fig:flow}
+%	\end{figure}
+%}
 
@@
 	\vspace{-15px}
-	\caption{Variability for 25 different viral populations of SIV, encoded with a VSUP.
+	\caption{Variability for 12 different viral populations of SIV, encoded with a VSUP.
 The magnitude of mutation rate, as well as the generally poorer quality of data towards the end of the sequence, is more readily visible than with the traditional bivariate color map.

### 2017-09-14 study/src/script.js

@@ -273,12 +273,12 @@ function makeTaskOneStimuli() {
   var types =
   [
     {binned: "discrete", shape:"square", vsum:false},
-    {binned: "continuous", shape:"square", vsum:false},
+//    {binned: "continuous", shape:"square", vsum:false},
     {binned: "discrete", shape:"juxtaposed", vsum:false},
-    {binned: "continuous", shape:"juxtaposed", vsum:false},
-    {binned: "discrete", shape:"arc", vsum:false},
-    {binned: "continuous", shape:"arc", vsum:false},
-    {binned: "discrete", shape:"square", vsum:true},
+//    {binned: "continuous", shape:"juxtaposed", vsum:false},
+//    {binned: "discrete", shape:"arc", vsum:false},
+//    {binned: "continuous", shape:"arc", vsum:false},
+//    {binned: "discrete", shape:"square", vsum:true},
     {binned: "discrete", shape:"arc", vsum:true}
   ];
 

### 2017-09-14 paper/main.tex
Prepping for study run.

@@
 
-To our knowledge, there are no established standards for visualizing uncertainty in heatmaps.
+There are many methods for visualizing uncertainty.
+MacEachren et al.
+~\cite{maceachren1992visualizing} presents a theoretical survey of uncertainty visualization generally.
+Relevant to our domain of heatmaps and thematic maps is work in GIS and cartography; for a survey of uncertainty visualization techniques in GIS, see Kinkeldey et al.
+\cite{kinkeldey2014assess,kinkeldey2017evaluating}.
+Yet, despite these wide variety of methods, to our knowledge, there are no established standards for visualizing uncertainty in heatmaps.
+
 Reviewing the state of the art in uncertainty visualization in the field, Greithe \ea~\cite{griethe2006visualization} and Brodlie \ea~\cite{brodlie2012review} present lists of potential techniques for conveying uncertainty in concert with data.
@@
 
-\subsection{Visual Variables for Representing Uncertainty}
+\subsection{Visual Variables for Uncertainty}
 
-The decision to explicitly encode uncertainty increases the dimensionality of the data, and so requires the use of (at least one) additional visual channel ~\cite{brodlie2012review}.
+% 1d uvis: olston2002visualizing, wickham201140, neyman1937outline
+
+One challenge to conveying uncertainty information is that the decision to explicitly encode uncertainty increases the dimensionality of the data, and so requires the use of (at least one) additional visual channel ~\cite{brodlie2012review}.
 Uncertainty therefore inherently increases the visual complexity of a visualization.
@@
 
-%TODO stuff that needs to get cited here:
-
-% uncertainty vis surveys: kinkeldey2017evaluating,maceachren1992visualizing,brodlie2012review
-% 2d uvis: griethe2006visualization, maceachren1992visualizing,maceachren1998visualizing
-% 1d uvis: olston2002visualizing, wickham201140, neyman1937outline
-% ignoreability: moritz2017trust
-
-%Foundational work in psychophysics and graphical perception make use of the concept of a ``just noticeable difference'' (JND)\cite{cleveland1984graphical}.
-A 50\% JND is a difference between two stimuli that is detectable 50\% of the time.
-For scales and legends, we would want a significantly higher standard of unambiguity.
-That is, the different levels of visual variables used in visualizations should be readily and reliably distinguishable.
-Complicating this goal is that a JND as measured in a controlled environment with cooperative participants can be considerably lower than JNDs measured in the sorts of heterogeneous conditions found in real environments~\cite{szafir2014adapting}.
-
-\subsection{Bivariate Maps}
-%TODO cites, fill out section
-
-While uncertainty can be visualized entirely separately from the rest of the data (for instance, in a juxtaposed chart), this runs the risk that uncertainty information becomes \emph{ignorable}\cite{moritz2017trust}.
-For our focus on \emph{integrating} uncertainty information into analysis, we focus instead on the construction of \emph{bivariate maps}, where value and uncertainty information is displayed simultaneously in the same visualization.
-
-% examples: roth2010value,ware2009quantitative
-% weaving: lucchesi2017visualizing
-% theory: trumbo1981theory,bernard2015survey,robertson1986generation,dunn1989dynamic,wainer1980empirical
-
-Given these considerations, and the perceptual integrality of color channels~\cite{garner1970integrality, ware2012information}, the construction of bivariate color maps is difficult.
-Viewers need to be able to independently assess the variables of interest, but the resulting bivariate map should also be expressive and easy to interpret.
-Bivariate maps with these constraints are often limited to relatively few categories (say, a 3x3 matrix as in \figref{fig:example}) \cite{robertson1986generation,trumbo1981theory}.
-In general, the quality of bivariate color maps is a multivariate measure involving consideration of not just the component color channels, but also the interpolation scheme and the color of the surround~\cite{bernard2015survey}.
-However, even simple bivariate maps can be difficult for a general audience to interpret \cite{wainer1980empirical}.
-
-\subsection{Quantizing Color Scales}
-
-%TODO cites, fill out section
-% binning stuff: [data] battersby2016shapes, [legends] padilla2017evaluating,muller1979perception,dobson1973choropleth,dunn1989dynamic
+\subsection{Color Maps}
 
@@
 
-%Dunn~\cite{dunn1989dynamic} attempts to circumvent these issues by dynamically allocating bins to bivariate maps.
-As the number of distinct, separable colors is limited, the dynamic approach allocates bins to regions of the highest utility.
-For instance, placing the dividing line between bins in areas aligned with a trend line allows the color to encode information about the sign of the residual with respect to each variable.
-This insight, that the bins in a bivariate map are a limited resource and so should be ``spent'' wisely, drives the creation of VSUPs, which allocate more bins in a bivariate map to regions with less uncertainty.
+A design consideration when constructing color maps is whether to quantize data into a discrete set of colors, or to encode data using a continuous mapping from value to color.
+While continuous color maps afford greater fidelity in presenting values~\cite{muller1979perception}, non-linearity in human color perception introduces errors in extracting numeric values from continuous colors~\cite{borland2007rainbow}.
+Quantizing a color map is therefore an exercise in balancing \emph{perceptual} error and \emph{quantization} error~\cite{dobson1973choropleth}.
+Discrete maps offer a finer control over this balance, which can result in better performance in tasks involving heatmaps~\cite{padilla2017evaluating}.
+As such, we limit our discussion to color maps with discrete sets of colors.
+
+\subsection{Bivariate Maps}
+
+While uncertainty can be visualized entirely separately from the rest of the data (for instance, in a juxtaposed chart), this runs the risk that uncertainty information becomes \emph{ignorable}\cite{moritz2017trust}.
+It also complicates analysis, as interpreting the value and uncertainty of a given data point requires consulting two separate charts.
+For our goal on integrating uncertainty information into analysis, we focus instead on the construction of \emph{bivariate maps}, where value and uncertainty information is displayed simultaneously in the same visualization.
+
+Bivariate maps can, in principle, be constructed from the combination of any two visual variables (such as shape and color, or size and texture).
+E.g., Ware's ``textons'' overlay glyphs on top of regions to simultaneously encode two quantitative values~\cite{ware2009quantitative}.
+However, in this work, our focus is on heatmaps and other sorts of thematic maps.
+In visualizations like heatmaps, common visual channels such as position and size are already in use to encode other information (such as spatial location and extent), and color is reserved for value.
+Bivariate maps in these settings therefore typically rely on a visual channel related to color for encoding a secondary variable, such as alpha~\cite{roth2010value} or pixel noise~\cite{lucchesi2017visualizing}.
+
+Due to the perceptual integrality of visual channels~\cite{garner1970integrality, ware2012information}, the construction of color-based bivariate maps is difficult.
+Viewers need to be able to independently assess the variables of interest, but the resulting bivariate map should also be expressive and easy to interpret.
+Bivariate maps with these constraints are often limited to relatively few categories (say, a 4x4 matrix as in \figref{fig:example}) \cite{robertson1986generation,trumbo1981theory}.
+For VSUPs, we adapt an insight from Dunn~\cite{dunn1989dynamic} that these limited categories ought to be assigned with regards to their importance to analytics tasks, rather than uniformly.
+
+In general, the quality of bivariate color maps is a multivariate measure involving consideration of not just the component color channels, but also the interpolation scheme and the color of the surround~\cite{bernard2015survey}.
+However, even simple bivariate maps can be difficult for a general audience to interpret \cite{wainer1980empirical}.
 
@@
 
-\flowFig
+%TODO need some example figure here, showing tree structure perhaps?
+%\flowFig
 
-Value-Suppressing Uncertainty Palettes (VSUPs) are a technique for creating bivariate maps of data \emph{value} and \emph{uncertainty}.
+Value-Suppressing Uncertainty Palettes (VSUPs) are a technique for creating bivariate maps of data \emph{value} and \emph{uncertainty}. VSUPs make two central assumptions about bivariate maps:
 
-%In some cases, this second assumption is violated.
+\begin{enumerate}
+	\item There is a limited \emph{budget} of potential outputs in a bivariate map.
+	\item Differences among \emph{certain} data are more significant than differences among \emph{uncertain} data.
+\end{enumerate}
+
+In some cases, these assumptions are violated.
 For instance, an analyst might be interested in ``long tail risks'' or other ``black swan events,'' where the impact of a value, no matter how uncertain, must be considered and planned for~\cite{taleb2011black}.
@@
 However, for many information fusion tasks, the assumption is that uncertainty is related to data quality, or the uncertainty of data values~\cite{riveiro2007evaluation}.
-As with null hypothesis significance testing, the aim would then be to avoid Type I errors, where patterns arising from statistical noise or measurement error are interpreted as significant signals.
 
-%If both of these assumptions hold, then it follows that the designer of a bivariate map should allocate more mark types to certain values, and fewer mark types to uncertain values.
+However, if both of these assumptions hold, then it follows that the designer of a bivariate map should allocate more mark types to certain values, and fewer mark types to uncertain values.
 VSUPs codify this decision by \emph{reducing the number of mark categories for representing value as uncertainty increases.} This means that data encoded using a VSUP will make visible only the largest of differences in uncertain data, but highlight comparatively small differences in value when uncertainty is low.
@@
 
-\subsection{Method}
-
-\flowFig
-
-The general VSUP construction algorithm is as follows:
-
-\begin{enumerate}
-	\item \textbf{Select a visual variable to encode value}. In this paper, this variable is position along some established color scale.
-\end{enumerate}
-
-\figref{fig:flow} shows this process in detail, and how it can be used to generate a bivariate map for presenting uncertainty.
-JavaScript code for generating VSUPs given arbitrary D3~\cite{bostock2011d3} quantitative scale functions is available at \url{URL-REMOVED-FOR-REVIEW}.
+VSUPs rely on an underlying \textbf{quantization tree} that governs how values are discretized.
+Maximally uncertain values are all mapped to a singular ``root'' node.
+As uncertainty decreases, the tree branches into leaves, which evenly divide the data domain.
+The data value of a parent is the midpoint of all of its children.
+These leaves can then branch again, until a designer-specified stopping point.
+The layers act to the quantize the uncertainty domain, and the nodes the data domain.
+\figref{fig:example}, for example, is a VSUP with a tree with a branching factor of 2, and 4 layers, resulting in a total of 15 different output colors.
 
@@
 
+Javascript code for generating VSUPs for use in D3 charts is available at \url{URL-REMOVED-FOR-REVIEW}.
+
 \subsection{Design Considerations}
 
-\performanceFig
+%\performanceFig
 
 There are multiple choices that designers must make before creating a VSUP.
-In particular, they must choose 1) which visual channels map to value and uncertainty, 2) a threshold of perceptual distinguishability (and, implicitly, a corresponding metric), and 3) a degradation function to determine how quickly uncertain values are aliased.
+In particular, they must choose 1) which visual channels map to value and uncertainty, 2) structure of the underlying VSUP tree.
 
@@
 We recommend channels with both a strong semiotic connection to uncertainty as well as a relatively large number of perceptually distinguishable levels.
-Here, we focus on value/saturation (or, equivalently, transparency against a white background) and size.
+Here, we focus on value/saturation (or, equivalently, transparency against a white background).
 Correll \& Gleicher~\cite{correll2013error} show that even audiences without statistical backgrounds can interpret uncertainty information encoded in these channels.
 
-Another reason for limiting our discussion of bivariate mappings to color and size channels is that the perceptual distinguishability of these channels has been well-studied by prior work.
-CIELAB and other color spaces are designed to approximate perceptual uniformity\,---\,that is, distance in the color space should universally correspond with perceptual distance.
-Recent work~\cite{stone2014engineering,szafir2014adapting} has investigated the discriminability of colors in such spaces not only under laboratory conditions, but also among the disparate display conditions and color vision proficiency of crowdworkers.
-Similarly, formative work by Cleveland \& McGill~\cite{cleveland1984graphical} assesses the discriminability of marks of different height, and this too has been extended both to a larger population of crowdworkers, and to the area channel specifically~\cite{heer2010crowdsourcing,talbot2014four}.
-
-We use values from these works as benchmarks for selecting thresholds of discriminability.
-The JNDs of these channels provides a useful minimum; in practice, we would wish to create maps with marks that are far more reliably distinguishable than the 50\% or 75\% discriminability rate of a JND.
-\figref{fig:performance} shows how these benchmarks can be used to select reasonable thresholds.
-Other factors, such as nameability and aesthetic preference, can be added in as additional weights impacting the desirability of certain color palettes, as in Gramazio \ea~\cite{gramazio2017colorgorical}.
-
-How quickly the number of bins decreases is another parameter of VSUPs.
-If this degradation is too fast, then we may only have a few discriminable levels of uncertainty.
-If this degradation is too slow, then we can be overwhelmed with colors.
-For instance, if we only decrement by one bin per level of uncertainty, there would be $1,485$ different colors in a VSUP with a tolerance threshold of $1$ unit in CIELAB space, far too much complexity for a discrete color mapping.
-In the examples in this paper, we halve the number of bins at each increasing level of uncertainty.
-When the number of initial bins at the highly certain region is a power of $2$ (as in Fig.
-\ref{fig:example}), this imparts a tree structure to the resulting VSUP.
-As uncertainty increases, bins in the previous level map to one and only one ``parent'' in the next uncertainty level, resulting in fewer crossings of color boundaries among rows.
+%TODO Add in discussion of tree structures, including figure showing that VSUPs have colors that are further apart than similar 2d bivariate maps.
+Also wortth discussing discretization issues here.
+Make new performance fig.
 
@@
 We use standard error $\left(\sigma / \sqrt{n}\right)$ as our measure of uncertainty.
-Starting from an minimum discriminability threshold of $18$ units in CIELAB space, we created two alternate heatmaps of this delay information.
+We created two alternate heatmaps of this delay information.
 \figref{fig:airline2d} shows a traditional bivariate map, whereas \figref{fig:airlineVsum} shows a VSUP generated under the same constraints.
@@
 The VSUP, in contrast, has sufficient color resolution to show a quasi-linear trend.
-The additional color resolution for uncertainty information in VSUPs also promotes caution when considering outliers: the small number of flights on Saturdays, or departing late at night, make standard error for these delays relatively high, and the connected values more suspect than the rest of the chart.
+Uncertainty is high only for a few bins for which there are small number of flights, either on the weekends or late at night.
+The VSUP does not obscure any significant trends in the high uncertainty data, despite having fewer color categories overall.
 
+\subsubsection{Viral Mutation}
+%TODO update this figure
 \viralFig
@@
 	\item An \textbf{identification} task, where we gave participants charts with value and uncertainty information, and asked them to locate specific regions.
-E.g., ``click on the region of the chart with highest uncertainty \emph{and} lowest value.''
+E.g., ``click on the region of the chart with a value of $0.1$ and an uncertainty of $0.2$.''
 	\item A \textbf{prediction} task, inspired by the game Battleship, where we gave participants charts with both \emph{forecast} and \emph{forecast uncertainty} information, and asked the participants to place tokens on the board in order to optimize expected value.
-E.g., ``place your $4$ ships on safe locations on the board.''
+E.g., ``place your $5$ ships on safe locations on the board.''
 \end{enumerate}
 
-We discuss the two experimental tasks in more detail in the following sections. Experimental materials, including data tables, are available at \url{URL-REMOVED-FOR-REVIEW}.
-
-For each task, participants saw one of three types of heatmaps.
-Either two \textbf{juxtaposed} maps side-by-side, with separate encodings for value and uncertainty, a traditional \textbf{2D bivariate} map, where value and uncertainty are orthogonally encoded, or a \textbf{VSUP}.
+%TODO finalize factor levels, regenerate conditions figure
+The experiment was a within-subjects design with 1 factor (encoding type) with $N$ levels.
 \figref{fig:conditions} shows examples of these factor levels.
-Other factors included the size of the map (either a 4x4 or 8x8 grid), and (for the prediction task), the question framing (attack or defense).
-Secondary factors were evenly blocked across stimuli, resulting in a 3x2 within subjects design for the identification task, and a 3x2x2 within subjects design for the prediction task.
-Participants reported fatigue effects in piloting, and so we limited the number of replicates in this study: participants saw 18 stimuli (6 for each map type) for the identification task, and 12 (4 for each map type) for the prediction task.
-
-In all cases, we generated color maps based on the requirement that no two colors in the resulting color scheme were closer than 18 units of Euclidean distance in CIELAB space.
-This threshold results in a $9$-color bivariate map and $15$-color VSUP for all color scales used in this study.
-Given the size of the marks used in our study, this threshold distance is sufficient for a predicted discriminability rate of over 95\%~\cite{stone2014engineering}.
-Other choices of threshold, and differing univariate scales, would result in different maps (see Fig.~\ref{fig:performance}).
+We had 8 replications for each level for the identification task, and 4 for the prediction task, for a total of $N$ stimuli for the first task, and $N$ for the second.
+We discuss the two experimental tasks in more detail in the following sections.
+Experimental materials, including data tables and stimuli generation code, are available at \url{URL-REMOVED-FOR-REVIEW}.
 
@@
 
-For the identification task, we gave participants bivariate representations of uncertainty, and asked them to identify a location meeting a given criteria, e.g., ``Click on the location with the greatest uncertainty.'' There were three potential criteria, mapping to the three vertices of a VSUP: 1) high value, low uncertainty, 2) low value, low uncertainty, and 3) high uncertainty.
-Each map contained four locations, one of each combination of maximum and minimum value and uncertainty.
-The correct answer was always one of these locations.
-Other locations in the grid were created pseudo-randomly, with less extreme ranges of uncertainty and probability.
+For the identification task, we gave participants a heatmap, and asked them to click on a region of the heatmap that encodes a particular value,uncertainty pair.
+To support comparability across conditions, we chose a set of target pairs such that each target mapped to a unique color across all encoding types.
+This resulted in (4+4+2+1)=11 valid targets.
+For each trial, these 11 targets were placed randomly in a heatmap; the remaining 13 cells of the heatmap were distractors randomly sampled from this target list.
+
+We selected this task both as a way of assessing the ability of people to successfully encode and decode values using a bivariate map, and to provide training for the prediction task (since we re-used the bivariate maps in that task).
 We measured performance both in terms of accuracy (did the participant select the correct point) as well as response time.
@@
 Greis \ea~\cite{greis2016decision} employ these game-like experimental tasks to assess how different visual designs communicate uncertainty information, which can be abstract or complex, to the general audience.
-In our task, the participant and a (fictional) adversary have to place tokens representing ships on a spatial grid.
-Without direct access to the location of the opponent's ships, the player then has to place tokens representing missile strikes on their opponent's grid.
+In our task, the participant and a (fictional) adversary have to place tokens representing ships on a spatial grid, with the expectation that certain squares will be hit by missiles.
 Players have to place all their tokens before continuing.
-If a player fires a missile at a square containing a ship, then that ship is sunk.
-The objective is to maximize the number of enemy ships hit, while minimizing the number of your own ships that are hit.
-In our task, participants were given a chart encoding predictions of the location of enemy ships (in the attacking case) or potential missile strikes (in the defending case).
-The \emph{value} component of the prediction was the likelihood of an item being present in a particular region.
+The objective is to minimize the number of your own ships that are hit.
+In our task, participants were given a map representing the predictions of missile strikes in each location on the grid.
+The \emph{value} component of the prediction was the likelihood of a missile strike in a particular region.
 The \emph{uncertainty} component was the confidence in this prediction.
 
-We included both a \textbf{gain} framing (the attacker's task: given predictions of where your opponent's ships are, where should you launch your missiles?), and a \textbf{loss} framing (the defender's task: given predictions of where your opponent's missiles will fall, where should you place your ships?) to account for systematic differences in human judgments under uncertainty.
+%TODO stimuli creation
+
+We selected this task in order to promote risk-averse behavior.
 Tversky \& Kahneman~\cite{tversky1985framing} illustrate that framings in terms of gains or losses produce reliably different outcomes.
-In general, the function of the value of gains is concave, but is convex for losses~\cite{kahneman1979prospect}.
-That is, avoiding a large loss is often more valuable than striving for a similarly large gain.
-To avoid the potential for (incorrect) transfer learning, the attacking and defending tasks used different value color maps (Plasma and Cool) than the identification task, and from the opposite framing in this task.
-\figref{fig:taskTwoConditions} shows these differing color maps and framings.
-
-Each map contained a minimum number of extreme locations with each combination of very high and very low value and uncertainty.
-High expected value, high uncertainty locations were ``risky.'' As a balance to the risky locations, there were also an equal number of ``safer'' locations, with slightly less expected value, but also less uncertainty.
-The remaining locations of the map varied pseudo-randomly in the middle ranges.
-
+In particular, there is greater perceived value in avoiding large losses as opposed to striving for a large gain~\cite{kahneman1979prospect}.
 The ideal strategy from a value-maximizing standpoint would be to place tokens on areas with the highest expected value, ignoring the uncertainty information.
@@
 We therefore measured the distribution of both \emph{value} and \emph{uncertainty} of the tokens placed by the participants.
-
-%Show:
-% 1) Juxtaposition makes people ignore/underweight uncertainty (even when they don't screw up the search task)
-% 2) VSUPs make people, surprise surprise, bad at distinguishing highly uncertain values
-% 3) But, they make uncertainty better-integrated
-%    i) lower weights on uncertain values in decision-making
-%    ii) a priori better discrimination
-
-%conditions:
-% 1) juxtaposed d map, u map
-% 2) binned d/u map as per a priori jnds
-% 3) vsm
-% 4) continuous d/u map?
 
@@
 As the experimental tasks required multi-hue color perception, we presented participants with a set of Ishihara plates~\cite{hardy1945tests} as a pretest, and excluded participants who either misidentified the values in the plates, or who self-reported as having a color vision deficiency (CVD) in the post-test.
-Based on piloting, we paid participants \$2 dollars for participation, for a target rate of \$8/hour.
+Based on piloting, we paid participants \$3 dollars for participation, for a target rate of \$8/hour.
 After completion of the main tasks, we solicited demographic information.
-We recruited a total of 24 participants for this experiment: 4 female, 19 male, and 1 who declined to state ($M_{\text{age}}$= 32, $SD_{\text{age}}$ = 7.5).
-All reported having either college degrees, or some college experience.
-Average self-reported experience with interpreting graphs and charts was low (1.9 on a 5 point Likert scale, with 1 indicating ``No experience''), and no participant reported experience greater than 3 out of 5.
+
+%TODO add participant info
+We recruited a total of $N$ participants for this experiment: $N$ female, $N$ male, and $N$ who declined to state ($M_{\text{age}}$= $N$, $SD_{\text{age}}$ = $N$).
 
 \subsubsection{Identification Task}
-\accuracyFig
-\responseTimeFig
 
-For the identification task, our primary measures of performance were accuracy and response time. Our hypothesis was that VSUPs would exhibit better performance than the other charts.
-
-We performed a one-way ANOVA investigating the effect of chart type (juxtaposed, full bivariate, or VSUP) on accuracy.
-We did not find a significant effect ($F(2,859) = 2.4$, $p=0.09$), although VSUPs had the highest accuracy.
-\figref{fig:accuracy1} illustrates this result.
-
-We also performed a one-way ANOVA on the effect of chart type on response time.
-In this case, we found a significant main effect ($(F,2,859) = 5.0$, $p<0.001$).
-A Tukey's HSD found that participants were significantly slower at performing the task on juxtaposed maps, as opposed to the other map types.
-\figref{fig:rt1} illustrates this result.
-
-These results only partially support our hypothesis.
-There was no significant difference in terms of accuracy between different maps, partially due to the simplicity of the identification task: participants only had to find a single target in a heatmap, with extreme values in both data and uncertainty axes.
-However, our results do provide evidence that juxtaposed maps, by introducing an additional search task to the problem of assessing the value and uncertainty of a location, require significantly more time to interpret.
+%TODO add results and figures
+%\accuracyFig
+%\responseTimeFig
 
@@
 
-\uncertaintyFig
+%TODO add results and figures
+%\uncertaintyFig
 
@@
 Our hypothesis is that VSUPs, by communicating uncertainty information at the expense of value information, would promote more conservatives guesses than the other chart types.
-
-We performed a one-way ANOVA of chart type on average value of guesses.
-We found no significant effect ($F(2,861)=0.27$, $p=0.8$), indicating that participants were generating similar expected values from their guesses across conditions.
-
-However, an ANOVA of chart type on average \emph{uncertainty} of guesses did find a significant main effect ($F(2,861)=61$, $p<0.001$).
-A post-hoc Tukey's HSD confirmed that participants had guesses with significantly less uncertainty with VSUPs than with both other chart types.
-Additionally, participants' guesses on juxtaposed charts were significantly riskier than with the other chart types.
-\figref{fig:uncertainty2} illustrates this result.
-
-As with the identification task, a one-way ANOVA also indicated that chart type had a significant effect on response time ($F(2,861)=13$, $p<0.001$).
-A Tukey's HSD confirmed that, as with the previous task, response time was significantly longer with juxtaposed maps (trimmed mean response time of $23$ seconds) compared to the other chart types (both with trimmed mean response time of $18$ seconds).
-
-Despite the documented effect of framing in prior work, a one-way ANOVA of task frame on expected value found no significant main effect ($F(1,862)=0.01$, $p=0.9$), and the trimmed means were quite similar (expected value of $0.76$ per guess for attackers, versus $0.75$ per guess for defenders).
-
-Our results again partially support our hypotheses.
-We did not observe statistically different patterns of \emph{expected value}, but did find that VSUPs produce different patterns of \emph{acceptable uncertainty}.
-That is, VSUPs promote avoidance of uncertain values, without notably impacting the expected value.
-Juxtaposed maps, by contrast, make the retrieval of uncertainty information significantly more difficult; it is possible that participants ignore or downweight it when making predictions, resulting in riskier predictions overall.
-By making the consideration of uncertainty values integral to the design of the chart, VSUPs appear to promote more cautious decisions.
 
@@
 In general, VSUPs excel in scenarios where we wish to make firm guarantees about the discriminability of different marks in a chart, but must also represent, in as much detail as possible, two, potentially orthogonal variables of interest.
-VSUPs have improved data resolution over na\"ive bivariate maps by assuming that uncertain values ought to have less visual impact that highly certain values.
+VSUPs have improved data resolution over other bivariate maps by assuming that uncertain values ought to have less visual impact that highly certain values.
 VSUPs therefore act both as a kind of filtering device as well as a bivariate representation.

### 2017-09-14 study/src/script.js

@@ -31,6 +31,18 @@ var main = d3.select("#fcontainer");
 var done = false;
 var questionNum = 1;
 
+var types =
+[
+  {binned: "discrete", shape:"square", vsum:false},
+//    {binned: "continuous", shape:"square", vsum:false},
+  {binned: "discrete", shape:"juxtaposed", vsum:false},
+//    {binned: "continuous", shape:"juxtaposed", vsum:false},
+//    {binned: "discrete", shape:"arc", vsum:false},
+//    {binned: "continuous", shape:"arc", vsum:false},
+//    {binned: "discrete", shape:"square", vsum:true},
+  {binned: "discrete", shape:"arc", vsum:true}
+];
+
 var taskOneStimuli = [];
 var taskTwoStimuli = [];
 
@@ -270,18 +282,6 @@ function finishTutorialTwo() {
 function makeTaskOneStimuli() {
   var stimuli = [];
 
-  var types =
-  [
-    {binned: "discrete", shape:"square", vsum:false},
-//    {binned: "continuous", shape:"square", vsum:false},
-    {binned: "discrete", shape:"juxtaposed", vsum:false},
-//    {binned: "continuous", shape:"juxtaposed", vsum:false},
-//    {binned: "discrete", shape:"arc", vsum:false},
-//    {binned: "continuous", shape:"arc", vsum:false},
-//    {binned: "discrete", shape:"square", vsum:true},
-    {binned: "discrete", shape:"arc", vsum:true}
-  ];
-
   var replicates = 8;
 
   for (var type of types) {
@@ -297,19 +297,7 @@ function makeTaskOneStimuli() {
 function makeTaskTwoStimuli() {
   var stimuli = [];
 
-  var types =
-  [
-    {binned: "discrete", shape:"square", vsum:false},
-    {binned: "continuous", shape:"square", vsum:false},
-    {binned: "discrete", shape:"juxtaposed", vsum:false},
-    {binned: "continuous", shape:"juxtaposed", vsum:false},
-    {binned: "discrete", shape:"arc", vsum:false},
-    {binned: "continuous", shape:"arc", vsum:false},
-    {binned: "discrete", shape:"square", vsum:true},
-    {binned: "discrete", shape:"arc", vsum:true}
-  ];
-
-  var replicates = 4;
+  var replicates = 6;
 
   for (var type of types) {
     for (var i = 0;i < replicates;i++) {

### 2017-09-14 study/src/script.js
Make VSUM field non-boolean for writing

@@ -530,6 +530,7 @@ function answerTaskOne() {
   answerData.vError = vError;
   answerData.uError = uError;
   answerData.error = vError + uError;
+  answerData.vsum = stim.vsum ? "yes" : "no";
 
   delete answerData.question;
   writeAnswerTaskOne(answerData);
@@ -774,6 +775,7 @@ function answerTaskTwo() {
   answerData.meanU = dl.mean(data, "u");
   answerData.stdV = dl.stdev(data, "v");
   answerData.stdU = dl.stdev(data, "u");
+  answerData.vsum = stim.vsum ? "yes" : "no";
 
   writeAnswerTaskTwo(answerData);
 }

### 2017-09-14 study/src/script.js
Linting

@@ -157,7 +157,7 @@ function consent() {
     readyBtn.text("PREVIEW");
   }
 
-  window.scrollTo(0,0);
+  window.scrollTo(0, 0);
 
 }
 
@@ -237,7 +237,7 @@ function tutorial() {
   main.selectAll("*").remove();
 
   main.append("iframe")
-  .attr("height","100%")
+  .attr("height", "100%")
   .attr("src", "tutorial.html");
 
   main.append("button")
@@ -248,7 +248,7 @@ function tutorial() {
   .text("Ready")
   .on("click", finishTutorial);
 
-  window.scrollTo(0,0);
+  window.scrollTo(0, 0);
 }
 
 function finishTutorial() {
@@ -769,8 +769,8 @@ function answerTaskTwo() {
   }
 
   d3.select("#tokenBar").selectAll("image").each(function(d) { data.push({"v":d.v, "u":d.u});});
-  answerData.Vs = data.map(function(obj) { return obj.v;}).toString();
-  answerData.Us = data.map(function(obj) { return obj.u;}).toString();
+  answerData.Vs = "\"" + data.map(function(obj) { return obj.v;}).toString() + "\"";
+  answerData.Us = "\"" + data.map(function(obj) { return obj.u;}).toString() + "\"";
   answerData.meanV = dl.mean(data, "v");
   answerData.meanU = dl.mean(data, "u");
   answerData.stdV = dl.stdev(data, "v");

### 2017-09-14 paper/main.tex
Prepping for experiment one

@@
 \section{Introduction}
+\exampleFig
 Uncertainty is an inescapable component of collecting, presenting, and using data.
@@
 
-\exampleFig
 \section{Related Work}
@@
 %TODO finalize factor levels, regenerate conditions figure
-The experiment was a within-subjects design with 1 factor (encoding type) with $N$ levels.
+The experiment was a within-subjects design with 1 factor (encoding type) with $3$ levels.
 \figref{fig:conditions} shows examples of these factor levels.
-We had 8 replications for each level for the identification task, and 4 for the prediction task, for a total of $N$ stimuli for the first task, and $N$ for the second.
+We had 8 replications for each level for the identification task, and 6 for the prediction task, for a total of $24$ stimuli for the first task, and $18$ for the second.
 We discuss the two experimental tasks in more detail in the following sections.

### 2017-09-14 paper/main.tex
Update abstract and intro.

--- 2017-09-14 paper/main.tex
@@
 \section{Introduction}
-Uncertainty is an inescapable component of collecting, presenting, and using data.
+Uncertainty is an inescapable component of collecting, analyzing, and presenting data.
 A common goal in the communication of uncertainty is promoting \emph{uncertainty-aware decisions}: the audience should be aware of the risks and rewards of certain decisions, modulate their confidence in their conclusions, and perhaps refrain from making a decision at all if there is too much uncertainty.
@@
 Simultaneous presentation of uncertainty and value necessitates the construction of a bivariate map\,---\,a relation, in terms of visual variables, between 2-tuples $(\text{value}, \text{uncertainty})$ and mark properties.
-The design of bivariate maps is difficult and, due to the interference and interplay between different visual variables, often leads to maps with few discriminable categories.
+Due to the interference and interplay between different visual variables, bivariate maps may suffer from limited discriminability.
 
 In this paper, we contribute \textbf{Value-Suppressing Uncertainty Palettes} (VSUPs) for integrating data and uncertainty information in visualizations.
-VSUP intentionally alias together data values with high uncertainty, affording greater discriminability as uncertainty decreases.
-If traditional bivariate maps are a two-dimensional square, with differing outputs for each combination of value and uncertainty, VSUP can be thought of as arcs: as uncertainty increases, values are mapped to smaller and smaller sets of outputs, culminating in a singularity where all inputs are mapped to an identical, highly uncertain mark regardless of their data value.
-\figref{fig:example} shows examples of both a more traditional bivariate map and a VSUP.
+VSUPs intentionally alias together data values with high uncertainty, affording greater discriminability as uncertainty decreases.
+Traditional bivariate maps might be thought of as a 2D square, with differing outputs for each combination of value and uncertainty.
+In contrast, VSUPs can be conceptualized as arcs: as uncertainty increases, values are mapped to smaller and smaller sets of outputs, culminating in a singularity where all inputs are mapped to an identical, highly uncertain mark regardless of data value.
+\figref{fig:example} shows examples of both a traditional bivariate map and a VSUP.
 
-We describe the motivations behind the use of VSUPs, examples of their utility for decision-making under uncertainty, and assess VSUPs in a crowdsourced experiment.
-Our experimental results indicate that VSUPs create close integration between uncertainty and data, promoting rapid but cautious decision-making.
+We describe the motivations for VSUPs, provide examples of their utility for decision-making under uncertainty, and assess VSUPs in a crowdsourced experiment.
+Our results indicate that VSUPs create close integration between uncertainty and data, promoting rapid but cautious decision-making.
 
 \exampleFig
+
 \section{Related Work}

### 2017-09-14 paper/main.tex
Resolve merge conflict.

@@
 \section{Introduction}
+
 Uncertainty is an inescapable component of collecting, analyzing, and presenting data.
+A common goal in the communication of uncertainty is promoting \emph{uncertainty-aware decisions}: the audience should be aware of the risks and rewards of certain decisions, modulate their confidence in their conclusions, and perhaps refrain from making a decision at all if there is too much uncertainty.
+A way that designers can contribute to this goal is by ensuring that uncertainty information is \emph{well-integrated} with the rest of the data.
+That is, it should be difficult to discount or ignore the uncertainty in a dataset.
+
+\exampleFig
+
+Uncertainty is an inescapable component of collecting, presenting, and using data.
 A common goal in the communication of uncertainty is promoting \emph{uncertainty-aware decisions}: the audience should be aware of the risks and rewards of certain decisions, modulate their confidence in their conclusions, and perhaps refrain from making a decision at all if there is too much uncertainty.
@@
 Our results indicate that VSUPs create close integration between uncertainty and data, promoting rapid but cautious decision-making.
-
-\exampleFig
 

### 2017-09-14 modules/src/continuous.js

@@ -1,8 +1,9 @@
 import * as d3 from "d3";
 
-export function continuousSquare(m_size, m_scale){
+export function continuousSquare(m_size, m_scale, m_id){
   var el = null,
     size = m_size,
+    id = m_id,
     scale = m_scale,
     context,
     canvas;
@@ -42,6 +43,10 @@ export function continuousSquare(m_size, m_scale){
     canvas
       .attr("width", size)
       .attr("height", size);
+
+    if (id) {
+      canvas.attr("id", id);
+    }
 
     var cnode = canvas.node();
     context = cnode.getContext("2d");
@@ -83,6 +88,19 @@ export function continuousSquare(m_size, m_scale){
       }
       else {
         square.make();
+      }
+      return square;
+    }
+  }
+
+  square.id = function(newId) {
+    if (!arguments.length) {
+      return id;
+    }
+    else {
+      id = newId;
+      if (canvas) {
+        canvas.attr("id", id);
       }
       return square;
     }

### 2017-09-14 study/src/script.js

@@ -31,18 +31,6 @@ var main = d3.select("#fcontainer");
 var done = false;
 var questionNum = 1;
 
-var types =
-[
-  {binned: "discrete", shape:"square", vsum:false},
-//    {binned: "continuous", shape:"square", vsum:false},
-  {binned: "discrete", shape:"juxtaposed", vsum:false},
-//    {binned: "continuous", shape:"juxtaposed", vsum:false},
-//    {binned: "discrete", shape:"arc", vsum:false},
-//    {binned: "continuous", shape:"arc", vsum:false},
-//    {binned: "discrete", shape:"square", vsum:true},
-  {binned: "discrete", shape:"arc", vsum:true}
-];
-
 var taskOneStimuli = [];
 var taskTwoStimuli = [];
 
@@ -61,6 +49,19 @@ var taskTwoTokens = 0;
 // 7. VSUM square
 // 8. VSUM arc
 
+var types =
+[
+  {binned: "discrete", shape:"square", vsum:false},
+  {binned: "continuous", shape:"square", vsum:false},
+  {binned: "discrete", shape:"juxtaposed", vsum:false},
+  {binned: "continuous", shape:"juxtaposed", vsum:false},
+  {binned: "discrete", shape:"arc", vsum:false},
+  {binned: "continuous", shape:"arc", vsum:false},
+  {binned: "discrete", shape:"square", vsum:true},
+  {binned: "discrete", shape:"arc", vsum:true}
+];
+
+//quantization schema
 var Qvsum = bvu.quantization().branching(2).layers(4);
 var Q2d = bvu.squareQuantization().n(4);
 
@@ -282,7 +283,7 @@ function finishTutorialTwo() {
 function makeTaskOneStimuli() {
   var stimuli = [];
 
-  var replicates = 8;
+  var replicates = experiment=="Exp1" ? 8 : 4;
 
   for (var type of types) {
     for (var i = 0;i < replicates;i++) {
@@ -559,7 +560,12 @@ function doneAnswerTaskOne() {
   d3.select("#questionNum").html(questionNum);
 
   if (done) {
-    tutorialTwo();
+    if (experiment=="Exp1") {
+      finishTask();
+    }
+    else {
+      tutorialTwo();
+    }
   }
   else {
     initializeTaskOne();

### 2017-09-14 paper/main.tex
Revisions / comments for related work.

@@
 There are many methods for visualizing uncertainty.
-MacEachren et al.
-~\cite{maceachren1992visualizing} presents a theoretical survey of uncertainty visualization generally.
-Relevant to our domain of heatmaps and thematic maps is work in GIS and cartography; for a survey of uncertainty visualization techniques in GIS, see Kinkeldey et al.
-\cite{kinkeldey2014assess,kinkeldey2017evaluating}.
-Yet, despite these wide variety of methods, to our knowledge, there are no established standards for visualizing uncertainty in heatmaps.
+MacEachren \ea~\cite{maceachren1992visualizing} present a theoretical survey of uncertainty visualization.
+Relevant to our domain of heatmaps and thematic maps \textbf{JH: where did this come from?
+There is no mention of focusing on this domain in the intro.
+The whole carto/GIS stuff seems to come out of left field.} is work in GIS and cartography; for a survey of uncertainty visualization techniques in GIS, see Kinkeldey \ea~\cite{kinkeldey2014assess,kinkeldey2017evaluating}.
+Yet, despite a wide variety of methods, to our knowledge, there are no established standards for visualizing uncertainty in heatmaps.
 
@@
 
-One challenge to conveying uncertainty information is that the decision to explicitly encode uncertainty increases the dimensionality of the data, and so requires the use of (at least one) additional visual channel ~\cite{brodlie2012review}.
+One challenge of uncertainty visualization is that the decision to explicitly encode uncertainty increases the dimensionality of the data, and so requires the use of (at least one) additional visual channel~\cite{brodlie2012review}.
 Uncertainty therefore inherently increases the visual complexity of a visualization.
 When the data are already complex to convey, and many of the more common or accurate visual variables are in use, allocating an additional dimension is non-trivial.
-As the number of dimensions increases, finding visual variables that are both perceptually accurate (in either estimation of quantity or discrimination of category) as well as perceptually separable from all the other encoding channels, becomes more and more difficult.
+As the number of dimensions increases, finding visual variables that are both perceptually accurate (in either estimation of quantity or discrimination of category) as well as perceptually separable from all the other encoding channels, becomes increasingly difficult.
 
@@
 Unfortunately, Boukhelifa \ea~\cite{boukhelifa2012evaluating} find that many visual variables habitually used for conveying uncertainty, such as blur and value, are also difficult to estimate.
-This results in a preference/performance gap where designers must choose between encoding uncertainty in a way that is intuitive but error prone, or use higher fidelity channels that may be more difficult to interpret.
+This results in a preference/performance gap where designers must choose between encoding uncertainty in a way that is intuitive but error prone, or use higher fidelity channels that may not intuitively convey uncertainty.
 
@@
 In these situations, data value is often encoded using color.
-Colors in such univariate quantitative color maps should be sufficiently far apart as to be perceptually distinguishable~\cite{ware1988color}, and vary in lightness as well as hue to afford an implicit ordering of value~\cite{borland2007rainbow,rogowitz2001blair}.
-Different choices of color maps can highlight different features of the data, and should be chosen with care based on the data types and important features to be visualized~\cite{rogowitz1996not}.
+Colors in univariate quantitative color maps should be sufficiently far apart as to be perceptually distinguishable~\cite{ware1988color}, and vary in lightness as well as hue to afford an implicit ordering of value~\cite{borland2007rainbow,rogowitz2001blair}.
+Different choices of color maps can highlight different features of the data, and should be chosen with care~\cite{rogowitz1996not}.
 
-A design consideration when constructing color maps is whether to quantize data into a discrete set of colors, or to encode data using a continuous mapping from value to color.
+A recurring design consideration when constructing color maps is whether to quantize data into a discrete set of colors, or to encode data using a continuous mapping.
 While continuous color maps afford greater fidelity in presenting values~\cite{muller1979perception}, non-linearity in human color perception introduces errors in extracting numeric values from continuous colors~\cite{borland2007rainbow}.
 Quantizing a color map is therefore an exercise in balancing \emph{perceptual} error and \emph{quantization} error~\cite{dobson1973choropleth}.
-Discrete maps offer a finer control over this balance, which can result in better performance in tasks involving heatmaps~\cite{padilla2017evaluating}.
+Discrete maps offer finer control over this balance, which can result in better performance in tasks involving heatmaps~\cite{padilla2017evaluating}.
 As such, we limit our discussion to color maps with discrete sets of colors.
@@
 
-While uncertainty can be visualized entirely separately from the rest of the data (for instance, in a juxtaposed chart), this runs the risk that uncertainty information becomes \emph{ignorable}\cite{moritz2017trust}.
+While uncertainty can be visualized entirely separately from the rest of the data (for instance, in a juxtaposed chart), this runs the risk that uncertainty information becomes \emph{ignorable}~\cite{moritz2017trust}.
 It also complicates analysis, as interpreting the value and uncertainty of a given data point requires consulting two separate charts.
-For our goal on integrating uncertainty information into analysis, we focus instead on the construction of \emph{bivariate maps}, where value and uncertainty information is displayed simultaneously in the same visualization.
+To integrate uncertainty information into an analysis, we focus on the construction of \emph{bivariate maps}, where value and uncertainty information is displayed simultaneously in a spatially co-located manner.
 
 Bivariate maps can, in principle, be constructed from the combination of any two visual variables (such as shape and color, or size and texture).
-E.g., Ware's ``textons'' overlay glyphs on top of regions to simultaneously encode two quantitative values~\cite{ware2009quantitative}.
+For example Ware's ``textons''~\cite{ware2009quantitative} overlay glyphs on top of regions to simultaneously encode two quantitative values.
 However, in this work, our focus is on heatmaps and other sorts of thematic maps.
-In visualizations like heatmaps, common visual channels such as position and size are already in use to encode other information (such as spatial location and extent), and color is reserved for value.
-Bivariate maps in these settings therefore typically rely on a visual channel related to color for encoding a secondary variable, such as alpha~\cite{roth2010value} or pixel noise~\cite{lucchesi2017visualizing}.
+In visualizations like heatmaps, common visual channels such as position and size are already in use to encode other information (such as spatial location and extent), and color is reserved for value \textbf{JH: this is a bit redundant with the previous subsection on color maps.
+It feels a bit artificial having these two sections separate.
+Color is a subset of bivariate map options...
+and we start talking about it again in the next paragraph.}.
+Bivariate maps in these settings therefore typically rely on a visual channel related to color for encoding a secondary variable, such as opacity (alpha)~\cite{roth2010value} or pixel noise~\cite{lucchesi2017visualizing}.
 
-Due to the perceptual integrality of visual channels~\cite{garner1970integrality, ware2012information}, the construction of color-based bivariate maps is difficult.
-Viewers need to be able to independently assess the variables of interest, but the resulting bivariate map should also be expressive and easy to interpret.
+Due to the perceptual integrality of visual channels~\cite{garner1970integrality, ware2012information}, \textbf{JH: We mention integrality but never give a brief definition or explanation.
+I wouldn't assume that all CHI readers are familiar with this.} the construction of color-based bivariate maps is difficult.
+Viewers need to be able to independently assess the variables of interest \textbf{JH: They do?
+I thought the whole point of VSUPs is to confound indepenent assessment of value!}, but the resulting bivariate map should also be expressive and easy to interpret.
 Bivariate maps with these constraints are often limited to relatively few categories (say, a 4x4 matrix as in \figref{fig:example}) \cite{robertson1986generation,trumbo1981theory}.
-For VSUPs, we adapt an insight from Dunn~\cite{dunn1989dynamic} that these limited categories ought to be assigned with regards to their importance to analytics tasks, rather than uniformly.
+For VSUPs, we adapt an insight from Dunn~\cite{dunn1989dynamic} that these limited categories ought to be assigned with regards to their importance to analysis tasks, rather than uniformly.
 
 In general, the quality of bivariate color maps is a multivariate measure involving consideration of not just the component color channels, but also the interpolation scheme and the color of the surround~\cite{bernard2015survey}.
-However, even simple bivariate maps can be difficult for a general audience to interpret \cite{wainer1980empirical}.
+However, even simple bivariate maps can be difficult for a general audience to interpret~\cite{wainer1980empirical} \textbf{JH: It seems there is a lot ``hiding'' behind this citation.
+In what ways is it hard?
+In what ways do people fail?
+Any more flavor here, even if very brief?
+And might this motivate something akin to VSUPs?}.
+
+\textbf{JH: This would be a natural place to put some connective prose that transitions to VSUPs...}
 

### 2017-09-14 paper/main.tex
Revise / comment VSUP design section.

@@
 \begin{enumerate}
-	\item There is a limited \emph{budget} of potential outputs in a bivariate map.
-	\item Differences among \emph{certain} data are more significant than differences among \emph{uncertain} data.
+	\item There is a limited \emph{budget} for perceptual discriminability.
+	\item Differences among \emph{certain} data are more germane than differences among \emph{uncertain} data.
 \end{enumerate}
@@
 VSUPs codify this decision by \emph{reducing the number of mark categories for representing value as uncertainty increases.} This means that data encoded using a VSUP will make visible only the largest of differences in uncertain data, but highlight comparatively small differences in value when uncertainty is low.
-VSUPs therefore act as both a filtering mechanism (in that values with too much uncertainty are all mapped to the same glyph), as well as an implicit test of effect size (in that smaller and smaller changes in value are visible as uncertainty decreases).
-This strategy of dampening the low quality or high uncertainty values in maps in order to focus on more informative regions has measurable benefits, including the removal of statistically spurious visual patterns, and the highlighting of regions of interest~\cite{correll2017surprise}.
+VSUPs therefore act as both a filtering mechanism (in that values with too much uncertainty are mapped to the same glyph), as well as an implicit test of effect size (in that smaller and smaller changes in value are visible as uncertainty decreases).
+This strategy of dampening low quality or highly uncertain values in maps in order to focus on more informative regions has measurable benefits, including the removal of statistically spurious visual patterns, and the highlighting of regions of interest~\cite{correll2017surprise}.
 
 VSUPs rely on an underlying \textbf{quantization tree} that governs how values are discretized.
-Maximally uncertain values are all mapped to a singular ``root'' node.
+Maximally uncertain values are mapped to a singular ``root'' node.
 As uncertainty decreases, the tree branches into leaves, which evenly divide the data domain.
 The data value of a parent is the midpoint of all of its children.
-These leaves can then branch again, until a designer-specified stopping point.
-The layers act to the quantize the uncertainty domain, and the nodes the data domain.
-\figref{fig:example}, for example, is a VSUP with a tree with a branching factor of 2, and 4 layers, resulting in a total of 15 different output colors.
+These leaves can then branch again, up to a designer-specified stopping point.
+The layers quantize the uncertainty domain, and sibling nodes quantize the data domain.
+\figref{fig:example} shows a VSUP with a tree with a branching factor of 2 and 4 layers, resulting in an output range with 15 distinct colors.
 
-Correll \ea~\cite{correll2015layercake,correll2011visualizing} use a precursor of VSUPs in their LayerCake genomics visualization tool, where marks representing genomic data with increasing uncertainty are mapped to a smaller and smaller set of increasingly grey colors, creating the effect of uncertain values retreating into ``confidence fog'' while highly certain values remain prominent.
+Correll \ea~\cite{correll2015layercake,correll2011visualizing} use a precursor of VSUPs in their LayerCake genomics visualization tool, where marks representing genomic data with increasing uncertainty are mapped to a smaller and smaller set of increasingly grey colors, creating the effect of uncertain values retreating into a ``confidence fog'' while highly certain values remain prominent.
 Other bivariate visualizations implicitly alias together uncertain values.
@@
 
-Javascript code for generating VSUPs for use in D3 charts is available at \url{URL-REMOVED-FOR-REVIEW}.
+\textbf{JH: The precending paragraph might be better placed in related work!}
+
+JavaScript code for generating VSUPs for use in D3 charts is available at \url{URL-REMOVED-FOR-REVIEW}.
 
@@
 There are multiple choices that designers must make before creating a VSUP.
-In particular, they must choose 1) which visual channels map to value and uncertainty, 2) structure of the underlying VSUP tree.
+Specifically, they must choose (1) which visual channels map to value and uncertainty and (2) the structure of the underlying VSUP tree.
 
@@
 In heatmaps, for instance, value is regularly encoded using color, while vertical and horizontal position are reserved for location.
-In this paper, we primarily deal with the case where value is encoded using color.
+\textbf{JH: This is the third time we've pointed this out...} In this paper, we primarily deal with the case where value is encoded using color.
 While there are some similar conventions for representing uncertainty, generally designers have more freedom in choosing a visual channel for uncertainty~\cite{maceachren1992visualizing}.
-We recommend channels with both a strong semiotic connection to uncertainty as well as a relatively large number of perceptually distinguishable levels.
+\textbf{This feels like we are revisiting introductory and related work material again!} We recommend channels with both a strong semiotic connection to uncertainty as well as a relatively large number of perceptually distinguishable levels.
 Here, we focus on value/saturation (or, equivalently, transparency against a white background).
-Correll \& Gleicher~\cite{correll2013error} show that even audiences without statistical backgrounds can interpret uncertainty information encoded in these channels.
+Correll \& Gleicher~\cite{correll2013error} provide evidence that even audiences without statistical training can successfully interpret uncertainty information encoded in these channels.
+
+\textbf{We should say something about the choice of value color map.
+Generally speaking, single-hue luminance ramps must be used with caution if they ramp too strongly towards white.
+This will be confused with manipulations of opacity or saturation.
+Viridis might be attractive due to its avoidance of desaturated colors.}
 
@@
 In fact, these regions have systematically low coverage; the lack of reliable data for these regions should discourage analysts from giving their apparent pattern equal consideration from the  stronger signals earlier in the genome.
+
+\textbf{A few thoughts about this section based on the current figures.
+(1) Fig.
+3 in particular has two graphics that look pretty similar.
+Is there anything we should do to draw attention to the most salient differences between conditions?
+At minimum, we have the text, but we could consider annotations or even a third vis that is a heatmap of differences between charts.
+(2) A third example using a different visual form might be useful.
+For example, a node-link network with value + uncertainty associated with each node.}
 

### 2017-09-14 study/src/script.js
Remove event pile ups.

@@ -514,6 +514,10 @@ function revealTaskOne() {
 
 
 function answerTaskOne() {
+  //prevent answers from piling up.
+  d3.select("#map").selectAll("rect")
+    .on("click", null);
+
   var rt = (new Date()).getTime() - startTime;
   var d = d3.select(this).datum();
   var stim = taskOneStimuli[questionNum - 1];

### 2017-09-14 study/src/script.js
Remove double-firing of task two events

@@ -769,6 +769,7 @@ function removeToken(d) {
 }
 
 function answerTaskTwo() {
+  d3.select("#answer").attr("disabled", "disabled");
   var data = [];
   var stimProp;
   var rt = (new Date()).getTime() - startTime;

### 2017-09-16 paper/main.tex
Updated evaluation section to reflect what we actually ran.

@@
 \section{Introduction}
+\exampleFig
 
@@
 That is, it should be difficult to discount or ignore the uncertainty in a dataset.
-
-\exampleFig
 
@@
 %TODO Add in discussion of tree structures, including figure showing that VSUPs have colors that are further apart than similar 2d bivariate maps.
-Also wortth discussing discretization issues here.
+Also worth discussing discretization issues here.
 Make new performance fig.
@@
 
-We performed a crowdsourced experiment on Amazon's Mechanical Turk to evaluate the effectiveness of VSUPs for integrating uncertainty and value information in visualizations.
+We performed two crowdsourced experiments on Amazon's Mechanical Turk to evaluate the effectiveness of VSUPs for integrating uncertainty and value information in visualizations.
 This focus on integration meant that we limited our experimental tasks to scenarios where the participants needed to consider both value and uncertainty before making a decision.
-We gave participants two main tasks:
+The experiments consisted of two separate tasks:
 
 \begin{enumerate}
-	\item An \textbf{identification} task, where we gave participants charts with value and uncertainty information, and asked them to locate specific regions.
+	\item An experiment with an \textbf{identification} task, where we gave participants charts with value and uncertainty information, and asked them to locate specific regions.
 E.g., ``click on the region of the chart with a value of $0.1$ and an uncertainty of $0.2$.''
-	\item A \textbf{prediction} task, inspired by the game Battleship, where we gave participants charts with both \emph{forecast} and \emph{forecast uncertainty} information, and asked the participants to place tokens on the board in order to optimize expected value.
+	\item An experiment with a \textbf{prediction} task, inspired by the game Battleship, where we gave participants charts with both \emph{forecast} and \emph{forecast uncertainty} information, and asked the participants to place tokens on the board in order to optimize expected value.
 E.g., ``place your $5$ ships on safe locations on the board.''
@@
 
-%TODO finalize factor levels, regenerate conditions figure
-The experiment was a within-subjects design with 1 factor (encoding type) with $3$ levels.
-\figref{fig:conditions} shows examples of these factor levels.
-We had 8 replications for each level for the identification task, and 6 for the prediction task, for a total of $24$ stimuli for the first task, and $18$ for the second.
-We discuss the two experimental tasks in more detail in the following sections.
+We limited our population to Turkers from the United States, with a prior task approval rate of at least 90\%.
+As the experimental tasks required multi-hue color perception, we presented participants with a set of Ishihara plates~\cite{hardy1945tests} as a pretest, and excluded participants who either misidentified the values in the plates, or who self-reported as having a color vision deficiency (CVD) in the post-test.
+Based on piloting, we paid participants \$2 dollars for participation, for a target rate of \$8/hour.
+After completion of the main tasks, we solicited demographic information, including a risk aversion assay from Mandrik \& Bao~\cite{mandrik2005exploring}.
+We recruited a total of $48$ participants for this experiment ($24$ for each experiment): $N$ female, $N$ male, and $N$ who declined to state ($M_{\text{age}}$= $N$, $SD_{\text{age}}$ = $N$).
+Participants from the first experiment were excluded from participating in the second.
+
+Rather than standard means, Cleveland \& McGill~\cite{cleveland1984graphical} use 25\% trimmed means to report effect sizes in graphical perception tasks.
+Trimmed means remove the first and last quartiles and then compute the mean from the remaining points.
+This filtering is important for graphical perception tasks (especially crowdsourced tasks) where error distributions may be long-tailed~\cite{heer2010crowdsourcing}.
+Trimmed means violate the sampling assumptions required for many null-hypothesis significance tests.
+Thus, while we report effect sizes in terms of trimmed means, we perform inferential statistical tests on the standard means.
+Asterisks in our charts denote significant differences, as reported by a post-hoc Tukey's test of Honest Significant Difference (HSD).
+
 Experimental materials, including data tables and stimuli generation code, are available at \url{URL-REMOVED-FOR-REVIEW}.
 
-\subsection{Identification Task}
+\subsection{Identification Experiment}
 
-For the identification task, we gave participants a heatmap, and asked them to click on a region of the heatmap that encodes a particular value,uncertainty pair.
+We wished to determine what effects, if any, our various design considerations had on performance at a basic analysis task: using a bivariate legend to interpret a heatmap.
+There are a number of design decisions to be made even after one has determined which two visual channels will be used to encode value and uncertainty.
+In particular:
+\begin{enumerate}
+	\item Whether or not to \emph{juxtapose} or \emph{superimpose} the data and uncertainty maps.
+	\item Whether or not to \emph{discretize} each visual channel, or encode values \emph{continuously}.
+	\item Whether or not to use a \emph{square} legend, which gives equal area to each bin, or a \emph{wedge} legend, which reduces in size as uncertainty increases.
+	\item If discretizing, whether to employ a \emph{VSUP} or a \emph{standard} quantization scheme.
+\end{enumerate}
+
+For the identification task, we gave participants a $5$x$5$ heatmap, and asked them to click on a region of the heatmap that encodes a particular value,uncertainty pair.
 To support comparability across conditions, we chose a set of target pairs such that each target mapped to a unique color across all encoding types.
@@
 
-\subsection{Prediction Task}
+Each of the 24 participants saw the data encoded as each of 8 types of bivariate encoding and legend, with 8 replications, for a total of 64 stimuli. \figref{fig:conditions} shows these factor levels.
+
+\subsubsection{Hypotheses}
+
+We had two initial hypotheses, based on prior work on bivariate maps:
+\begin{enumerate}
+	\item Juxtaposed maps, by introducing a second search task to the identification task (searching for the proper values in two, rather than one map), would have poorer performance than the other conditions.
+	\item Continuous maps required more specific and accurate color matching between value and color, and so would have poorer performance than the other conditions.
+\end{enumerate}
+
+We had no strong hypotheses about the shape of the legend (square vs.
+wedge), but, as the wedge shape makes more uncertain bins in the legend smaller (and thus potentially harder to disambiguate), we included it as a check against VSUPs which explicitly alias uncertain values.
+For this identification task, which did not involve reasoning with uncertainty, but simply matching colors to a color legend, we also did not have strong hypotheses about performance differences between VSUPs and traditional discrete bivariate maps.
+
+\subsubsection{Results}
+
+\subsection{Prediction Experiment}
 
@@
 Greis \ea~\cite{greis2016decision} employ these game-like experimental tasks to assess how different visual designs communicate uncertainty information, which can be abstract or complex, to the general audience.
-In our task, the participant and a (fictional) adversary have to place tokens representing ships on a spatial grid, with the expectation that certain squares will be hit by missiles.
+In our task, the participant and a (fictional) adversary have to place tokens representing ships on a $5$x$5$ spatial grid, with the expectation that certain squares will be hit by missiles.
 Players have to place all their tokens before continuing.
@@
 In our task, participants were given a map representing the predictions of missile strikes in each location on the grid.
-The \emph{value} component of the prediction was the likelihood of a missile strike in a particular region.
+The \emph{value} component of the prediction was the likelihood of a missile strike in a particular region: the ship's \emph{safety} if placed on the square.
 The \emph{uncertainty} component was the confidence in this prediction.
+Other studies of uncertainty representation, such as in Cox et al.~\cite{cox2013visualizing}, have used "prediction + prediction uncertainty" stimuli to elicit differences in decision-making between visualizations of uncertainty.
 
-%TODO stimuli creation
+Our stimuli were created by randomly sampling from values that fell within each of the $16$ bins of the $4$x$4$ 2D bivariate color map.
+This resulted in 16 samples.
+The remaining 9 samples were distractors, with low safety and high certainty.
 
@@
 In particular, there is greater perceived value in avoiding large losses as opposed to striving for a large gain~\cite{kahneman1979prospect}.
+
 The ideal strategy from a value-maximizing standpoint would be to place tokens on areas with the highest expected value, ignoring the uncertainty information.
 However, as with roulette and other similar games of chance, the variability in expected value is relevant when considering where to place bets~\cite{mlodinow2009drunkard}.
-Over the short term, a player may seek out high expected value but risky locations, or, if risk averse, accept lower expected value but lower variability locations.
+A risky player would choose guesses with high value, regardless of the uncertainty of those points.
+A more conservative guesser might eschew high-risk, high-reward locations, resulting in a lower average value of guesses, but also lower uncertainty.
 We therefore measured the distribution of both \emph{value} and \emph{uncertainty} of the tokens placed by the participants.
 
-\subsection{Hypotheses}
+Our results from the prior study indicate indicate that discrete, non-juxtaposed maps outperformed the other bivariate maps we selected, so we limited our study to only 4 types: square and arc bivariate maps.
+These in turn were either VSUPs or traditional bivariate maps, for a 2 (square or arc legend) x 2 (VSUP or standard) factorial design, with 6 replications, for a total of 24 stimuli for each of our 24 participants.
+\figref{fig:conditions} shows these factor levels.
+Prior to the main task, we included a short replication of the identification task from the prior experiment (with 12 stimuli) for training purposes.
 
-Across both the identification and prediction tasks, we had several hypotheses, stemming from our belief that VSUPs promote better \emph{integration} between uncertainty and value information, and also encourage \emph{caution} by highlighting the ambiguity or untrustworthiness introduced by uncertain data.
+\subsubsection{Hypotheses}
+
+We had two hypotheses, stemming from our belief that VSUPs promote better \emph{integration} between uncertainty and value information, and also encourage \emph{caution} by highlighting the ambiguity or untrustworthiness introduced by uncertain data.
 In particular:
 \begin{enumerate}
-	\item Participants would be \textbf{faster and more accurate} when completing the identification task using a VSUP as compared to the other map types.
-Juxtaposed maps, by adding an additional search task to the lookup task, would have especially poor performance compared to VSUPs.
-	\item Participants would choose targets with \textbf{less uncertainty} in the prediction task when using a VSUP compared to the other maps types.
-This would result in a tradeoff where they would also choose targets with \textbf{less expected value} than the other map types.
-Again, juxtaposed maps, by increasing the difficulty of retrieving both variables simultaneously, would highlight this difference.
+	\item Participants would choose targets with \textbf{lower uncertainty} in the prediction task when using a VSUP. 
+	\item This would result in a tradeoff where they would also choose targets with \textbf{lower expected value}.
 \end{enumerate}
 
-\subsection{Results}
+As with the prior experiment, we had no strong hypotheses for square vs. wedge-shaped legends, but included both as a check against the potential implicit VSUP-like properties of wedge-shaped legends.
 
-Rather than standard means, Cleveland \& McGill~\cite{cleveland1984graphical} use 25\% trimmed means to report effect sizes in graphical perception tasks.
-Trimmed means remove the first and last quartiles and then compute the mean from the remaining points.
-This filtering is important for graphical perception tasks (especially crowdsourced tasks) where error distributions may be long-tailed~\cite{heer2010crowdsourcing}.
-Trimmed means violate the sampling assumptions required for many null-hypothesis significance tests.
-Thus, while we report effect sizes in terms of trimmed means, we perform inferential statistical tests on the standard means.
-Asterisks in our charts denote significant differences, as the reported by a post-hoc Tukey's test of Honest Significant Difference (HSD).
-
-\subsubsection{Participants}
-
-We limited our population to Turkers from the United States, with a prior task approval rate of at least 90\%.
-As the experimental tasks required multi-hue color perception, we presented participants with a set of Ishihara plates~\cite{hardy1945tests} as a pretest, and excluded participants who either misidentified the values in the plates, or who self-reported as having a color vision deficiency (CVD) in the post-test.
-Based on piloting, we paid participants \$3 dollars for participation, for a target rate of \$8/hour.
-After completion of the main tasks, we solicited demographic information.
-
-%TODO add participant info
-We recruited a total of $N$ participants for this experiment: $N$ female, $N$ male, and $N$ who declined to state ($M_{\text{age}}$= $N$, $SD_{\text{age}}$ = $N$).
-
-\subsubsection{Identification Task}
-
-%TODO add results and figures
-%\accuracyFig
-%\responseTimeFig
-
-\subsubsection{Prediction Task}
-
-%TODO add results and figures
-%\uncertaintyFig
-
-For the prediction task, where there is no dominating strategy, our primary measure was the distribution of guesses, in terms of both their average value, and their average uncertainty.
-A risky player would choose guesses with high value, regardless of the uncertainty of those points.
-A more conservative guesser might eschew high-risk, high-reward locations, resulting in a lower average value of guesses, but also lower uncertainty.
-Our hypothesis is that VSUPs, by communicating uncertainty information at the expense of value information, would promote more conservatives guesses than the other chart types.
+\subsubsection{Results}
 

### 2017-09-16 paper/main.tex
Added experiment 1 results.

@@
 
+We performed a repeated measures ANOVA on our results to measure the effect of juxtaposed versus superimposed maps, and continuous versus discrete legends, on accuracy.
+Participant ID was included as a random effect.
+Figure %TODO
+shows the performance of each of our 8 conditions.
+
+Our results support our first hypothesis.
+Juxtaposition was a significant effect ($F(1,166)=5.5$,$p=0.02$).
+A post-hoc t-test confirmed ($t=2.5$, $p=0.01$)that superimposed charts ($M=0.58$, $SD=0.5$) performed significantly better than juxtaposed ($M=0.51$, $SD=0.49$) ones.
+For our identification task, juxtaposition requires at least two search tasks (one to locate squares with the correct data value, another to locate ones with the correct uncertainty value), which introduces additional error compared to conditions where value and uncertainty are displayed simultaneously.
+
+Our results also support our second hypothesis.
+Discretization was a significant effect ($F(1,166)=30$,$p<0.01$).
+A post-hoc t-test confirmed ($t=6.1$, $p<0.01$) that charts with discrete bins ($M=0.63$, $SD=0.48$) performed significantly better than charts with continuous color maps ($M=0.47$, $SD=0.5$).
+The lack of quantization bias in continuous maps is countered by the perceptual error in precisely estimating value from color.
+Relying on a discrete set of output colors simplifies this task.
+
+We performed a second ANOVA amongst just the superimposed, discrete charts, to determine the effect of legend shape (wedge or square) and quantization scheme (VSUP or standard) on performance, with participant ID as a random factor.
+Neither the legend shape ($F(1,70)=0.03$, $p=0.084$) nor the quantization scheme ($F(1,70)=1.4$, $p=0.24$) were significant effects.
+
 \subsection{Prediction Experiment}
@@
 For the prediction task, we gave participants the rules of a game like Battleship.
-Greis \ea~\cite{greis2016decision} employ these game-like experimental tasks to assess how different visual designs communicate uncertainty information, which can be abstract or complex, to the general audience.
+Greis et al.
+\ea~\cite{greis2016decision} employ these game-like experimental tasks to assess how different visual designs communicate uncertainty information, which can be abstract or complex, to the general audience.
 In our task, the participant and a (fictional) adversary have to place tokens representing ships on a $5$x$5$ spatial grid, with the expectation that certain squares will be hit by missiles.
@@
 
-The ideal strategy from a value-maximizing standpoint would be to place tokens on areas with the highest expected value, ignoring the uncertainty information.
-However, as with roulette and other similar games of chance, the variability in expected value is relevant when considering where to place bets~\cite{mlodinow2009drunkard}.
-A risky player would choose guesses with high value, regardless of the uncertainty of those points.
+The ideal strategy from a value-maximizing standpoint would be to place tokens on areas with the highest expected value (lowest danger), ignoring the uncertainty information.
+However, as with roulette and other similar games of chance, the \emph{variability} in expected value is relevant when considering where to place bets~\cite{mlodinow2009drunkard}.
+A risky player would choose guesses with high expected value, regardless of the uncertainty of those points.
 A more conservative guesser might eschew high-risk, high-reward locations, resulting in a lower average value of guesses, but also lower uncertainty.
@@
 	\item Participants would choose targets with \textbf{lower uncertainty} in the prediction task when using a VSUP. 
-	\item This would result in a tradeoff where they would also choose targets with \textbf{lower expected value}.
+	\item This would result in a tradeoff where they would also choose targets with \textbf{lower expected value, but lower variance in expected value} when using a VSUP.
 \end{enumerate}

### 2017-09-16 study/data/analysis.js
Add experiment two data.

@@ -13,8 +13,10 @@ function process() {
 
   var demoTable = dl.csv("demo.csv");
   var dataOneTable = dl.csv("dataOne.csv");
+  var dataTwoTable = dl.csv("dataTwo.csv");
 
   var cleanedOne = [];
+  var cleanedTwo = [];
   var valids = [];
   var turkId;
   var row;
@@ -25,6 +27,9 @@ function process() {
     valids = dataOneTable.filter(function(x){ return x.workerId == turkId;});
     valids.forEach(function(d){ d.workerId = i+1;});
     cleanedOne = cleanedOne.concat(valids);
+    valids = dataTwoTable.filter(function(x){ return x.workerId == turkId;});
+    valids.forEach(function(d){ d.workerId = i+1;});
+    cleanedTwo = cleanedTwo.concat(valids);
   }
 
   for(var i = 0;i<cleanedOne.length;i++) {
@@ -36,7 +41,16 @@ function process() {
     writeRow(cleanedOne[i]);
   }
 
+  for(var i = 0;i<cleanedTwo.length;i++) {
+    cleanedTwo[i].condition = cleanedTwo[i].binned + cleanedTwo[i].shape;
+    if(cleanedTwo[i].vsum == "yes") {
+      cleanedTwo[i].condition = "vsum" + cleanedTwo[i].condition;
+    }
+    writeRow(cleanedTwo[i]);
+  }
+
   dataOne = cleanedOne;
+  dataTwo = cleanedTwo;
 }
 
 function writeRow(row) {

### 2017-09-16 paper/main.tex
Dealt with all JH edits except for the difficult one.

@@
 Uncertainty is an inescapable component of collecting, analyzing, and presenting data.
-A common goal in the communication of uncertainty is promoting \emph{uncertainty-aware decisions}: the audience should be aware of the risks and rewards of certain decisions, modulate their confidence in their conclusions, and perhaps refrain from making a decision at all if there is too much uncertainty.
-A way that designers can contribute to this goal is by ensuring that uncertainty information is \emph{well-integrated} with the rest of the data.
-That is, it should be difficult to discount or ignore the uncertainty in a dataset.
-
-Uncertainty is an inescapable component of collecting, presenting, and using data.
 A common goal in the communication of uncertainty is promoting \emph{uncertainty-aware decisions}: the audience should be aware of the risks and rewards of certain decisions, modulate their confidence in their conclusions, and perhaps refrain from making a decision at all if there is too much uncertainty.
@@
 We describe the motivations for VSUPs, provide examples of their utility for decision-making under uncertainty, and assess VSUPs in a crowdsourced experiment.
-Our results indicate that VSUPs create close integration between uncertainty and data, promoting rapid but cautious decision-making.
+Our results indicate that VSUPs create close integration between uncertainty and data, resulting in better performance at data tasks with uncertainty components.
 
@@
 
-There are many methods for visualizing uncertainty.
+%There are many methods for visualizing uncertainty.
 MacEachren \ea~\cite{maceachren1992visualizing} present a theoretical survey of uncertainty visualization.
@@
 
-\subsection{Color Maps}
-
-For spatial visualizations like choropleth maps, heatmaps, and treemaps, visual channels such as position and length are reserved for data variables other than value (such as geographic location or relative size).
-In these situations, data value is often encoded using color.
-Colors in univariate quantitative color maps should be sufficiently far apart as to be perceptually distinguishable~\cite{ware1988color}, and vary in lightness as well as hue to afford an implicit ordering of value~\cite{borland2007rainbow,rogowitz2001blair}.
-Different choices of color maps can highlight different features of the data, and should be chosen with care~\cite{rogowitz1996not}.
-
-A recurring design consideration when constructing color maps is whether to quantize data into a discrete set of colors, or to encode data using a continuous mapping.
-While continuous color maps afford greater fidelity in presenting values~\cite{muller1979perception}, non-linearity in human color perception introduces errors in extracting numeric values from continuous colors~\cite{borland2007rainbow}.
-Quantizing a color map is therefore an exercise in balancing \emph{perceptual} error and \emph{quantization} error~\cite{dobson1973choropleth}.
-Discrete maps offer finer control over this balance, which can result in better performance in tasks involving heatmaps~\cite{padilla2017evaluating}.
-As such, we limit our discussion to color maps with discrete sets of colors.
-
 \subsection{Bivariate Maps}
@@
 For example Ware's ``textons''~\cite{ware2009quantitative} overlay glyphs on top of regions to simultaneously encode two quantitative values.
-However, in this work, our focus is on heatmaps and other sorts of thematic maps.
-In visualizations like heatmaps, common visual channels such as position and size are already in use to encode other information (such as spatial location and extent), and color is reserved for value \textbf{JH: this is a bit redundant with the previous subsection on color maps.
-It feels a bit artificial having these two sections separate.
-Color is a subset of bivariate map options...
-and we start talking about it again in the next paragraph.}.
-Bivariate maps in these settings therefore typically rely on a visual channel related to color for encoding a secondary variable, such as opacity (alpha)~\cite{roth2010value} or pixel noise~\cite{lucchesi2017visualizing}.
+However, for spatial visualizations like choropleth maps, heatmaps, and treemaps, visual channels such as position and length are reserved for data variables other than value (such as geographic location or relative size).
+In these situations, data value is often encoded using color.
+Bivariate maps in these settings therefore typically rely on a visual channel related to color for encoding a secondary variable, such as opacity~\cite{roth2010value} or pixel noise~\cite{lucchesi2017visualizing}.
 
-Due to the perceptual integrality of visual channels~\cite{garner1970integrality, ware2012information}, \textbf{JH: We mention integrality but never give a brief definition or explanation.
-I wouldn't assume that all CHI readers are familiar with this.} the construction of color-based bivariate maps is difficult.
-Viewers need to be able to independently assess the variables of interest \textbf{JH: They do?
-I thought the whole point of VSUPs is to confound indepenent assessment of value!}, but the resulting bivariate map should also be expressive and easy to interpret.
-Bivariate maps with these constraints are often limited to relatively few categories (say, a 4x4 matrix as in \figref{fig:example}) \cite{robertson1986generation,trumbo1981theory}.
+Colors in univariate quantitative color maps should be sufficiently far apart as to be perceptually distinguishable~\cite{ware1988color}, and vary in lightness as well as hue to afford an implicit ordering of value~\cite{borland2007rainbow,rogowitz2001blair}.
+Different choices of color maps can highlight different features of the data, and should be chosen with care~\cite{rogowitz1996not}.
+These principles extend to bivariate color maps, with the additional consideration that our perception of color channels lacks \emph{orthogonality}.
+That is, while we can perceive e.g., the height of a bar in a bar chart independently of its width, a particular property of a color (e.g., its redness, saturation, transparency, etc.) affect how its other properties are perceived \cite{ware2012information,garner1970integrality}.
+Many color channels such as hue and saturation are therefore perceptually \emph{integral}, which complicates their use in bivariate maps.
+
+Another recurring design consideration when constructing color maps is whether to quantize data into a discrete set of colors, or to encode data using a continuous mapping.
+While continuous color maps afford greater fidelity in presenting values~\cite{muller1979perception}, non-linearity in human color perception introduces errors in extracting numeric values from continuous colors~\cite{borland2007rainbow}.
+Quantizing a color map is therefore an exercise in balancing \emph{perceptual} error and \emph{quantization} error~\cite{dobson1973choropleth}.
+Discrete maps offer finer control over this balance, which can result in better performance in tasks involving heatmaps~\cite{padilla2017evaluating}.
+
+In general, the quality of bivariate color maps is a multivariate measure involving consideration of not just the component color channels, but also the interpolation scheme and the color of the surround~\cite{bernard2015survey}.
+However, even simple bivariate maps can be difficult for a general audience to interpret: due to their additional complexity, Wainer \& Francolini~\cite{wainer1980empirical} reported high levels of error for participants even for ``elementary'' graph reading tasks, and additionally that bivariate legends were more difficult to memorize and internalize than their univariate equivalents.
+Therefore, bivariate maps in practice are often limited to a small set of output colors (say, a 4x4 matrix as in \figref{fig:example}) \cite{robertson1986generation,trumbo1981theory}.
 For VSUPs, we adapt an insight from Dunn~\cite{dunn1989dynamic} that these limited categories ought to be assigned with regards to their importance to analysis tasks, rather than uniformly.
 
-In general, the quality of bivariate color maps is a multivariate measure involving consideration of not just the component color channels, but also the interpolation scheme and the color of the surround~\cite{bernard2015survey}.
-However, even simple bivariate maps can be difficult for a general audience to interpret~\cite{wainer1980empirical} \textbf{JH: It seems there is a lot ``hiding'' behind this citation.
-In what ways is it hard?
-In what ways do people fail?
-Any more flavor here, even if very brief?
-And might this motivate something akin to VSUPs?}.
+Given that bivariate maps, for reasons of either practicality or perceptual fidelity, have only a limited budget of outputs, we adapt an insight from Dunn~\cite{dunn1989dynamic} that these limited categories ought to be assigned with regards to their importance to analysis tasks, rather than uniformly.
+Our Value-Suppressing Uncertainty Palettes apply this insight to data with uncertainty, where, frequently, uncertain information ought to have less valence on the decision-making process, or at least less prominence in the analysis, than certain information.
 
-\textbf{JH: This would be a natural place to put some connective prose that transitions to VSUPs...}
+Correll \ea~\cite{correll2015layercake,correll2011visualizing} use a precursor of VSUPs in their LayerCake genomics visualization tool, where marks representing genomic data with increasing uncertainty are mapped to a smaller and smaller set of increasingly grey colors, creating the effect of uncertain values retreating into a ``confidence fog'' while highly certain values remain prominent.
+Other bivariate visualizations implicitly alias together uncertain values through perceptual integrality.
+For instance, if uncertainty is encoded by transparency, a maximally uncertain glyph may be entirely transparent, and so impossible to distinguish from any other maximally uncertain glyph.
+Other channels with a semiotic connection to uncertainty, such as saturation, value, blur, or size, have similar deleterious effects on the disambiguation of colors and shapes.
+In both cases, the property of aliasing is \emph{ad hoc}, and places no guarantees on the discriminability of colors.
+The binning and degradation approach of VSUPs makes the choice to alias values explicit to both the designer and the viewer, and results in a bivariate mapping with known perceptual properties.
 
@@
 
-Correll \ea~\cite{correll2015layercake,correll2011visualizing} use a precursor of VSUPs in their LayerCake genomics visualization tool, where marks representing genomic data with increasing uncertainty are mapped to a smaller and smaller set of increasingly grey colors, creating the effect of uncertain values retreating into a ``confidence fog'' while highly certain values remain prominent.
-Other bivariate visualizations implicitly alias together uncertain values.
-For instance, if uncertainty is encoded by transparency, a maximally uncertain glyph may be entirely transparent, and so impossible to distinguish from any other maximally uncertain glyph.
-Other channels with a semiotic connection to uncertainty, such as saturation, value, blur, or size, also have deleterious effects on the disambiguation of colors and shapes.
-In both cases, the property of aliasing is \emph{ad hoc}, and places no guarantees on the discriminability of colors.
-The binning and degradation approach of VSUPs makes the choice to alias values explicit to both the designer and the viewer, and results in a bivariate mapping with known perceptual properties.
-
-\textbf{JH: The precending paragraph might be better placed in related work!}
+Since the visual channel used to encode uncertainty (e.g., saturation, lightness, transparency) often reduces the ability of people to distinguish colors, VSUPs have the added benefit of creating bivariate palettes that are perceptually easier to distinguish, by reducing color resolution in precisely the areas where color disambiguation would be most difficult anyway.
+For instance, the two closest colors in the VSUP in \figref{fig:example} are 1.4 units farther apart in CIELAB color space.
 
@@
 
-%\performanceFig
+There are multiple choices that designers must make before creating a VSUP.
+Specifically, they must choose (1) which visual channels map to value (1), uncertainty (2), and the structure of the underlying VSUP tree (3).
 
-There are multiple choices that designers must make before creating a VSUP.
-Specifically, they must choose (1) which visual channels map to value and uncertainty and (2) the structure of the underlying VSUP tree.
-
-Often, the graphical conventions of a chart reserve variables such as position and area for other uses, and determine the visual channel to be used for value.
-In heatmaps, for instance, value is regularly encoded using color, while vertical and horizontal position are reserved for location.
-\textbf{JH: This is the third time we've pointed this out...} In this paper, we primarily deal with the case where value is encoded using color.
-While there are some similar conventions for representing uncertainty, generally designers have more freedom in choosing a visual channel for uncertainty~\cite{maceachren1992visualizing}.
-\textbf{This feels like we are revisiting introductory and related work material again!} We recommend channels with both a strong semiotic connection to uncertainty as well as a relatively large number of perceptually distinguishable levels.
-Here, we focus on value/saturation (or, equivalently, transparency against a white background).
+When selecting a visual channel to represent uncertainty, we recommend channels with both a strong semiotic connection to uncertainty (such as those recommended by MacEachren et al.~\cite{maceachren2012visual}) as well as a relatively large number of perceptually distinguishable levels (such as those evaluated by Boukhelifa et al.
+\cite{boukhelifa2012evaluating}).
+Here, we focus on a combination of increasing color value and decreasing saturation.
 Correll \& Gleicher~\cite{correll2013error} provide evidence that even audiences without statistical training can successfully interpret uncertainty information encoded in these channels.
 
-\textbf{We should say something about the choice of value color map.
-Generally speaking, single-hue luminance ramps must be used with caution if they ramp too strongly towards white.
-This will be confused with manipulations of opacity or saturation.
-Viridis might be attractive due to its avoidance of desaturated colors.}
+When selecting a visual channel to represent value, the designer should consider the perceptual integrality of the value and uncertainty visual channels.
+For instance, we use the Viridis~\cite{viridis} color map frequently in this paper, as many other standard color ramps (such as sequential ColorBrewer~\cite{harrower2003colorbrewer} ramps), following usual best practices, interpolate in both hue and value~\cite{ware1988color}.
+This interpolation in value interferes with our uncertainty encoding, introducing ambiguity.
+Ramps with similar saturation and value among colors, such as Virids, are less ambiguous.
 
-%TODO Add in discussion of tree structures, including figure showing that VSUPs have colors that are further apart than similar 2d bivariate maps.
-Also worth discussing discretization issues here.
-Make new performance fig.
+When determining the quantization tree, the design should be careful to avoid having too many output colors.
+For reasons of memorability and complexity, Wainer\& Francolini~\cite{wainer1980empirical} suggest no more than 16.
+For a binary quantization tree, this results in a maximum tree depth of 4, as seen in \figref{fig:example}.
+In general, for tasks that require finer-grade distinctions between uncertainty, the depth of the tree should be increased.
+For tasks involving finer-grade distinctions between value, the branching factor of the tree should be increased.
+Beyond these two parameters, non-uniform binning would allow the designer to target particular distributions or important subregions of the data that would otherwise be difficult to visualize in a limited palette, as in Dunn's~\cite{dunn1989dynamic} dynamic bivariate maps.
+As an additional form of control, prior tools using VSUP-like techniques, such as LayerCake~\cite{correll2015layercake}, allow the analyst to interactively control the size of bins in the tree through dragging on the color legend.
 
@@
 \section{Evaluation}
-
+%TODO update conditions fig
 \conditionFig
@@
 After completion of the main tasks, we solicited demographic information, including a risk aversion assay from Mandrik \& Bao~\cite{mandrik2005exploring}.
-We recruited a total of $48$ participants for this experiment ($24$ for each experiment): $N$ female, $N$ male, and $N$ who declined to state ($M_{\text{age}}$= $N$, $SD_{\text{age}}$ = $N$).
+We recruited a total of $48$ participants for this experiment ($24$ for each experiment): $24$ female, $24$ male, ($M_{\text{age}}$= $34$, $SD_{\text{age}}$ = $9.1$).
 Participants from the first experiment were excluded from participating in the second.
@@
 Prior to the main task, we included a short replication of the identification task from the prior experiment (with 12 stimuli) for training purposes.
+Poor performance on this training was used for exclusion purposes.
+Three people with low accuracy ($6\%$, $6\%$, and $25\%$ accuracy compared to a mean of $70\%$ ) were excluded.
 
@@
 	\item Participants would choose targets with \textbf{lower uncertainty} in the prediction task when using a VSUP. 
-	\item This would result in a tradeoff where they would also choose targets with \textbf{lower expected value, but lower variance in expected value} when using a VSUP.
+	\item This would result in a tradeoff where they would also choose targets with \textbf{lower expected value} when using a VSUP.
 \end{enumerate}
@@
 \subsubsection{Results} 
+\taskTwoViolin
+Consistent with our first experiment, we found no significant effect of legend shape on either uncertainty ($F(1,61) = 0.01$, $p=0.92$) or value ($(1,61)=3.1$, $p=0.08$) of guesses.
+This seems to indicate that both wedge- or square-shaped legends promote similar patterns of decision-making.
+In this paper, and in general, we prefer to employ the wedge-shaped legend for VSUPs, and the square-shaped legend for traditional maps, as it makes the conceptual differences between the two more apparent.
+
+Our results failed to support our first hypothesis.
+We performed a repeated measures ANOVA on our results to measure the effect of VSUP versus standard quantizations, and square- versus wedge-shaped legends, on average uncertainty in bets.
+We found no significant effect of quantization scheme on average guess uncertainty ($F(1,61)=0.05$, $p=0.83$).
+This indicates that there does not appear to be a uniform pattern of risk aversion between the two scale types, as we had surmised.
+
+Our results supported our second hypothesis.
+A similar ANOVA on average danger value of guesses found that quantization scheme was a significant effect ($F(1,61)=17$, $p<0.01$).
+A post-hoc t-test confirmed ($t=2.3$, $p=0.02$) that guesses made with VSUPs had significantly higher average danger ($M=0.32$, $SD=0.17$) (and so lower expected value) than traditional maps ($M=0.29$, $SD=0.18$).
+Since there was no significant difference in average guess uncertainty, and the distribution of danger and uncertainty was identical across conditions, this seems to indicate that participants were using VSUPs moderating their guesses: avoiding highly uncertain regions, but willing to ``gamble'' more on regions of middling certainty.
+\figref{fig:taskTwoViolin} shows the comparative distribution of uncertainty in participant guesses.
 
@@
 \subsection{Limitations and Future Work}
-% When should you use these things?
-In general, VSUPs excel in scenarios where we wish to make firm guarantees about the discriminability of different marks in a chart, but must also represent, in as much detail as possible, two, potentially orthogonal variables of interest.
-VSUPs have improved data resolution over other bivariate maps by assuming that uncertain values ought to have less visual impact that highly certain values.
+
+VSUPs assume that uncertain values ought to have less visual impact that highly certain values.
 VSUPs therefore act both as a kind of filtering device as well as a bivariate representation.

### 2017-09-16 paper/figures.tex
Dealt with all JH edits except for the difficult one.
Update conditions figure

@@
 		\centering
-		\begin{subfigure}{0.3\textwidth}
-			\centering
-			\includegraphics[width=.45\textwidth]{conditions1.pdf}
-			\caption{Traditional Bivariate Map}
-		\end{subfigure}
-		~
-		\begin{subfigure}{0.3\textwidth}
-			\centering
-			\includegraphics[width=\textwidth]{conditions2.pdf}
-			\caption{Juxtaposed Univariate Maps}
-		\end{subfigure}
-		~
-		\begin{subfigure}{0.3\textwidth}
-			\centering
-			\includegraphics[width=0.45\textwidth]{conditions3.pdf}
-			\caption{Value-Suppressing Uncertainty Palette}
-		\end{subfigure}
-		\caption{The three graph types in our evaluation.
-Bivariate color ramps ensure that there are orthogonal mappings for each combination of value and uncertainty, but, since color channels are not separable, can afford only a few discrete colors before color categories become unacceptably close, perceptually.
-Juxtaposed maps, by keeping these channels separate, afford a larger range of colors, but require a visual search task to recover both value and uncertainty in a region.
-VSUPs have many of the advantages of bivariate maps, but assign more color categories to certain data, at the expense of ambiguity when uncertainty is high.}
+		\includegraphics[width=0.95\textwidth]{allConditions.pdf}
+		\caption{The 8 conditions from our first experiment.
+Juxtaposed maps require a participants to make an error-prone connection between areas in two separate maps in order to make a decision that integrates value and uncertainty.
+Traditional bivariate maps integrate both value and uncertainty.
+VSUPs improve on traditional bivariate maps by reducing color resolution as uncertainty increases, discouraging conclusions based on noisy or imprecise data.}
 		\label{fig:conditions}
@@
 		\centering
-		\includegraphics[width=0.45\columnwidth]{defend.png}
+		\includegraphics[width=0.4\columnwidth]{defend.png}
 		\caption{The prediction task. The participant has a list of locations, and ought to place their ships on locations with low probability of attack, and high certainty in this probability.}
@@
 
+\newcommand{\taskTwoViolin}{
+	\begin{figure}
+		\centering
+		\includegraphics[width=0.75\columnwidth]{violin.pdf}
+		\caption{A violin plot of results from our prediction experiment.
+While, on average, participants selected squares with similar uncertainty overall, participants using VSUPs avoiding were dissuaded from making guesses with high uncertainty, instead spreading their risk over other categories.
+The confidence intervals are bootstrapped 95\% CIs of trimmed means.}
+		\label{fig:taskTwoViolin}
+	\end{figure}
+}
 %\newcommand{\performanceFig}{

### 2017-09-17 paper/figures.tex
Created exp1 figure

@@
 
+\newcommand{\taskOneFig}{
+	\begin{figure}
+		\centering
+		\includegraphics[width=0.95\columnwidth]{task1.png}
+		\caption{Accuracy at our identification experiment.
+For examples of each condition, see \protect\figref{fig:conditions}.
+Juxtaposing two univariate maps for both value and uncertainty requires an error-prone search task for identification tasks.
+Continuous rather than discrete bivariate maps requires an error-prone color encoding and estimation task.
+Discrete bivariate maps, both VSUPs and otherwise, avoid these issues.
+The confidence intervals are bootstrapped 95\% CIs of trimmed means.}
+	\end{figure}
+}
+
 \newcommand{\taskTwoFig}{
@@
 		\caption{A violin plot of results from our prediction experiment.
-While, on average, participants selected squares with similar uncertainty overall, participants using VSUPs avoiding were dissuaded from making guesses with high uncertainty, instead spreading their risk over other categories.
+While, on average, participants selected squares with similar uncertainty overall, participants using VSUPs were dissuaded from making guesses with high uncertainty, instead spreading their risk over other categories.
 The confidence intervals are bootstrapped 95\% CIs of trimmed means.}

### 2017-09-17 paper/main.tex
Created exp1 figure
Fresh commit to deal with paper weirdness.
Jeff's revisions.

@@
 Therefore, bivariate maps in practice are often limited to a small set of output colors (say, a 4x4 matrix as in \figref{fig:example}) \cite{robertson1986generation,trumbo1981theory}.
-For VSUPs, we adapt an insight from Dunn~\cite{dunn1989dynamic} that these limited categories ought to be assigned with regards to their importance to analysis tasks, rather than uniformly.
 
 Given that bivariate maps, for reasons of either practicality or perceptual fidelity, have only a limited budget of outputs, we adapt an insight from Dunn~\cite{dunn1989dynamic} that these limited categories ought to be assigned with regards to their importance to analysis tasks, rather than uniformly.
-Our Value-Suppressing Uncertainty Palettes apply this insight to data with uncertainty, where, frequently, uncertain information ought to have less valence on the decision-making process, or at least less prominence in the analysis, than certain information.
+Our Value-Suppressing Uncertainty Palettes apply this insight to data with uncertainty, where, frequently, uncertain information ought to be weighted less heavily in a decision-making process, or at least less prominence in an analysis, than certain information.
 
@@
 
-However, if both of these assumptions hold, then it follows that the designer of a bivariate map should allocate more mark types to certain values, and fewer mark types to uncertain values.
+If both of these assumptions hold, then it follows that the designer of a bivariate map should allocate more mark types to certain values, and fewer mark types to uncertain values.
 VSUPs codify this decision by \emph{reducing the number of mark categories for representing value as uncertainty increases.} This means that data encoded using a VSUP will make visible only the largest of differences in uncertain data, but highlight comparatively small differences in value when uncertainty is low.
@@
 There are multiple choices that designers must make before creating a VSUP.
-Specifically, they must choose (1) which visual channels map to value (1), uncertainty (2), and the structure of the underlying VSUP tree (3).
+Specifically, they must choose which visual channels map to value (1), uncertainty (2), and the structure of the underlying VSUP tree (3).
 
@@
 This interpolation in value interferes with our uncertainty encoding, introducing ambiguity.
-Ramps with similar saturation and value among colors, such as Virids, are less ambiguous.
+Ramps with similar saturation and value among colors, such as Viridis, are less ambiguous.
 
@@
 
-We present a set of examples showing how VSUPs can be applied to realworld datasets.
+We present a set of examples showing how VSUPs can be applied to real-world datasets.
 The decision of VSUPs to allocate color bins asymmetrically amongst value and uncertainty regions affords greater discriminability in regions of the chart that warrant closer scrutiny, while discouraging exploration of regions with noisy or unreliable signals.
@@
 
-Both maps illustrate a similar temporal trend: flight delays are shorter towards the beginning of the day, increasing on average over time.
+Both maps illustrate a similar trend: flight delays are shorter towards the beginning of the day, increasing on average over time.
 The traditional bivariate map affords only a coarse examination of this trend: two visible ``blocks'' of color, early in the day, and later in the afternoon.
@@
 
+Next generation sequencing allows biologists to collect and analyze large amounts of genomics data.
+However, these techniques often create data quality issues, as they must align large numbers of potentially ambiguous ``reads'' of relatively small numbers of base pairs.
 Using a dataset from a group of virologists interested in the population dynamics of viral mutation~\cite{o2012conditional}, we visualized the variability in viral genomes across 25 populations of the simian immunodeficiency virus (SIV).
-Next generation sequencing (NGS) allows biologists to collect and analyze large amounts of genomics data.
-However, these techniques often create data quality issues, as they must align large numbers of potentially ambiguous ``reads'' of relatively small numbers of base pairs.
-In the dataset here, it is important not only to identify hotspots in the viral genome (locations with high rates of mutation or variability), but to not be distracted by false positives.
+In the dataset here, it is important not only to identify hotspots in the viral genome (locations with high rates of mutation or variability), but also to not be distracted by false positives.
 The traditional bivariate map (\figref{fig:viral2d}) affords the analysis of only a few, large hot spots.
-Extending the ``confidence fog'' metaphor used in viral genome browsers like LayerCake~\cite{correll2015layercake}, VSUPs (\figref{fig:viralVsum}) enable biologists to explore interesting patterns while not being distracted by noise.
+
+VSUPs (\figref{fig:viralVsum}) enable biologists to explore interesting patterns while deemphasizing noise.
 Smaller scale ``hot spots'' that are mapped to the same magenta color in the traditional map are visible as bright orange lines.
-Pink hot spots towards the end of the genome (around reference nucleotide $9,500$) that could be plausibly considered potential regions of interest in the 2D map, are encoded as highly uncertain and of ambiguous value in the VSUPs.
+Pink hot spots towards the end of the genome (around reference nucleotide $9,500$) could be considered potential regions of interest in the 2D map; in the VSUP these are encoded as highly uncertain and of ambiguous value.
 In fact, these regions have systematically low coverage; the lack of reliable data for these regions should discourage analysts from giving their apparent pattern equal consideration from the stronger signals earlier in the genome.
 
-\textbf{A few thoughts about this section based on the current figures.
-(1) Fig.
-3 in particular has two graphics that look pretty similar.
-Is there anything we should do to draw attention to the most salient differences between conditions?
-At minimum, we have the text, but we could consider annotations or even a third vis that is a heatmap of differences between charts.
-(2) A third example using a different visual form might be useful.
-For example, a node-link network with value + uncertainty associated with each node.}
+\textbf{A third example using a different visual form might be useful. For example, a node-link network with value + uncertainty associated with each node.}
 
@@
 This focus on integration meant that we limited our experimental tasks to scenarios where the participants needed to consider both value and uncertainty before making a decision.
-The experiments consisted of two separate tasks:
+Each experiment focused on one of two separate tasks:
 
 \begin{enumerate}
-	\item An experiment with an \textbf{identification} task, where we gave participants charts with value and uncertainty information, and asked them to locate specific regions.
+	\item An \textbf{identification} task, where we gave participants charts with value and uncertainty information, and asked them to locate specific regions.
 E.g., ``click on the region of the chart with a value of $0.1$ and an uncertainty of $0.2$.''
-	\item An experiment with a \textbf{prediction} task, inspired by the game Battleship, where we gave participants charts with both \emph{forecast} and \emph{forecast uncertainty} information, and asked the participants to place tokens on the board in order to optimize expected value.
+	\item A \textbf{prediction} task, inspired by the game Battleship, where we gave participants charts with both \emph{forecast} and \emph{forecast uncertainty} information, and asked the participants to place tokens on the board in order to optimize expected value.
 E.g., ``place your $5$ ships on safe locations on the board.''
@@
 We limited our population to Turkers from the United States, with a prior task approval rate of at least 90\%.
-As the experimental tasks required multi-hue color perception, we presented participants with a set of Ishihara plates~\cite{hardy1945tests} as a pretest, and excluded participants who either misidentified the values in the plates, or who self-reported as having a color vision deficiency (CVD) in the post-test.
+As the experimental tasks required multi-hue color perception, we presented participants with a set of Ishihara plates~\cite{hardy1945tests} as a pre-test, and excluded participants who either misidentified the values in the plates, or who self-reported as having a color vision deficiency (CVD) in the post-test.
 Based on piloting, we paid participants \$2 dollars for participation, for a target rate of \$8/hour.
 After completion of the main tasks, we solicited demographic information, including a risk aversion assay from Mandrik \& Bao~\cite{mandrik2005exploring}.
-We recruited a total of $48$ participants for this experiment ($24$ for each experiment): $24$ female, $24$ male, ($M_{\text{age}}$= $34$, $SD_{\text{age}}$ = $9.1$).
+We recruited a total of $48$ participants ($24$ for each experiment): $24$ female, $24$ male, ($M_{\text{age}}$= $34$, $SD_{\text{age}}$ = $9.1$).
 Participants from the first experiment were excluded from participating in the second.
@@
 
-For the identification task, we gave participants a $5$x$5$ heatmap, and asked them to click on a region of the heatmap that encodes a particular value,uncertainty pair.
+For the identification task, we gave participants a $5$x$5$ heatmap, and asked them to click on a region of the heatmap that encodes a particular (value, uncertainty) pair.
 To support comparability across conditions, we chose a set of target pairs such that each target mapped to a unique color across all encoding types.
@@
 
-We had two initial hypotheses, based on prior work on bivariate maps:
+We had two initial hypotheses, informed by prior work:
+
 \begin{enumerate}
-	\item Juxtaposed maps, by introducing a second search task to the identification task (searching for the proper values in two, rather than one map), would have poorer performance than the other conditions.
-	\item Continuous maps required more specific and accurate color matching between value and color, and so would have poorer performance than the other conditions.
+	\item Juxtaposed maps, by introducing a second search task to the identification task (searching for the proper values in two, rather than one map), would have poorer performance than other conditions.
+	\item Continuous maps require more specific and accurate color matching between value and color, and so would have poorer performance than discrete conditions.
 \end{enumerate}
@@
 \subsubsection{Results}
+\taskOneFig
 
@@
 Participant ID was included as a random effect.
-Figure %TODO
-shows the performance of each of our 8 conditions.
+\figref{fig:taskOne} shows the performance of each of our 8 conditions.
 
-Our results support our first hypothesis.
+The results support our first hypothesis.
 Juxtaposition was a significant effect ($F(1,166)=5.5$,$p=0.02$).
-A post-hoc t-test confirmed ($t=2.5$, $p=0.01$)that superimposed charts ($M=0.58$, $SD=0.5$) performed significantly better than juxtaposed ($M=0.51$, $SD=0.49$) ones.
+A post-hoc t-test confirmed ($t=2.5$, $p=0.01$) that superimposed charts ($M=0.58$, $SD=0.5$) performed significantly better than juxtaposed ($M=0.51$, $SD=0.49$) ones.
 For our identification task, juxtaposition requires at least two search tasks (one to locate squares with the correct data value, another to locate ones with the correct uncertainty value), which introduces additional error compared to conditions where value and uncertainty are displayed simultaneously.
 
-Our results also support our second hypothesis.
+The results also support our second hypothesis.
 Discretization was a significant effect ($F(1,166)=30$,$p<0.01$).
@@
 
-We performed a second ANOVA amongst just the superimposed, discrete charts, to determine the effect of legend shape (wedge or square) and quantization scheme (VSUP or standard) on performance, with participant ID as a random factor.
-Neither the legend shape ($F(1,70)=0.03$, $p=0.084$) nor the quantization scheme ($F(1,70)=1.4$, $p=0.24$) were significant effects.
+We performed a second ANOVA among the superimposed discrete charts to determine the effect of legend shape (wedge or square) and quantization scheme (VSUP or standard) on performance, with participant ID as a random factor.
+We did not find a significant effect for either the legend shape ($F(1,70)=0.03$, $p=0.084$) or the quantization scheme ($F(1,70)=1.4$, $p=0.24$).
 
@@
 
-For the prediction task, we gave participants the rules of a game like Battleship.
-Greis et al.
-\ea~\cite{greis2016decision} employ these game-like experimental tasks to assess how different visual designs communicate uncertainty information, which can be abstract or complex, to the general audience.
+For the prediction task, we gave participants the rules of a game similar to Battleship.
+Greis \ea~\cite{greis2016decision} employ these game-like experimental tasks to assess how different visual designs communicate uncertainty information, which can be abstract or complex, to the general audience.
 In our task, the participant and a (fictional) adversary have to place tokens representing ships on a $5$x$5$ spatial grid, with the expectation that certain squares will be hit by missiles.
@@
 
-Our results from the prior study indicate indicate that discrete, non-juxtaposed maps outperformed the other bivariate maps we selected, so we limited our study to only 4 types: square and arc bivariate maps.
+Our results from the prior study indicate that discrete, non-juxtaposed maps outperformed the other bivariate maps we selected, so we limited our study to only 4 types: square and arc bivariate maps.
 These in turn were either VSUPs or traditional bivariate maps, for a 2 (square or arc legend) x 2 (VSUP or standard) factorial design, with 6 replications, for a total of 24 stimuli for each of our 24 participants.
@@
 Poor performance on this training was used for exclusion purposes.
-Three people with low accuracy ($6\%$, $6\%$, and $25\%$ accuracy compared to a mean of $70\%$ ) were excluded.
+Three people with low accuracy ($6\%$, $6\%$, and $25\%$ accuracy compared to a mean of $70\%$) were excluded.
 
@@
 
-We had two hypotheses, stemming from our belief that VSUPs promote better \emph{integration} between uncertainty and value information, and also encourage \emph{caution} by highlighting the ambiguity or untrustworthiness introduced by uncertain data.
+We had two hypotheses, stemming from our belief that VSUPs promote better \emph{integration} between uncertainty and value information, and encourage \emph{caution} by highlighting the ambiguity introduced by uncertain data.
 In particular:
@@
 \taskTwoViolin
-Consistent with our first experiment, we found no significant effect of legend shape on either uncertainty ($F(1,61) = 0.01$, $p=0.92$) or value ($(1,61)=3.1$, $p=0.08$) of guesses.
-This seems to indicate that both wedge- or square-shaped legends promote similar patterns of decision-making.
-In this paper, and in general, we prefer to employ the wedge-shaped legend for VSUPs, and the square-shaped legend for traditional maps, as it makes the conceptual differences between the two more apparent.
+Consistent with our first experiment, we found no significant effect of legend shape on either uncertainty ($F(1,61) = 0.01$, $p=0.92$) or value ($F(1,61)=3.1$, $p=0.08$) of guesses.
+This result suggests that wedge- and square-shaped legends promote similar patterns of decision-making.
+Overall, we prefer to employ the wedge-shaped legend for VSUPs, and the square-shaped legend for traditional maps, as it makes the conceptual differences between the two more apparent.
 
-Our results failed to support our first hypothesis.
+The results fail to support our first hypothesis.
 We performed a repeated measures ANOVA on our results to measure the effect of VSUP versus standard quantizations, and square- versus wedge-shaped legends, on average uncertainty in bets.
 We found no significant effect of quantization scheme on average guess uncertainty ($F(1,61)=0.05$, $p=0.83$).
-This indicates that there does not appear to be a uniform pattern of risk aversion between the two scale types, as we had surmised.
+This result indicates that there does not appear to be a uniform pattern of risk aversion between the two scale types.
 
-Our results supported our second hypothesis.
-A similar ANOVA on average danger value of guesses found that quantization scheme was a significant effect ($F(1,61)=17$, $p<0.01$).
+The results support our second hypothesis.
+An ANOVA for average danger value of guesses found a significant effect of quantization scheme ($F(1,61)=17$, $p<0.01$).
 A post-hoc t-test confirmed ($t=2.3$, $p=0.02$) that guesses made with VSUPs had significantly higher average danger ($M=0.32$, $SD=0.17$) (and so lower expected value) than traditional maps ($M=0.29$, $SD=0.18$).
-Since there was no significant difference in average guess uncertainty, and the distribution of danger and uncertainty was identical across conditions, this seems to indicate that participants were using VSUPs moderating their guesses: avoiding highly uncertain regions, but willing to ``gamble'' more on regions of middling certainty.
+Since there was no significant difference in average guess uncertainty, and the distribution of danger and uncertainty was identical across conditions, this result suggests that VSUPs led participants to moderate their guesses: avoiding highly uncertain regions, but willing to ``gamble'' more on regions of middling certainty.
 \figref{fig:taskTwoViolin} shows the comparative distribution of uncertainty in participant guesses.
@@
 
-The results from our initial evaluation of VSUPs are promising, showing that even a general audience can make use of uncertainty information in a reasonable way.
+The results from our initial evaluation of VSUPs are promising, showing that a general audience can make use of uncertainty information in a reasonable way.
 The way that uncertainty information is presented can have a measurable impact on decision-making.
-The VSUP strategy of only ``showing'' differences when uncertainty is sufficiently low is somewhat analogous to the inferential statistics such as effects tests.
+The VSUP strategy of only emphasizing differences when uncertainty is sufficiently low is somewhat analogous to the inferential statistics such as effects tests.
 VSUPs might therefore be used to promote more caution in judgments, and lead analysts away from spurious signals in data.
@@
 
-Often, uncertainty, data quality, or confidence are considered separate information from the data itself, relegated to tooltips or visually distant supplemental charts.
-We believe, in contrast, that uncertainty information ought to be directly integrated with charts of data itself.
+Uncertainty, data quality, or confidence values are often considered separately from the data itself, relegated to tooltips or visually distant supplemental charts.
+We contend, in contrast, that uncertainty information ought to be directly integrated within a shared chart.
 This integration introduces additional complexity in the design and presentation of data.
-Value-Suppressing Uncertainty Palettes represent one strategy for dealing with this complexity, by assigning marks properties in a way that supports the disambiguation of values in data where uncertainty is low, but suppresses these judgments when uncertainty is high.
+Value-Suppressing Uncertainty Palettes represent one strategy for dealing with this complexity, by assigning mark properties in a way that supports the disambiguation of values in data where uncertainty is low, but suppresses these judgments when uncertainty is high.
 This decision of how to allocate visual variables promotes patterns of decision-making that make responsible use of uncertainty information, discouraging comparison of values in unreliable regions of the data, and promoting comparison in regions of high certainty.

### 2017-09-17 paper/figures.tex
Tweak exp1 fig
Jeff's revisions.

@@
 		\includegraphics[width=0.95\textwidth]{allConditions.pdf}
-		\caption{The 8 conditions from our first experiment.
+		\caption{The 8 conditions from the identification experiment.
 Juxtaposed maps require a participants to make an error-prone connection between areas in two separate maps in order to make a decision that integrates value and uncertainty.
 Traditional bivariate maps integrate both value and uncertainty.
-VSUPs improve on traditional bivariate maps by reducing color resolution as uncertainty increases, discouraging conclusions based on noisy or imprecise data.}
+VSUPs attempt to improve on traditional bivariate maps by reducing color resolution as uncertainty increases, discouraging conclusions based on noisy or imprecise data.}
 		\label{fig:conditions}
@@
 		\includegraphics[width=0.95\columnwidth]{task1.png}
-		\caption{Accuracy at our identification experiment.
+		\caption{Accuracy results for the identification experiment.
 For examples of each condition, see \protect\figref{fig:conditions}.
@@
 The confidence intervals are bootstrapped 95\% CIs of trimmed means.}
+		\label{fig:taskOne}
 	\end{figure}

### 2017-09-17 paper/main.tex
First round of JH changes

@@
 VSUPs rely on an underlying \textbf{quantization tree} that governs how values are discretized.
-Maximally uncertain values are mapped to a singular ``root'' node.
+Values about a certain uncertainty threshold are mapped to a singular ``root'' node.
 As uncertainty decreases, the tree branches into leaves, which evenly divide the data domain.
@@
 Since the visual channel used to encode uncertainty (e.g., saturation, lightness, transparency) often reduces the ability of people to distinguish colors, VSUPs have the added benefit of creating bivariate palettes that are perceptually easier to distinguish, by reducing color resolution in precisely the areas where color disambiguation would be most difficult anyway.
-For instance, the two closest colors in the VSUP in \figref{fig:example} are 1.4 units farther apart in CIELAB color space.
+For instance, the two closest colors in the VSUP in \figref{fig:example} are 1.4 units farther apart in CIELAB color space than the two closest colors in the traditional bivariate map.
 
-JavaScript code for generating VSUPs for use in D3 charts is available at \url{URL-REMOVED-FOR-REVIEW}.
+JavaScript code for generating VSUPs for use in D3~\cite{bostock2011d3} charts is available at \url{URL-REMOVED-FOR-REVIEW}.
 
@@
 When selecting a visual channel to represent value, the designer should consider the perceptual integrality of the value and uncertainty visual channels.
-For instance, we use the Viridis~\cite{viridis} color map frequently in this paper, as many other standard color ramps (such as sequential ColorBrewer~\cite{harrower2003colorbrewer} ramps), following usual best practices, interpolate in both hue and value~\cite{ware1988color}.
+For instance, we use the Viridis~\cite{viridis} color map frequently in this paper, as many other standard color ramps (such as sequential ColorBrewer~\cite{harrower2003colorbrewer} ramps), following usual best practices, interpolate in both hue and luminance~\cite{ware1988color}.
 This interpolation in value interferes with our uncertainty encoding, introducing ambiguity.
-Ramps with similar saturation and value among colors, such as Viridis, are less ambiguous.
+Ramps such as Viridis, which avoid very light and very dark colors, reduce this ambiguity.
 
@@
 For a binary quantization tree, this results in a maximum tree depth of 4, as seen in \figref{fig:example}.
-In general, for tasks that require finer-grade distinctions between uncertainty, the depth of the tree should be increased.
-For tasks involving finer-grade distinctions between value, the branching factor of the tree should be increased.
+In general, for tasks that require finer-grade distinctions between uncertainty, the depth of the tree could be increased.
+For tasks involving finer-grade distinctions between value, the branching factor of the tree could be increased.
+However, these factors increase the number of colors, and so designers should moderate between legend complexity and data fidelity.
 Beyond these two parameters, non-uniform binning would allow the designer to target particular distributions or important subregions of the data that would otherwise be difficult to visualize in a limited palette, as in Dunn's~\cite{dunn1989dynamic} dynamic bivariate maps.
@@
 
-\subsection{Examples}
+\section{Examples}
 
 We present a set of examples showing how VSUPs can be applied to real-world datasets.
-The decision of VSUPs to allocate color bins asymmetrically amongst value and uncertainty regions affords greater discriminability in regions of the chart that warrant closer scrutiny, while discouraging exploration of regions with noisy or unreliable signals.
+VSUPs allocate color bins asymmetrically amongst value and uncertainty regions, affording greater discriminability in regions of the chart that warrant closer scrutiny, while discouraging exploration of regions with noisy or unreliable signals.
 
-\subsubsection{Air Travel Delay}
+\subsection{Air Travel Delay}
 
@@
 
-\subsubsection{Viral Mutation}
+\subsection{Viral Mutation}
 %TODO update this figure
@@
 
-Rather than standard means, Cleveland \& McGill~\cite{cleveland1984graphical} use 25\% trimmed means to report effect sizes in graphical perception tasks.
+%Rather than standard means, Cleveland \& McGill~\cite{cleveland1984graphical} use 25\% trimmed means to report effect sizes in graphical perception tasks.
 Trimmed means remove the first and last quartiles and then compute the mean from the remaining points.
@@
 This resulted in (4+4+2+1)=11 valid targets.
-For each trial, these 11 targets were placed randomly in a heatmap; the remaining 13 cells of the heatmap were distractors randomly sampled from this target list.
+For each trial, we placed one example of each of these 11 targets randomly in a heatmap; the remaining 13 cells of the heatmap were randomly sampled from this target list.
 
@@
 For our identification task, juxtaposition requires at least two search tasks (one to locate squares with the correct data value, another to locate ones with the correct uncertainty value), which introduces additional error compared to conditions where value and uncertainty are displayed simultaneously.
+This is reflected also in response times, which, although highly variable, were highest in juxtaposed conditions ($M=19$, $SD=65$), than the other conditions ($M=13$, $SD=38$), although this difference was not significant ($t=1.64$, $p=0.1$).
 
@@
 
-Our results from the prior study indicate that discrete, non-juxtaposed maps outperformed the other bivariate maps we selected, so we limited our study to only 4 types: square and arc bivariate maps.
-These in turn were either VSUPs or traditional bivariate maps, for a 2 (square or arc legend) x 2 (VSUP or standard) factorial design, with 6 replications, for a total of 24 stimuli for each of our 24 participants.
+Our results from the prior study indicate that discrete, non-juxtaposed maps outperformed the other bivariate maps we selected, so we limited our study to only 4 types: square and wedge bivariate maps.
+These in turn were either VSUPs or traditional bivariate maps, for a 2 (square or wedge legend) x 2 (VSUP or standard) factorial design, with 6 replications, for a total of 24 stimuli for each of our 24 participants.
 \figref{fig:conditions} shows these factor levels.
@@
 This additional step before information fusion is reflected in longer response times, and patterns of decision-making that are less mindful of risk.
-Given the deleterious effects of latency on data analysis~\cite{liu2014effects}, designers ought to carefully consider the tradeoffs between expressiveness of visual encoding, and usability.
 

### 2017-09-17 modules/src/legend.js
Support domains, add virology example, delete old code

@@ -13,7 +13,7 @@ export function simpleLegend(m_scale, m_size, m_height, m_format, m_title, m_x, 
       scale = m_scale ? m_scale : null,
       size = m_size ? m_size : 200,
       height = m_height ? m_height : 30,
-      fmat = m_format ? m_format : ".2f",
+      fmat = m_format || null,
       x = m_x ? m_x : 0,
       y = m_y ? m_y : 0;
 
@@ -54,7 +54,7 @@ export function simpleLegend(m_scale, m_size, m_height, m_format, m_title, m_x, 
     }
     axis
       .attr("transform", "translate(0, " + height + ")")
-      .call(d3.axisBottom(axisScale).tickFormat(d3.format(fmat)));
+      .call(d3.axisBottom(axisScale).tickFormat(d3.format(fmat || "")));
 
     var label = el.select("g.legend > text");
     if (label.empty()) {
@@ -147,17 +147,18 @@ export function simpleLegend(m_scale, m_size, m_height, m_format, m_title, m_x, 
   return legend;
 }
 
-export function heatmapLegend(data, m_scale, m_size, m_format, m_utitle, m_vtitle, m_x, m_y) {
+export function heatmapLegend(m_scale, m_size, m_format, m_utitle, m_vtitle, m_x, m_y) {
   var el = null,
     utitle = m_utitle ? m_utitle : "Uncertainty",
     vtitle = m_vtitle ? m_vtitle : "Value",
     scale = m_scale ? m_scale : null,
     size = m_size ? m_size : 200,
-    fmat = m_format ? m_format : ".2f",
+    fmat = m_format || null,
     x = m_x ? m_x : 0,
-    y = m_y ? m_y : 0;
+    y = m_y ? m_y : 0,
+    data = null;
 
-  var heatmap = simpleHeatmap(data, m_scale, m_size, null, 0, 0);
+  var heatmap = simpleHeatmap();
 
   var legend = function(nel) {
     el = nel;
@@ -171,7 +172,17 @@ export function heatmapLegend(data, m_scale, m_size, m_format, m_utitle, m_vtitl
       return;
     }
 
-    heatmap.data(data);
+    var tmp = data;
+    if (!tmp) {
+      tmp = scale.quantize().data();
+    }
+
+    var inverted = [];
+    for (var i = 0; i < tmp.length; i++) {
+      inverted[i] = tmp[tmp.length - i - 1];
+    }
+
[…]

### 2017-09-17 modules/src/scale.js

@@ -3,9 +3,9 @@ import * as d3 from "d3";
 // Returns a color, based on a bivariate data point,
 // a quantization function, and a color interpolator
 // can be extended to different modes (just saturation, say)
-export function simpleScale(m_mode, m_range, m_scale) {
+export function simpleScale(m_mode, m_range, m_quantization) {
   var range = m_range,
-      scale = m_scale ? m_scale : function(v, u) {
+      quantization = m_quantization ? m_quantization : function(v, u) {
         var data = u != undefined ? {v:v, u:u} : {v:v.v, u:v.u};
         return data;
       },
@@ -18,9 +18,23 @@ export function simpleScale(m_mode, m_range, m_scale) {
   }
 
   function map(value, uncertainty) {
-    var data = scale(value, uncertainty);
-    var vcolor = range(data.v);
-    var uScale = d3.scaleLinear().domain([0, 1]).range([0, 1]);
+    var data = quantization(value, uncertainty);
+
+    var uDom = [0, 1];
+    var vDom = [0, 1];
+
+    if (quantization.uncertaintyDomain) {
+      uDom = quantization.uncertaintyDomain();
+    }
+    if (quantization.valueDomain) {
+      vDom = quantization.valueDomain();
+    }
+
+    var uScale = d3.scaleLinear().domain(uDom).range([0, 1]);
+    var vScale = d3.scaleLinear().domain(vDom).range([0, 1]);
+
+    var vcolor = range(vScale(data.v));
+
     switch (mode) {
       case "usl":
       default:
@@ -78,12 +92,12 @@ export function simpleScale(m_mode, m_range, m_scale) {
     }
   }
 
-  map.quantize = function(newScale) {
+  map.quantize = function(newQuantization) {
     if (!arguments.length) {
-      return scale;
+      return quantization;
     }
     else {
-      scale = newScale;
+      quantization = newQuantization;
       return map;
     }
   }

### 2017-09-17 old_code/heatmap/heatmap.js

@@ -1,883 +1,0 @@
-//Chart creation functions
-
-// for floating point errors
-var epsylon = 0.0000001;
-
-function makeHeatmap(svg, x, y, size, data, z, name){
-  //creates an svg heatmap with a 2d matrix of data, and a mapping function z
-  var w = size/data[0].length;
-  var h = size/data.length;
-
-  var heatmap = svg.append("g")
-             .attr("transform","translate("+x+","+y+")");
-
-  if(name){
-    heatmap.attr("id",name);
-  }
-
-  heatmap.selectAll("g")
-  .data(data)
-  .enter()
-  .append("g")
-  .selectAll("rect")
-    .data(function(d,i){ return d.map(function(val){ return {r:i, v:val};});})
-    .enter()
-    .append("rect")
-    .datum(function(d,i){ d.c = i; return d; })
-    .attr("x", function(d){ return d.c*w;})
-    .attr("y", function(d){ return d.r*h;})
-    .attr("width",w)
-    .attr("height",h)
-    .attr("fill", function(d){ return z(d.v);});
-}
-
-function makeHexmap(svg, x, y, size, data, z, maxSize, name){
-  var w = size/data[0].length;
-  var h = size/data.length;
-
-  var heatmap = svg.append("g")
-             .attr("transform","translate("+(x+w/2)+","+(y+h/2)+")");
-
-  if(name){
-    heatmap.attr("id",name);
-  }
-
-  heatmap.selectAll("g")
-  .data(data)
-  .enter()
-  .append("g")
-  .selectAll("path")
-    .data(function(d,i){ return d.map(function(val){ return {r:i, v:val};});})
-    .enter()
-    .append("path")
-    .datum(function(d,i){ d.c = i; return d; })
-    .attr("d", function(d){ return makeHexagon(z(d.v).s,d.c*w,d.r*h);})
-    .attr("fill", function(d){ return z(d.v).c;});
-}
-
-function makeArcmap(svg, x, y, size, data, z, name){
-  //creates an svg "wedge" map
-  var w = size/data[0].length;
-  var h = size/data.length;
-
-  var arcmap = svg.append("g")
-             .attr("transform","translate("+(x+size/2)+","+(y+size)+")");
-
-  if(name){
-    arcmap.attr("id",name);
-  }
-
-  arcmap.selectAll("g")
[…]

### 2017-09-17 paper/main.tex
Second round of JH edits

@@
 This resulted in 16 samples.
-The remaining 9 samples were distractors, with low safety and high certainty.
+The remaining 9 samples were ``bad'' values, with low safety and high certainty.
+This created 4 quartile categories of uncertainty, with a skew towards the highest quartile.
+This stimuli design meant that, while participants had at least one ``safe'' square (low danger with high certainty), they were forced to make at least some guesses in other quartiles.
 
@@
 \begin{enumerate}
-	\item Participants would choose targets with \textbf{lower uncertainty} in the prediction task when using a VSUP. 
+	\item Participants would avoid targets with \textbf{high uncertainty} when using a VSUP.
 	\item This would result in a tradeoff where they would also choose targets with \textbf{lower expected value} when using a VSUP.
@@
 \subsubsection{Results} 
-\taskTwoViolin
+\taskTwoQuartiles
+
 Consistent with our first experiment, we found no significant effect of legend shape on either uncertainty ($F(1,61) = 0.01$, $p=0.92$) or value ($F(1,61)=3.1$, $p=0.08$) of guesses.
@@
 
-The results fail to support our first hypothesis.
+The results partially support our first hypothesis.
 We performed a repeated measures ANOVA on our results to measure the effect of VSUP versus standard quantizations, and square- versus wedge-shaped legends, on average uncertainty in bets.
@@
 This result indicates that there does not appear to be a uniform pattern of risk aversion between the two scale types.
+However, given our quartile-based stimuli design, the central tendency of uncertainty would not necessarily capture caution in guesses.
+To capture differences in this non-normal distribution, and to test our belief that VSUP users would avoid highly uncertain regions, we performed a one-sided two-sample, Kolmogorov-Smirnov test on the distributions of uncertainty in guesses in the VSUP and traditional bivariate conditions.
+We found a significant difference ($D = 0.5$, $p=0.03$), and an inspection of the distribution shows that participants using VSUPs were less likely to make guesses with high uncertainty.
+\figref{fig:taskTwoQuartiles} shows this result.
 
@@
 A post-hoc t-test confirmed ($t=2.3$, $p=0.02$) that guesses made with VSUPs had significantly higher average danger ($M=0.32$, $SD=0.17$) (and so lower expected value) than traditional maps ($M=0.29$, $SD=0.18$).
-Since there was no significant difference in average guess uncertainty, and the distribution of danger and uncertainty was identical across conditions, this result suggests that VSUPs led participants to moderate their guesses: avoiding highly uncertain regions, but willing to ``gamble'' more on regions of middling certainty.
-\figref{fig:taskTwoViolin} shows the comparative distribution of uncertainty in participant guesses.
+
+These two results, taken together, suggest that VSUPs led participants to moderate their guesses: avoiding highly uncertain regions, but willing to ``gamble'' more on regions of middling certainty.
+In our design, where these gambles were usually poor (due to the predominance of ``bad'' cells), this moderation resulted in lower expected values.
+However, when value and uncertainty are orthogonal, this moderation can be beneficial, or at least not harmful.
+Indeed, for guesses with middling and low uncertainty (quartiles 1-3), where this orthogonality was present, there was no significant difference ($t=1.56$, $p=0.11$) in average danger between VSUP ($M=0.30$, $SD=0.24$) and traditional ($M=0.29$, $SD=024$) conditions.
 

### 2017-09-17 paper/figures.tex
Second round of JH edits

@@
 		\label{fig:taskTwoViolin}
+	\end{figure}
+}
+
+\newcommand{\taskTwoQuartiles}{
+	\begin{figure}
+		\centering
+		\includegraphics[width=0.75\columnwidth]{exp2histogram.pdf}
+		\caption{A histogram of the distribution of uncertainty across all participants guesses from the prediction experiment.
+While participants generally favored low uncertainty in guesses, our stimuli design encouraged them to make at least some guesses with higher uncertainty.
+VSUPs encouraged more guesses with middling uncertainty, and discouraged highly uncertain guesses.}
+		\label{fig:taskTwoQuartiles}
 	\end{figure}

### 2017-09-17 paper/figures.tex
Updated virus example, made more JH edits.

@@
 		\caption{A histogram of the distribution of uncertainty across all participants guesses from the prediction experiment.
+For each trial, participants made 5 guesses.
 While participants generally favored low uncertainty in guesses, our stimuli design encouraged them to make at least some guesses with higher uncertainty.
@@
 	\caption{Variability for 12 different viral populations of SIV, encoded with a VSUP.
-The magnitude of mutation rate, as well as the generally poorer quality of data towards the end of the sequence, is more readily visible than with the traditional bivariate color map.
+By suppressing uncertain information, the VSUP version removes the apparent spike in value at the second column, which is an artifact of sampling error due to poor coverage in that region.
+It also suppresses apparent inter-species variability in the solid orange spike at location ~9,500, which, again, is a function of poor coverage rather than a genuine per-strain difference.
 }

### 2017-09-17 paper/main.tex
Updated virus example, made more JH edits.

@@
 In the dataset here, it is important not only to identify hotspots in the viral genome (locations with high rates of mutation or variability), but also to not be distracted by false positives.
-The traditional bivariate map (\figref{fig:viral2d}) affords the analysis of only a few, large hot spots.
+The traditional bivariate map (\figref{fig:viral2d}) has many locations with apparent inter-population variance.
 
 VSUPs (\figref{fig:viralVsum}) enable biologists to explore interesting patterns while deemphasizing noise.
-Smaller scale ``hot spots'' that are mapped to the same magenta color in the traditional map are visible as bright orange lines.
-Pink hot spots towards the end of the genome (around reference nucleotide $9,500$) could be considered potential regions of interest in the 2D map; in the VSUP these are encoded as highly uncertain and of ambiguous value.
-In fact, these regions have systematically low coverage; the lack of reliable data for these regions should discourage analysts from giving their apparent pattern equal consideration from the stronger signals earlier in the genome.
-
-\textbf{A third example using a different visual form might be useful. For example, a node-link network with value + uncertainty associated with each node.}
+Orange hot spots towards the beginning (the second column) and end of the genome (around reference nucleotide $9,500$) could be considered potential regions of interest in the 2D map; in the VSUP these are encoded as highly uncertain and of ambiguous value.
+In fact, these regions have systematically low coverage; the lack of reliable data for these regions should discourage analysts from giving their apparent pattern equal consideration from the stronger signals elsewhere in the genome.
 
@@
 After completion of the main tasks, we solicited demographic information, including a risk aversion assay from Mandrik \& Bao~\cite{mandrik2005exploring}.
-We recruited a total of $48$ participants ($24$ for each experiment): $24$ female, $24$ male, ($M_{\text{age}}$= $34$, $SD_{\text{age}}$ = $9.1$).
 Participants from the first experiment were excluded from participating in the second.
-
-%Rather than standard means, Cleveland \& McGill~\cite{cleveland1984graphical} use 25\% trimmed means to report effect sizes in graphical perception tasks.
-Trimmed means remove the first and last quartiles and then compute the mean from the remaining points.
-This filtering is important for graphical perception tasks (especially crowdsourced tasks) where error distributions may be long-tailed~\cite{heer2010crowdsourcing}.
-Trimmed means violate the sampling assumptions required for many null-hypothesis significance tests.
-Thus, while we report effect sizes in terms of trimmed means, we perform inferential statistical tests on the standard means.
-Asterisks in our charts denote significant differences, as reported by a post-hoc Tukey's test of Honest Significant Difference (HSD).
 
@@
 
-Each of the 24 participants saw the data encoded as each of 8 types of bivariate encoding and legend, with 8 replications, for a total of 64 stimuli. \figref{fig:conditions} shows these factor levels.
+We recruited 24 participants for this task: $12$ female, $12$ male, ($M_{\text{age}}$= $31$, $SD_{\text{age}}$ = $7.8$).
+Each of the participants saw the data encoded as each of 8 types of bivariate encoding and legend, with 8 replications, for a total of 64 stimuli.
+\figref{fig:conditions} shows these factor levels.
 
@@
 
+ We recruited 24 participants for this task: $12$ female, $12$ male, ($M_{\text{age}}$= $37$, $SD_{\text{age}}$ = $9.8$).
 Our results from the prior study indicate that discrete, non-juxtaposed maps outperformed the other bivariate maps we selected, so we limited our study to only 4 types: square and wedge bivariate maps.
-These in turn were either VSUPs or traditional bivariate maps, for a 2 (square or wedge legend) x 2 (VSUP or standard) factorial design, with 6 replications, for a total of 24 stimuli for each of our 24 participants.
-\figref{fig:conditions} shows these factor levels.
-Prior to the main task, we included a short replication of the identification task from the prior experiment (with 12 stimuli) for training purposes.
-Poor performance on this training was used for exclusion purposes.
-Three people with low accuracy ($6\%$, $6\%$, and $25\%$ accuracy compared to a mean of $70\%$) were excluded.
+These in turn were either VSUPs or traditional bivariate maps, for a 2 (square or wedge legend) x 2 (VSUP or standard) factorial design, with 6 replications, for a total of 24 stimuli.
+Prior to the main task, we included a short replication of the identification task from the prior experiment (with 12 stimuli) for training and exclusion purposes.
+3 people with unacceptably low accuracy on the training task ($6\%$, $6\%$, and $25\%$ accuracy compared to a mean of $70\%$) were excluded from analysis.
 
@@
 The results support our second hypothesis.
-An ANOVA for average danger value of guesses found a significant effect of quantization scheme ($F(1,61)=17$, $p<0.01$).
+A similiar ANOVA for average danger value of guesses found a significant effect of quantization scheme ($F(1,59)=17$, $p<0.01$).
 A post-hoc t-test confirmed ($t=2.3$, $p=0.02$) that guesses made with VSUPs had significantly higher average danger ($M=0.32$, $SD=0.17$) (and so lower expected value) than traditional maps ($M=0.29$, $SD=0.18$).

### 2017-09-18 paper/main.tex
Just In Case submitted version.

@@
 These principles extend to bivariate color maps, with the additional consideration that our perception of color channels lacks \emph{orthogonality}.
-That is, while we can perceive e.g., the height of a bar in a bar chart independently of its width, a particular property of a color (e.g., its redness, saturation, transparency, etc.) affect how its other properties are perceived \cite{ware2012information,garner1970integrality}.
+That is, while we can perceive e.g., the height of a bar in a bar chart independently of its width, a particular property of a color (e.g., its redness, saturation, transparency, etc.) affect how its other properties are perceived \cite{garner1970integrality,ware2012information}.
 Many color channels such as hue and saturation are therefore perceptually \emph{integral}, which complicates their use in bivariate maps.
@@
 However, even simple bivariate maps can be difficult for a general audience to interpret: due to their additional complexity, Wainer \& Francolini~\cite{wainer1980empirical} reported high levels of error for participants even for ``elementary'' graph reading tasks, and additionally that bivariate legends were more difficult to memorize and internalize than their univariate equivalents.
-Therefore, bivariate maps in practice are often limited to a small set of output colors (say, a 4x4 matrix as in \figref{fig:example}) \cite{robertson1986generation,trumbo1981theory}.
+Therefore, bivariate maps in practice are often limited to a small set of output colors~\cite{robertson1986generation,trumbo1981theory} (say, a 4x4 matrix as in \figref{fig:example}).
 
 Given that bivariate maps, for reasons of either practicality or perceptual fidelity, have only a limited budget of outputs, we adapt an insight from Dunn~\cite{dunn1989dynamic} that these limited categories ought to be assigned with regards to their importance to analysis tasks, rather than uniformly.
-Our Value-Suppressing Uncertainty Palettes apply this insight to data with uncertainty, where, frequently, uncertain information ought to be weighted less heavily in a decision-making process, or at least less prominence in an analysis, than certain information.
+Our Value-Suppressing Uncertainty Palettes apply this insight to data with uncertainty.
+Frequently, uncertain information ought to be weighted less heavily in a decision-making process, or at least have less prominence in an analysis, than certain information.
 
@@
 When selecting a visual channel to represent value, the designer should consider the perceptual integrality of the value and uncertainty visual channels.
-For instance, we use the Viridis~\cite{viridis} color map frequently in this paper, as many other standard color ramps (such as sequential ColorBrewer~\cite{harrower2003colorbrewer} ramps), following usual best practices, interpolate in both hue and luminance~\cite{ware1988color}.
-This interpolation in value interferes with our uncertainty encoding, introducing ambiguity.
+For instance, we use the Viridis~\cite{viridis} color map frequently in this paper, as many standard color ramps (such as sequential ColorBrewer~\cite{harrower2003colorbrewer} ramps), following usual best practices, interpolate in both hue and luminance~\cite{ware1988color}.
+This interpolation in luminance interferes with our uncertainty encoding, introducing ambiguity.
 Ramps such as Viridis, which avoid very light and very dark colors, reduce this ambiguity.
@@
 When determining the quantization tree, the design should be careful to avoid having too many output colors.
-For reasons of memorability and complexity, Wainer\& Francolini~\cite{wainer1980empirical} suggest no more than 16.
+For reasons of memorability and complexity, Wainer \& Francolini~\cite{wainer1980empirical} suggest no more than 16 distinct outputs in bivariate maps.
 For a binary quantization tree, this results in a maximum tree depth of 4, as seen in \figref{fig:example}.
@@
 Beyond these two parameters, non-uniform binning would allow the designer to target particular distributions or important subregions of the data that would otherwise be difficult to visualize in a limited palette, as in Dunn's~\cite{dunn1989dynamic} dynamic bivariate maps.
-As an additional form of control, prior tools using VSUP-like techniques, such as LayerCake~\cite{correll2015layercake}, allow the analyst to interactively control the size of bins in the tree through dragging on the color legend.
+As an additional form of control, prior tools using VSUP-like techniques, such as LayerCake~\cite{correll2015layercake}, allow the analyst to control the bins in the tree through interaction with the color legend.
 
@@
 For each trial, we placed one example of each of these 11 targets randomly in a heatmap; the remaining 13 cells of the heatmap were randomly sampled from this target list.
+Thus there could be multiple correct locations in each trial, deterring a strategy of looking for ``oddball'' colors.
 
@@
 For our identification task, juxtaposition requires at least two search tasks (one to locate squares with the correct data value, another to locate ones with the correct uncertainty value), which introduces additional error compared to conditions where value and uncertainty are displayed simultaneously.
-This is reflected also in response times, which, although highly variable, were highest in juxtaposed conditions ($M=19$, $SD=65$), than the other conditions ($M=13$, $SD=38$), although this difference was not significant ($t=1.64$, $p=0.1$).
+This is reflected also in response times, which, although highly variable, were highest in juxtaposed conditions ($M=19$s, $SD=65$), than the other conditions ($M=13$s, $SD=38$), although this difference was not significant ($t=1.64$, $p=0.1$).
 
@@
 In our task, participants were given a map representing the predictions of missile strikes in each location on the grid.
-The \emph{value} component of the prediction was the likelihood of a missile strike in a particular region: the ship's \emph{safety} if placed on the square.
+The \emph{value} component was the ship's \emph{danger} if placed on the square.
 The \emph{uncertainty} component was the confidence in this prediction.
@@
 To capture differences in this non-normal distribution, and to test our belief that VSUP users would avoid highly uncertain regions, we performed a one-sided two-sample, Kolmogorov-Smirnov test on the distributions of uncertainty in guesses in the VSUP and traditional bivariate conditions.
-We found a significant difference ($D = 0.5$, $p=0.03$), and an inspection of the distribution shows that participants using VSUPs were less likely to make guesses with high uncertainty.
+We found a significant difference ($D = 0.5$, $p=0.03$), and an inspection of the distribution shows that participants using VSUPs were less likely to make guesses with very high uncertainty.
 \figref{fig:taskTwoQuartiles} shows this result.
@@
 However, when value and uncertainty are orthogonal, this moderation can be beneficial, or at least not harmful.
-Indeed, for guesses with middling and low uncertainty (quartiles 1-3), where this orthogonality was present, there was no significant difference ($t=1.56$, $p=0.11$) in average danger between VSUP ($M=0.30$, $SD=0.24$) and traditional ($M=0.29$, $SD=024$) conditions.
+Indeed, for guesses with middling and low uncertainty (quartiles 1-3), where this orthogonality was present, there was no significant difference ($t=1.56$, $p=0.11$) in average danger between VSUP ($M=0.30$, $SD=0.24$) and traditional ($M=0.29$, $SD=0.24$) conditions.
 
@@
 
+VSUPs, like other mappings involving quantization, are also highly sensitive to the binning scheme used.
+Uncertain data with values that fall very close to bin boundaries may end up having large color differences, even though they may be close in value.
+The comparison to effects tests is therefore imperfect, and a particular item's encoding in a VSUP cannot be used as a proxy for a test of statistical significance.
+As with other quantized forms of data analysis, including histograms and geospatial binning, analysts and designers should consider the distribution of values before quantizing.
+
 \subsection{Conclusion}

### 2017-09-19 paper/main.tex
Last minute edits

@@
 \cite{boukhelifa2012evaluating}).
-Here, we focus on a combination of increasing color value and decreasing saturation.
+Here, we focus on a combination of increasing luminance and decreasing saturation.
 Correll \& Gleicher~\cite{correll2013error} provide evidence that even audiences without statistical training can successfully interpret uncertainty information encoded in these channels.
@@
 However, these techniques often create data quality issues, as they must align large numbers of potentially ambiguous ``reads'' of relatively small numbers of base pairs.
-Using a dataset from a group of virologists interested in the population dynamics of viral mutation~\cite{o2012conditional}, we visualized the variability in viral genomes across 25 populations of the simian immunodeficiency virus (SIV).
+Using a dataset from a group of virologists interested in the population dynamics of viral mutation~\cite{o2012conditional}, we visualized the variability in viral genomes across 12 populations of the simian immunodeficiency virus (SIV).
 In the dataset here, it is important not only to identify hotspots in the viral genome (locations with high rates of mutation or variability), but also to not be distracted by false positives.
@@
 As the experimental tasks required multi-hue color perception, we presented participants with a set of Ishihara plates~\cite{hardy1945tests} as a pre-test, and excluded participants who either misidentified the values in the plates, or who self-reported as having a color vision deficiency (CVD) in the post-test.
-Based on piloting, we paid participants \$2 dollars for participation, for a target rate of \$8/hour.
+Based on piloting, we paid participants \$2 for participation, for a target rate of \$8/hour.
 After completion of the main tasks, we solicited demographic information, including a risk aversion assay from Mandrik \& Bao~\cite{mandrik2005exploring}.
@@
 For our identification task, juxtaposition requires at least two search tasks (one to locate squares with the correct data value, another to locate ones with the correct uncertainty value), which introduces additional error compared to conditions where value and uncertainty are displayed simultaneously.
-This is reflected also in response times, which, although highly variable, were highest in juxtaposed conditions ($M=19$s, $SD=65$), than the other conditions ($M=13$s, $SD=38$), although this difference was not significant ($t=1.64$, $p=0.1$).
+%This is reflected also in response times, which, although highly variable, were higher in juxtaposed conditions ($M=19$s, $SD=65$), than the other conditions ($M=13$s, $SD=38$), although this difference was not significant ($t=1.64$, $p=0.1$).
 
@@
 We performed a second ANOVA among the superimposed discrete charts to determine the effect of legend shape (wedge or square) and quantization scheme (VSUP or standard) on performance, with participant ID as a random factor.
-We did not find a significant effect for either the legend shape ($F(1,70)=0.03$, $p=0.084$) or the quantization scheme ($F(1,70)=1.4$, $p=0.24$).
+We did not find a significant effect for either the legend shape ($F(1,70)=0.04$, $p=0.84$) or the quantization scheme ($F(1,70)=1.4$, $p=0.24$).
 

### 2017-09-19 paper/main.tex
JH edits.

@@
 We describe the motivations for VSUPs, provide examples of their utility for decision-making under uncertainty, and assess VSUPs in a crowdsourced experiment.
-Our results indicate that VSUPs create close integration between uncertainty and data, resulting in better performance at data tasks with uncertainty components.
+Our results indicate that VSUPs create close integration between uncertainty and data, resulting in a demonstrably effect of decision-making, leading viewers to make more cautious choices that avoid high uncertainty.
 
@@
 E.g., ``click on the region of the chart with a value of $0.1$ and an uncertainty of $0.2$.''
-	\item A \textbf{prediction} task, inspired by the game Battleship, where we gave participants charts with both \emph{forecast} and \emph{forecast uncertainty} information, and asked the participants to place tokens on the board in order to optimize expected value.
-E.g., ``place your $5$ ships on safe locations on the board.''
+	\item A \textbf{prediction} task, inspired by the game Battleship, where we gave participants charts with both \emph{forecast} and \emph{forecast uncertainty} information, and asked the participants to place tokens on the board in order to minimize danger.
+E.g., ``place your $5$ ships on the safest locations on the board.''
 \end{enumerate}
@@
 
-The ideal strategy from a value-maximizing standpoint would be to place tokens on areas with the highest expected value (lowest danger), ignoring the uncertainty information.
+The ideal strategy from a value-maximizing standpoint would be to place tokens on areas with the lowest predicted danger (highest expected value), ignoring the uncertainty information.
 However, as with roulette and other similar games of chance, the \emph{variability} in expected value is relevant when considering where to place bets~\cite{mlodinow2009drunkard}.
@@
 	\item Participants would avoid targets with \textbf{high uncertainty} when using a VSUP.
-	\item This would result in a tradeoff where they would also choose targets with \textbf{lower expected value} when using a VSUP.
+	\item This would result in a tradeoff where they would also choose targets with lower expected value \textbf{higher danger} when using a VSUP.
 \end{enumerate}
@@
 These two results, taken together, suggest that VSUPs led participants to moderate their guesses: avoiding highly uncertain regions, but willing to ``gamble'' more on regions of middling certainty.
-In our design, where these gambles were usually poor (due to the predominance of ``bad'' cells), this moderation resulted in lower expected values.
-However, when value and uncertainty are orthogonal, this moderation can be beneficial, or at least not harmful.
+In our design, where these gambles were usually poor (due to the predominance of ``bad'' cells), this moderation resulted in guesses with higher average danger.
+However, when expected value and uncertainty are orthogonal, this moderation can be beneficial, or at least not harmful.
 Indeed, for guesses with middling and low uncertainty (quartiles 1-3), where this orthogonality was present, there was no significant difference ($t=1.56$, $p=0.11$) in average danger between VSUP ($M=0.30$, $SD=0.24$) and traditional ($M=0.29$, $SD=0.24$) conditions.

### 2017-12-06 paper/main.tex
Some bibliography fixes.

@@
 \section{Value-Suppressing Uncertainty Palettes}
-
-%TODO need some example figure here, showing tree structure perhaps?
-%\flowFig
-
 Value-Suppressing Uncertainty Palettes (VSUPs) are a technique for creating bivariate maps of data \emph{value} and \emph{uncertainty}. VSUPs make two central assumptions about bivariate maps:
@@
 
-JavaScript code for generating VSUPs for use in D3~\cite{bostock2011d3} charts is available at \url{URL-REMOVED-FOR-REVIEW}.
+JavaScript code for generating VSUPs for use in D3~\cite{bostock2011d3} charts is available at \url{https://github.com/uwdata/spatial-sampling}.
 
@@
 Here, we focus on a combination of increasing luminance and decreasing saturation.
-Correll \& Gleicher~\cite{correll2013error} provide evidence that even audiences without statistical training can successfully interpret uncertainty information encoded in these channels.
+Correll \& Gleicher~\cite{correll2014error} provide evidence that even audiences without statistical training can successfully interpret uncertainty information encoded in these channels.
 
@@
 
-\subsection{Viral Mutation}
-%TODO update this figure
-\viralFig
+%\subsection{Viral Mutation}
+%\viralFig
 
-Next generation sequencing allows biologists to collect and analyze large amounts of genomics data.
+%Next generation sequencing allows biologists to collect and analyze large amounts of genomics data.
 However, these techniques often create data quality issues, as they must align large numbers of potentially ambiguous ``reads'' of relatively small numbers of base pairs.
@@
 
-VSUPs (\figref{fig:viralVsum}) enable biologists to explore interesting patterns while deemphasizing noise.
+%VSUPs (\figref{fig:viralVsum}) enable biologists to explore interesting patterns while deemphasizing noise.
 Orange hot spots towards the beginning (the second column) and end of the genome (around reference nucleotide $9,500$) could be considered potential regions of interest in the 2D map; in the VSUP these are encoded as highly uncertain and of ambiguous value.
@@
 
+\subsection{2016 Pre-Election Polling}
+
+Using a public data set from Wikipedia~\cite{wiki}, we collected the statewide polls that were conducted closest to the 2016 U.S.
+presidential election, we created a choropleth map showing which candidate was leading in the polls.
+We calculated uncertainty as the size of the lead in terms of margin of error (lead/margin).
+The AP style guide for reporters suggests that a candidate leading by more than two margins of error is said to be ``significantly ahead'' in the polls~\cite{french1988associated}.
+As the lead becomes closer to the margin of error, there is more doubt about the electoral outcome.
+
+By providing more color categories when the margin of error was relatively small, the VSUP map shows that, in many of the states leaning towards Hillary Clinton, her lead was quite small.
+The VSUP map, by limiting inferences about which candidate was leading when the uncertainty is high, also prevents the viewer from making strong claims about the outcome of the election, even if a state is leaning one way or the other.
+Given the relative volatility in polling data, this caution was warranted.
+An additional benefit to the VSUP map is that, as the diverging blue-red color map passes through a central whitepoint, it can be difficult to distinguish between uncertainty levels (which also interpolates to white) in the 2D map.
+The VSUP's tree-like structure separates these problematic colors.
+
 \section{Evaluation}
-%TODO update conditions fig
 \conditionFig
@@
 
-Experimental materials, including data tables and stimuli generation code, are available at \url{URL-REMOVED-FOR-REVIEW}.
+Experimental materials, including data tables and stimuli generation code, are available at \url{https://github.com/uwdata/spatial-sampling}.
 

### 2017-12-06 paper/figures.tex
Just In Case submitted version.
Some bibliography fixes.

@@
 		\centering
-		\includegraphics[width=0.95\columnwidth]{task1.png}
+		\includegraphics[width=0.95\columnwidth]{task1.pdf}
 		\caption{Accuracy results for the identification experiment.
@@
 		\includegraphics[width=0.75\columnwidth]{exp2histogram.pdf}
-		\caption{A histogram of the distribution of uncertainty across all participants guesses from the prediction experiment.
+		\caption{Overlaid histograms of the distribution of uncertainty across all participants guesses from the prediction experiment.
 For each trial, participants made 5 guesses.
@@
 
+\newcommand{\pollFig}{
+	\begin{figure}[t]
+		\centering
+		\includegraphics[width=\columnwidth]{polling-2d.pdf}
+		\vspace{-15px}
+		\caption{Polling data prior to the 2016 U.S.
+Presidential Election, encoded in a traditional 2D bivariate map.
+Many states appear to be ``leaning'' towards Clinton, even when the margin of error of the polls is quite large.
+The blue-red color scale that is semantically appropriate for the topic is also not isoluminant, making it difficult to see just how much uncertainty there is in states with small leads for either candidate.}
+		\label{fig:polling2d}
+
+		\vspace{10px}
+
+		\centering
+		\includegraphics[width=\columnwidth]{polling-vsum.pdf}
+		\vspace{-15px}
+		\caption{Polling data prior to the 2016 U.S.
+Presidential Election, encoded with a VSUP.
+The larger amount of color bins for high-certainty regions shows that the leads for Hillary Clinton were, in many states, quite small.
+The collapsing of bins as uncertainty increases dissuades viewers from making confident predictions about states where the outcome was very much in doubt.
+}
+		\label{fig:pollingVsum}
+	\end{figure}
+}
+
 \newcommand{\responseTimeFig}{

### 2017-12-07 paper/main.tex
First typo pass.

@@
 We describe the motivations for VSUPs, provide examples of their utility for decision-making under uncertainty, and assess VSUPs in a crowdsourced experiment.
-Our results indicate that VSUPs create close integration between uncertainty and data, resulting in a demonstrably effect of decision-making, leading viewers to make more cautious choices that avoid high uncertainty.
+Our results indicate that VSUPs create close integration between uncertainty and data.
+This integration has a demonstrable effect on decision-making, leading viewers to make more cautious choices that avoid high uncertainty.
 
@@
 Our Value-Suppressing Uncertainty Palettes apply this insight to data with uncertainty.
-Frequently, uncertain information ought to be weighted less heavily in a decision-making process, or at least have less prominence in an analysis, than certain information.
+Frequently, uncertain information ought to be weighted less heavily in a decision-making process, or at least have less prominence in an analysis, than information with high certainty.
 
@@
 VSUPs rely on an underlying \textbf{quantization tree} that governs how values are discretized.
-Values about a certain uncertainty threshold are mapped to a singular ``root'' node.
+Values above a certain uncertainty threshold are mapped to a singular ``root'' node.
 As uncertainty decreases, the tree branches into leaves, which evenly divide the data domain.
@@
 By providing more color categories when the margin of error was relatively small, the VSUP map shows that, in many of the states leaning towards Hillary Clinton, her lead was quite small.
-The VSUP map, by limiting inferences about which candidate was leading when the uncertainty is high, also prevents the viewer from making strong claims about the outcome of the election, even if a state is leaning one way or the other.
-Given the relative volatility in polling data, this caution was warranted.
+The VSUP map, by limiting inferences about which candidate was leading when the uncertainty is high, also prevents the viewer from making strong claims about the outcome of the election in highly volatile states, even if a state is leaning one way or the other.
+Given the relative volatility in polling data, this caution is warranted.
 An additional benefit to the VSUP map is that, as the diverging blue-red color map passes through a central whitepoint, it can be difficult to distinguish between uncertainty levels (which also interpolates to white) in the 2D map.
-The VSUP's tree-like structure separates these problematic colors.
+The VSUP's tree-like structure separates these problematic colors: the closest colors in the 2D scale are 2.9 units apart in CIELab space, but the closest colors are 6.2 units apart in the VSUP scale.
 
@@
 Juxtaposition was a significant effect ($F(1,166)=5.5$,$p=0.02$).
-A post-hoc t-test confirmed ($t=2.5$, $p=0.01$) that superimposed charts ($M=0.58$, $SD=0.5$) performed significantly better than juxtaposed ($M=0.51$, $SD=0.49$) ones.
+A post-hoc t-test confirmed ($t=2.5$, $p=0.01$) that superimposed charts ($M=0.58$, $SD=0.50$) performed significantly better than juxtaposed ($M=0.51$, $SD=0.49$) ones.
 For our identification task, juxtaposition requires at least two search tasks (one to locate squares with the correct data value, another to locate ones with the correct uncertainty value), which introduces additional error compared to conditions where value and uncertainty are displayed simultaneously.
@@
 
-\sizeFig
+%\sizeFig
 
@@
 VSUPs therefore act both as a kind of filtering device as well as a bivariate representation.
-This assumption holds in scenarios like the virology dataset (\figref{fig:viralVsum}), where the expectation is that highly uncertain data is unreliable or should otherwise be downweighted in analysis.
+This assumption holds in scenarios like the virology dataset (\figref{fig:pollingVsum]}), where the expectation is that highly uncertain data is unreliable or should otherwise be downweighted in analysis.
 If the analyst has a different interpretation of uncertainty, or wishes to quickly and orthogonally analyze the distributions of uncertainty and value, other strategies, such as juxtaposed maps, may be more appropriate.
@@
 Where these models do not exist, or where the perceptual interaction is too complex to efficiently model, experimental work remains to be done before VSUPs can be considered a feasible design strategy.
-We are currently experimenting with the creation of VSUPs using these other channels, as in \figref{fig:size}.
+We are currently experimenting with the creation of VSUPs using these other channels, such as color and size.
 

### 2017-12-07 modules/src/legend.js
Add study result 2D heatmap

@@ -180,6 +180,7 @@ export function heatmapLegend(m_scale, m_size, m_format, m_utitle, m_vtitle, m_x
       inverted[i] = tmp[tmp.length - i - 1];
     }
 
+    heatmap.y(1);  // don't hide x-axis
     heatmap.data(inverted);
     heatmap.scale(scale);
     heatmap.size(size);

### 2017-12-07 paper/figures.tex
Add study result 2D heatmap
Better name for figure
Exp2 task figure change

@@
 		\includegraphics[width=0.4\columnwidth]{defend.png}
-		\caption{The prediction task. The participant has a list of locations, and ought to place their ships on locations with low probability of attack, and high certainty in this probability.}
+		\caption{The prediction task.
+The participant has a list of locations, and ought to place their ships on locations with low probability of attack, and high certainty in this probability.
+Ships above the heatmap have yet to be placed.}
 		\label{fig:taskTwoConditions}
@@
 		\includegraphics[width=0.75\columnwidth]{exp2histogram.pdf}
-		\caption{Overlaid histograms of the distribution of uncertainty across all participants guesses from the prediction experiment.
+		\caption{2D histogram of number of guesses from the prediction experiment for the cases when participants saw a traditional bivariate maps (left) and VSUPs (right).
 For each trial, participants made 5 guesses.

### 2017-12-08 paper/figures.tex
Larger figure
Added poll example

@@
 
-\newcommand{\taskTwoQuartiles}{
+\newcommand{\taskTwoHeatmap}{
 	\begin{figure}
 		\centering
-		\includegraphics[width=0.75\columnwidth]{exp2histogram.pdf}
-		\caption{2D histogram of number of guesses from the prediction experiment for the cases when participants saw a traditional bivariate maps (left) and VSUPs (right).
-For each trial, participants made 5 guesses.
-While participants generally favored low uncertainty in guesses, our stimuli design encouraged them to make at least some guesses with higher uncertainty.
-VSUPs encouraged more guesses with middling uncertainty, and discouraged highly uncertain guesses.}
-		\label{fig:taskTwoQuartiles}
+		\includegraphics[width=0.9\columnwidth]{exp2histogram.pdf}
+		\caption{2D histogram of placements from the prediction experiment.
+For each trial, participants were asked to place 5 tokens on heatmaps encoding both predicted safety, and uncertainty in these predictions.
+There were intentionally more tokens to place than there were ``safe and certain'' locations (top left corner).
+When looking at traditional 2D maps (left), participants favored safe but uncertain locations (bottom left corner).
+When looking at VSUPs (right), participants favored locations that were less safe, but more certain.)}
+		\label{fig:taskTwoHeatmap}
 	\end{figure}
@@
 Presidential Election, encoded in a traditional 2D bivariate map.
-Many states appear to be ``leaning'' towards Clinton, even when the margin of error of the polls is quite large.
-The blue-red color scale that is semantically appropriate for the topic is also not isoluminant, making it difficult to see just how much uncertainty there is in states with small leads for either candidate.}
+The redness and blueness show the polling lead for Trump and Clinton, respectively.
+But many polls had high margins of error, creating uncertainty about the election results.}
 		\label{fig:polling2d}
@@
 		\vspace{-15px}
-		\caption{Polling data prior to the 2016 U.S.
-Presidential Election, encoded with a VSUP.
-The larger amount of color bins for high-certainty regions shows that the leads for Hillary Clinton were, in many states, quite small.
-The collapsing of bins as uncertainty increases dissuades viewers from making confident predictions about states where the outcome was very much in doubt.
-}
+		\caption{Polling data prior to the 2016 U.S. Presidential Election, encoded with a VSUP. }
 		\label{fig:pollingVsum}

### 2017-12-08 paper/main.tex
Added poll example

@@
 
+\pollFig
+
 Using a public data set from Wikipedia~\cite{wiki}, we collected the statewide polls that were conducted closest to the 2016 U.S.
@@
 In particular, there is greater perceived value in avoiding large losses as opposed to striving for a large gain~\cite{kahneman1979prospect}.
+Our results from the prior study indicate that discrete, non-juxtaposed maps outperformed the other bivariate maps we selected, so we limited our study to only 4 types: square and wedge bivariate maps.
+While other map types might produce different patterns of decision-making (for instance, the juxtaposed map might encourage participants to ignore uncertainty information altogether), their low accuracy for our previous information fusion tasks led us to discard them, as it would be difficult to distinguish between different patterns of predictions caused by the design, and different patterns caused by simply misreading the heatmap.
+Having fewer conditions also afforded a within-subjects design that controlled for the variation in inter-personal differences in strategies and risk-aversion, while limiting the potential effects of learning and fatigue on predictions from large numbers of stimuli.
 
@@
  We recruited 24 participants for this task: $12$ female, $12$ male, ($M_{\text{age}}$= $37$, $SD_{\text{age}}$ = $9.8$).
-Our results from the prior study indicate that discrete, non-juxtaposed maps outperformed the other bivariate maps we selected, so we limited our study to only 4 types: square and wedge bivariate maps.
-These in turn were either VSUPs or traditional bivariate maps, for a 2 (square or wedge legend) x 2 (VSUP or standard) factorial design, with 6 replications, for a total of 24 stimuli.
+Our selected square and wedge maps were either VSUPs or traditional bivariate maps, for a 2 (square or wedge legend) x 2 (VSUP or standard) factorial design, with 6 replications, for a total of 24 stimuli.
 Prior to the main task, we included a short replication of the identification task from the prior experiment (with 12 stimuli) for training and exclusion purposes.
@@
 \subsubsection{Results} 
-\taskTwoQuartiles
+\taskTwoHeatmap
 
@@
 We found a significant difference ($D = 0.5$, $p=0.03$), and an inspection of the distribution shows that participants using VSUPs were less likely to make guesses with very high uncertainty.
-\figref{fig:taskTwoQuartiles} shows this result.
+\figref{fig:taskTwoHeatmap} shows this result.
 
@@
 
+%exp1
+%continuous bad
+%juxtaposed bad
+
+%exp2
+%vsup: more cautious
+
 The results from our initial evaluation of VSUPs are promising, showing that a general audience can make use of uncertainty information in a reasonable way.
@@
 VSUPs might therefore be used to promote more caution in judgments, and lead analysts away from spurious signals in data.
+
+% results to discuss:
+% bivariate maps: performance can be bad! esp. compared to univariate maps
+% juxtaposition bad
+% encourages caution
+% different patterns of risk, same people
 
@@
 VSUPs therefore act both as a kind of filtering device as well as a bivariate representation.
-This assumption holds in scenarios like the virology dataset (\figref{fig:pollingVsum]}), where the expectation is that highly uncertain data is unreliable or should otherwise be downweighted in analysis.
+This assumption holds in scenarios like the polling dataset (\figref{fig:pollingVsum}), where the expectation is that highly uncertain data is unreliable or should otherwise be downweighted in analysis.
 If the analyst has a different interpretation of uncertainty, or wishes to quickly and orthogonally analyze the distributions of uncertainty and value, other strategies, such as juxtaposed maps, may be more appropriate.

### 2017-12-09 paper/main.tex
Outlining for new discussion section.

@@
 
+Our experimental results have a number of implications for designers wishing to employ bivariate maps for encoding uncertainty.
+Both of our experiments show that the way that uncertainty information is presented can have a measurable impact of decision-making.
+VSUPs make a number of design choices in presenting uncertainty.
+Firstly, they present uncertainty information simultaneously (superimposed) with value, rather than either of the other two strategies for comparing data in visualizations~\cite{gleicher2011visual}, juxtaposition or superposition.
+Secondly, they encode value and uncertainty in discrete, quantized bins, rather than continuously.
+Lastly, they alias and suppress uncertain, rather than certain, values.
+Our experiments provide specific performance information about the effects of each of these choices on information fusion tasks.
+
+\subsection{Discrete vs Continuous Encodings}
+
+\subsection{Juxtaposition vs. Superposition}
+
+\subsection{Risk Aversion vs. Risk Seeking}
 %\sizeFig
@@
 
-The results from our initial evaluation of VSUPs are promising, showing that a general audience can make use of uncertainty information in a reasonable way.
+%The results from our initial evaluation of VSUPs are promising, showing that a general audience can make use of uncertainty information in a reasonable way.
 The way that uncertainty information is presented can have a measurable impact on decision-making.
@@
 
-Bivariate maps have an inherent limitation in that visual channels are often difficult to attend to separately or orthogonally.
+%Bivariate maps have an inherent limitation in that visual channels are often difficult to attend to separately or orthogonally.
 Bivariate maps are also complex to interpret.

### 2017-12-11 paper/main.tex
Fix small typos

@@
 
-We selected this task both as a way of assessing the ability of people to successfully encode and decode values using a bivariate map, and to provide training for the prediction task (since we re-used the bivariate maps in that task).
+We selected this task both as a way of assessing the ability of people to successfully encode and decode values using a bivariate map, and to provide training for the prediction task (since we reused the bivariate maps in that task).
 We measured performance both in terms of accuracy (did the participant select the correct point) as well as response time.
@@
 While other map types might produce different patterns of decision-making (for instance, the juxtaposed map might encourage participants to ignore uncertainty information altogether), their low accuracy for our previous information fusion tasks led us to discard them, as it would be difficult to distinguish between different patterns of predictions caused by the design, and different patterns caused by simply misreading the heatmap.
-Having fewer conditions also afforded a within-subjects design that controlled for the variation in inter-personal differences in strategies and risk-aversion, while limiting the potential effects of learning and fatigue on predictions from large numbers of stimuli.
+Having fewer conditions also afforded a within-subjects design that controlled for the variation in interpersonal differences in strategies and risk-aversion, while limiting the potential effects of learning and fatigue on predictions from large numbers of stimuli.
 
@@
 The results support our second hypothesis.
-A similiar ANOVA for average danger value of guesses found a significant effect of quantization scheme ($F(1,59)=17$, $p<0.01$).
+A similar ANOVA for average danger value of guesses found a significant effect of quantization scheme ($F(1,59)=17$, $p<0.01$).
 A post-hoc t-test confirmed ($t=2.3$, $p=0.02$) that guesses made with VSUPs had significantly higher average danger ($M=0.32$, $SD=0.17$) (and so lower expected value) than traditional maps ($M=0.29$, $SD=0.18$).

### 2017-12-11 study/src/script.js
Replace all VSUM with VSUP

@@ -46,30 +46,30 @@ var taskTwoTokens = 0;
 // 4. Continuous juxtaposed
 // 5. Discrete arc
 // 6. Continuous arc
-// 7. VSUM square
-// 8. VSUM arc
+// 7. VSUP square
+// 8. VSUP arc
 
 
 //The top types for exp2
 var types =
 [
-  {binned: "discrete", shape:"square", vsum:false},
-//  {binned: "continuous", shape:"square", vsum:false},
-//  {binned: "discrete", shape:"juxtaposed", vsum:false},
-//  {binned: "continuous", shape:"juxtaposed", vsum:false},
-  {binned: "discrete", shape:"arc", vsum:false},
-//  {binned: "continuous", shape:"arc", vsum:false},
-  {binned: "discrete", shape:"square", vsum:true},
-  {binned: "discrete", shape:"arc", vsum:true}
+  {binned: "discrete", shape:"square", vsup:false},
+//  {binned: "continuous", shape:"square", vsup:false},
+//  {binned: "discrete", shape:"juxtaposed", vsup:false},
+//  {binned: "continuous", shape:"juxtaposed", vsup:false},
+  {binned: "discrete", shape:"arc", vsup:false},
+//  {binned: "continuous", shape:"arc", vsup:false},
+  {binned: "discrete", shape:"square", vsup:true},
+  {binned: "discrete", shape:"arc", vsup:true}
 ];
 
 //quantization schema
-var Qvsum = bvu.quantization().branching(2).layers(4);
+var Qvsup = bvu.quantization().branching(2).layers(4);
 var Q2d = bvu.squareQuantization().n(4);
 
 
 //bivariate scales
-var scaleVsum = bvu.scale().quantize(Qvsum).range(d3.interpolateViridis);
+var scalevsup = bvu.scale().quantize(Qvsup).range(d3.interpolateViridis);
 var scale2d = bvu.scale().quantize(Q2d).range(d3.interpolateViridis);
 var scaleContinuous = function(d) {
   var c = d3.interpolateViridis(d.v);
@@ -109,10 +109,10 @@ var validQs = [
   {v:0.4, u:0.8},
 ];
 
-var vsumLegend = [];
-var temp = Qvsum.tree();
+var vsupLegend = [];
+var temp = Qvsup.tree();
 for (var i = 0;i < temp.length;i++) {
-  vsumLegend[i] = temp[temp.length - i - 1];
+  vsupLegend[i] = temp[temp.length - i - 1];
 }
 
[…]

### 2017-12-11 modules/src/heatmap.js
Replace all VSUM with VSUP

@@ -25,7 +25,7 @@ export function simpleHeatmap(data, m_scale, m_size, m_id, m_x, m_y) {
       heatmap.svgGroup = heatmap.el.append("g");
     }
 
-    heatmap.svgGroup.attr("transform", "translate(" + x + "," + y + ")");
+    heatmap.svgGroup.attr("transform", "translate(" + x + "," + y + ")").style("opacity", 0.6);
 
     heatmap.svgGroup.selectAll("g")
       .data(data)

### 2017-12-11 modules/src/heatmap.js
Undo opacity

@@ -25,7 +25,7 @@ export function simpleHeatmap(data, m_scale, m_size, m_id, m_x, m_y) {
       heatmap.svgGroup = heatmap.el.append("g");
     }
 
-    heatmap.svgGroup.attr("transform", "translate(" + x + "," + y + ")").style("opacity", 0.6);
+    heatmap.svgGroup.attr("transform", "translate(" + x + "," + y + ")");
 
     heatmap.svgGroup.selectAll("g")
       .data(data)

### 2017-12-11 modules/src/scale.js
Messed with polling example, set up for npm publishing.

@@ -38,7 +38,7 @@ export function simpleScale(m_mode, m_range, m_quantization) {
     switch (mode) {
       case "usl":
       default:
-        vcolor = d3.interpolateLab(vcolor, "#ddd")(uScale(data.u));
+        vcolor = d3.interpolateLab(vcolor, "#fff")(uScale(data.u));
       break;
     }
     return vcolor;

### 2017-12-11 paper/main.tex

@@
 We created two alternate heatmaps of this delay information.
-\figref{fig:airline2d} shows a traditional bivariate map, whereas \figref{fig:airlineVsum} shows a VSUP generated under the same constraints.
+\figref{fig:airline2d} shows a traditional bivariate map, whereas \figref{fig:airlineVsup} shows a VSUP generated under the same constraints.
 
@@
 
-%VSUPs (\figref{fig:viralVsum}) enable biologists to explore interesting patterns while deemphasizing noise.
+%VSUPs (\figref{fig:viralVsup}) enable biologists to explore interesting patterns while deemphasizing noise.
 Orange hot spots towards the beginning (the second column) and end of the genome (around reference nucleotide $9,500$) could be considered potential regions of interest in the 2D map; in the VSUP these are encoded as highly uncertain and of ambiguous value.
@@
 VSUPs therefore act both as a kind of filtering device as well as a bivariate representation.
-This assumption holds in scenarios like the polling dataset (\figref{fig:pollingVsum}), where the expectation is that highly uncertain data is unreliable or should otherwise be downweighted in analysis.
+This assumption holds in scenarios like the polling dataset (\figref{fig:pollingVsup}), where the expectation is that highly uncertain data is unreliable or should otherwise be downweighted in analysis.
 If the analyst has a different interpretation of uncertainty, or wishes to quickly and orthogonally analyze the distributions of uncertainty and value, other strategies, such as juxtaposed maps, may be more appropriate.

### 2017-12-12 paper/main.tex
Start of filling out discussion section mk2

@@
 VSUPs make a number of design choices in presenting uncertainty.
-Firstly, they present uncertainty information simultaneously (superimposed) with value, rather than either of the other two strategies for comparing data in visualizations~\cite{gleicher2011visual}, juxtaposition or superposition.
+Firstly, they present uncertainty information simultaneously (superimposed) with value, rather than in juxtaposed charts.
 Secondly, they encode value and uncertainty in discrete, quantized bins, rather than continuously.
@@
 
+\subsection{Juxtaposition vs. Superposition}
+
+Gleicher et al.~\cite{gleicher2011visual} propose three major categories for visualizations that support comparison: juxtaposition (where two variables are visualized in adjacent charts), superposition (where the two variables are encoded simultaneously in a single chart) and explicit encoding (where the two variables are reduced to a single calculated variable, which is directly visualized).
+Juxtaposition has a number of advantages for bivariate data of this sort: firstly, there are no issues of the integrality of different components of color perception, as there are separate maps for both uncertainty and value.
+Secondly, juxtaposed displays allow viewers to separately consider patterns in value and uncertainty.
+
 \subsection{Discrete vs Continuous Encodings}
 
-\subsection{Juxtaposition vs. Superposition}
+When using color to encode data, designers must make the decision whether or not to separate the data into discrete classes of color outputs, or to employ a continuous and class-less scale.
+This, in turn, introduces a trade-off between \emph{quantization error} (the sampling error introduced by binning the data) and \emph{perceptual error} (the error introduced by the viewer as they decode the color).
+For bivariate color maps, where often the color properties used to encode value and uncertainty are not perceptually separable, this perceptual error can be quite high.
+We observed this high error in our identification task, where continuous maps, despite, in theory, having greater expressivity in encoding color and uncertainty, had significantly worse accuracy than the discretized bivariate maps.
+This is despite the fact that we had comparatively few categories (16 in total).
 
 \subsection{Risk Aversion vs. Risk Seeking}
+
+It is often difficult for the general audience to successfully make use of uncertainty information.
+In order to minimize the complexity of uncertainty, uncertainty information is often visualized in a binary way, where values either meet some explicit uncertainty standard, or do not.
+These binary encodings for uncertain information, such as error bars for estimates, or ``cones of uncertainty'' for hurricane paths, result in binary patterns of decision-making that do not accurately reflect the underlying uncertainty in the data~\cite{cox2013visualizing,correll2014error}.
+
+We designed VSUPs to promote risk-averse decision-making.
+In the limit of uncertainty VSUPs have only one output color, and so explicitly prevent any decoding of value.
+This narrowing of options is reflected not only in the quantization scheme, but also in the shape of the legend, narrowing as the uncertainty increases.
 %\sizeFig

### 2017-12-22 paper/main.tex
Acknowledgments
Probably should add this section before I forget.
Added missing cite
More discussion writing.

@@
 \cite{boukhelifa2012evaluating}).
-Here, we focus on a combination of increasing luminance and decreasing saturation.
+Here, we focus on a combination of increasing luminance and decreasing saturation, as in Hengl \& Toomanian~\cite{hengl2006maps}.
 Correll \& Gleicher~\cite{correll2014error} provide evidence that even audiences without statistical training can successfully interpret uncertainty information encoded in these channels.
@@
 Our experimental results have a number of implications for designers wishing to employ bivariate maps for encoding uncertainty.
-Both of our experiments show that the way that uncertainty information is presented can have a measurable impact of decision-making.
+Both of our experiments show that the way that uncertainty information is presented can have a measurable impact on decision-making.
 VSUPs make a number of design choices in presenting uncertainty.
 Firstly, they present uncertainty information simultaneously (superimposed) with value, rather than in juxtaposed charts.
-Secondly, they encode value and uncertainty in discrete, quantized bins, rather than continuously.
-Lastly, they alias and suppress uncertain, rather than certain, values.
-Our experiments provide specific performance information about the effects of each of these choices on information fusion tasks.
+Secondly, they encode value and uncertainty via discrete, quantized bins, rather than continuously.
+Lastly, they are designed to alias and suppress uncertain values, rather than certain values.
+Our experiments provide evidence about the potential impact of each of these choices on performance for information fusion tasks.
+We will discuss the impact of our experimental findings on the consideration of each of these design decisions in turn.
 
@@
 Secondly, juxtaposed displays allow viewers to separately consider patterns in value and uncertainty.
+The literature also suggests that superposition, which requires a bivariate scale, results in poor accuracy even for basic identification tasks~\cite{wainer1980empirical}.
 
-\subsection{Discrete vs Continuous Encodings}
+However, our results indicate that juxtaposed visualizations have a substantial drawback for information fusion tasks.
+
+\subsection{Discrete vs. Continuous Encodings}
 
@@
 This, in turn, introduces a trade-off between \emph{quantization error} (the sampling error introduced by binning the data) and \emph{perceptual error} (the error introduced by the viewer as they decode the color).
-For bivariate color maps, where often the color properties used to encode value and uncertainty are not perceptually separable, this perceptual error can be quite high.
-We observed this high error in our identification task, where continuous maps, despite, in theory, having greater expressivity in encoding color and uncertainty, had significantly worse accuracy than the discretized bivariate maps.
-This is despite the fact that we had comparatively few categories (16 in total).
+There is also a cost in terms of expressivity and consistency: a designer might wish for quantitative, continuous information to be represented continuously~\cite{mackinlay1986automating}.
+
+However, for bivariate color maps, where often the color properties used to encode value and uncertainty are not perceptually separable, error in decoding values can be quite high.
+In our identification experiment, this perceptual error was high.
+Continuous maps performed the worst of all of our other bivariate map types (47\% vs.
+63\% accuracy for discrete maps).
+This is likely due to the non-linearity of
 
@@
 This narrowing of options is reflected not only in the quantization scheme, but also in the shape of the legend, narrowing as the uncertainty increases.
+This is despite the fact that the continuous maps were constructed via interpolations in CIELAB colorspace, which was designed for perceptual uniformity~\cite{mahy1994evaluation}.
 %\sizeFig
@@
 This decision of how to allocate visual variables promotes patterns of decision-making that make responsible use of uncertainty information, discouraging comparison of values in unreliable regions of the data, and promoting comparison in regions of high certainty.
+
+\section{Acknowledgments}
+
+This work was supported by a Moore Foundation Data-Driven Discovery Investigator Award.

### 2017-12-22 paper/main.tex
EOD 2/22

@@
 In our identification experiment, this perceptual error was high.
-Continuous maps performed the worst of all of our other bivariate map types (47\% vs.
+Continuous maps performed the worst of all of our other bivariate map types (47\% accuracy vs.
 63\% accuracy for discrete maps).
-This is likely due to the non-linearity of
+We believe that the main driver for this poor performance
 
@@
 This is despite the fact that the continuous maps were constructed via interpolations in CIELAB colorspace, which was designed for perceptual uniformity~\cite{mahy1994evaluation}.
+The integrality of the value and uncertainty channels
 %\sizeFig

### 2018-01-03 paper/main.tex
Fleshed out discussion section.

@@
 Other analysis tasks (such as filtering out outliers), require increased, rather than decreased, discriminability when uncertainty is high.
-However, for many information fusion tasks, the assumption is that uncertainty is related to data quality, or the uncertainty of data values~\cite{riveiro2007evaluation}.
+However, for many information fusion tasks, the assumption is that uncertainty is related to data quality, or the variability of data values~\cite{riveiro2007evaluation}.
 
@@
 
-Using a public data set from Wikipedia~\cite{wiki}, we collected the statewide polls that were conducted closest to the 2016 U.S.
+Using a public data set from Wikipedia~\cite{wiki} of the statewide polls that were conducted closest to the 2016 U.S.
 presidential election, we created a choropleth map showing which candidate was leading in the polls.
@@
 Given the relative volatility in polling data, this caution is warranted.
-An additional benefit to the VSUP map is that, as the diverging blue-red color map passes through a central whitepoint, it can be difficult to distinguish between uncertainty levels (which also interpolates to white) in the 2D map.
-The VSUP's tree-like structure separates these problematic colors: the closest colors in the 2D scale are 2.9 units apart in CIELab space, but the closest colors are 6.2 units apart in the VSUP scale.
 
@@
 	\item Participants would avoid targets with \textbf{high uncertainty} when using a VSUP.
-	\item This would result in a tradeoff where they would also choose targets with lower expected value \textbf{higher danger} when using a VSUP.
+	\item This would result in a tradeoff where they would also choose targets with \textbf{higher danger} when using a VSUP.
 \end{enumerate}
@@
 
-These two results, taken together, suggest that VSUPs led participants to moderate their guesses: avoiding highly uncertain regions, but willing to ``gamble'' more on regions of middling certainty.
+These two results, taken together, suggest that VSUPs led participants to moderate their guesses: avoiding highly uncertain regions, but willing to ``gamble'' more on regions of middling certainty (as in \figref{fig:taskTwoHeatmap}).
 In our design, where these gambles were usually poor (due to the predominance of ``bad'' cells), this moderation resulted in guesses with higher average danger.
@@
 However, our results indicate that juxtaposed visualizations have a substantial drawback for information fusion tasks.
+Integrating information from juxtaposed heatmaps is error-prone, as it requires searching for connections in two separate charts.
+In our experiments, the error introduced by juxtaposition outweighed the benefits of having simpler univariate scales rather than bivariate ones (accuracy of 58\% vs.
+51\% for superimposed and juxtaposed charts, respectively).
+In our experimental data, where value and uncertainty were not well-correlated, there were no reliable visual landmarks in common amongst the value and uncertainty charts, which may have made it difficult for participants to perform quick correspondence between value in one chart, and uncertainty in the other.
+Given this error-prone information fusion, we recommend either the use of superposition or explicit encoding for information fusion tasks involving uncertainty.
+At the very least, designers should use interaction (such as highlighting across charts) to make comparison in dense juxtaposed charts easier.
 
@@
 
-However, for bivariate color maps, where often the color properties used to encode value and uncertainty are not perceptually separable, error in decoding values can be quite high.
-In our identification experiment, this perceptual error was high.
+However, for bivariate color maps, where often the color channels for each variable are not perceptually separable, it can be difficult to decode the color into its component parts.
+In our identification experiment this perceptual decoding error was high.
 Continuous maps performed the worst of all of our other bivariate map types (47\% accuracy vs.
 63\% accuracy for discrete maps).
-We believe that the main driver for this poor performance
+Echoing the results of Padilla et al.~\cite{padilla2017evaluating}, we hold that the error introduced by quantization, which is known and bounded, has performance benefits even for continuous data, if the perceptual error is sufficiently high.
+Bivariate maps, where the integrality of channels raise this perceptual error, are a prime example of such a scenario.
+We advise designers to avoid continuous bivariate color scales when possible, and instead employ discrete scales with a relatively small number of categories.
 
-\subsection{Risk Aversion vs. Risk Seeking}
+\subsection{Risk Averse vs. Risk Seeking Decision-Making}
 
 It is often difficult for the general audience to successfully make use of uncertainty information.
-In order to minimize the complexity of uncertainty, uncertainty information is often visualized in a binary way, where values either meet some explicit uncertainty standard, or do not.
+In order to minimize the complexity of uncertainty, uncertain information is often visualized in a binary way, where values either meet some explicit uncertainty standard, or do not.
 These binary encodings for uncertain information, such as error bars for estimates, or ``cones of uncertainty'' for hurricane paths, result in binary patterns of decision-making that do not accurately reflect the underlying uncertainty in the data~\cite{cox2013visualizing,correll2014error}.
+On the other hand, presenting more detailed uncertainty information may overwhelm or confuse the viewer~\cite{greis2017designing}.
 
-We designed VSUPs to promote risk-averse decision-making.
-In the limit of uncertainty VSUPs have only one output color, and so explicitly prevent any decoding of value.
+VSUPs are in a middle ground between sparse but simple, and dense but complex presentations of uncertainty.
+At high levels of uncertainty, VSUPs have only one output color, and so explicitly prevent any decoding of value.
 This narrowing of options is reflected not only in the quantization scheme, but also in the shape of the legend, narrowing as the uncertainty increases.
-This is despite the fact that the continuous maps were constructed via interpolations in CIELAB colorspace, which was designed for perceptual uniformity~\cite{mahy1994evaluation}.
-The integrality of the value and uncertainty channels
+We selected this design to promote \emph{risk-averse} decision-making.
+That is, people should be hesitant to make strong predictions or decisions about values when uncertainty is high.
+Our prediction experiment, drawing from Kahneman \& Tversky's prospect theory~\cite{kahneman1979prospect}, was further designed to elicit avoidance of uncertain data, by framing the predictions in terms of minimizing losses, rather than maximizing gains.
+This framing was reflected in participants mostly making safe and certain guesses (low danger, high certainty).
+Where the differences in decision-making occurred were where the participants had exhausted the set of safe and certain locations, and were forced to value either certainty or expected value.
+A risk seeker would choose uncertain locations if the reward were high enough, whereas a risk-avoider would select locations with less uncertainty, even with a lower payoff.
+Participants using VSUPs were likelier to make these risk-avoiding decisions than when they were using traditional bivariate maps.
+Beyond VSUPs, this result indicates that the designers of visualizations can have fine-grained control over the types of decision-making their visualizations promote.
+Visualization designers should consider the not just fidelity to the underlying dataset, but also the utility of the decisions their designs support.
 %\sizeFig

### 2018-01-04 paper/figures.tex
Replace all VSUM with VSUP
Fleshed out discussion section.
More JH edits, prep for camera ready.

@@
 %Figure list:
-% Vsum vs. traditional 2D heatmap vs. juxtaposed maps
+% Vsup vs. traditional 2D heatmap vs. juxtaposed maps
 % Chart of color bins w/r/t CIELAB threshold, with iconic maps at intervals
@@
 		\caption{The 8 conditions from the identification experiment.
-Juxtaposed maps require a participants to make an error-prone connection between areas in two separate maps in order to make a decision that integrates value and uncertainty.
+Juxtaposed maps require participants to make an error-prone connection between areas in two separate maps in order to make a decision that integrates value and uncertainty.
 Traditional bivariate maps integrate both value and uncertainty.
@@
 \newcommand{\taskOneFig}{
-	\begin{figure}
+	\begin{figure}[h!]
 		\centering
@@
 	\centering
-	\includegraphics[width=\columnwidth]{airline-vsum}
+	\includegraphics[width=\columnwidth]{airline-vsup}
 	\vspace{-15px}
 	\caption{Flight delay data encoded with a VSUP.}
-	\label{fig:airlineVsum}
+	\label{fig:airlineVsup}
 \end{figure}
@@
 	\centering
-	\includegraphics[width=\columnwidth]{viral-vsum.pdf}
+	\includegraphics[width=\columnwidth]{viral-vsup.pdf}
 	\vspace{-15px}
@@
 }
-	\label{fig:viralVsum}
+	\label{fig:viralVsup}
 \end{figure}
@@
 		\centering
-		\includegraphics[width=\columnwidth]{polling-vsum.pdf}
+		\includegraphics[width=\columnwidth]{polling-vsup.pdf}
 		\vspace{-15px}
 		\caption{Polling data prior to the 2016 U.S. Presidential Election, encoded with a VSUP. }
-		\label{fig:pollingVsum}
+		\label{fig:pollingVsup}
 	\end{figure}

### 2018-01-04 paper/main.tex
More JH edits, prep for camera ready.

@@
 Our Value-Suppressing Uncertainty Palettes apply this insight to data with uncertainty.
-Frequently, uncertain information ought to be weighted less heavily in a decision-making process, or at least have less prominence in an analysis, than information with high certainty.
+Frequently, uncertain information ought to be weighted less heavily in a decision-making process, or have less prominence in an analysis, than information with high certainty.
 
@@
 
-JavaScript code for generating VSUPs for use in D3~\cite{bostock2011d3} charts is available at \url{https://github.com/uwdata/spatial-sampling}.
+JavaScript code for generating VSUPs for use in D3~\cite{bostock2011d3} charts is available at \url{https://github.com/uwdata/vsup}.
 
@@
 The AP style guide for reporters suggests that a candidate leading by more than two margins of error is said to be ``significantly ahead'' in the polls~\cite{french1988associated}.
-As the lead becomes closer to the margin of error, there is more doubt about the electoral outcome.
+As the lead becomes closer to the margin of error, there is more doubt about the electoral outcome, and smaller leads are less definitive.
 
@@
 
-Experimental materials, including data tables and stimuli generation code, are available at \url{https://github.com/uwdata/spatial-sampling}.
+Experimental materials, including data tables and stimuli generation code, are available at \url{https://github.com/uwdata/vsup-paper}.
 
@@
 The \emph{uncertainty} component was the confidence in this prediction.
-Other studies of uncertainty representation, such as in Cox et al.~\cite{cox2013visualizing}, have used "prediction + prediction uncertainty" stimuli to elicit differences in decision-making between visualizations of uncertainty.
+Other studies of uncertainty representation, such as in Cox et al.~\cite{cox2013visualizing}, have used ``prediction + prediction uncertainty'' stimuli to elicit differences in decision-making between visualizations of uncertainty.
 
@@
 While other map types might produce different patterns of decision-making (for instance, the juxtaposed map might encourage participants to ignore uncertainty information altogether), their low accuracy for our previous information fusion tasks led us to discard them, as it would be difficult to distinguish between different patterns of predictions caused by the design, and different patterns caused by simply misreading the heatmap.
-Having fewer conditions also afforded a within-subjects design that controlled for the variation in interpersonal differences in strategies and risk-aversion, while limiting the potential effects of learning and fatigue on predictions from large numbers of stimuli.
+Having fewer conditions also afforded a within-subjects design that controlled for the variation in interpersonal differences in strategies and risk-aversion, while limiting the potential effects of learning and fatigue from large numbers of stimuli.
 
@@
 Beyond VSUPs, this result indicates that the designers of visualizations can have fine-grained control over the types of decision-making their visualizations promote.
-Visualization designers should consider the not just fidelity to the underlying dataset, but also the utility of the decisions their designs support.
+Visualization designers should consider not just fidelity to the underlying dataset, but also the decision strategies their designs support.
 %\sizeFig
@@
 
-VSUPs assume that uncertain values ought to have less visual impact that highly certain values.
+VSUPs assume that uncertain values ought to have less visual impact than highly certain values.
 VSUPs therefore act both as a kind of filtering device as well as a bivariate representation.
